### 2016-07-27 definitions.tex
opening TACAS17 submission

\documentclass{article}

\usepackage{lipsum}
\usepackage{titlesec}
\usepackage{import}
\usepackage{tikz}
\usepackage{xspace}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{listings}
\usepackage{url}
\usepackage{galois}
\usepackage{color} 
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wasysym} 
\usepackage{graphicx}

\begin{document}

\title{CDCL lifted to Programs}
\vspace{0.3cm}
%\author{Rajdeep Mukherjee}
\date{}
\maketitle
\pagenumbering{arabic}

\maketitle
Let us consider a program $\mathcal{P}$. Let $\mathcal{P}$ be translated 
to a Static Single Assignment (SSA) representation. These SSA representations 
can be of type equality ($E$), constraints ($C$) or assertions ($A$). 
Let us consider an example of a program and its corresponding SSA constraints as
shown in Figure~\ref{figure:ssa}.

\begin{figure}[htbp]
\begin{tabular}{l|l}
\hline
Program & SSA \\
\hline
\begin{lstlisting}[mathescape=true,language=C]
void main()
{
 int x;
 if(x==0 || x==1)
 {
   x = x * 2; 
   assert(x != 1);
 }
}
\end{lstlisting}
&
\begin{lstlisting}[mathescape=true,language=C]

(E) x#16 == nondet_symbol(ssa::nondet16.1)
(E) cond#17 == (!(x#16 == 0) && !(x#16 == 1))
(E) x#18 == 2 * x#16
(E) guard#18 == (!cond#17 && guard#0)
(A) !(x#18 == 1) || !guard#18
(E) x#phi20 == (guard#18 ? x#18 : x#16)
(E) guard#20 == (cond#17 && guard#0 || guard#18)
\end{lstlisting}
\\
\hline
\end{tabular}
\caption{Translation of C program into SSA}
\label{figure:ssa}
\end{figure}

\textbf{CDCL view:}
Given a set of formula $Form$ and a set of structure $Struct$, 
a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
\times Form)$, which specifies which structure satisfies which formula. A 
concrete domain the power set of structures which gives us lattice. 
We define two transformers for a CDCL solver. \\
\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:\\
$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\

\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
that contradict a formula $\varphi$. This is formally shows as follows:\\
$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
\\

\textbf{Lifting CDCL to reason about Programs:}
Treat program as a logical formula. Give a programs $P$ and a trace $\pi$, \\
$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
$P$.

\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
set of traces generated by the program and is erroneous. This can be
characterised in two different ways. We can start with the initial state and
compute least fix-point with strongest post-condition or start from error state 
and compute the greatest fix-point with weakest pre-condition.

\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
set of program traces that are safe, that is, all program traces that are not
generated by the program.

\textit{Interaction between amod and aconf:} If the "search" for finding an 
erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
"proof" stage, we get back refinement of strongest post-condition which keeps in 
mind satisfiability information.

\textbf{Program representation as CNF:}
A CNF formulation of a program is a conjunction of all 
SSA constriants of type $E$, $C$ and negation of $A$.

\textbf{Literal:} A Literal is a meet irreducible which are minimum complementable
elements which specify that for certain program variable, there is a certain
bound. An example of literal is $x \leq 0$.

Let us define the notion of literals with respect to interval domain. 
An interval $([l:u])$ is defined by a lower bound $(l)$ and a upper bound $(u)$.
Let I and J be two intervals. Then we say $I \leq J$ or (I leq J) if the following
holds. \\
$(I.u \leq J.u \wedge I.l \geq J.l)$

We say I disjoint J if the following holds.\\
$(I.u < J.l \vee  I.l > J.u)$

Let x and x' be the intervals of a variable $p$ corresponding
to the present abstract value and the current clause respectively.

\textbf{Satisfiable literal:} 
We say x' is satisfiable literal if the following holds: \\
$(x \leq x')$

\textbf{Unsatisfiable literal:}
We say x' is satisfiable literal if the following holds: \\
$!(x \leq x') \wedge$ $!$(x disjoint x')

\textbf{Contradicting Literal:}
We say x' is contradicting literal if the following holds: \\
$!(x \leq x') \wedge$ (x disjoint x')

\textbf{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.

\textbf{Satisfiable clause:}
If at least one meet irreducible is satisfiable, then a clause is said to 
be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
abstract value be $x \in [0,3]$, then $C$ is a satisfiable clause.

\textbf{Unsatisfiable clause:}
If no meet irreducibles in a clause are satisfiable or some meet
irreducible are not satisfiable and the rest are contradicting, we call a 
clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
Let the abstract value be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
then $C$ is unsatisfiable.

\textbf{Conflicting clause:}
If all the meet irreducibles in a clause are contradicting, then the clause is said to
be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
abstract value contain the value of $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[17,32]$, then $C$ is a conflicting clause.

\textbf{Unit clause:}
If all meet irreducible but one is contradicting in a clause, we call the clause
to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
abstract value contain the value of $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[10,12]$, then $C$ is a unit clause where the unit literal is $z$. 
After we apply the unit rule, the clause becomes $C=(x<4 \vee y>10 \vee x < 9)$, 
and now C is satisfying.

\textbf{Deductions:} A deduction is an application of unit rule. A 
unit rule is the best abstract transformer (strongest post-condition or 
weakest pre-condition). Boolean constraint propagation (BCP) is the 
repeated application of unit rule. This corresponds to computing the 
greatest fixed point

\textbf{Characteristics of Conflict clause}
\begin{enumerate}
\item A conflict clause must include asserting cuts. An asserting cut is a cut
that contain exactly one node at the current decision level. Assertion cuts yields 
clauses that can be used to derive new information after backtracking.

\item A conflict clause must be UNIT after backtracking.

\item There can be multiple cuts and hence multiple UIPs. In other words, there
can be multiple incomparable reasons for a conflict. But conflict analysis
procedure choses one that is asserting.

\item The conflict clause should be made false by the current partial assignment
and thus exclude an assignment leading to conflict. 
\end{enumerate}

\textbf{Backjumping}
The backjumping level is defined by the literal of the conflict clause assigned
at the level that is the closest to the conflict one. In other words, the
backjumping level is the level closest to the root (decision level 0)  where the
conflict clause is still unit. If a conflict clause is "globally"  unit, then
the backjumping level is the root of the search tree. The root here means
decision level 0.

\end{document}

### 2016-07-27 invariant.tex
opening TACAS17 submission

\documentclass[a4paper,conference]{llncs}

\usepackage{colortbl}

\newcommand{\dkcmt}[1]{{\color{red} [{#1}]}}
\newcommand{\rmcmt}[1]{{\color{magenta} [{#1}]}}
\newcommand{\sjcmt}[1]{{\color{blue} [{#1}]}}
\newcommand{\tmcmt}[1]{{\color{orange} [{#1}]}}
\usepackage{xspace}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,positioning,decorations.markings,chains,fit,shapes.multipart}
\usepackage{import}
\usepackage{times}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathptm}
\usepackage{rotating}
\usepackage{color}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{alltt}
\usepackage{mdwlist}
\usepackage{booktabs}
\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
\usepackage{hyperref}
\makeatother

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\hector}{\tool{hector}}
\newcommand{\slec}{\tool{slec}}
\newcommand{\hwcbmcv}{\tool{hw-cbmc}}
\newcommand{\hwcbmcver}{\tool{hw-cbmc 5.4}}
\newcommand{\verifoxver}{\tool{VerifOx 0.1}}
\newcommand{\verifox}{\tool{VerifOx}}
\newcommand{\klee}{\tool{klee}}
\newcommand{\kleever}{\tool{klee 1.0.0}}
\newcommand{\symbiotic}{\tool{symbiotic}}

\newcommand{\Omit}[1]{}
\newcommand{\scs}[1]{\scriptsize{#1}}

\input{symbols}
\input{refs}

\begin{document}

\title{Invariant Generation using ACDL}
\author{}
\institute{}

\maketitle

\section{Algorithm for Invariant Generation}
\Omit{
We propose a two-solver approach for proving programs with loops using ACDL. 
Let us call these solvers: \em{proof solver} (PS) and \em{invariant solver} (IS).
The purpose of proof solver is to determine the safety of the program in the
presence of the negation of assertion. Whereas, the purpose of invariant solver is 
to determine the inductivenss of the candidate invariant produced by the proof 
solver in the presence of the actual assertion.
}
We present a program verification algorithm, ACDCL-P, inspired 
from the propositional-level CDCL algorithm. ACDCL-P generates a candidate 
invariant following decisions and deductions. The inductiveness of the 
candidate invariants are then checked using an Abstract CDCL solver, ACDCL. ACDCL 
solver returns SAT if the input formula is satisfiable and UNSAT otherwise.

\section{Communication between ACDCL and ACDCL-P}
The deduction information in the trail corresponding to 
decisions that participate in the invariant can be 
communicated from ACDCL-P to ACDCL.

\subsection{Highlights of ACDCL-P}
\begin{enumerate}
\item decision heuristics on cond variables enable us to prove many 
programs without actually deciding on the actual program variables. This 
leads to significantly less number of decisions and less number of conflicts 
compared to the SAT solver. Further, decisions on conditional program 
variables can be seen as simulating path-based symbolic execution. But when 
a decision on conditional varaible leads to BOTTOM, ACDCL-P tried to learn 
the reason for failure, and hence can also be seen similar to IMPACT like
algorithm. 
\item Last UIP allows us to learn clauses that has more information in it. 
\item Compared to pure AI based tools, ACDCL-P performs program and property 
driven partitioning with simpler domains to prove programs with relational 
properties (branch7). AI based tool would require richer domains like 
relational (equality) domain to prove such programs. 
\item Compared to cdfpl, our tool proves more programs correctly and works on 
non-numeric programs as well.
\item Chaotic propagations help ACDCL-P to generate stronger deductions and reach 
fixpoint early in propagation stage.
\item Compared to cdfpl, acdl solver is agnostic to abstract domain. cdfpl
stores the value of variables (from interval domain) in every decision 
and deduction stage in a container "value". So, cdfpl keeps track of literals 
and their intervals at every stage which makes it difficult to extend it to other 
richer domains. However, ACDCL-P treat literals as meet irreducibles. Hence, 
ACDCL-P can be easily extended to other richer domains.    
\item ACDCL-P handles loops by automatic invariant generation.
\end{enumerate}

\subsection{Difference between cdfpl and ACDCL-P}
\begin{enumerate}
\item cdfpl stores deductions from conflict clause into the trail. It does not 
store the deductions from the AI phase into the propagation trail. Hence, each 
entry in the trail has a corresponding reason (conflict clause responsible for 
deducing the literal in the entry). Thus, it is easier to compute the UIP in
cdfpl since it retains only the clause based deduction information in the trail. 
Thus, it can perform resolution proof easily to compute the UIP. The resolution 
proof in cdfpl is implemented in a way that it iterates through the marked
portion of the trail (contradicted portion in the current decision segment of
the trail with respect to the conflict clause) to find the corresponding clause 
(predecessor clause) for each marked entry in the DL segment of the trail 
(Note that the predecessor clause is the one which inferred this literal in the 
trail). The predecessor clause is then iterated to find if a literal in the 
clause is contradicted in the current decision level or previous decision level 
and accordingly the learnt clause and the backtrack level is updated.

\item Whereas, ACDCL-P stores the deduction information from the AI phase into the 
trail. In such case, the entries in the trail does not have a corresponding
reason clause. Thus, it is not possible to perform resolution proof to find the 
UIP since the propagation trail does not store such information.   
\end{enumerate}

\textbf{Lemma 1:}Normalization is an equivalent transformation. \\ 
\textbf{Lemma 2:}Normalizing a learnt clause is UNIT preserving. \\
Proof: Prove that the above lemma holds both for monotonic ($x \leq 2 \land x \leq 1 
\land x \leq 0)$ and non-monotonic decisions $(x \leq 20 \land x \geq 10 \land
x \leq 15 \land x \leq 13)$. \\
\textbf{Lemma 3:}Normalizing a learnt clause does not affect the backtrack
level. \\
\begin{enumerate}
\item Step 1: Construct the learnt clause, L=${\neg{d1} \lor \neg{d2} \lor \neg{UIP}}$
\item Step 2: Normalize the learnt clause without UIP, normalize L'=${\neg{d1} \lor \neg{d2}}$
\item Step 3: Add the UIP to the normalized learnt clause, L\_final=L' $\lor \neg{UIP}$
\item Step 4: Backtrack to blevel=$max\{d1, d2\}$
\item Step 5: Let $val$ be the current partial assignment after backtracking
\item Step 6: Normalize the current partial assignment, $val$
\item Step 7: Check unit rule for (L\_final, val) -- The learnt clause should now be UNIT !
\end{enumerate}

\subsection{Some ideas}
\begin{enumerate}
\item Can we infer the smallest UNSAT core when we reach BOTTOM 
as the potential candidate for Invariant. If the UNSAT core for 
a proof is indeed an invariant, then it will be UNSAT core for other proofs 
as well. 
\end{enumerate}

\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{safe}{safe}
\SetKwData{unsafe}{unsafe}
\SetKwData{sat}{sat}
\SetKwData{unsat}{unsat}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{KwNot}{not}
\begin{small}
\Input{Formula $\mathcal{F}$}
\Output{The status (\sat or \unsat or \unknown)}
%\tcc{The initial state}
%$ACDL \leftarrow \tuple{\neg(assert(c)),loopselect=nondet}$ \nllabel{algline:initPS}\;
\assign result=\deduce(F)\;
\uIf{$result == \perp$} {
  \return \unsat \;
}
\uElse
{
  \uIf{is\_complete()} {
    \return \sat \;
  }
}
\While{$true$} 
{
  \assign decision\_variable = decision\_heuristics($F$)\;
  \assign decision\_container = decision\_container $\cup$ decision\_variable\; 
  \uIf{decision\_variable == NULL}
  {
    return \unknown\; 
  }
  \uElse
  {
    \assign result=\deduce(F)\;
    \uIf{result == UNKNOWN}
    {
      \uIf{is\_complete()} {
        \return \sat \;
      }
      \uElse 
      {
        \continue\;
      }
    }
    \uElse
    {
      \Do{result == PASS} {
        \uIf{$\neg$ conflict\_analysis()} {
          return \unsat \;  
        }
        \assign result=\deduce(F)\;
      }
    }
  }
}
\end{small}
\caption{ACDCL Solver \label{Alg:acdl}}
\end{algorithm2e}

\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{safe}{safe}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{KwNot}{not}
\begin{small}
\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
%\tcc{The initial state}
$ACDCL-P \leftarrow \tuple{\neg(assert(c)),loopselect=nondet}$ \nllabel{algline:initPS}\;
\assign result=\deduce(P)\;
\uIf{$result == \perp$} {
  \return \safe \;
}
\uElse
{
  \uIf{is\_complete()} {
    \return \unsafe \;
  }
}
\While{$true$} 
{
  \assign decision\_variable = decision\_heuristics($P$)\;
  \assign decision\_container = decision\_container $\cup$ decision\_variable\; 
  \uIf{decision\_variable == NULL}
  {
    return \unknown\; 
  }
  \uElse
  {
    \assign result=\deduce(P)\;
    \uIf{result == UNKNOWN}
    {
      \uIf{is\_complete()} {
        \return \unsafe \;
      }
      \uElse 
      {
        \continue\;
      }
    }
    \uElse
    {
      \Do{result == PASS} {
        \uIf {loop\_variable $\in$ decision\_container} {
          \assign $Inv=(\bigcap_{i=1}^{n} d_{i} \rightarrow
              \bigcap_{j=1}^{m} l_{j})$ where $d_i$ $\in$ \{non\_loop\_variable 
            $\cup$ loop\_variable\}, $l_j$ $\in$ \{loop\_variable\} \nllabel{algline:inv}\;
            \assign Inv\_formula = $Inv$ $\wedge$ $P$ $\wedge$ $Inv'$, where $Inv'$
            is the invariant after the application of transition system\;
        $inv\_result \leftarrow$ run($ACDCL$, $Inv\_formula$) \; 
        \uIf{inv\_result == UNSAT \nllabel{algline:statusIS}} {
          \learn $Inv$ \;
        }
        }
        \uIf{$\neg$ conflict\_analysis()} {
          return \safe \;  
        }
        \assign result=\deduce(P)\;
      }
    }
  }
}
\end{small}
\caption{ACDCL-P Verifier with Invariant Checking\label{Alg:verifox}}
\end{algorithm2e}

\end{document}

### 2016-08-06 invariant.tex
gamma completeness and spurious check algorithm

@@
 
+\begin{algorithm2e}[t]
+\DontPrintSemicolon
+\SetKwFunction{decide}{Decide}
+\SetKwFunction{deduce}{deduce}
+\SetKwFunction{backtrack}{backtrack}
+\SetKwFunction{print}{print}
+\SetKwFunction{return}{return}
+\SetKwFunction{nondet}{nondet}
+\SetKwFunction{run}{run}
+\SetKwFunction{learn}{learn}
+\SetKwFunction{continue}{continue}
+\SetKwFunction{assign}{assign}
+\SetKwRepeat{Do}{do}{while}
+%\SetKwFunction{assume}{assume}
+%\SetKwFunction{isf}{isFeasible}
+\SetKwData{safe}{safe}
+\SetKwData{unsafe}{unsafe}
+\SetKwData{true}{true}
+\SetKwData{false}{false}
+\SetKwInOut{Input}{input}
+\SetKwInOut{Output}{output}
+\SetKw{KwNot}{not}
+\begin{small}
+\Input{A counterexample trace $t$ of Program $P$}
+\Output{The status (\true) if $t$ is not spurious and $t$ is complete 
+or (\false) if $t$ is spurious or not complete}
+\uIf{loopselect $\in$ t} {
+   \uIf{loopselect==false} {
+      \uIf{gamma\_complete(t)} {
+        return \true \;
+      }
+      \uElse 
+      {
+        return \false \;
+      }
+   }
+   \uElse {
+     return \false \;
+   }
+}
+\end{small}
+\caption{Spurious Counterexample Checking \label{Alg:spurious-check}}
+\end{algorithm2e}
+
+\begin{algorithm2e}[t]
+\DontPrintSemicolon
+\SetKwFunction{decide}{Decide}
+\SetKwFunction{deduce}{deduce}
+\SetKwFunction{backtrack}{backtrack}
+\SetKwFunction{print}{print}
+\SetKwFunction{return}{return}
+\SetKwFunction{nondet}{nondet}
+\SetKwFunction{run}{run}
+\SetKwFunction{learn}{learn}
+\SetKwFunction{continue}{continue}
+\SetKwFunction{assign}{assign}
+\SetKwRepeat{Do}{do}{while}
+\SetKwData{safe}{safe}
+\SetKwData{unsafe}{unsafe}
+\SetKwData{true}{true}
+\SetKwData{false}{false}
+\SetKwInOut{Input}{input}
+\SetKwInOut{Output}{output}
+\SetKw{KwNot}{not}
+\begin{small}
+\Input{A counterexample trace $t$ of Program $P$}
+\Output{The status (\true) if $t$ is complete, else \false if $t$ is not complete} 
+\For{elements $e \in t$} {
+  \uIf{$e$ is non-singleton} {
+    return \false \;  
+  }
+}
+return \true \;  
+\end{small}
+\caption{Gamma Completeness Check \label{Alg:gamma-complete}}
+\end{algorithm2e}
+
 \end{document}

### 2016-08-06 invariant.tex
minor algo modifications

@@
 \input{symbols}
-\input{refs}
 
@@
 %$ACDL \leftarrow \tuple{\neg(assert(c)),loopselect=nondet}$ \nllabel{algline:initPS}\;
-\assign result=\deduce(F)\;
+result=\deduce(F)\;
 \uIf{$result == \perp$} {
@@
 {
-  \uIf{is\_complete()} {
+  $t$ = get\_counterexample\_trace\_from\_P \;
+  \uIf{spurious\_check(t)}
+  {
     \return \sat \;
@@
 {
-  \assign decision\_variable = decision\_heuristics($F$)\;
-  \assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
+  meet\_irreducible = decision\_heuristics($F$)\;
+  decision\_container = decision\_container $\cup$ meet\_irreducible\; 
+  \uIf{meet\_irreducible == NULL}
   {
@@
     {
-      \uIf{is\_complete()} {
+      $t$ = get\_counterexample\_trace\_from\_P \;
+      \uIf{spurious\_check(t)}
+      {
         \return \sat \;
@@
 $ACDCL-P \leftarrow \tuple{\neg(assert(c)),loopselect=nondet}$ \nllabel{algline:initPS}\;
-\assign result=\deduce(P)\;
+result=\deduce(P)\;
 \uIf{$result == \perp$} {
@@
 {
-  \uIf{is\_complete()} {
+  $t$ = get\_counterexample\_trace\_from\_P \;
+  \uIf{spurious\_check(t)} {
     \return \unsafe \;
@@
 {
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  \assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
+   meet\_irreducible = decision\_heuristics($P$)\;
+   decision\_container = decision\_container $\cup$ meet\_irreducible \; 
+  \uIf{meet\_irreducible == NULL}
   {
@@
   {
-    \assign result=\deduce(P)\;
+    result=\deduce(P)\;
     \uIf{result == UNKNOWN}
     {
-      \uIf{is\_complete()} {
+      $t$ = get\_counterexample\_trace\_from\_P \;
+      \uIf{spurious\_check(t)} {
         \return \unsafe \;
@@
         \uIf {loop\_variable $\in$ decision\_container} {
-          \assign $Inv=(\bigcap_{i=1}^{n} d_{i} \rightarrow
+              $Inv=(\bigcap_{i=1}^{n} d_{i} \rightarrow
               \bigcap_{j=1}^{m} l_{j})$ where $d_i$ $\in$ \{non\_loop\_variable 
             $\cup$ loop\_variable\}, $l_j$ $\in$ \{loop\_variable\} \nllabel{algline:inv}\;
-            \assign Inv\_formula = $Inv$ $\wedge$ $P$ $\wedge$ $Inv'$, where $Inv'$
+             Inv\_formula = $Inv$ $\wedge$ $P$ $\wedge$ $\neg{Inv'}$, where $Inv'$
             is the invariant after the application of transition system\;
@@
         }
-        \assign result=\deduce(P)\;
+        result=\deduce(P)\;
       }
@@
 \end{small}
-\caption{Spurious Counterexample Checking \label{Alg:spurious-check}}
+\caption{Spurious Counterexample Check \label{Alg:spurious-check}}
 \end{algorithm2e}

### 2016-08-07 invariant.tex
highlights of ACDCL

@@
 ACDCL-P can be easily extended to other richer domains.    
+\item  ACDL computes learnt clause with UIP for octagon domains -- first time 
+extension of learning in ACDL to richer relational domains.   
 \item ACDCL-P handles loops by automatic invariant generation.
@@
 UIP since the propagation trail does not store such information.
+
+\item unitness is guaranteed in propagation in CDCL — that is all deduction are 
+made through unit clause. Whereas, unitness is not guaranteed in propagation in 
+ACDCL — that is deduction in AI phase are not made through unit clause, but deductions
+made through conflict clause are only unit. Thus, resolution proof is not possible.
+
 \end{enumerate}
@@
 \Output{The status (\sat or \unsat or \unknown)}
-%\tcc{The initial state}
-%$ACDL \leftarrow \tuple{\neg(assert(c)),loopselect=nondet}$ \nllabel{algline:initPS}\;
 result=\deduce(F)\;
@@
 {
-  $t$ = get\_counterexample\_trace\_from\_P \;
+  $t$ = get\_satisfying\_assignment($F$) \;
   \uIf{spurious\_check(t)}
@@
     {
-      $t$ = get\_counterexample\_trace\_from\_P \;
+      $t$ = get\_satisfying\_assignment($F$) \;
       \uIf{spurious\_check(t)}

### 2016-09-01 paper.tex
initial draft of TACAS17

\documentclass[a4paper,conference]{llncs}

\usepackage{import}
\usepackage{times}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{mathptm}
\usepackage{rotating}
\usepackage{color}
\usepackage{listings}
\usepackage{verbatim}
\usepackage{comment}
\usepackage{alltt}
\usepackage{graphicx}
\usepackage{psfrag}
\usepackage{epsfig}
\usepackage{wasysym} 
\usepackage{subfigure}
\usepackage{paralist}
\usepackage{mdwlist}
\usepackage{multirow}
\usepackage{dingbat}
% \usepackage{hyperref}
\usepackage{colortbl}
\usepackage{booktabs}
\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
\usepackage{hyperref}
\usepackage{dcolumn}
%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
\makeatother

\newcommand{\tool}[1]{\textsc{#1}\xspace}
\newcommand{\symex}{\tool{path-symex}}
\newcommand{\cbmcv}{\tool{cbmc 5.0}}
\newcommand{\Omit}[1]{}

\input{symbols}

\begin{document}

\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
\maketitle
\begin{abstract}
%
{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
for abstract reasoning using ACDCL.

%
\end{abstract}

\section{Introduction}

\section{Lifting CDCL to Programs}
The conflict driven clause learning algorithm can be seen as an 
alternation between over-approximate deduction with boolean 
constraint propagation, and under-approximate abduction, with 
conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
shows that satisfiability solvers are static analyzers. They present 
a framework that strictly generalizes CDCL algorithm to lattices and 
abstract transformers. \\
Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
However, we extend the framework to use it as a backend reasoning 
engine for high-level reasoning of hardwares circuits represented as programs.
To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
$P$. CDCL algorithm has two main phases which feedback information to each other --
{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
satisfying assignment to the input formula. If the search fails, meaning that the current 
partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
reason for such failure to block the model search from entering into same conflicting 
search-space in future. More formally, given a set of formula $Form$ and a set of 
structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
\times Form)$, which specifies which structure satisfies which formula. A 
concrete domain is the power set of structures which gives us lattice. 
We define two transformers for a CDCL solver. \\
\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\

\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
that contradict a formula $\varphi$. This is formally shows as follows:
$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
\\
We now define abstract interpretation analogues to model search and conflict analysis.   \\
\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
set of traces generated by the program and is erroneous. This can be
characterised in two different ways. We can start with the initial state and
compute least fix-point with strongest post-condition or start from error state 
and compute the greatest fix-point with weakest pre-condition. \\

\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
set of program traces that are safe, that is, all program traces that are not
generated by the program. \\

\textit{Interaction between amod and aconf:} If the search for finding an 
erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
proof stage, we get back refinement of strongest post-condition which keeps in 
mind satisfiability information.

Most lattice structure used in static analysis have meet and join operations 
but lack negations. The meet operation precisely model conjunction, and
the join operation over-approximate disjunction. Thus, the precision loss in 
static analyzers is often due to imprecise join operations. The loss in precision is
overcome by using richer abstract domains or equipping the analysis with disjunction --
which is often very expensive since the analysis exhibit case enumeration behavior.

On the other hand, the combination of precise conjunction in partial assignments domain 
with learning allow CDCL solvers to reason about disjunction without 
enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
the analysis and prevent enumeration behavior.

\section{Definitions}
\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
elements which specify that for certain program variable, there is a certain
bound. An example of literal is $x \leq 0$.

Let us define the notion of literals with respect to interval domain. 
An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
Let I and J be two intervals. The relation between I and J is defined as follows:\\
We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.

Let $x$, $x'$ be the intervals of a variable $p$ in current 
partial assignment and the current clause respectively. 
We define a literal and a clause to be satisfiable or unsatisfiable 
or contradicting with respect to an interval abstraction as follows. \\
\textit{Satisfiable literal:} 
We say $x'$ is satisfiable literal if $(x \leq x')$.

\textit{Unsatisfiable literal:}
We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).

\textit{Contradicting Literal:}
We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).

\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.

\textit{Satisfiable clause:}
If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.

\textit{Conflicting clause:}
If all the meet irreducibles in a clause are contradicting, then the clause is said to
be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[17,32]$, then $C$ is a conflicting clause.

\textit{Unsatisfiable clause:}
If no meet irreducibles in a clause are satisfiable or some meet
irreducibles are not satisfiable and the rest are contradicting, we call a 
clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
then $C$ is a unsatisfiable clause.

\textit{Unit clause:}
If all meet irreducible but one is contradicting in a clause, we call the clause
to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[10,12]$, then $C$ is a unit clause where the unit literal is $z$.

\textit{Unit Rule:} A unit rule is the best abstract transformer 
(strongest post-condition or weakest pre-condition).

\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
application of unit rule. This corresponds to computing the greatest fixed point.

\textit{Characteristics of Conflict clause}
\begin{enumerate}
\item A conflict clause must include asserting cuts. An asserting cut is a cut
that contain exactly one node at the current decision level. Assertion cuts yields 
clauses that can be used to derive new information after backtracking.

\item A conflict clause must be UNIT after backtracking.

%\item There can be multiple cuts and hence multiple UIPs. In other words, there
%can be multiple incomparable reasons for a conflict. But conflict analysis
%procedure choses one that is asserting.

%\item The conflict clause should be made false by the current partial assignment
%and thus exclude an assignment leading to conflict. 
\end{enumerate}

\textit{Backjumping:}
The backjumping level is defined by the literal of the conflict clause assigned
at the level that is the closest to the conflict one. In other words, the
backjumping level is the level closest to the root (decision level 0)  where the
conflict clause is still unit. If a conflict clause is "globally"  unit, then
the backjumping level is the root of the search tree.

\section{Abstract Conflict Driven Clause Learning}
\begin{algorithm2e}[t]
\DontPrintSemicolon
\SetKwFunction{decide}{Decide}
\SetKwFunction{deduce}{deduce}
\SetKwFunction{backtrack}{backtrack}
\SetKwFunction{print}{print}
\SetKwFunction{return}{return}
\SetKwFunction{nondet}{nondet}
\SetKwFunction{run}{run}
\SetKwFunction{learn}{learn}
\SetKwFunction{continue}{continue}
\SetKwFunction{assign}{assign}
\SetKwRepeat{Do}{do}{while}
%\SetKwFunction{assume}{assume}
%\SetKwFunction{isf}{isFeasible}
\SetKwData{safe}{safe}
\SetKwData{unsafe}{unsafe}
\SetKwData{unknown}{unknown}
\SetKwData{true}{true}
\SetKwInOut{Input}{input}
\SetKwInOut{Output}{output}
\SetKw{KwNot}{not}
\begin{small}
\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
\assign result=\deduce(P)\;
\uIf{$result == \perp$} {
  \return \safe \;
}
\uElse
{
  \uIf{is\_gamma\_complete()} {
    \return \unsafe \;
  }
}
\While{$true$} 
{
  \assign decision\_variable = decision\_heuristics($P$)\;
  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
  \uIf{decision\_variable == NULL}
  {
    return \unknown\; 
  }
  \uElse
  {
    \assign result=\deduce(P)\;
    \uIf{result == UNKNOWN}
    {
      \uIf{is\_gamma\_complete()} {
        \return \unsafe \;
      }
      \uElse 
      {
        \continue\;
      }
    }
    \uElse
    {
      \Do{$result == \perp$} {
        \uIf{$\neg$ conflict\_analysis()} {
          return \safe \;  
        }
        \assign result=\deduce(P)\;
      }
    }
  }
}
\end{small}
\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
\end{algorithm2e}
%
Static program analysis based on abstract
interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
verify certain classes of properties for safety-critical systems.  In
abstract interpretation, a given program is analysed with respect to a set
of given abstract domains.  However, ACDCL is a novel program analysis 
technique that embeds an abstract domain inside the CDCL algorithm.
From the abstract interpretation point of view, ACDCL is an abstract 
interpreter that uses decision and learning to increase transformer precision. 
From the decision procedure perspective, ACDCL is a SAT solver for program 
analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
range of intervals and learning generates program analysis constraints (not assumptions) 
that preserve the error reachability.  Moreover, ACDCL implicitly 
perform program and property driven trace partitioning to increase the precision
of the analysis.  We observe that the abstract domains necessary for 
hardware property checking are -- booleans, constants, interval, equality 
and octagons domains.

Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
to BCP step in SAT solvers where it computes least fix-point with 
strongest post-condition using forward analysis. If the result of $deduce(P)$ 
is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
real counterexample. If the {\em gamma completeness} check is successful, then the 
counterexample is real. Else, the algorithm enters into the while($true$)
loop and heuristically picks a meet irreducible for decision. For example, assuming 
interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
greatest fixed-point. Note that the widening operation in abstract interpretation 
jumps above a least fixed-point, so decisions can be viewed as 
dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
A learnt clause must include asserting cuts which guarantees
derivation of new information after backtracking. The clause learning and backtracking continues 
as long as the result of deduction is BOTTOM ($\perp$) or the 
analysis backtracks to decision level 0. If no further backtrack is possible, then the 
algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
the above process is repeated until a real counterexample is obtained or the algorithm 
backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
However, we are developing a new technique for automatic invariant generation using 
ACDCL for unbounded proofs.

%
\subsection{Example Demonstrating Working of ACDCL}
\begin{figure}[htp]
\centering
\vspace*{0.3cm}
\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
\end{figure}

\lstdefinestyle{mystyle}{
    numbers=left,                    
    numbersep=5pt,                  
    tabsize=2,
    mathescape=true,
    language=C
}
\lstset{style=mystyle}
%\vspace{-2mm}
Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
Astr{\'e}e fails to verify the safety using interval and trace-partition domain
due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
manually annotating the code with partition directives at $n1$, to prove safety.  
In general, the imprecision is either intended by the tool because such high precision analysis is 
normally not required for runtime error analysis or the imprecision is unavoidable 
due to the complexity of the application under analysis.

On the other hand, ACDCL uses decision and learning to generate the proof 
using simple interval domain. The analysis associates an interval with each location and variable.  
Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
starting from $n6$ by computing the weakest precondition for every statement 
in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
and suffices to prove safety.  The deductions made during fixed point
iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
discarding all assumptions. Interval analysis is run with the learnt constraint and 
can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
ACDCL can exploit the program structure by making decisions on interesting program 
variables (for example variables in conditional branches or loop variables). Experimental 
evidence shows that this leads to significantly less number of decisions compared 
to the propositional solver. Compared to classical abstract interpretation 
based tools, ACDCL automatically performs program and property driven trace partitioning 
using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
precise than classical abstract interpretation.

\bibliographystyle{splncs03}
\bibliography{biblio.bib}

\end{document}

### 2016-09-01 paper.tex
structure created

@@
 \usepackage{mathptm}
-\usepackage{rotating}
+%\usepackage{rotating}
 \usepackage{color}
@@
 \usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
+%\usepackage{mdwlist}
+%\usepackage{multirow}
 \usepackage{dingbat}
 % \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
+%\usepackage{colortbl}
+%\usepackage{booktabs}
 \usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
 \usepackage{hyperref}
-\usepackage{dcolumn}
+%\usepackage{dcolumn}
 %\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
@@
 
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
+%\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
+\title{Automated Reasoning of Relational properties using \\
+Abstact Conflict Driven Clause Learning}
 \maketitle
@@
 In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
+of a program by embedding relational domains like octagon domain inside the ACDCL solver.  Experimental 
+evidence shows significantly less number of decisions and learning iterations required by ACDCL solver to 
+prove programs with relational properties compared to bit-blasting and applying propositional SAT solvers -- 
+thereby emphasizing the need for abstract reasoning using ACDCL. 
 %
@@
 perform program and property driven trace partitioning to increase the precision
-of the analysis.  We observe that the abstract domains necessary for 
+of the analysis.  
+\Omit{
+We observe that the abstract domains necessary for 
 hardware property checking are -- booleans, constants, interval, equality 
 and octagons domains.
+}
 
@@
 
+\section{Deduction Phase}
+Overapproximating.
+\subsection{Forward analysis with strongest post-condition}
+\subsection{Chaotic iteration strategy} 
+\section{Decision Heuristics}
+Berkmin, longest-range, random, ordered, octagons 
+\section{Abstract Domains}
+Support for Interval, Octagons
+\section{Clause Learning}
+1.DPLL style -- chronological backtracking \\
+2. CDCL style -- non-chronological backtracking \\
+  a. first-uip \\
+  b. last-uip \\
+
 \bibliographystyle{splncs03}

### 2016-09-05 paper.tex
lifting uip to relational domains

@@
 
+**********************************************
+\section{Lifting First UIP to Octagon domain}
+**********************************************
+unit-ness guarantee in octagon domain:
+  Popped stmt: y23=1+y21
+   Abstract value:
+   D1: y23-y21 < 2 &&
+   D2: y23+y21 > 0 &&
+   D3: y21 < 1
+   After backtracking, apply unit rule 
+   y23 > 1 -- deduction from unit rule
+   Value inconsistent !!
+
+ Note: 
+ 1> cannot make reasoning at literal level for relational domain because literals are dependant on each other.
+As soon as literals denote relation between first-order variables, the pure reasoning on boolean skeleton is not sufficient.
+
+2> Intervals are orthogonal half spaces similar to booleans.
+
+3> After backtracking, the application of unit rule is done as follows for relational domains:
+  Pass the learnt clause (as statement) and the abstrat value to the domain to make deductions.
+
 \bibliographystyle{splncs03}

### 2016-09-22 algorithm.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
-\documentclass[a4paper,conference]{llncs}
-
-\usepackage{import}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mathptm}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{comment}
-\usepackage{alltt}
-\usepackage{graphicx}
-\usepackage{psfrag}
-\usepackage{epsfig}
-\usepackage{wasysym} 
-\usepackage{subfigure}
-\usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
-\usepackage{dingbat}
-% \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
-\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
-\usepackage{hyperref}
-\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
-
-\newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
-\newcommand{\cbmcv}{\tool{cbmc 5.0}}
-\newcommand{\Omit}[1]{}
-
-\input{symbols}
-
-\begin{document}
-
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
-\maketitle
-\begin{abstract}
+\section{Lifting CDCL to Programs}\label{algorithm}
 %
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
-%
-\end{abstract}
-
-\section{Introduction}
-
-\section{Lifting CDCL to Programs}
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
-
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
-
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
-
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
-
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
-
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
-
-\section{Definitions}
-\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
-elements which specify that for certain program variable, there is a certain
-bound. An example of literal is $x \leq 0$.
-
-Let us define the notion of literals with respect to interval domain. 
-An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
-Let I and J be two intervals. The relation between I and J is defined as follows:\\
-We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
-We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.
-
-Let $x$, $x'$ be the intervals of a variable $p$ in current 
-partial assignment and the current clause respectively. 
-We define a literal and a clause to be satisfiable or unsatisfiable 
-or contradicting with respect to an interval abstraction as follows. \\
-\textit{Satisfiable literal:} 
-We say $x'$ is satisfiable literal if $(x \leq x')$.
-
-\textit{Unsatisfiable literal:}
-We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
-
-\textit{Contradicting Literal:}
-We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).
-
-\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
-example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.
-
-\textit{Satisfiable clause:}
-If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
-be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
-current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.
-
-\textit{Conflicting clause:}
-If all the meet irreducibles in a clause are contradicting, then the clause is said to
-be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[17,32]$, then $C$ is a conflicting clause.
-
-\textit{Unsatisfiable clause:}
-If no meet irreducibles in a clause are satisfiable or some meet
-irreducibles are not satisfiable and the rest are contradicting, we call a 
-clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
-Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
-then $C$ is a unsatisfiable clause.
-
-\textit{Unit clause:}
-If all meet irreducible but one is contradicting in a clause, we call the clause
-to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[10,12]$, then $C$ is a unit clause where the unit literal is $z$.
-
-\textit{Unit Rule:} A unit rule is the best abstract transformer 
-(strongest post-condition or weakest pre-condition).
-
-\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
-application of unit rule. This corresponds to computing the greatest fixed point.
-
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
 \begin{algorithm2e}[t]
@@
 ACDCL for unbounded proofs.
-
-%
-\subsection{Example Demonstrating Working of ACDCL}
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-%\vspace{-2mm}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio.bib}
-
-\end{document}

### 2016-09-22 decision.tex
separate sections uploaded

\section{Decision Heuristics}\label{decision}
Berkmin, longest-range, random, ordered, octagons

### 2016-09-22 deductions.tex
separate sections uploaded

\section{Propagation Strategy}\label{deduction}
Let $L$ be the number of constraints generated by a domain $D$
for a program with $N$ variables. For $D$ = {\em Interval}, the 
maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
$2*(N^2)$ for $D$ = {\em Octagons}. Note that the conjunction of all 
constraints in the set $L$ forms an octagon.

Overapproximating.
\subsection{Forward Iteration Strategy}

\subsection{Backward Iteration Strategy}

\subsection{Chaotic Iteration Strategy}

### 2016-09-22 definition.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
-\documentclass[a4paper,conference]{llncs}
-
-\usepackage{import}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mathptm}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{comment}
-\usepackage{alltt}
-\usepackage{graphicx}
-\usepackage{psfrag}
-\usepackage{epsfig}
-\usepackage{wasysym} 
-\usepackage{subfigure}
-\usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
-\usepackage{dingbat}
-% \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
-\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
-\usepackage{hyperref}
-\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
-
-\newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
-\newcommand{\cbmcv}{\tool{cbmc 5.0}}
-\newcommand{\Omit}[1]{}
-
-\input{symbols}
-
-\begin{document}
-
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
-\maketitle
-\begin{abstract}
+\section{Definitions}\label{definition}
 %
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
-%
-\end{abstract}
-
-\section{Introduction}
-
-\section{Lifting CDCL to Programs}
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
-
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
-
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
-
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
-
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
-
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
-
-\section{Definitions}
 \textit{Literal:} A Literal is a meet irreducible which are minimum complementable
@@
 the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwRepeat{Do}{do}{while}
-%\SetKwFunction{assume}{assume}
-%\SetKwFunction{isf}{isFeasible}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
-  \return \safe \;
-}
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
-}
-\While{$true$} 
-{
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
-  }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
-        \uIf{$\neg$ conflict\_analysis()} {
-          return \safe \;  
-        }
-        \assign result=\deduce(P)\;
-      }
-    }
-  }
-}
-\end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
-\end{algorithm2e}
-%
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
-
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
-%
-\subsection{Example Demonstrating Working of ACDCL}
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-%\vspace{-2mm}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio.bib}
-
-\end{document}

### 2016-09-22 domains.tex
separate sections uploaded

\section{Abstract Domains}\label{domain}
Support for Interval, Octagons

### 2016-09-22 example.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
-\documentclass[a4paper,conference]{llncs}
-
-\usepackage{import}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mathptm}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{comment}
-\usepackage{alltt}
-\usepackage{graphicx}
-\usepackage{psfrag}
-\usepackage{epsfig}
-\usepackage{wasysym} 
-\usepackage{subfigure}
-\usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
-\usepackage{dingbat}
-% \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
-\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
-\usepackage{hyperref}
-\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
-
-\newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
-\newcommand{\cbmcv}{\tool{cbmc 5.0}}
-\newcommand{\Omit}[1]{}
-
-\input{symbols}
-
-\begin{document}
-
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
-\maketitle
-\begin{abstract}
-%
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
-%
-\end{abstract}
-
-\section{Introduction}
-
-\section{Lifting CDCL to Programs}
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
-
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
-
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
-
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
-
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
-
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
-
-\section{Definitions}
-\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
-elements which specify that for certain program variable, there is a certain
-bound. An example of literal is $x \leq 0$.
-
-Let us define the notion of literals with respect to interval domain. 
-An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
-Let I and J be two intervals. The relation between I and J is defined as follows:\\
-We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
-We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.
-
-Let $x$, $x'$ be the intervals of a variable $p$ in current 
-partial assignment and the current clause respectively. 
-We define a literal and a clause to be satisfiable or unsatisfiable 
-or contradicting with respect to an interval abstraction as follows. \\
-\textit{Satisfiable literal:} 
-We say $x'$ is satisfiable literal if $(x \leq x')$.
-
-\textit{Unsatisfiable literal:}
-We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
-
-\textit{Contradicting Literal:}
-We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).
-
-\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
-example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.
-
-\textit{Satisfiable clause:}
-If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
-be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
-current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.
-
-\textit{Conflicting clause:}
-If all the meet irreducibles in a clause are contradicting, then the clause is said to
-be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[17,32]$, then $C$ is a conflicting clause.
-
-\textit{Unsatisfiable clause:}
-If no meet irreducibles in a clause are satisfiable or some meet
-irreducibles are not satisfiable and the rest are contradicting, we call a 
-clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
-Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
-then $C$ is a unsatisfiable clause.
-
-\textit{Unit clause:}
-If all meet irreducible but one is contradicting in a clause, we call the clause
-to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[10,12]$, then $C$ is a unit clause where the unit literal is $z$.
-
-\textit{Unit Rule:} A unit rule is the best abstract transformer 
-(strongest post-condition or weakest pre-condition).
-
-\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
-application of unit rule. This corresponds to computing the greatest fixed point.
-
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwRepeat{Do}{do}{while}
-%\SetKwFunction{assume}{assume}
-%\SetKwFunction{isf}{isFeasible}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
-  \return \safe \;
-}
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
-}
-\While{$true$} 
-{
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
-  }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
-        \uIf{$\neg$ conflict\_analysis()} {
-          return \safe \;  
-        }
-        \assign result=\deduce(P)\;
-      }
-    }
-  }
-}
-\end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
-\end{algorithm2e}
-%
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
-
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
+\section{Working Example}\label{example}
 %
@@
 \lstset{style=mystyle}
-%\vspace{-2mm}
 Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
@@
 precise than classical abstract interpretation.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio.bib}
-
-\end{document}

### 2016-09-22 experiments.tex
separate sections uploaded

\section{Experimental Results}\label{experiments}

### 2016-09-22 introduction.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
-\documentclass[a4paper,conference]{llncs}
-
-\usepackage{import}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mathptm}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{comment}
-\usepackage{alltt}
-\usepackage{graphicx}
-\usepackage{psfrag}
-\usepackage{epsfig}
-\usepackage{wasysym} 
-\usepackage{subfigure}
-\usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
-\usepackage{dingbat}
-% \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
-\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
-\usepackage{hyperref}
-\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
-
-\newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
-\newcommand{\cbmcv}{\tool{cbmc 5.0}}
-\newcommand{\Omit}[1]{}
-
-\input{symbols}
-
-\begin{document}
-
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
-\maketitle
-\begin{abstract}
+\section{Introduction}\label{introduction}
 %
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
-%
-\end{abstract}
-
-\section{Introduction}
-
-\section{Lifting CDCL to Programs}
 The conflict driven clause learning algorithm can be seen as an 
@@
 the analysis and prevent enumeration behavior.
-
-\section{Definitions}
-\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
-elements which specify that for certain program variable, there is a certain
-bound. An example of literal is $x \leq 0$.
-
-Let us define the notion of literals with respect to interval domain. 
-An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
-Let I and J be two intervals. The relation between I and J is defined as follows:\\
-We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
-We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.
-
-Let $x$, $x'$ be the intervals of a variable $p$ in current 
-partial assignment and the current clause respectively. 
-We define a literal and a clause to be satisfiable or unsatisfiable 
-or contradicting with respect to an interval abstraction as follows. \\
-\textit{Satisfiable literal:} 
-We say $x'$ is satisfiable literal if $(x \leq x')$.
-
-\textit{Unsatisfiable literal:}
-We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
-
-\textit{Contradicting Literal:}
-We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).
-
-\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
-example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.
-
-\textit{Satisfiable clause:}
-If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
-be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
-current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.
-
-\textit{Conflicting clause:}
-If all the meet irreducibles in a clause are contradicting, then the clause is said to
-be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[17,32]$, then $C$ is a conflicting clause.
-
-\textit{Unsatisfiable clause:}
-If no meet irreducibles in a clause are satisfiable or some meet
-irreducibles are not satisfiable and the rest are contradicting, we call a 
-clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
-Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
-then $C$ is a unsatisfiable clause.
-
-\textit{Unit clause:}
-If all meet irreducible but one is contradicting in a clause, we call the clause
-to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[10,12]$, then $C$ is a unit clause where the unit literal is $z$.
-
-\textit{Unit Rule:} A unit rule is the best abstract transformer 
-(strongest post-condition or weakest pre-condition).
-
-\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
-application of unit rule. This corresponds to computing the greatest fixed point.
-
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwRepeat{Do}{do}{while}
-%\SetKwFunction{assume}{assume}
-%\SetKwFunction{isf}{isFeasible}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
-  \return \safe \;
-}
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
-}
-\While{$true$} 
-{
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
-  }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
-        \uIf{$\neg$ conflict\_analysis()} {
-          return \safe \;  
-        }
-        \assign result=\deduce(P)\;
-      }
-    }
-  }
-}
-\end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
-\end{algorithm2e}
-%
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
-
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
-%
-\subsection{Example Demonstrating Working of ACDCL}
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-%\vspace{-2mm}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio.bib}
-
-\end{document}

### 2016-09-22 learning.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
-\documentclass[a4paper,conference]{llncs}
+\section{Clause Learning with ACDL}\label{learning}
+1.DPLL style -- chronological backtracking \\
+2. CDCL style -- non-chronological backtracking \\
+  a. first-uip \\
+  b. last-uip \\
 
-\usepackage{import}
-\usepackage{times}
-\usepackage{graphicx}
-\usepackage{cite}
-\usepackage{amsfonts}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{mathptm}
-\usepackage{rotating}
-\usepackage{color}
-\usepackage{listings}
-\usepackage{verbatim}
-\usepackage{comment}
-\usepackage{alltt}
-\usepackage{graphicx}
-\usepackage{psfrag}
-\usepackage{epsfig}
-\usepackage{wasysym} 
-\usepackage{subfigure}
-\usepackage{paralist}
-\usepackage{mdwlist}
-\usepackage{multirow}
-\usepackage{dingbat}
-% \usepackage{hyperref}
-\usepackage{colortbl}
-\usepackage{booktabs}
-\usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
-\usepackage{hyperref}
-\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
-\makeatother
+\Omit{
+**********************************************
+\subsection{Lifting First UIP to Octagon domain}
+**********************************************
+unit-ness guarantee in octagon domain:
+  Popped stmt: y23=1+y21
+   Abstract value:
+   D1: y23-y21 < 2 &&
+   D2: y23+y21 > 0 &&
+   D3: y21 < 1
+   After backtracking, apply unit rule 
+   y23 > 1 -- deduction from unit rule
+   Value inconsistent !!
 
-\newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
-\newcommand{\cbmcv}{\tool{cbmc 5.0}}
-\newcommand{\Omit}[1]{}
+ Note: 
+ 1> cannot make reasoning at literal level for relational domain because literals are dependant on each other.
+As soon as literals denote relation between first-order variables, the pure reasoning on boolean skeleton is not sufficient.
 
-\input{symbols}
+2> Intervals are orthogonal half spaces similar to booleans.
 
-\begin{document}
-
-\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
-\maketitle
-\begin{abstract}
-%
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program using richer abstract domains.  Experimental evidence shows significantly less number of 
-decisions and learning iterations compared to propositional SAT solvers -- thereby emphasizing the need
-for abstract reasoning using ACDCL.
-
-%
-\end{abstract}
-
-\section{Introduction}
-
-\section{Lifting CDCL to Programs}
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
-
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
-
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
-
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
-
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
-
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
-
-\section{Definitions}
-\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
-elements which specify that for certain program variable, there is a certain
-bound. An example of literal is $x \leq 0$.
-
-Let us define the notion of literals with respect to interval domain. 
-An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
-Let I and J be two intervals. The relation between I and J is defined as follows:\\
-We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
-We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.
-
-Let $x$, $x'$ be the intervals of a variable $p$ in current 
-partial assignment and the current clause respectively. 
-We define a literal and a clause to be satisfiable or unsatisfiable 
-or contradicting with respect to an interval abstraction as follows. \\
-\textit{Satisfiable literal:} 
-We say $x'$ is satisfiable literal if $(x \leq x')$.
-
-\textit{Unsatisfiable literal:}
-We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
-
-\textit{Contradicting Literal:}
-We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).
-
-\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
-example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.
-
-\textit{Satisfiable clause:}
-If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
-be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
-current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.
-
-\textit{Conflicting clause:}
-If all the meet irreducibles in a clause are contradicting, then the clause is said to
-be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[17,32]$, then $C$ is a conflicting clause.
-
-\textit{Unsatisfiable clause:}
-If no meet irreducibles in a clause are satisfiable or some meet
-irreducibles are not satisfiable and the rest are contradicting, we call a 
-clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
-Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
-then $C$ is a unsatisfiable clause.
-
-\textit{Unit clause:}
-If all meet irreducible but one is contradicting in a clause, we call the clause
-to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[10,12]$, then $C$ is a unit clause where the unit literal is $z$.
-
-\textit{Unit Rule:} A unit rule is the best abstract transformer 
-(strongest post-condition or weakest pre-condition).
-
-\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
-application of unit rule. This corresponds to computing the greatest fixed point.
-
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwRepeat{Do}{do}{while}
-%\SetKwFunction{assume}{assume}
-%\SetKwFunction{isf}{isFeasible}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
-  \return \safe \;
+3> After backtracking, the application of unit rule is done as follows for relational domains:
+  Pass the learnt clause (as statement) and the abstrat value to the domain to make deductions.
 }
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
-}
-\While{$true$} 
-{
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
-  }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
-        \uIf{$\neg$ conflict\_analysis()} {
-          return \safe \;  
-        }
-        \assign result=\deduce(P)\;
-      }
-    }
-  }
-}
-\end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
-\end{algorithm2e}
-%
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
-
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
-%
-\subsection{Example Demonstrating Working of ACDCL}
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-%\vspace{-2mm}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
-
-\bibliographystyle{splncs03}
-\bibliography{biblio.bib}
-
-\end{document}

### 2016-09-22 paper.tex
separate sections uploaded

--- 2016-09-01 paper.tex
@@
 \usepackage{mathptm}
-%\usepackage{rotating}
 \usepackage{color}
@@
 \usepackage{alltt}
-\usepackage{graphicx}
 \usepackage{psfrag}
@@
 \usepackage{paralist}
-%\usepackage{mdwlist}
-%\usepackage{multirow}
 \usepackage{dingbat}
-% \usepackage{hyperref}
-%\usepackage{colortbl}
-%\usepackage{booktabs}
 \usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
 \usepackage{hyperref}
-%\usepackage{dcolumn}
-%\newcolumntype{"}{@{\hskip\tabcolsep\vrule width 1pt\hskip\tabcolsep}}
 \makeatother
@@
 \newcommand{\tool}[1]{\textsc{#1}\xspace}
-\newcommand{\symex}{\tool{path-symex}}
 \newcommand{\cbmcv}{\tool{cbmc 5.0}}
@@
 
-%\title{Equipping Abstact Conflict Driven Clause Learning \\ Solver with Relational Domains}
 \title{Automated Reasoning of Relational properties using \\
 Abstact Conflict Driven Clause Learning}
+
+\Omit{
+\author{Rajdeep Mukherjee\inst{1} \and
+Peter Schrammel\inst{2} \and
+Daniel Kroening\inst{1}}
+
+\institute{University of Oxford, UK \and University of Sussex, UK \\
+\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
+\email{P.Schrammel@sussex.ac.uk}}
+}
 \maketitle
+
+%===============================================================================
 \begin{abstract}
+%===============================================================================
 %
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program analysis that embeds an 
-abstract domain inside the Conflict Driven Clause Learning (CDCL) algorithm. From decision procedure 
-perspective, ACDCL is a SAT solver for program analysis constraints and is a strict generalisation of 
-propositional CDCL solvers. From Abstract Interpretation point of view, ACDCL can be viewed as an 
-abstract interpreter that uses decision and learning to increase the precision of abstract transformers. 
-In this paper, we present a technique to generalize ACDCL solvers to reason about relational properties 
-of a program by embedding relational domains like octagon domain inside the ACDCL solver.  Experimental 
-evidence shows significantly less number of decisions and learning iterations required by ACDCL solver to 
-prove programs with relational properties compared to bit-blasting and applying propositional SAT solvers -- 
-thereby emphasizing the need for abstract reasoning using ACDCL. 
+{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program 
+analysis that embeds an abstract domain inside the Conflict Driven Clause 
+Learning (CDCL) algorithm. From decision procedure perspective, ACDCL is a 
+SAT solver for program analysis constraints and is a strict generalisation 
+of propositional CDCL solvers. From Abstract Interpretation point of view, 
+ACDCL can be viewed as an abstract interpreter that uses decision and 
+learning to increase the precision of abstract transformers. In this paper, 
+we present a technique to generalize ACDCL solvers to reason about relational 
+properties of a program by embedding relational domains like octagon domain 
+inside the ACDCL solver.  Experimental evidence shows significantly less 
+number of decisions and learning iterations required by ACDCL solver to 
+prove programs with relational properties compared to bit-blasting and 
+applying propositional SAT solvers -- thereby emphasizing the need for 
+abstract reasoning using ACDCL. 
 %
@@
 
-\section{Introduction}
+%===============================================================================
+\input{introduction}
+%===============================================================================
 
-\section{Lifting CDCL to Programs}
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
+%===============================================================================
+\input{background}
+%===============================================================================
 
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
+%===============================================================================
+\input{contributions}
+%===============================================================================
 
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
+%===============================================================================
+\input{definition}
+%===============================================================================
 
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
+%===============================================================================
+\input{example}
+%===============================================================================
 
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
+%===============================================================================
+\input{algorithm}
+%===============================================================================
 
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
+%===============================================================================
+\input{deductions}
+%===============================================================================
 
-\section{Definitions}
-\textit{Literal:} A Literal is a meet irreducible which are minimum complementable
-elements which specify that for certain program variable, there is a certain
-bound. An example of literal is $x \leq 0$.
+%===============================================================================
+\input{decision}
+%===============================================================================
 
-Let us define the notion of literals with respect to interval domain. 
-An interval $([l:u])$ is defined by a lower bound $(l)$ and an upper bound $(u)$.
-Let I and J be two intervals. The relation between I and J is defined as follows:\\
-We say $I \leq J$ or {\em (I leq J)} if $(I.u \leq J.u \wedge I.l \geq J.l)$\\
-We say {\em I disjoint J} if $(I.u < J.l \vee  I.l > J.u)$.
+%===============================================================================
+\input{learning}
+%===============================================================================
 
-Let $x$, $x'$ be the intervals of a variable $p$ in current 
-partial assignment and the current clause respectively. 
-We define a literal and a clause to be satisfiable or unsatisfiable 
-or contradicting with respect to an interval abstraction as follows. \\
-\textit{Satisfiable literal:} 
-We say $x'$ is satisfiable literal if $(x \leq x')$.
+%===============================================================================
+\input{experiments}
+%===============================================================================
 
-\textit{Unsatisfiable literal:}
-We say $x'$ is satisfiable literal if $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
-
-\textit{Contradicting Literal:}
-We say $x'$ is contradicting literal if $!(x \leq x') \wedge$ (x disjoint $x'$).
-
-\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. An 
-example clause is $(x \geq 0 \vee y \geq 5 \vee y \leq 10 \vee y \leq 7)$.
-
-\textit{Satisfiable clause:}
-If at least one meet irreducible in a clause is satisfiable, then a clause is said to 
-be satisfiable. For example, consider a clause $C=(x<4 \vee y>10)$. Let the 
-current partial assignment be $x \in [0,3]$, then $C$ is a satisfiable clause.
-
-\textit{Conflicting clause:}
-If all the meet irreducibles in a clause are contradicting, then the clause is said to
-be conflicting. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[17,32]$, then $C$ is a conflicting clause.
-
-\textit{Unsatisfiable clause:}
-If no meet irreducibles in a clause are satisfiable or some meet
-irreducibles are not satisfiable and the rest are contradicting, we call a 
-clause unsatisfiable. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
-Let the current partial assignment be $x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, 
-then $C$ is a unsatisfiable clause.
-
-\textit{Unit clause:}
-If all meet irreducible but one is contradicting in a clause, we call the clause
-to be unit. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
-current partial assignment be $x \in [5,13]$ and $y \in [-2,9]$ and $z \in
-[10,12]$, then $C$ is a unit clause where the unit literal is $z$.
-
-\textit{Unit Rule:} A unit rule is the best abstract transformer 
-(strongest post-condition or weakest pre-condition).
-
-\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
-application of unit rule. This corresponds to computing the greatest fixed point.
-
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
-
-\section{Abstract Conflict Driven Clause Learning}
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwRepeat{Do}{do}{while}
-%\SetKwFunction{assume}{assume}
-%\SetKwFunction{isf}{isFeasible}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
-  \return \safe \;
-}
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
-}
-\While{$true$} 
-{
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
-  }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
-        \uIf{$\neg$ conflict\_analysis()} {
-          return \safe \;  
-        }
-        \assign result=\deduce(P)\;
-      }
-    }
-  }
-}
-\end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
-\end{algorithm2e}
-%
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  
-\Omit{
-We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
-}
-
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
-%
-\subsection{Example Demonstrating Working of ACDCL}
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-%\vspace{-2mm}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
-
-\section{Deduction Phase}
-Overapproximating.
-\subsection{Forward analysis with strongest post-condition}
-\subsection{Chaotic iteration strategy} 
-\section{Decision Heuristics}
-Berkmin, longest-range, random, ordered, octagons 
-\section{Abstract Domains}
-Support for Interval, Octagons
-\section{Clause Learning}
-1.DPLL style -- chronological backtracking \\
-2. CDCL style -- non-chronological backtracking \\
-  a. first-uip \\
-  b. last-uip \\
+%===============================================================================
+\input{conclusion}
+%===============================================================================
 

### 2016-09-22 deductions.tex
deductions text updated

@@
 \subsection{Chaotic Iteration Strategy}
+
+Let us consider an example. 
+int id(int a)
+{
+  return a;
+}
+
+void main() {
+ int x,y,z;
+ _Bool c;
+ if (c)
+   x = -1;
+ else
+   x = 2;
+ int k = id(x);
+ z = x * k;
+ assert((z!=-2));
+}
+
+With forward iteration strategy using octagon domian, the program 
+is proved without any decision. However, with the forward strategy
+using interval domain, we just required one decision to prove the program.

### 2016-09-22 deductions.tex
deductions

@@
 Overapproximating.
+
+begin{figure}[t]
+\tiny
+\begin{tabular}{l|l}
+\hline
+Main Module & Interface Functions \\
+\hline
+\begin{lstlisting}
+int main() {
+ unsigned x, y;
+ __CPROVER_assume(x==y);
+ x++;
+ assert(x==y+1);
+}
+\end{lstlisting}
+&
+\begin{lstlisting}[mathescape=true,language=C]
+$cond#19 == (x#16 == y#18)
+$guard#20 == ($cond#19 && $guard#0)
+x#20 == 1u + x#16
+x#20 == 1u + y#18 || !$guard#20
+$guard#0 == TRUE
+\end{lstlisting}
+\\
+\hline
+\end{tabular}
+\caption{C Program ad its corresponding SSA}
+\label{sw-ssa}
+\end{figure}
+
 \subsection{Forward Iteration Strategy}
+For forward iteration on Figure~\ref{sw-ssa}, we need atleast 
+one decision to deduce BOTTOM and prove safety. The deduction 
+happens from right to left in forward iteration strategy.
+
+The initialization of live variable and subsequent update depends on the iteration strategy.
 
@@
 \subsection{Chaotic Iteration Strategy}  
+The set of octagonal equilities corresponding to Figurei~\ref{sw-ssa}
+is shown below. For chaotic, no decision is required to deduce BOTTOM.
+x#20 > 1                                                                      
+-x20 - y18 < -2                                                               
+-x16 - x20 < -2                                                               
+y18 - x20 < 0                                                                 
+x16 - x20 < 0                                                                 
+y18 > 0                                                                       
+x16 > 0                                                                       
+-x20 - y18 < 0                                                                
+x16 + y18 > 1                                                                 
+y18 - x16 < 1                                                                 
+x20 - y18 < 2                                                                 
+x16 - y18 < 1                                                                 
+cond19                                                                        
+x16 + y18 > 0                                                                 
+guard20                                                                       
+guard0                                                                        
+x16 + x20 > 0                                                                 
+x20 - x16 < 2
 

### 2016-09-23 deductions.tex
deduction text updated

@@
 
-begin{figure}[t]
-\tiny
+\begin{figure}[t]
+\scriptsize
 \begin{tabular}{l|l}
 \hline
-Main Module & Interface Functions \\
+C program & SSA \\
 \hline
-\begin{lstlisting}
+\begin{lstlisting}[mathescape=true,language=C]
 int main() {
@@
 \begin{lstlisting}[mathescape=true,language=C]
-$cond#19 == (x#16 == y#18)
-$guard#20 == ($cond#19 && $guard#0)
-x#20 == 1u + x#16
-x#20 == 1u + y#18 || !$guard#20
-$guard#0 == TRUE
+cond == (x == y)
+guard1 == (cond19 && guard0)
+x' == 1u + x
+x' == 1u + y || !guard1
+guard0 == TRUE
 \end{lstlisting}
@@
 is shown below. For chaotic, no decision is required to deduce BOTTOM.
-x#20 > 1                                                                      
--x20 - y18 < -2                                                               
--x16 - x20 < -2                                                               
-y18 - x20 < 0                                                                 
-x16 - x20 < 0                                                                 
-y18 > 0                                                                       
-x16 > 0                                                                       
--x20 - y18 < 0                                                                
-x16 + y18 > 1                                                                 
-y18 - x16 < 1                                                                 
-x20 - y18 < 2                                                                 
-x16 - y18 < 1                                                                 
-cond19                                                                        
-x16 + y18 > 0                                                                 
-guard20                                                                       
-guard0                                                                        
-x16 + x20 > 0                                                                 
-x20 - x16 < 2
+\begin{figure}[t]
+\scriptsize
+\begin{tabular}{l}
+\hline
+Octagonal Inequalities \\
+\hline
+\begin{lstlisting}
+x' > 1                                                                      
+-x' - y < -2                                                               
+-x - x' < -2                                                               
+y - x' < 0                                                                 
+x - x' < 0                                                                 
+y > 0                                                                       
+x > 0                                                                       
+-x' - y < 0                                                                
+x + y > 1                                                                 
+y - x < 1                                                                 
+x' - y < 2                                                                 
+x - y < 1                                                                 
+x + y > 0                                                                 
+x + x' > 0                                                                 
+x' - x < 2  
+\end{lstlisting}
+\\
+\hline
+\end{tabular}
+\caption{Octagonal Inequalities}
+\label{sw-ssa}
+\end{figure}
 
@@
  int x,y,z;
- _Bool c;
+ \_Bool c;
  if (c)

### 2016-09-23 contributions.tex
contribution

\section{Contributions}\label{contribution}
The goal of this paper is to instantiate ACDCL with 
different relational and non-relational numerical abstract 
domains and evaluate the performance of our solver with 
propositional-level solvers. Experimental evaluation on different
benchmarks shows that ACDCL is able to suitably exploit the 
expressivity of richer abstract domains as well as precision 
of CDCL architecture. This is attributed to the intelligent decision
heuristics which exploits the high-level structure of the problem 
combined with stronger deductions and clause learning mechanism aided 
by the richer abstract domains.

### 2016-09-24 domains.cpp
domain implementation

@@ -1,0 +1,12 @@
+\section{Abstract Domains}\lable{domain}
+The commonly used library for numerical abstract domains  
+is the APRON C library~\cite{apron}.  This library is 
+used for the static analysis of the numerical variables 
+of a program by abstract interpretation. APRON provides a 
+C API interface to various abstract domains and libraries 
+such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
+{\em Linear Equalities} library.  The aim of such analysis is 
+to compute invariants over numerical variables in the 
+program~\cite{se2011}. 
+
+\subsection{Octagon Abstract Domain}

### 2016-09-25 conclusion.tex
paper update

--- 2016-09-23 contributions.tex
@@
-\section{Contributions}\label{contribution}
-The goal of this paper is to instantiate ACDCL with 
-different relational and non-relational numerical abstract 
-domains and evaluate the performance of our solver with 
-propositional-level solvers. Experimental evaluation on different
-benchmarks shows that ACDCL is able to suitably exploit the 
-expressivity of richer abstract domains as well as precision 
-of CDCL architecture. This is attributed to the intelligent decision
-heuristics which exploits the high-level structure of the problem 
-combined with stronger deductions and clause learning mechanism aided 
-by the richer abstract domains.
+\section{Conclusions}
+In this paper, we instantiate ACDCL solver with different numerical domains -- 
+intervals and octagons.  Experimental evaluation on different benchmarks shows that
+ACDCL is able to suitably exploit the expressivity of richer
+abstract domains as well as precision of CDCL architecture.
+This is attributed to the intelligent decision heuristics
+which exploits the high-level structure of the problem
+combined with stronger deductions and clause learning
+mechanism aided by the richer abstract domains.

### 2016-09-25 contributions.tex
paper update

@@
-\section{Contributions}\label{contribution}
+\section{Contributions}
 The goal of this paper is to instantiate ACDCL with 
 different relational and non-relational numerical abstract 
-domains and evaluate the performance of our solver with 
-propositional-level solvers. Experimental evaluation on different
-benchmarks shows that ACDCL is able to suitably exploit the 
-expressivity of richer abstract domains as well as precision 
-of CDCL architecture. This is attributed to the intelligent decision
-heuristics which exploits the high-level structure of the problem 
-combined with stronger deductions and clause learning mechanism aided 
-by the richer abstract domains.
+domains and evaluate the performance of our program analyzer 
+with program verifiers that uses propositional-level solvers.
+
+Experimental evaluation on different benchmarks shows that 
+ACDCL is able to suitably exploit the expressivity of richer 
+abstract domains as well as precision of CDCL architecture. 
+This is attributed to the intelligent decision heuristics 
+which exploits the high-level structure of the problem 
+combined with stronger deductions and clause learning 
+mechanism aided by the richer abstract domains.

### 2016-09-25 deductions.tex
paper update

@@
-\section{Propagation Strategy}\label{deduction}
+\section{Propagation Strategy}
 Let $L$ be the number of constraints generated by a domain $D$
@@
 \caption{C Program ad its corresponding SSA}
-\label{sw-ssa}
+\label{swssa}
 \end{figure}
@@
 \subsection{Forward Iteration Strategy}
-For forward iteration on Figure~\ref{sw-ssa}, we need atleast 
+For forward iteration on Figure~\ref{swssa}, we need atleast 
 one decision to deduce BOTTOM and prove safety. The deduction 
@@
 \caption{Octagonal Inequalities}
-\label{sw-ssa}
+\label{octagons}
 \end{figure}

### 2016-09-25 domains.tex
paper update

@@
-\section{Abstract Domains}\label{domain}
-Support for Interval, Octagons
+\section{Abstract Domains}
+The commonly used library for numerical abstract domains  
+is the APRON C library~\cite{apron}.  This library is 
+used for the static analysis of the numerical variables 
+of a program by abstract interpretation. APRON provides a 
+C API interface to various abstract domains and libraries 
+such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
+{\em Linear Equalities} library.  The aim of such analysis is 
+to compute invariants over numerical variables in the 
+program~\cite{se2011}.
+
+\subsection{Interval Abstract Domain}
+Domain:
+Operators:
+Semantics of elements:
+
+\subsection{Octagon Abstract Domain}
+Domain:
+Operators:
+Semantics of elements:
+
+\subsection{Conditions for Absract Domains}
+An important criterion to support clause learning in 
+partial assignments domain is that elements in the 
+domain must have complementation property.  Most 
+numerical abstract domain does not support 
+complementation operator.  Similar to the interval 
+domain, octagons does not have complementation 
+property, but the decomposition of octagons into 
+half spaces are complementable.
+
+APRON library is specialized for the abstract interpretation 
+framework which does not require complementation 
+operator.  So, we implement our own template-based 
+octagon domain which supports complementation operator.
+
+The octagon in example can be written as a conjunction of:
+\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\]
+
+The complementation of the above octagon can be written as disjunction of:
+\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-09-25 paper.tex
paper update

@@
 
-\title{Automated Reasoning of Relational properties using \\
-Abstact Conflict Driven Clause Learning}
+\title{Lifting CDCL to Numerical Domains}
 
-\Omit{
+\author{Rajdeep Mukherjee Peter Schrammel   Daniel Kroening}
+
+\institute{University of Oxford, UK \and University of Sussex, UK \\
+\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
+\email{p.schrammel@sussex.ac.uk}}
+\Omit {
 \author{Rajdeep Mukherjee\inst{1} \and
@@
 \email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
-\email{P.Schrammel@sussex.ac.uk}}
+\email{p.schrammel@sussex.ac.uk}}
 }
+
 \maketitle
@@
 %
-{\em Abstract Conflict Driven Clause Learning} (ACDCL) is a novel program 
-analysis that embeds an abstract domain inside the Conflict Driven Clause 
-Learning (CDCL) algorithm. From decision procedure perspective, ACDCL is a 
-SAT solver for program analysis constraints and is a strict generalisation 
-of propositional CDCL solvers. From Abstract Interpretation point of view, 
-ACDCL can be viewed as an abstract interpreter that uses decision and 
-learning to increase the precision of abstract transformers. In this paper, 
-we present a technique to generalize ACDCL solvers to reason about relational 
-properties of a program by embedding relational domains like octagon domain 
-inside the ACDCL solver.  Experimental evidence shows significantly less 
-number of decisions and learning iterations required by ACDCL solver to 
-prove programs with relational properties compared to bit-blasting and 
-applying propositional SAT solvers -- thereby emphasizing the need for 
-abstract reasoning using ACDCL. 
+The conflict driven clause learning algorithm can be seen as an 
+alternation between over-approximate deduction with boolean 
+constraint propagation, and under-approximate abduction, with 
+conflict analysis.  {\em Abstract Conflict Driven Clause Learning} 
+(ACDCL) is a novel program analysis that embeds an abstract domain 
+inside the Conflict Driven Clause Learning (CDCL) algorithm. From 
+decision procedure perspective, ACDCL is a SAT solver for program 
+analysis constraints and is a strict generalisation of propositional CDCL 
+solvers. From Abstract Interpretation point of view, ACDCL can be 
+viewed as an abstract interpreter that uses decision and learning to 
+increase the precision of abstract transformers.  In this paper, we 
+instantiate ACDCL with different relational and non-relational numerical 
+abstract domains and evaluate the performance of our program analyzer
+with program verifiers that uses propositional-level solvers.  Experimental 
+evidence on a set of \rmcmt{N} benchmarks shows significant drop in number 
+of decisions and backjumping iterations in addition to stronger 
+deductions and learned clause aided by the richer abstract domains.
 %
 \end{abstract}
-
 %===============================================================================
@@
 \input{algorithm}
+%===============================================================================
+
+%===============================================================================
+\input{domains}
 %===============================================================================

### 2016-09-25 domains.cpp

@@ -1,12 +1,0 @@
-\section{Abstract Domains}\lable{domain}
-The commonly used library for numerical abstract domains  
-is the APRON C library~\cite{apron}.  This library is 
-used for the static analysis of the numerical variables 
-of a program by abstract interpretation. APRON provides a 
-C API interface to various abstract domains and libraries 
-such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
-{\em Linear Equalities} library.  The aim of such analysis is 
-to compute invariants over numerical variables in the 
-program~\cite{se2011}. 
-
-\subsection{Octagon Abstract Domain}

### 2016-09-26 domains.tex
domain

@@
 \subsection{Conditions for Absract Domains}
-An important criterion to support clause learning in 
-partial assignments domain is that elements in the 
-domain must have complementation property.  Most 
-numerical abstract domain does not support 
-complementation operator.  Similar to the interval 
-domain, octagons does not have complementation 
-property, but the decomposition of octagons into 
-half spaces are complementable.
+One of the requirements for a clause learning SAT solver 
+is that each element of the partial assignments domain 
+needs to have a decomposition into precisely complementable 
+elements.  This property of the underlying domain helps 
+guiding the search away from the conflicting region of the search 
+space.
 
+Thus, clause learning SAT solvers requires a weak complementation property
+of the abstraction.  However, numerical abstract domain,  such as 
+intervals and octagons are intersections of complementable half-spaces. 
 APRON library is specialized for the abstract interpretation 
 framework which does not require complementation 
-operator.  So, we implement our own template-based 
+operator.  To this end, we implement our own template-based 
 octagon domain which supports complementation operator.
 
-The octagon in example can be written as a conjunction of:
+For example, the octagon in example can be written as a conjunction of:
 \[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\]

### 2016-09-26 paper.tex
domain

@@
 
-\author{Rajdeep Mukherjee Peter Schrammel   Daniel Kroening}
+\author{Rajdeep Mukherjee\inst{1} Peter Schrammel   Daniel Kroening}
 
@@
 %
-The conflict driven clause learning algorithm can be seen as an 
+The Conflict Driven Clause Learning (CDCL) algorithm can be seen as an 
 alternation between over-approximate deduction with boolean 
@@
 (ACDCL) is a novel program analysis that embeds an abstract domain 
-inside the Conflict Driven Clause Learning (CDCL) algorithm. From 
-decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is a strict generalisation of propositional CDCL 
-solvers. From Abstract Interpretation point of view, ACDCL can be 
-viewed as an abstract interpreter that uses decision and learning to 
-increase the precision of abstract transformers.  In this paper, we 
-instantiate ACDCL with different relational and non-relational numerical 
-abstract domains and evaluate the performance of our program analyzer
-with program verifiers that uses propositional-level solvers.  Experimental 
-evidence on a set of \rmcmt{N} benchmarks shows significant drop in number 
-of decisions and backjumping iterations in addition to stronger 
+inside the CDCL algorithm. From decision procedure perspective, 
+ACDCL is a SAT solver for program analysis constraints and is a 
+strict generalisation of propositional CDCL solvers. From Abstract 
+Interpretation point of view, ACDCL can be viewed as an abstract 
+interpreter that uses decision and learning to increase the precision 
+of abstract transformers.  In this paper, we instantiate ACDCL with 
+a relational and a non-relational numerical abstract domain and 
+evaluate the performance of our analyzer with program verifiers that 
+uses propositional-level solvers.  Experimental evidence on a set of 
+\rmcmt{N} benchmarks \rmcmt{drawn from X} shows significant reduction 
+in the number of decisions and backjumping iterations along with stronger 
 deductions and learned clause aided by the richer abstract domains.

### 2016-09-26 domains.tex
domain text

@@
 \section{Abstract Domains}
+\subsection{Program Models}
+
 The commonly used library for numerical abstract domains  
@@
 \subsection{Interval Abstract Domain}
-Domain:
-Operators:
-Semantics of elements:
+The {\em Interval} abstract domain is a lattice 
+$ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
+$ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
+the set of intervals of type $[l,u]$ over numeric data 
+type with $l \leq u$. The least element is $\bot$ and the 
+greatest element is $\top$ which maps all variables to their
+minimum ($min$) and maximum ($max$) values.  An interval 
+$\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
+partial order $\sle_{I}$ over elements in the set $Itv$ is 
+given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
+A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
+x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
+$\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
+A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
+x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
+$\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
+The galois connection between the concrete environment and intervals is 
+given as follows.
+\[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
+\[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
+For a program with $N$ variables, there are a total of 
+$2*N$ intervals. The intervals restrict the set of values 
+taken by program variables at each program point, starting 
+from $[-\infty, \infty]$.
 
 \subsection{Octagon Abstract Domain}
+The {\em Octagon} abstract domain is a lattice 
+$OctDom = (OctElm, \sle_{O}, \join_{O}, \meet_{O})$, where
+$OctElm: (Var \times Var \rightarrow (\mathbb{R} \cup \{\infty\})) \cup \bot$. 
+The least element is $\bot$ that contains all unsatisfiable 
+set of inequalities and the greatest element is $\top$ which 
+maps the bounds of all octagonal inequalities to $\infty$. 
+The partial order $\sle_{O}$ over elements in the set $OctElm$ is 
+given by $O_1 \sle_{O} O_2$ iff the bounds of each inequalities in $0_1$ 
+is included (by $\leq$ order) in the bounds of corresponding inequalities 
+in $O_2$, that is, the octagons are ordered by the inclusion relations.
+
+The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
+(v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
+and is computed by taking piece-wise maximum of bounds of corresponding 
+octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
+However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
+(v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
+computed by taking piece-wise minimum of bounds of corresponding 
+octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.
+
+For building program analyzers using octagon domains, the domain also 
+supports few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.
+The closure operator is often used to reduce the degree of over-approximation 
+resulting from the join operation. Closure always produces unique octagonal 
+inequalities, but it is very expensive to compute because of cubic complexity 
+in the number of program variables.   
+The octagon domain is a relational numberic abstract domain that  
 Domain:
@@
 needs to have a decomposition into precisely complementable 
-elements.  This property of the underlying domain helps 
-guiding the search away from the conflicting region of the search 
-space.
+elements.  This property of the underlying domain helps to 
+guide the search away from the conflicting region of the 
+search space.
 

### 2016-09-27 domains.tex
octagon domain description

@@
 \subsection{Program Models}
-
-The commonly used library for numerical abstract domains  
-is the APRON C library~\cite{apron}.  This library is 
-used for the static analysis of the numerical variables 
-of a program by abstract interpretation. APRON provides a 
-C API interface to various abstract domains and libraries 
-such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
-{\em Linear Equalities} library.  The aim of such analysis is 
-to compute invariants over numerical variables in the 
-program~\cite{se2011}.
 
@@
 computed by taking piece-wise minimum of bounds of corresponding 
-octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.
+octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
+octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
+the octagons that abstract the same concrete values.
+
+The octagon domain is a relational abstract domain that permits $2n^2$ 
+linear inequalities between $n$ program variables.  The octagonal 
+inequalities are of two types: binary or unary inequalities as shown below.
+\[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
+An element in octagon domain, $OctElm$, is a conjunction of such 
+inequalities and is called an octagon.  Figure~\ref{octagon} shows an 
+example octagon and its associated inequalities.
 
 For building program analyzers using octagon domains, the domain also 
-supports few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.
-The closure operator is often used to reduce the degree of over-approximation 
-resulting from the join operation. Closure always produces unique octagonal 
-inequalities, but it is very expensive to compute because of cubic complexity 
-in the number of program variables.   
-The octagon domain is a relational numberic abstract domain that  
-Domain:
-Operators:
-Semantics of elements:
+provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
+The widening operator is used to accelerate convergence for loops in the program
+and has a quadratic complexity in the number of variables.  If the bound of an 
+octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
+to $\infty$.  However, the closure operator is often used to reduce the degree 
+of over-approximation resulting from the join operation. Closure always produces 
+unique octagonal inequalities, but it is very expensive to compute because of 
+cubic complexity in the number of program variables.
 
-\subsection{Conditions for Absract Domains}
+The abstract transformer, $apost$, captures the effect of different program 
+statements in the abstract domain. The transformer is precise for octagonal 
+assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
+as shown below.
+\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \]  
+\[apost_{x:=y+z}(b) = \langle \top \rangle \]
+
+\subsection{Properties of Domain Elements for Decision and Learning}
 One of the requirements for a clause learning SAT solver 
@@
 intervals and octagons are intersections of complementable half-spaces. 
+The commonly used library for numerical abstract domains  
+is the APRON C library~\cite{apron}.  This library is 
+used for the static analysis of the numerical variables 
+of a program by abstract interpretation. APRON provides a 
+C API interface to various abstract domains and libraries 
+such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
+{\em Linear Equalities} library.  The aim of such analysis is 
+to compute invariants over numerical variables in the 
+program~\cite{se2011}.
+
 APRON library is specialized for the abstract interpretation 

### 2016-09-27 domains.tex
program model

@@
 \section{Abstract Domains}
-\subsection{Program Models}
-
+\textbf{Programs:}
+%
+A {\em program} $\mathcal{P}$ is formally defined as follows.
+\[
+\begin{array}[t]{@{}lll}
+\mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
+\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
+\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
+\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
+\end{array}
+\]
+Consider sets of expressions $Exp$ and Boolean expressions $BExp$
+over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
+can take numeric values in $Val$.  A procedure is denoted as 
+$\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
+is an assignment, conditional, sequential concatenation or a loop.
+\rmcmt{post}
 \subsection{Interval Abstract Domain}

### 2016-09-27 domains.tex
more on program model

@@
-\section{Abstract Domains}
+\section{Program and Abstract Domains}
+\subsection{Concrete Semantics}
 \textbf{Programs:}
@@
 $\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
-is an assignment, conditional, sequential concatenation or a loop.
-\rmcmt{post}
+is an assignment, conditional, sequential concatenation or a loop. \\
+
+\textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
+where $Loc$ is the set of locations with an unique start 
+location $(init)$ and error location $(err)$, $E$ is the set 
+of control-flow edges that are labelled with the set 
+$lbl \in Statement$.  For the purpose of illustration, we 
+assume that all the procedures are inlined.
+
+\textbf{Concrete Domain:} The {\em concrete} domain is a lattice 
+of concrete environments, $Env = Var \rightarrow Val$, and is 
+defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
+A transformer, $post_{stmt}$, of concrete domain defines 
+the effect of a statement $stmt$ on the concrete domain, 
+$post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.
+
+A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
+where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
+of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
+$0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
+if $\sigma_{i+1} \in post_{i}(\sigma_{i})$. \\
+
 \subsection{Interval Abstract Domain}
@@
 
-The abstract transformer, $apost$, captures the effect of different program 
+The abstract transformer, $apost_{stmt}$, captures the effect of different program 
 statements in the abstract domain. The transformer is precise for octagonal 

### 2016-09-29 domains.tex
octagon figure added
domain text updated

@@
 \section{Program and Abstract Domains}
-\subsection{Concrete Semantics}
-\textbf{Programs:}
+\section{Programs:}
 %
@@
 
-\textbf{Concrete Domain:} The {\em concrete} domain is a lattice 
+\subsection{Concrete Semantics}
+The {\em concrete} domain is a lattice 
 of concrete environments, $Env = Var \rightarrow Val$, and is 
@@
 $0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
-if $\sigma_{i+1} \in post_{i}(\sigma_{i})$. \\
+if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.
 
@@
 example octagon and its associated inequalities.   
-
+%
+\begin{wrapfigure}{r}{5.5cm}
+\caption{Am example of Octagon}\label{octagon}
+\scalebox{.55}{\import{figures/}{octagons.pspdftex}}
+\end{wrapfigure} 
+%
 For building program analyzers using octagon domains, the domain also 
@@
 as shown below.
-\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \]  
-\[apost_{x:=y+z}(b) = \langle \top \rangle \]
+\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
 
@@
 octagon domain which supports complementation operator.  
-
-For example, the octagon in example can be written as a conjunction of:
+For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
 \[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
-
 The complementation of the above octagon can be written as disjunction of:

### 2016-09-29 cdcl.tex
CDCL

--- 2016-09-22 introduction.tex
@@
-\section{Introduction}\label{introduction}
-%
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
+\section{Lifting CDCL to numeric domains}
+The work of~\cite{dhk2013-popl,sas12,tacas12} shows that 
+satisfiability solvers are static analyzers. They present 
+a framework that strictly generalizes CDCL algorithm to 
+lattices and abstract transformers. We refer the reader 
+to~\cite{dhk2013-popl} for a more formal perspective alongwith 
+soundness and completeness proofs.  In this section, we focus 
+on the practical aspect of lifting CDCL algorithm to numeric 
+abstract domains.
+
+We treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
 $\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
+$P$. 
+CDCL algorithm has two main phases which feedback information to 
+each other -- {\em model search} and {\em conflict analysis}. The model 
+search procedure tries to find a satisfying assignment to the input 
+formula. If the search fails, meaning that the current partial assignment 
+leads to a {\em conflict}, the conflict analysis phase learns the reason 
+for such failure to block the model search from entering into same 
+conflicting search-space in future. More formally, given a set of 
+formula $Form$ and a set of structure $Struct$, a semantic entailment 
+relation, $\models \in \mathcal{P}(Struct \times Form)$, identifies the 
+structure that satisfies a formula. A concrete domain is the power set 
+of structures which gives us lattice. 
 We define two transformers for a CDCL solver. \\
@@
 mind satisfiability information.
-
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
-
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.

### 2016-09-29 deductions.tex
CDCL

@@
 \section{Propagation Strategy}
-Let $L$ be the number of constraints generated by a domain $D$
-for a program with $N$ variables. For $D$ = {\em Interval}, the 
+For a program with $N$ variables, let $L$ be the total number of 
+meet irreducibles returned by a domain $D$.  For $D$ = {\em Interval}, the 
 maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em Octagons}. Note that the conjunction of all 
-constraints in the set $L$ forms an octagon.
+$2*(N^2)$ for $D$ = {\em Octagons}. Note that an octagon is the conjunction 
+of all octagonal inequalities in the set $L$.
 
@@
 \scriptsize
-\begin{tabular}{l|l}
+\begin{tabular}{l|l|l}
 \hline
-C program & SSA \\
+C program & SSA & Octagon \\
 \hline
@@
 &
-\begin{lstlisting}[mathescape=true,language=C]
-cond == (x == y)
-guard1 == (cond19 && guard0)
-x' == 1u + x
-x' == 1u + y || !guard1
-guard0 == TRUE
-\end{lstlisting}
+\begin{minipage}{4.40cm}
+$\begin{array}{l@{\,\,}c@{\,\,}l}
+SSA &\iff& ((guard0 == TRUE) \land \\
+    &    & (cond == (x == y)) \land \\
+    &    & (guard1 == (cond \&\& guard0)) \land \\
+    &    & (x' == 1u + x) \land \\
+    &    & (x' == 1u + y || !guard1))
+\end{array}$
+\end{minipage}
+&
+\begin{minipage}{3.75cm}
+$\begin{array}{l@{\,\,}c@{\,\,}l}
+C &\iff& ((x' > 1) \land (-x'-y < -2) \land \\
+  &    & (-x-x' < -2) \land (y-x' < 0) \land \\                                                                
+  &    & (x-x' < 0) \land (y > 0) \land \\
+  &    & (x > 0) \land (-x'-y < 0) \land \\
+  &    & (x+y > 1) \land (y-x < 1) \land \\
+  &    & (x'-y < 2) \land (x-y < 1) \land \\
+  &    & (x+y > 0) \land (x+x' > 0) \land \\
+  &    & (x'-x < 2))
+\end{array}$
+\end{minipage}
 \\
@@
 \end{tabular}
-\caption{C Program ad its corresponding SSA}
+\caption{C Program, corresponding SSA and Octagonal Inequalities}
 \label{swssa}
 \end{figure}
-
+%
 \subsection{Forward Iteration Strategy}
@@
 is shown below. For chaotic, no decision is required to deduce BOTTOM.
+
+\Omit {
+Let us consider an example. 
 \begin{figure}[t]
 \scriptsize
-\begin{tabular}{l}
+\begin{tabular}{l|l|l}
 \hline
-Octagonal Inequalities \\
+C program & SSA & Octagons \\
 \hline
-\begin{lstlisting}
-x' > 1                                                                      
--x' - y < -2                                                               
--x - x' < -2                                                               
-y - x' < 0                                                                 
-x - x' < 0                                                                 
-y > 0                                                                       
-x > 0                                                                       
--x' - y < 0                                                                
-x + y > 1                                                                 
-y - x < 1                                                                 
-x' - y < 2                                                                 
-x - y < 1                                                                 
-x + y > 0                                                                 
-x + x' > 0                                                                 
-x' - x < 2  
-\end{lstlisting}
-\\
-\hline
-\end{tabular}
-\caption{Octagonal Inequalities}
-\label{octagons}
-\end{figure}
-
-Let us consider an example. 
-int id(int a)
-{
+\begin{lstlisting}[mathescape=true,language=C]
+int id(int a) {
   return a;
 }
-
-void main() {
+int main() {
  int x,y,z;
- \_Bool c;
+ _Bool c;
  if (c)
@@
 }
+\end{lstlisting}
+&
+\begin{lstlisting}[mathescape=true,language=C]
+cond == (x == y)
+guard1 == (cond19 && guard0)
+x' == 1u + x
+x' == 1u + y || !guard1
+guard0 == TRUE
+\end{lstlisting}
+&
+\begin{lstlisting}[mathescape=true,language=C]
+
+\end{lstlisting}
+\\
+\hline
+\end{tabular}
+\caption{C Program, corresponding SSA and Octagonal inequalities}
+\label{swssa}
+\end{figure}
+}
 

### 2016-09-29 introduction.tex
paper update
CDCL

@@
-\section{Introduction}\label{introduction}
+\section{Introduction}
 %
-The conflict driven clause learning algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis. The work of~\cite{dhk2013-popl,sas12,tacas12} 
-shows that satisfiability solvers are static analyzers. They present 
-a framework that strictly generalizes CDCL algorithm to lattices and 
-abstract transformers. \\
-Our work is fundamentally based on the work of~\cite{dhk2013-popl}. 
-However, we extend the framework to use it as a backend reasoning 
-engine for high-level reasoning of hardwares circuits represented as programs.
-To this end, we treat programs as a logical formula. Give a programs $P$ and a trace $\pi$,
-$\pi \models P$ iff trace $\pi$ is an erroneous trace generated by program
-$P$. CDCL algorithm has two main phases which feedback information to each other --
-{\em model search} and {\em conflict analysis}. The model search procedure tries to find a 
-satisfying assignment to the input formula. If the search fails, meaning that the current 
-partial assignment leads to a {\em conflict}, then the conflict analysis phase learns the 
-reason for such failure to block the model search from entering into same conflicting 
-search-space in future. More formally, given a set of formula $Form$ and a set of 
-structure $Struct$, a semantic entailment relation is given as, $\models \in \mathcal{P}(Struct
-\times Form)$, which specifies which structure satisfies which formula. A 
-concrete domain is the power set of structures which gives us lattice. 
-We define two transformers for a CDCL solver. \\
-\textit{Model Transformer (mod):} Given a set of structure $S$, return all structures 
-in $S$ that satisfy a formula $\varphi$. This is formally shows as follows:
-$mod_{\varphi}(S) = \{\sigma | \sigma \in S \wedge \sigma \models \varphi\}$ \\
-
-\textit{Conflict Transformer (conf):} Start with a set $S$, and add to $S$ all structures 
-that contradict a formula $\varphi$. This is formally shows as follows:
-$conf_{\varphi}(S) = \{\sigma | \sigma \in S \vee \sigma \not\models \varphi\}$
-\\
-We now define abstract interpretation analogues to model search and conflict analysis.   \\
-\textit{Abstract Model Transformer (amod):} Given a program P, $amod$ gives the 
-set of traces generated by the program and is erroneous. This can be
-characterised in two different ways. We can start with the initial state and
-compute least fix-point with strongest post-condition or start from error state 
-and compute the greatest fix-point with weakest pre-condition. \\
-
-\textit{Abstract Conflict Transformer (aconf):} Given a program P, $aconf$ gives the 
-set of program traces that are safe, that is, all program traces that are not
-generated by the program. \\
-
-\textit{Interaction between amod and aconf:} If the search for finding an 
-erroneous trace $(amod)$ fails, we get partial safety proofs which correspond 
-to conflict in SAT solver. From the conflict collection transformer $(aconf)$ or 
-proof stage, we get back refinement of strongest post-condition which keeps in 
-mind satisfiability information.
 

### 2016-09-29 cdcl.tex
cdcl

@@
 lattices and abstract transformers. We refer the reader 
-to~\cite{dhk2013-popl} for a more formal perspective alongwith 
-soundness and completeness proofs.  In this section, we focus 
-on the practical aspect of lifting CDCL algorithm to numeric 
-abstract domains.
+to~\cite{dhk2013-popl} for a more formal perspective of 
+abstract CDCL.  In this section, we focus on the practical 
+aspect of lifting CDCL algorithm to numeric (relational and 
+non-relational) abstract domains.
 
@@
 $P$. 
-CDCL algorithm has two main phases which feedback information to 
-each other -- {\em model search} and {\em conflict analysis}. The model 
+CDCL algorithm has two main phases which alternates between each 
+other -- {\em model search} and {\em conflict analysis}. The model 
 search procedure tries to find a satisfying assignment to the input 
-formula. If the search fails, meaning that the current partial assignment 
-leads to a {\em conflict}, the conflict analysis phase learns the reason 
-for such failure to block the model search from entering into same 
-conflicting search-space in future. More formally, given a set of 
-formula $Form$ and a set of structure $Struct$, a semantic entailment 
-relation, $\models \in \mathcal{P}(Struct \times Form)$, identifies the 
-structure that satisfies a formula. A concrete domain is the power set 
-of structures which gives us lattice. 
+formula. If the search fails, then the current partial assignment is 
+conflicting, that is, it contains only countermodels.  The conflict 
+analysis phase derives a generalized reason for conflict which learns a new 
+lemma in the form of a clause. The learned clause is used to block the 
+model search from entering into same conflicting search-space in future.
+
+More formally, given a set of formula $Form$ and a set of structure $Struct$, a semantic entailment relation, $\models \in \mathcal{P}(Struct \times Form)$, identifies the structure that satisfies a formula.
+A concrete domain is the power set of structures which gives us lattice.
 We define two transformers for a CDCL solver. \\

### 2016-10-02 cdcl.tex
definitions added

@@
 other -- {\em model search} and {\em conflict analysis}. The model 
-search procedure tries to find a satisfying assignment to the input 
-formula. If the search fails, then the current partial assignment is 
+search procedure refines a partial assignment until either a satisfying 
+assignment to the input formula is found or a conflict is encountered.  
+If the search fails, then the current partial assignment is 
 conflicting, that is, it contains only countermodels.  The conflict 
-analysis phase derives a generalized reason for conflict which learns a new 
-lemma in the form of a clause. The learned clause is used to block the 
-model search from entering into same conflicting search-space in future.
+analysis phase derives a generalized reason from the conflicting partial 
+assignment which learns a new lemma in the form of a clause. The learned 
+clause is used to block the model search from entering into same conflicting 
+search-space in future.
 

### 2016-10-02 domains.tex
CDCL
octagon figure
definitions added

@@
 if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.
+
+\begin{definition}{(Meet Irreducibles)} A meet irreducible is a minimum complementable 
+element that cannot be expressed in terms of the meet over a set of other 
+elements. Formally, a {\em meet irreducible} $M$ in a complete lattice 
+structure $A$ is an element $M \in L$ that has the following property.
+\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
+\end{definition}
+
+In case of partial assignments, meet irreducibles are the singleton
+assignments, for example $\{x \mapsto true \}$, which represents the set 
+of all propositional assignments where the literal $x$ is mapped to true.  
+An important property of meet irreducibles in case of partial assignment is that 
+they have precise complements.  For example, the complement of $\{x \mapsto true \}$
+is a singleton element,$\{x \mapsto false \}$, in the partial assignment domain.
+
+\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
+irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
+irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
+is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
+complementable, then $A$ is said to be have complementable meet irreducibles.  
+\end{definition}
 
@@
 \begin{wrapfigure}{r}{5.5cm}
-\caption{Am example of Octagon}\label{octagon}
-\scalebox{.55}{\import{figures/}{octagons.pspdftex}}
+\caption{An example of Octagon}\label{octagon}
+\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
 \end{wrapfigure} 

### 2016-10-03 deductions.tex
definition of deduction transformer

@@
-\section{Propagation Strategy}
+\section{Deduction}
+\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$ is a sound approximation of a concrete model transformer
+$ded_{\varphi}$ given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+$m \in A$ is a meet irreducible.   
+\end{definition}
+
 For a program with $N$ variables, let $L$ be the total number of 
@@
 %
+\subsection{Algorithm for Deduction}
+
 \subsection{Forward Iteration Strategy}

### 2016-10-03 deductions.tex
deduction definitions complete

--- 2016-09-29 deductions.tex
@@
-\section{Propagation Strategy}
+\section{Deduction}
+\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$ is a sound approximation of a concrete model transformer
+$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+$m \in A$ is a meet irreducible.   
+\end{definition}
+
+Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
+\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+\rangle$.  An abstract deduction transformer is typically computed in the form 
+of strongest post-condition or a weakest pre-condition of a formula in the 
+abstract domain.
+
+A meet decomposition of the outcome of abstract deduction transformer 
+is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
+\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+precisely complementable.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
+
 For a program with $N$ variables, let $L$ be the total number of 
-meet irreducibles returned by a domain $D$.  For $D$ = {\em Interval}, the 
+meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
 maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em Octagons}. Note that an octagon is the conjunction 
+$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
 of all octagonal inequalities in the set $L$.
-
-Overapproximating.
-
+%
 \begin{figure}[t]

### 2016-10-03 deductions.tex
Iteration strategy

@@
 %
-\subsection{Algorithm for Deduction}
+\subsection{Iteration Strategy For Abstract Deduction Transformer}
+\begin{algorithm2e}[t]
+\DontPrintSemicolon
+\SetKwFunction{deduce}{deduce}
+\SetKwFunction{print}{print}
+\SetKwFunction{return}{return}
+\SetKwFunction{continue}{continue}
+\SetKwFunction{assign}{assign}
+\SetKwData{safe}{safe}
+\SetKwData{unsafe}{unsafe}
+\SetKwData{unknown}{unknown}
+\SetKwData{true}{true}
+\SetKwInOut{Input}{input}
+\SetKwInOut{Output}{output}
+\SetKw{KwNot}{not}
+\begin{small}
+\Input{Program $\mathcal{P}$, trail $\mathcal{T}$ and assertion $A$}
+\Output{The status (\safe or \unknown)}
+\assign worklist = initialize\_worklist() \;
+\While{$!worklist.empty()$} 
+{
+  \assign $\langle live\_vars, statement \rangle$ = worklist.pop() \;
+  \assign $p_{asgn}$ = $\bigwedge\{\mathcal{T}\}$ \; 
+  \assign $\mathit{meet\_irrd}$ = domain(statement, live\_vars, $p_{asgn}$)\;
+  // add deduction to trail \;
+  $\mathcal{T} = \mathcal{T} \join \mathit{meet\_irrd}$ \; 
+  \assign $\widetilde{p_{asgn}}$ = $\bigwedge\{\mathcal{T}\}$ \;
+
+  // check if the domain deduce BOTTOM \;
+  \uIf{$\widetilde{p_{asgn}} == \bot$} {
+    // empty the worklist \;
+    worklist.delete() \;
+    \return \safe \;
+  }
+  \uElse
+  {
+    \assign new\_vars = $\widetilde{p_{asgn}}.get\_variables()$ \;
+    worklist.update(new\_vars, statement, A) \; 
+  }
+}
+\return \unknown \;
+\end{small}
+\caption{Worklist-based Iteration Strategy\label{Alg:iteration}}
+\end{algorithm2e}
 

### 2016-10-04 deductions.tex
Model search

--- 2016-10-03 deductions.tex
@@
 \section{Deduction}
+In a typical abstract interpretation based galois-connection settings over 
+an over-approximate domain, every concrete elements have a unique over-approximate 
+representation in the abstract.  Likewise, every concrete transformer is
+over-approximated by a unique abstract transformer.  We now define a abstract 
+deduction transformer.
+
 \begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
@@
 %
-\subsection{Algorithm for Deduction}
+\subsection{Iteration Strategy For Abstract Deduction Transformer}
+Algorithm~\label{Alg:iteration} presents the modelSearch procedure. 
+our deduction high-level sketch for the abstract deduction transformer.
+%The inputs are Input{Set of $\widehat{ded_{\varphi}}$ for $\varphi$ from Program $\mathcal{P}$, trail $\mathcal{T}$ and assertion $A$}
+$statement \in S$, $statement$ is a meet irreducible deduction transformer.
+\begin{algorithm2e}[t]
+\DontPrintSemicolon
+\SetKwFunction{deduce}{deduce}
+\SetKwFunction{print}{print}
+\SetKwFunction{return}{return}
+\SetKwFunction{continue}{continue}
+\SetKwFunction{assign}{assign}
+\SetKwData{safe}{safe}
+\SetKwData{unsafe}{unsafe}
+\SetKwData{unknown}{unknown}
+\SetKwData{true}{true}
+\SetKwInOut{Input}{input}
+\SetKwInOut{Output}{output}
+\SetKwFor{Loop}{Loop}{}{}
+\SetKw{KwNot}{not}
+\begin{small}
+\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
+abstract value $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
+\Output{The status (\safe or \unsafe or \unknown)}
+worklist $\leftarrow$ initialize\_worklist($S$) \;
+\Loop{}
+{
+// Greatest fixed point \;
+\While{$!worklist.empty()$} 
+{
+  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
+  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
+  // check if the domain deduce BOTTOM \;
+  \uIf{$meet\_irrd == \bot$} {
+    $\mathcal{R}[\bot] \leftarrow s$ \;
+    // empty the worklist \;
+    worklist.delete() \;
+    \return \safe \;
+  }
+  \uElse
+  {
+    // add deduction to trail \;
+    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
+    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
+    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
+    // update the worklist \;
+    worklist.update(v, s) \; 
+  }
+}
+// check for counterexample\;
+\uIf{$v$ is gamma\_complete} {
+  \return \unsafe;
+}
+$d \leftarrow decide(v)$ \;
+\uIf{$v \sqsubseteq d$} {
+\return \unknown \;
+}
+// add decision to trail\;
+$\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
+$v \leftarrow v \meet d$ \; 
+$\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
+}
+\end{small}
+\caption{Abstract Model Search Procedure\label{Alg:model_search}}
+\end{algorithm2e}
 

### 2016-10-04 deductions.tex
description of algorithm

@@
 %
-\subsection{Iteration Strategy For Abstract Deduction Transformer}
-Algorithm~\label{Alg:iteration} presents the modelSearch procedure. 
-our deduction high-level sketch for the abstract deduction transformer.
-%The inputs are Input{Set of $\widehat{ded_{\varphi}}$ for $\varphi$ from Program $\mathcal{P}$, trail $\mathcal{T}$ and assertion $A$}
-$statement \in S$, $statement$ is a meet irreducible deduction transformer.
+\subsection{Abstract Model Search}
+Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
+In propositional case, a model search involves two steps -- 
+{\em deductions} using unit rule refines current partial assignments 
+and {\em decisions} to heuristically guess a value for an unassigned 
+literal.  A unit rule overapproximates a model transformer and deduction 
+computes a greatest fixed point over the partial assignments
+domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
+a greatest fixed point over meet irreducible deduction transformer in $S$.
+
+A {\em worklist} is initialized following a heuristics choice for iteration 
+strategies. A {\em forward} iteration strategy initializes the worklist with 
+transformers that has constants in the right-hand side. Whereas, a {\em backward} 
+iteration strategy starts with a set of assertions. On the other hand, a 
+{\em chaotic} iteration strategy initializes the worklist with set of all 
+transformers.  For every meet irreducible deduction transformer $s \in S$ 
+passed to the abstract domain, if the domain infers a new meet irreducible 
+$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
+trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
+$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
+is inferred by the domain, the algorithm terminates with UNSAT.  However,
+when a fixed-point is reached, the abstract value $v$ is checked for set of
+models.  This is done by checking whether the set of assignments in $v$ is 
+$\gamma$-complete~\cite{gamma}.  If $v$ is $\gamma$-complete, the abstract model 
+search terminates with SAT.  Otherwise, a decision is made by calling a 
+procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
+current abstract value $v$, then the corresponding propagation and reason trail,
+$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the algorithm terminates
+with UNKNOWN.      
+%    
 \begin{algorithm2e}[t]
@@
 \SetKwFunction{assign}{assign}
-\SetKwData{safe}{safe}
-\SetKwData{unsafe}{unsafe}
-\SetKwData{unknown}{unknown}
+\SetKwData{sat}{SAT}
+\SetKwData{unsat}{UNSAT}
+\SetKwData{unknown}{UNKNOWN}
 \SetKwData{true}{true}
@@
 abstract value $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
-\Output{The status (\safe or \unsafe or \unknown)}
+\Output{The status (\sat or \unsat or \unknown)}
 worklist $\leftarrow$ initialize\_worklist($S$) \;
@@
     worklist.delete() \;
-    \return \safe \;
+    \return \sat \;
   }
@@
 // check for counterexample\;
-\uIf{$v$ is gamma\_complete} {
-  \return \unsafe;
+\uIf{$v$ is $\gamma$-complete} {
+  \return \unsat;
 }
@@
 \end{small}
-\caption{Abstract Model Search Procedure\label{Alg:model_search}}
+\caption{Abstract Model Search Procedure\label{Alg:ms}}
 \end{algorithm2e}
-
+%
 \subsection{Forward Iteration Strategy}

### 2016-10-05 deductions.tex
git update

@@
 current abstract value $v$, then the corresponding propagation and reason trail,
-$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the algorithm terminates
-with UNKNOWN.      
+$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure terminates with UNKNOWN.      
 %    
@@
 %
-\subsection{Forward Iteration Strategy}
-For forward iteration on Figure~\ref{swssa}, we need atleast 
-one decision to deduce BOTTOM and prove safety. The deduction 
-happens from right to left in forward iteration strategy.
-
-The initialization of live variable and subsequent update depends on the iteration strategy.
-
-\subsection{Backward Iteration Strategy}
-
-\subsection{Chaotic Iteration Strategy}  
-The set of octagonal equilities corresponding to Figurei~\ref{sw-ssa}
-is shown below. For chaotic, no decision is required to deduce BOTTOM.
-
-\Omit {
-Let us consider an example. 
-\begin{figure}[t]
-\scriptsize
-\begin{tabular}{l|l|l}
-\hline
-C program & SSA & Octagons \\
-\hline
-\begin{lstlisting}[mathescape=true,language=C]
-int id(int a) {
-  return a;
-}
-int main() {
- int x,y,z;
- _Bool c;
- if (c)
-   x = -1;
- else
-   x = 2;
- int k = id(x);
- z = x * k;
- assert((z!=-2));
-}
-\end{lstlisting}
-&
-\begin{lstlisting}[mathescape=true,language=C]
-cond == (x == y)
-guard1 == (cond19 && guard0)
-x' == 1u + x
-x' == 1u + y || !guard1
-guard0 == TRUE
-\end{lstlisting}
-&
-\begin{lstlisting}[mathescape=true,language=C]
-
-\end{lstlisting}
-\\
-\hline
-\end{tabular}
-\caption{C Program, corresponding SSA and Octagonal inequalities}
-\label{swssa}
-\end{figure}
-}
-
-With forward iteration strategy using octagon domian, the program 
-is proved without any decision. However, with the forward strategy
-using interval domain, we just required one decision to prove the program.

### 2016-10-06 domains.tex
deduction definitions complete
clauses and literals

@@
 \section{Program and Abstract Domains}
-\section{Programs:}
+In this section, we define a program and the notion of concrete domain 
+and abstract domain. 
+\subsection{Programs:}
 %
@@
 elements. Formally, a {\em meet irreducible} $M$ in a complete lattice 
-structure $A$ is an element $M \in L$ that has the following property.
+structure $A$ is an element $M \in A$ that has the following property.
 \[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
@@
 unique octagonal inequalities, but it is very expensive to compute because of 
-cubic complexity in the number of program variables.
+cubic complexity in the number of program variables. An example of closure
+operation is shown below.
+\[((x-y \leq 4) \wedge (y-z \leq 5))^* \implies ((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]
 
@@
 \[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
+
+%========================
+\input{literals-clauses}
+%========================
 

### 2016-10-07 contributions.tex
decision

--- 2016-09-23 contributions.tex
@@
-\section{Contributions}\label{contribution}
-The goal of this paper is to instantiate ACDCL with 
-different relational and non-relational numerical abstract 
-domains and evaluate the performance of our solver with 
-propositional-level solvers. Experimental evaluation on different
-benchmarks shows that ACDCL is able to suitably exploit the 
-expressivity of richer abstract domains as well as precision 
-of CDCL architecture. This is attributed to the intelligent decision
-heuristics which exploits the high-level structure of the problem 
-combined with stronger deductions and clause learning mechanism aided 
-by the richer abstract domains.
+\section{Contributions}
+In this paper, we have the following contributions.
+\begin{enumerate}
+\item We present a procedure for lifting CDCL algorithm to 
+relational and non-relational numerical abstract domain.  
+To this end, we instatiate model search and conflict analysis 
+phases of CDCL algorithm over intervals and octagon abstract 
+domains.  We implement our own template-based numerical 
+domains that supports complementation property, essential for 
+such lifting.  To the best of our knowledge, this is the first 
+instantiation of CDCL algorithm with relational abstract domains.
+
+\item We present a new program analysis tool, {\em Titlis}, that 
+is a strict generalisation of propositional CDCL algorithm.  From 
+abstract interpretation perspective, {\em Titlis} uses decision 
+and learning to improve the precision of abstract transformers.  
+Thus, {\em Titlis} combines the efficiency of abstract domains 
+while being path-sensitive and bit-precise at the same time.
+
+\item  Experiments on several numerical and control-intensive benchmarks 
+shows that the proposed generalization of CDCL algorithm is able to 
+suitably exploit the expressivity of richer abstract domains as well 
+as precision of CDCL architecture.  This is attributed to the 
+intelligent decision heuristics which exploits the high-level 
+structure of the problem combined with stronger deductions and 
+clause learning mechanism aided by the richer abstract domains.
+\end{enumerate}

### 2016-10-07 decision.tex
paper update
literals
decision

@@
-\section{Decision Heuristics}\label{decision}
-Berkmin, longest-range, random, ordered, octagons
+\section{Decision}
+A decision $\mathcal{D}$ is a meet irreducible that refines the 
+current partial assignment.  A decision can be of the form 
+$\langle x \preceq c \rangle$ or $\langle x \succeq c \rangle$, 
+where $c$ is the bound and $x$ is a program variable.  A 
+decision $\mathcal{D}$ must always be consistent with respect 
+to the trail $\mathcal{T}$, that is, it must not contradict with 
+the elements in the trail.  Each variable in propositional CDCL 
+is assigned at most once at a particular decision level.  However, 
+a variable can be assigned several times in our lifting of CDCL to 
+numerical domains, each time with increasingly precise bounds.
+
+Let $\mathcal{V}$ be a set of all singletons and non-singletons 
+variables.  A singleton variable is one that has been assigned a 
+singleton value, for example, $\langle x:[1,1] \rangle$.  Otherwise, 
+the variable is non-singleton.  Given a set of variables $\mathcal{V}$, 
+a decision phase heuristically chooses a non-singleton branching 
+variable $v \in \mathcal{V}$, a bound $c$, and the polarity ($\succeq$ or 
+$\preceq$).
+
+We implemented several decision heuristics namely, {\em Berkmin}~\cite{}, 
+{\em longest-range}, {\em ordered}, {\em random}.  The {\em ordered} 
+decision heuristics creates an ordering among non-singleton variables, 
+making decision on conditional variables (variables that appear in 
+conditional branches) first before choosing numerical variables.  Thus, 
+this heuristics is path-sensitive which gives significant speed-up for 
+control-intensive benchmarks.

### 2016-10-07 contributions.tex
decision text

--- 2016-09-23 contributions.tex
@@
-\section{Contributions}\label{contribution}
-The goal of this paper is to instantiate ACDCL with 
-different relational and non-relational numerical abstract 
-domains and evaluate the performance of our solver with 
-propositional-level solvers. Experimental evaluation on different
-benchmarks shows that ACDCL is able to suitably exploit the 
-expressivity of richer abstract domains as well as precision 
-of CDCL architecture. This is attributed to the intelligent decision
-heuristics which exploits the high-level structure of the problem 
-combined with stronger deductions and clause learning mechanism aided 
-by the richer abstract domains.
+\section{Contributions}
+In this paper, we make the following contributions.
+\begin{enumerate}
+\item We present a novel program analysis technique based on 
+abstract conflict driven clause learning algorithm~\cite{dhk2013-popl}.
+To this end, we present a procedure for lifting CDCL algorithm to 
+richer lattice structure such as {\em interval} and {\em octagon} 
+abstract domains.
+
+\item To enable lifting of CDCL algorithm, we implement 
+our own template-based numerical abstract domains that 
+supports complementation property.  Our proposed technique 
+is embodied in our tool called, {\em Titlis}.  To the best 
+of our knowledge, this is the first instantiation of CDCL 
+algorithm with relational abstract domains.
+
+\item  Experiments on a set of benchmarks drawn from bit-vector 
+regression suite in sv-comp, numerical programs and few 
+control-intensive benchmarks from hardware domain shows that 
+the strict generalization of CDCL algorithm effectively combines 
+the efficiency of abstract domains while being path-sensitive and 
+bit-precise at the same time.  This is attributed to the intelligent 
+decision heuristics which exploits the high-level structure of the 
+problem combined with stronger deductions and clause learning 
+mechanism aided by the richer abstract domains.
+\end{enumerate}

### 2016-10-07 decision.tex
decision text

@@
 where $c$ is the bound and $x$ is a program variable.  A 
-decision $\mathcal{D}$ must always be consistent with respect 
-to the trail $\mathcal{T}$, that is, it must not contradict with 
+decision must always be consistent with respect to the 
+trail $\mathcal{T}$, that is, it must not contradict with 
 the elements in the trail.  Each variable in propositional CDCL 
@@
 
-We implemented several decision heuristics namely, {\em Berkmin}~\cite{}, 
-{\em longest-range}, {\em ordered}, {\em random}.  The {\em ordered} 
-decision heuristics creates an ordering among non-singleton variables, 
-making decision on conditional variables (variables that appear in 
-conditional branches) first before choosing numerical variables.  Thus, 
-this heuristics is path-sensitive which gives significant speed-up for 
-control-intensive benchmarks.
+{\em Titlis} support several decision heuristics namely, {\em ordered} 
+{\em longest-range}, {\em random}, {\em Berkmin}~\cite{} decision heuristics.  
+The {\em ordered} decision heuristics creates an ordering among non-singleton 
+variables, thereby making decisions on conditional variables (variables that 
+appear in conditional branches) first before choosing numerical variables.  
+This variable ordering is path-sensitive as it gives an effect of trace
+partitioning.  A {\em longest-range} heuristics simply keeps track of the
+interval $[l,u]$ range ($r=u-l$) of a variable and chooses a variable with 
+the longest range value.  This ensures a fairness policy in selecting a 
+variable since it guarantees that the intervals of variables are uniformly 
+restricted.  A {\em random} decision heuristics arbitrarily picks a variable 
+for making decision.  Whereas, a {\em berkmin} decision heuristics is inspired 
+from decision heuristics used in Berkmin~\cite{eugoldberg07} SAT solver, which 
+keeps track of the activity of a variable that participate in conflict clauses 
+as well as variables that actively contribute to conflicts but do not explicitly 
+appear in conflict clauses.  The set of conflict clauses is organized 
+chronologically with the top clause as the one deduced in the last.  A 
+branching variable is chosen among the free variables whose literals are 
+in the top unsatisfied conflict clause.  A similar decision heuristics is 
+also implemented in Chaff~\cite{chaff} SAT solver, that computes the activity 
+of a variable as the number of occurrences of that variable in conflict 
+clauses only.
+
+A bound of a variable is heuristically chosen to be an approximation of the 
+arithmetic average of the current bounds.  However, the polarity ($\preceq$ or
+$\succeq$) of a variable is chosen randomly.

### 2016-10-07 literals-clauses.tex
literals-clauses

\subsection{Literals and Clauses for Abstract CDCL}
%
\textit{Literal:} A Literal is a meet irreducible which specify that 
for certain program variable, there is a certain bound. An example of 
literal in interval domain is $x \in [0,5]$. An example of octagonal 
literal is given by $((x-y \leq 0) \wedge (y-x \leq 0))$.

Let I and J be two interval literals of the form $x \in [l, u]$ with $l$ 
as lower bound and $u$ as upper bound. We say $I \leq J$ or {\em (I leq J)} 
iff $(I.u \leq J.u \wedge I.l \geq J.l)$.  Whereas, {\em I disjoint J} iff 
$(I.u < J.l \vee  I.l > J.u)$.

Similarly, let $O_1$ and $O_2$ be two octagonal literals of the form 
$x-y \leq c$.  Then, $O_1 \leq O_2$ {\em ($O_1$ leq $O_2$)} iff 
$O_1.c \leq O_2.c$.  Two octagons are disjoint ({\em $O_1$ disjoint $O_2$}) 
iff $O_1:= (x-y < c)$ and $O_2:= (x-y > c)$.

\textit{Clause:} A clause is a disjunction of one or more meet irreducibles. 
An example clause is given by $(x \geq 0 \vee y \geq 5 \vee y+z \leq 10)$.

Let $x$, $x'$ be the literals in clause $C$ and current partial assignment 
$A$ respectively.  We call a literal $x \in C$ to be {\em satisfiable} with 
respect to the literal $x' \in A$ iff $(x \leq x')$.  We call $x$ to be  
{\em unsatisfiable} with respect to $x'$ iff $!(x \leq x') \wedge$ $!$(x disjoint $x'$).
A literal is said to be {\em contradicting} iff $!(x \leq x') \wedge$ (x disjoint $x'$).

Similarly, a clause may be catagorized into four classes -- satisfiable,
unsatisfiable, conflicting, unit. \\ 
\textit{Satisfiable clause:}
A clause $C$ is said to be {\em satisfiable} with respect to a current partial 
assignment $A$ if at least one literal in $C$ is satisfiable. For example, 
consider a clause $C=(x<4 \vee y>10)$ and let the current partial assignment 
be $A: x \in [0,3]$. Then $C$ is a satisfiable clause, where $x<4$ is a
satisfiable literal.

\textit{Conflicting clause:}
If all literals in a clause are contradicting, then the clause is said to
be {\em conflicting}. For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. 
Let the current partial assignment be $A: x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[17,32]$, then $C$ is a conflicting clause where all literals in $C$ are
contradicting with respect to $A$.

\textit{Unsatisfiable clause:}
A clause $C$ is said to be {\em unsatisfiable} with respect to a current partial 
assignment $A$ if there exists no satisfiable literal in $C$ or some literals in 
$C$ are unsatisfiable and the rest are contradicting. For example, consider a 
clause $C=(x<4 \vee y>10 \vee z<15)$.  Let the current partial assignment be 
$A: x \in [3,10]$, $y \in [8,10]$ and $z \in [12,20]$, then $C$ is a unsatisfiable 
clause, where $x<4$ and $z<15$ are unsatisfiable literal and $y>10$ is
contradicting literal.

\textit{Unit clause:}
Clause $C$ is unit if all literals but one is contradicting in $C$. 
For example, consider a clause $C=(x<4 \vee y>10 \vee z<15)$. Let the 
current partial assignment be $A: x \in [5,13]$ and $y \in [-2,9]$ and $z \in
[10,12]$, then $C$ is a unit clause where the unit literal is $z$.

\Omit {
\textit{Boolean constraint propagation (BCP):} BCP is the repeated 
application of unit rule. This corresponds to computing the greatest fixed point.

\textit{Characteristics of Conflict clause}
\begin{enumerate}
\item A conflict clause must include asserting cuts. An asserting cut is a cut
that contain exactly one node at the current decision level. Assertion cuts yields 
clauses that can be used to derive new information after backtracking.

\item A conflict clause must be UNIT after backtracking.

%\item There can be multiple cuts and hence multiple UIPs. In other words, there
%can be multiple incomparable reasons for a conflict. But conflict analysis
%procedure choses one that is asserting.

%\item The conflict clause should be made false by the current partial assignment
%and thus exclude an assignment leading to conflict. 
\end{enumerate}

\textit{Backjumping:}
The backjumping level is defined by the literal of the conflict clause assigned
at the level that is the closest to the conflict one. In other words, the
backjumping level is the level closest to the root (decision level 0)  where the
conflict clause is still unit. If a conflict clause is "globally"  unit, then
the backjumping level is the root of the search tree.
}

### 2016-10-08 algorithm.tex
paper update
minor changes

@@
-\section{Lifting CDCL to Programs}\label{algorithm}
+\section{Abstract Conflict Driven Clause Learning}
+\rmcmt{safety as satisfiability}
 %
@@
 %
+\Omit {
 Static program analysis based on abstract
@@
 of the analysis.  
-\Omit{
-We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
 }
-
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
+Given a \rmcmt{formula (F)} and an abstract valuation $v = \top$,  
+ACDCL returns {\em safe} if $F$ is unsatisfiable and P is safe. Else, 
+ACDCL returns a model that satisfies $F$ and P is {\em unsafe}.  The 
+procedure $deduce(P)$ is similar to BCP step in SAT solvers where 
+it computes least fix-point with strongest post-condition using 
+forward analysis. If the result of $deduce(P)$ 
 is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 

### 2016-10-08 deductions.tex
clauses and literals
decision
minor changes

@@
 %
-\subsection{Abstract Model Search}
-Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
-In propositional case, a model search involves two steps -- 
-{\em deductions} using unit rule refines current partial assignments 
-and {\em decisions} to heuristically guess a value for an unassigned 
-literal.  A unit rule overapproximates a model transformer and deduction 
-computes a greatest fixed point over the partial assignments
-domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
-a greatest fixed point over meet irreducible deduction transformer in $S$.
-
-A {\em worklist} is initialized following a heuristics choice for iteration 
-strategies. A {\em forward} iteration strategy initializes the worklist with 
-transformers that has constants in the right-hand side. Whereas, a {\em backward} 
-iteration strategy starts with a set of assertions. On the other hand, a 
-{\em chaotic} iteration strategy initializes the worklist with set of all 
-transformers.  For every meet irreducible deduction transformer $s \in S$ 
-passed to the abstract domain, if the domain infers a new meet irreducible 
-$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
-trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
-$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with UNSAT.  However,
-when a fixed-point is reached, the abstract value $v$ is checked for set of
-models.  This is done by checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{gamma}.  If $v$ is $\gamma$-complete, the abstract model 
-search terminates with SAT.  Otherwise, a decision is made by calling a 
-procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
-current abstract value $v$, then the corresponding propagation and reason trail,
-$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure terminates with UNKNOWN.      
 %    
-\begin{algorithm2e}[t]
+\begin{algorithm2e}[htbp]
 \DontPrintSemicolon
@@
 \end{small}
-\caption{Abstract Model Search Procedure\label{Alg:ms}}
+\caption{Abstract Model Search\label{Alg:ms}}
 \end{algorithm2e}
 %
+\subsection{Abstract Model Search}
+Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
+In propositional case, a model search involves two steps -- 
+{\em deductions} using unit rule refines current partial assignments 
+and {\em decisions} to heuristically guess a value for an unassigned 
+literal.  A unit rule overapproximates a model transformer and deduction 
+computes a greatest fixed point over the partial assignments
+domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
+a greatest fixed point over meet irreducible deduction transformer in $S$.
+
+A {\em worklist} is initialized following a heuristics choice for iteration 
+strategies. A {\em forward} iteration strategy initializes the worklist with 
+transformers that has constants in the right-hand side. Whereas, a {\em backward} 
+iteration strategy starts with a set of assertions. On the other hand, a 
+{\em chaotic} iteration strategy initializes the worklist with set of all 
+transformers.  For every meet irreducible deduction transformer $s \in S$ 
+passed to the abstract domain, if the domain infers a new meet irreducible 
+$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
+trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
+$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
+is inferred by the domain, the algorithm terminates with UNSAT.  However,
+when a fixed-point is reached, the abstract valuation $v$ is checked for set of
+models.  This is determined by checking whether the set of assignments in $v$ is 
+$\gamma$-complete~\cite{gamma}.  This can be achieved by checking if there is a 
+concrete solution $c$ such that the inequality $c(x) \subseteq \gamma \circ
+v(x)$ holds for every meet irreducible $x$.  If $v$ is $\gamma$-complete, that is, the 
+abstract valuation $v$ is abstractly satisfying, the abstract model 
+search terminates with SAT.  Otherwise, a decision is made by calling a 
+procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
+current abstract value $v$, then the corresponding propagation and reason trail,
+$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure 
+terminates with UNKNOWN.

### 2016-10-08 domains.tex
minor changes

@@
 
+\begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
+mapping of variables to an element of abstract domain, for example 
+a mapping of variable $x$ to an interval environment is given by 
+$\langle x \mapsto [2,5] \rangle$ or a mapping of $\{x,y\}$ to octagon 
+environment is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
+An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
+value or to $\bot$.  
+\end{definition}
+
 %========================

### 2016-10-08 learning.tex
paper update
Model search
title
minor changes

@@
-\section{Clause Learning with ACDL}\label{learning}
+\section{Clause Learning in Abstract Lattice}
+Conflict graph for Intervals
+Conflict graph for Octagons
+
+\textit{Characteristics of Conflict clause}
+\begin{enumerate}
+\item A conflict clause must include asserting cuts. An asserting cut is a cut
+that contain exactly one node at the current decision level. Assertion cuts yields 
+clauses that can be used to derive new information after backtracking.
+
+\item A conflict clause must be UNIT after backtracking.
+
+\item There can be multiple cuts and hence multiple UIPs. In other words, there
+can be multiple incomparable reasons for a conflict. But conflict analysis
+procedure choses one that is asserting.
+
+\item The conflict clause should be made false by the current partial assignment
+and thus exclude an assignment leading to conflict. 
+\end{enumerate}
+
+\textit{Backjumping:}
+The backjumping level is defined by the literal of the conflict clause assigned
+at the level that is the closest to the conflict one. In other words, the
+backjumping level is the level closest to the root (decision level 0)  where the
+conflict clause is still unit. If a conflict clause is "globally"  unit, then
+the backjumping level is the root of the search tree.
 1.DPLL style -- chronological backtracking \\

### 2016-10-08 literals-clauses.tex
minor changes

@@
 
-\textit{Characteristics of Conflict clause}
-\begin{enumerate}
-\item A conflict clause must include asserting cuts. An asserting cut is a cut
-that contain exactly one node at the current decision level. Assertion cuts yields 
-clauses that can be used to derive new information after backtracking.
-
-\item A conflict clause must be UNIT after backtracking.
-
-%\item There can be multiple cuts and hence multiple UIPs. In other words, there
-%can be multiple incomparable reasons for a conflict. But conflict analysis
-%procedure choses one that is asserting.
-
-%\item The conflict clause should be made false by the current partial assignment
-%and thus exclude an assignment leading to conflict. 
-\end{enumerate}
-
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
 }

### 2016-10-09 deductions.tex
introduction

@@
-\section{Deduction}
+\section{Abstract Model Search}
 In a typical abstract interpretation based galois-connection settings over 
@@
   $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
-  // check if the domain deduce BOTTOM \;
   \uIf{$meet\_irrd == \bot$} {
     $\mathcal{R}[\bot] \leftarrow s$ \;
-    // empty the worklist \;
     worklist.delete() \;
@@
   {
-    // add deduction to trail \;
     $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
@@
     $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    // update the worklist \;
     worklist.update(v, s) \; 
@@
 }
-// check for counterexample\;
 \uIf{$v$ is $\gamma$-complete} {
@@
 }
-// add decision to trail\;
 $\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
@@
 $\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
+worklist.update(v, d) \; 
 }
@@
 %
-\subsection{Abstract Model Search}
+\subsection{Algorithm for Abstract Model Search}
 Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
@@
 procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
-current abstract value $v$, then the corresponding propagation and reason trail,
-$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure 
+current abstract value $v$, then the worklist is updated with all transformers that are dependant on the current decision $d$ along with the update of the corresponding propagation and reason trail,
+$\mathcal{T}$ and $\mathcal{R}$. Else, the model search procedure 
 terminates with UNKNOWN.

### 2016-10-09 introduction.tex
definitions added
minor
introduction

@@
 %
+A standard static analyses for program safety aims to compute 
+invariants by applying transformers that refines an abstract 
+value.  Most lattice structure used in static analysis have 
+meet and join operations but lack negations.  The meet operation 
+precisely model conjunction, and the join operation over-approximate 
+disjunction. Thus, the precision loss in static analyzers is often 
+due to imprecise join operations. The loss in precision is overcome 
+by using richer abstract domains or equipping the analysis with disjunction --
+which is often very expensive since the analysis exhibit case enumeration 
+behavior.
 
-Most lattice structure used in static analysis have meet and join operations 
-but lack negations. The meet operation precisely model conjunction, and
-the join operation over-approximate disjunction. Thus, the precision loss in 
-static analyzers is often due to imprecise join operations. The loss in precision is
-overcome by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration behavior.
+On the other hand, the combination of precise conjunction in partial assignments 
+domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
+solvers to reason about disjunction without enumerating cases.  
+Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
+a formal mechanism for applying abstract interpretation to solve the 
+satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
+of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
+in terms of deduction and abduction transformers, fixed-point, interpolation 
+and extrapolation operators on lattices.  The partial assignments in the 
+CDCL algorithm are an abstract domain, unit rule is the best abstract 
+transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
+over the abstract domain, decision is dual widening and clause learning can be 
+viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
+lifts learning techniques in SAT solver to operate on non-distributive 
+lattice -- thus equipping the analysis with negation. This enables the 
+analyzer to refine the analysis and prevent enumeration behavior.
 
-On the other hand, the combination of precise conjunction in partial assignments domain 
-with learning allow CDCL solvers to reason about disjunction without 
-enumerating cases. Learning can be viewed as synthesizing an abstract transformer for 
-negation. ACDCL lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the analyzer to refine 
-the analysis and prevent enumeration behavior.
+The instantiation of ACDCL has so far been limited to interval domain only.  In this paper, 
+we identify specific properties of the CDCL algorithm that are necessary to 
+lift the propositional skeleton of CDCL to other richer lattice structures 
+and relational abstract domains such as octagons, polyhedra, zones and equalities.  
+To this end, we present a characterization of abstract model search procedure and 
+conflict analysis procedure that lift the abstract deduction transformer and the 
+first-UIP algorithm to relational lattice structures.

### 2016-10-09 contributions.tex
contribution

@@
 abstract domains.
-
-\item To enable lifting of CDCL algorithm, we implement 
-our own template-based numerical abstract domains that 
-supports complementation property.  Our proposed technique 
-is embodied in our tool called, {\em Titlis}.  To the best 
-of our knowledge, this is the first instantiation of CDCL 
-algorithm with relational abstract domains.
 

### 2016-10-09 introduction.tex
contribution

@@
 %
-A standard static analyses for program safety aims to compute 
+A standard static analyses~\cite{se2011} for program safety aims to compute 
 invariants by applying transformers that refines an abstract 
@@
 Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
-a formal mechanism for applying abstract interpretation to solve the 
+a formal mechanism for applying abstract interpretation~\cite{se2011} to solve the 
 satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
@@
 first-UIP algorithm to relational lattice structures.
+
+Clause learning SAT solvers requires a weak complementation property 
+of the partial assignments domain.  However,  the most commonly 
+used APRON~\cite{apron} library that implements abstract domains, is specialized 
+for abstract interpretation framework which does not require complementation 
+operator.  Thus, we implement our own template-based numerical abstract 
+domains that support complementation property.  Our proposed technique 
+is embodied in our tool called, {\em Titlis}.  To the best of our 
+knowledge, this is the first instantiation of CDCL 
+algorithm with relational abstract domains.

### 2016-10-09 algorithm.tex
acdcl algorithm

@@
 \section{Abstract Conflict Driven Clause Learning}
-\rmcmt{safety as satisfiability}
+In a typical abstract interpretation based galois-connection settings over 
+an over-approximate domain, every concrete elements have a unique over-approximate 
+representation in the abstract.  Likewise, every concrete transformer is
+over-approximated by a unique abstract transformer.  We now define a abstract 
+deduction transformer.
+
+\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$ is a sound approximation of a concrete model transformer
+$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+$m \in A$ is a meet irreducible.   
+\end{definition}
+
+Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
+\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+\rangle$.  An abstract deduction transformer is typically computed in the form 
+of strongest post-condition or a weakest pre-condition of a formula in the 
+abstract domain.
+
+A meet decomposition of the outcome of abstract deduction transformer 
+is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
+\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+precisely complementable.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
+
+For a program with $N$ variables, let $L$ be the total number of 
+meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
+maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
+$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
+of all octagonal inequalities in the set $L$.
+%\rmcmt{safety as satisfiability}
 %
@@
 %\SetKwFunction{isf}{isFeasible}
+\SetKwData{conflict}{CONFLICT}
 \SetKwData{safe}{safe}
@@
 \SetKwInOut{Output}{output}
+\SetKwFor{Loop}{Loop}{}{}
 \SetKw{KwNot}{not}
 \begin{small}
-\Input{Program $\mathcal{P}$ with properties specified with $assert(c)$}
+\Input{Set $S$ of abstract deduction transformer, 
+$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract domain $A$}
 \Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-\assign result=\deduce(P)\;
-\uIf{$result == \perp$} {
+$v \leftarrow \top$ \;
+$\mathcal{T} \leftarrow \epsilon$ \;
+$\mathcal{R} \leftarrow \emptyset$ \;
+$result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;
+\uIf{$result$ == \conflict} {
   \return \safe \;
-}
-\uElse
-{
-  \uIf{is\_gamma\_complete()} {
-    \return \unsafe \;
-  }
 }
@@
 {
-  \assign decision\_variable = decision\_heuristics($P$)\;
-  %\assign decision\_container = decision\_container $\cup$ decision\_variable\; 
-  \uIf{decision\_variable == NULL}
-  {
-    return \unknown\; 
+  \uIf{$v$ is $\gamma$-complete} {
+    \return \unsafe;
   }
-  \uElse
-  {
-    \assign result=\deduce(P)\;
-    \uIf{result == UNKNOWN}
-    {
-      \uIf{is\_gamma\_complete()} {
-        \return \unsafe \;
-      }
-      \uElse 
-      {
-        \continue\;
-      }
-    }
-    \uElse
-    {
-      \Do{$result == \perp$} {
+  $d \leftarrow decide(v)$ \;
+  $v \leftarrow v \meet d$ \; 
+  $\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
+  $\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
+  $result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
+  \Do{$result == \conflict$} {
+    learn() \;
     \uIf{$\neg$ conflict\_analysis()} {
@@
     }
-        \assign result=\deduce(P)\;
-      }
-    }
+    result $\leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;
   }
@@
 \end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
+\caption{Abstract Conflict Driven Clause Learning $\{ACDCL(S)\}$ \label{Alg:acdcl}}
 \end{algorithm2e}
@@
 ACDCL for unbounded proofs.
+
+A decision is made by calling a procedure $decide(v)$ which 
+returns a meet irreducible $d$.  If $d$ refines the 
+current abstract value $v$, then the worklist is updated with 
+all transformers that are dependant on the current decision 
+$d$ along with the update of the corresponding propagation 
+and reason trail, $\mathcal{T}$ and $\mathcal{R}$.

### 2016-10-09 deductions.tex
acdcl algorithm

--- 2016-10-05 deductions.tex
@@
-\section{Deduction}
-In a typical abstract interpretation based galois-connection settings over 
-an over-approximate domain, every concrete elements have a unique over-approximate 
-representation in the abstract.  Likewise, every concrete transformer is
-over-approximated by a unique abstract transformer.  We now define a abstract 
-deduction transformer.
-
-\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
-transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
-domain $A$ is a sound approximation of a concrete model transformer
-$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
-$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
-$m \in A$ is a meet irreducible.   
-\end{definition}
-
-Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
-an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
-\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
-\rangle$.  An abstract deduction transformer is typically computed in the form 
-of strongest post-condition or a weakest pre-condition of a formula in the 
-abstract domain.
-
-A meet decomposition of the outcome of abstract deduction transformer 
-is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
-\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
-$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
-\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
-precisely complementable.
-
-\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
-element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
-\end{definition}
-
-For a program with $N$ variables, let $L$ be the total number of 
-meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
-maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
-of all octagonal inequalities in the set $L$.
+\section{Abstract Model Search}
 %
@@
 %
-\subsection{Abstract Model Search}
-Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
-In propositional case, a model search involves two steps -- 
-{\em deductions} using unit rule refines current partial assignments 
-and {\em decisions} to heuristically guess a value for an unassigned 
+%    
+\begin{algorithm2e}[t]
+\DontPrintSemicolon
+\SetKwFunction{deduce}{deduce}
+\SetKwFunction{print}{print}
+\SetKwFunction{return}{return}
+\SetKwFunction{continue}{continue}
+\SetKwFunction{assign}{assign}
+\SetKwData{sat}{SAT}
+\SetKwData{conflict}{CONFLICT}
+\SetKwData{unsat}{UNSAT}
+\SetKwData{unknown}{UNKNOWN}
+\SetKwData{true}{true}
+\SetKwInOut{Input}{input}
+\SetKwInOut{Output}{output}
+\SetKwFor{Loop}{Loop}{}{}
+\SetKw{KwNot}{not}
+\begin{small}
+\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
+abstract valuation $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
+\Output{The status (\sat or \conflict or \unknown)}
+worklist $\leftarrow$ initialize\_worklist($S$) \;
+\Loop{}
+{
+// Greatest fixed point \;
+\While{$!worklist.empty()$} 
+{
+  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
+  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
+  \uIf{$meet\_irrd == \bot$} {
+    $\mathcal{R}[\bot] \leftarrow s$ \;
+    worklist.delete() \;
+    \return \conflict \;
+  }
+  \uElse
+  {
+    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
+    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
+    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
+    worklist.update(v, s) \; 
+  }
+}
+\uIf{$v$ is $\gamma$-complete} {
+  \return \sat;
+}
+\uElse {
+ \return \unknown \;
+}
+}
+\end{small}
+\caption{Abstract Model Search $\{deduce(S,v,\mathcal{T},\mathcal{R})\}$ \label{Alg:ms}}
+\end{algorithm2e}
+%  
+A model search procedure in a SAT solver involves two steps -- {\em deductions} 
+using unit rule refines current partial assignments and 
+{\em decisions} to heuristically guess a value for an unassigned 
 literal.  A unit rule overapproximates a model transformer and deduction 
 computes a greatest fixed point over the partial assignments
-domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
-a greatest fixed point over meet irreducible deduction transformer in $S$.
-
+domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
+that computes a greatest fixed point over meet irreducible deduction 
+transformer in $S$.  
+%
+\subsection{Algorithm for Deduction Phase}
+%
+Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.
 A {\em worklist} is initialized following a heuristics choice for iteration 
@@
 $\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with UNSAT.  However,
-when a fixed-point is reached, the abstract value $v$ is checked for set of
-models.  This is done by checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{gamma}.  If $v$ is $\gamma$-complete, the abstract model 
-search terminates with SAT.  Otherwise, a decision is made by calling a 
-procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
-current abstract value $v$, then the corresponding propagation and reason trail,
-$\mathcal{T}$ and $\mathcal{R}$, are updated. Else, the model search procedure terminates with UNKNOWN.      
-%    
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwData{sat}{SAT}
-\SetKwData{unsat}{UNSAT}
-\SetKwData{unknown}{UNKNOWN}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKwFor{Loop}{Loop}{}{}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
-abstract value $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
-\Output{The status (\sat or \unsat or \unknown)}
-worklist $\leftarrow$ initialize\_worklist($S$) \;
-\Loop{}
-{
-// Greatest fixed point \;
-\While{$!worklist.empty()$} 
-{
-  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
-  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
-  // check if the domain deduce BOTTOM \;
-  \uIf{$meet\_irrd == \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \;
-    // empty the worklist \;
-    worklist.delete() \;
-    \return \sat \;
-  }
-  \uElse
-  {
-    // add deduction to trail \;
-    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
-    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
-    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    // update the worklist \;
-    worklist.update(v, s) \; 
-  }
-}
-// check for counterexample\;
-\uIf{$v$ is $\gamma$-complete} {
-  \return \unsat;
-}
-$d \leftarrow decide(v)$ \;
-\uIf{$v \sqsubseteq d$} {
-\return \unknown \;
-}
-// add decision to trail\;
-$\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
-$v \leftarrow v \meet d$ \; 
-$\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
-}
-\end{small}
-\caption{Abstract Model Search Procedure\label{Alg:ms}}
-\end{algorithm2e}
-%
+is inferred by the domain, the algorithm terminates with CONFLICT.  However,
+when a fixed-point is reached, the abstract valuation $v$ is checked for set of
+models.  This is determined by checking whether the set of assignments in $v$ is 
+$\gamma$-complete~\cite{gamma}.  This can be achieved by checking if there is a 
+concrete solution $c$ such that the inequality $c(x) \subseteq \gamma \circ
+v(x)$ holds for every meet irreducible $x$.  If $v$ is $\gamma$-complete, that is, the 
+abstract valuation $v$ is abstractly satisfying, the abstract model 
+search terminates with SAT.  Otherwise, the algorithm returns UNKNOWN and the 
+model search procedure makes a new decision.

### 2016-10-09 domains.tex
acdcl algorithm

@@
 APRON library is specialized for the abstract interpretation 
-framework which does not require complementation 
-operator.  To this end, we implement our own template-based 
-octagon domain which supports complementation operator.  
-For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
+framework which does not require complementation property.  
+To this end, we implement our own template-based 
+interval, octagon and equality domains which supports 
+complementation operator.  For example, the octagon in 
+Figure~\ref{octagon} can be written as a conjunction of:
 \[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 

### 2016-10-09 introduction.tex
acdcl algorithm

@@
 %
-A standard static analyses~\cite{se2011} for program safety aims to compute 
+Static program analysis based on abstract 
+interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
+verify certain classes of properties for safety-critical systems.  A 
+standard static analyses~\cite{se2011} for program safety aims to compute 
 invariants by applying transformers that refines an abstract 
@@
 Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
-a formal mechanism for applying abstract interpretation~\cite{se2011} to solve the 
+a formal recipe for applying abstract interpretation~\cite{se2011} to solve the 
 satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
@@
 
-The instantiation of ACDCL has so far been limited to interval domain only.  In this paper, 
-we identify specific properties of the CDCL algorithm that are necessary to 
-lift the propositional skeleton of CDCL to other richer lattice structures 
-and relational abstract domains such as octagons, polyhedra, zones and equalities.  
-To this end, we present a characterization of abstract model search procedure and 
-conflict analysis procedure that lift the abstract deduction transformer and the 
+The instantiation of ACDCL has so far been limited to interval domains only.  
+In this paper, we identify specific properties of the CDCL algorithm that 
+are necessary to lift the propositional skeleton of CDCL to other richer 
+lattice structures and relational abstract domains such as octagons, 
+polyhedra, zones and equalities.  To this end, we present a 
+characterization of abstract model search procedure and conflict 
+analysis procedure that lift the abstract deduction transformer and the 
 first-UIP algorithm to relational lattice structures.

### 2016-10-09 algorithm.tex
acdcl algorithm

--- 2016-09-22 algorithm.tex
@@
-\section{Lifting CDCL to Programs}\label{algorithm}
+\section{Abstract Conflict Driven Clause Learning}
+In a typical abstract interpretation based galois-connection settings over 
+an over-approximate domain, every concrete elements have a unique over-approximate 
+representation in the abstract.  Likewise, every concrete transformer is
+over-approximated by a unique abstract transformer.  We now define a abstract 
+deduction transformer.
+
+\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$ is a sound approximation of a concrete model transformer
+$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+$m \in A$ is a meet irreducible.   
+\end{definition}
+
+Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
+\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+\rangle$.  An abstract deduction transformer is typically computed in the form 
+of strongest post-condition or a weakest pre-condition of a formula in the 
+abstract domain.
+
+A meet decomposition of the outcome of abstract deduction transformer 
+is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
+\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+precisely complementable.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
+
+For a program with $N$ variables, let $L$ be the total number of 
+meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
+maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
+$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
+of all octagonal inequalities in the set $L$.
+%\rmcmt{safety as satisfiability}
+\subsection{Static Analyses Equations for Safety}
+Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
 %
@@
 \end{small}
-\caption{Abstract Conflict Driven Clause Learning\label{Alg:acdcl}}
+\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}
 \end{algorithm2e}
 %
+\Omit {
 Static program analysis based on abstract
@@
 of the analysis.  
-\Omit{
-We observe that the abstract domains necessary for 
-hardware property checking are -- booleans, constants, interval, equality 
-and octagons domains.
 }
+Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
+The algorithm begins with a set $S$ of abstract deduction transformer 
+$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
+and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
+respectively.  The propagation trail stores all meet irreducibles inferred 
+by the abstact model search phase (deductions and decisions), while the 
+reason trail maps elements of propagation trail to transformers that were 
+used to drive them.
 
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a program $\mathcal{P}$ with properties specified as $assert(c)$,
-ACDCL terminate with {\em safe} if there exist no counterexample trace violating the 
-assertion or {\em unsafe} otherwise. The procedure $deduce(P)$ is similar 
-to BCP step in SAT solvers where it computes least fix-point with 
-strongest post-condition using forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
+The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
+fixed point over the transformers in F that refines the abstract valuation 
+$v$, similar to BCP step in SAT solvers.  If the result of $deduce$ 
+returns {\em CONFLICT} ($\perp$), the algorithm terminates with {\em safe}. 
+Else, the analysis enters into the while($true$) loop and makes a new decision 
+by a call to $decide(v)$ which returns a new meet irreducible $d$ that is
+consistent to the state of the solver.  If $d$ refines the current abstract 
+value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 
+and $\mathcal{R}$ are updated.
+
+For example, a decision in interval domain restrict the range of 
+intervals for variables, so the analysis jumps under a 
 greatest fixed-point. Note that the widening operation in abstract interpretation 
 jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
+{\em dual widening}.  The procedure $deduce$ is called next to infer new 
+set of meet irreducibles for the current decision.  The model search phase 
+alternates between the decision and deduction until the result of deduction 
+is either $\gamma$-complete~\cite{gamma} or the deduction leads to a {\em CONFLICT}.  
+If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
+algorithm returns the error trace $v$ and terminates with UNSAFE or the 
+current abstraction is insufficient to determine the satisfiability of $\varphi$.  
+However, in case of {\em CONFLICT}, the algorithm enters in to 
+$conflict\_analysis()$ phase to learn the reason for {\em conflict}.  
+There can be multiple incomparable reasons for conflict -- based on 
+the choice of Unique Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically 
+choose one reason.  A learnt clause must include asserting cuts which 
+guarantees derivation of new meet irreducibles after backtracking. The clause 
+learning and backtracking continues as long as the result of deduction 
+is conflicting ($\perp$), that is, the abstract valuation $v$ does not
+abstractly satisfy the formula.  If no further backtrack is possible, 
+then the algorithm terminates and $\varphi$ is {\em SAFE}. Else, the 
+algorithm makes a new decision and the above process is repeated 
+until a real counterexample is obtained or the algorithm backtracks to 
+decision level 0 after a conflict in which case it returns {\em SAFE}.

### 2016-10-09 algorithm.tex
static analysis equation

--- 2016-10-09 algorithm.tex
@@
 %\rmcmt{safety as satisfiability}
+\subsection{Static Analysis Equations for Safety}
+Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+perform safety analysis by computing fixed point to infer invariants 
+over program variables.  However, bounded model checking (BMC) tries to search 
+for a counterexample in a bounded execution trace, by symbolically executing a 
+program up to a finite bound.  
+Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+For a set $N$ of program analysis constraints defined over a set 
+of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+$\mathcal{P}$ and a concrete domain $\powerset{Env}$, the safety equation 
+is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+and the \rmcmt{final valuation} of constraint variables.
+\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
 %
@@
 \end{small}
-\caption{Abstract Conflict Driven Clause Learning $\{ACDCL(S)\}$ \label{Alg:acdcl}}
+\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}
 \end{algorithm2e}
@@
 }
+\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-Given a \rmcmt{formula (F)} and an abstract valuation $v = \top$,  
-ACDCL returns {\em safe} if $F$ is unsatisfiable and P is safe. Else, 
-ACDCL returns a model that satisfies $F$ and P is {\em unsafe}.  The 
-procedure $deduce(P)$ is similar to BCP step in SAT solvers where 
-it computes least fix-point with strongest post-condition using 
-forward analysis. If the result of $deduce(P)$ 
-is BOTTOM ($\perp$), the algorithm terminates with {\em safe}. Else, the 
-{\em gamma completeness} check~\cite{sas01} is performed to determine if it is a 
-real counterexample. If the {\em gamma completeness} check is successful, then the 
-counterexample is real. Else, the algorithm enters into the while($true$)
-loop and heuristically picks a meet irreducible for decision. For example, assuming 
-interval domain, decisions restrict the range of intervals for variables, so the analysis jumps under a 
+The algorithm begins with a set $S$ of abstract deduction transformer 
+$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
+and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
+respectively.  The propagation trail stores all meet irreducibles inferred 
+by the abstact model search phase (deductions and decisions), while the 
+reason trail maps elements of propagation trail to transformers that were 
+used to drive them.
+
+The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
+fixed point over the transformers in F that refines the abstract valuation 
+$v$, similar to BCP step in SAT solvers.  If the result of $deduce$ 
+returns {\em CONFLICT} ($\perp$), the algorithm terminates with {\em safe}. 
+Else, the analysis enters into the while($true$) loop and makes a new decision 
+by a call to $decide(v)$ which returns a new meet irreducible $d$ that is
+consistent to the state of the solver.  If $d$ refines the current abstract 
+value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 
+and $\mathcal{R}$ are updated.
+
+For example, a decision in interval domain restrict the range of 
+intervals for variables, so the analysis jumps under a 
 greatest fixed-point. Note that the widening operation in abstract interpretation 
 jumps above a least fixed-point, so decisions can be viewed as 
-dual widening. The procedure $deduce(P)$ is called next to deduce new facts for 
-current decision. The algorithm terminates with {\em unsafe} if the result of $deduce(P)$ 
-is {\em gamma complete}. Else, the algorithm enters in to $conflict\_analysis()$ phase to learn 
-the reason for {\em conflict}. There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP), ACDCL heuristically choose one. 
-A learnt clause must include asserting cuts which guarantees
-derivation of new information after backtracking. The clause learning and backtracking continues 
-as long as the result of deduction is BOTTOM ($\perp$) or the 
-analysis backtracks to decision level 0. If no further backtrack is possible, then the 
-algorithm terminates with {\em safe}. Else, the algorithm makes a new decision and 
-the above process is repeated until a real counterexample is obtained or the algorithm 
-backtracks to decision level 0 after a conflict in which case it returns {\em safe}. 
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
-However, we are developing a new technique for automatic invariant generation using 
-ACDCL for unbounded proofs.
-
-A decision is made by calling a procedure $decide(v)$ which 
-returns a meet irreducible $d$.  If $d$ refines the 
-current abstract value $v$, then the worklist is updated with 
-all transformers that are dependant on the current decision 
-$d$ along with the update of the corresponding propagation 
-and reason trail, $\mathcal{T}$ and $\mathcal{R}$.
+{\em dual widening}.  The procedure $deduce$ is called next to infer new 
+set of meet irreducibles for the current decision.  The model search phase 
+alternates between the decision and deduction until the result of deduction 
+is either $\gamma$-complete~\cite{gamma} or the deduction leads to a {\em CONFLICT}.  
+If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
+algorithm returns the error trace $v$ and terminates with UNSAFE or the 
+current abstraction is insufficient to determine the satisfiability of $\varphi$.  
+However, in case of {\em CONFLICT}, the algorithm enters in to 
+$conflict\_analysis()$ phase to learn the reason for {\em conflict}.  
+There can be multiple incomparable reasons for conflict -- based on 
+the choice of Unique Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically 
+choose one reason.  A learnt clause must include asserting cuts which 
+guarantees derivation of new meet irreducibles after backtracking. The clause 
+learning and backtracking continues as long as the result of deduction 
+is conflicting ($\perp$), that is, the abstract valuation $v$ does not
+abstractly satisfy the formula.  If no further backtrack is possible, 
+then the algorithm terminates and $\varphi$ is {\em SAFE}. Else, the 
+algorithm makes a new decision and the above process is repeated 
+until a real counterexample is obtained or the algorithm backtracks to 
+decision level 0 after a conflict in which case it returns {\em SAFE}.

### 2016-10-09 domains.tex
static analysis equation

@@
 
+\textbf{Program Analysis Constraints:}  A common approach to analyse 
+a program $\mathcal{P}$ is to represent $\mathcal{P}$ as set of 
+equations called, {\em program analysis constraints}.  The 
+program analysis constraints, as shown below, is an assignment, 
+or boolean expressions $BExp$ over program variables $Var$ or
+a non-deterministic choice (ternary experssion) between statements $s_1$ and
+$s_2$.  All the bounded loops are completely unrolled and represented using 
+following constraints. Similarly, an assertion is of type $BExp$.  
+\[c::= x:= exp | BExp | (exp) ? s_1, s_2\]
+For example, figure~\ref{swssa} presents the program analysis constraints for 
+a C program in Static Single Assignment (SSA) form.
+
 \subsection{Concrete Semantics}
@@
 
-\begin{definition}{(Meet Irreducibles)} A meet irreducible is a minimum complementable 
-element that cannot be expressed in terms of the meet over a set of other 
-elements. Formally, a {\em meet irreducible} $M$ in a complete lattice 
-structure $A$ is an element $M \in A$ that has the following property.
+\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
+in a complete lattice structure $A$ is a minimum complementable element 
+$M \in A$ that has the following property.
 \[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  

### 2016-10-09 domains.tex
minor

@@
 interval, octagon and equality domains which supports 
-complementation operator.  For example, the octagon in 
-Figure~\ref{octagon} can be written as a conjunction of:
-\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
-The complementation of the above octagon can be written as disjunction of:
+complementation operator.  
+%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
+%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+For example, the complementation of the octagon in Figure~\ref{octagon} can be written as disjunction of:
 \[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-10-10 example.tex
paper update
CDCL
decision
static analysis equation
example figure added

@@
-\section{Working Example}\label{example}
+\section{Motivating Example}
 %
-\subsection{Example Demonstrating Working of ACDCL}
+\begin{figure}[t]
+\scriptsize
+\begin{tabular}{l|l|l}
+\hline
+C program & CFG & Solver Statistics\\
+\hline
+\begin{lstlisting}[mathescape=true,language=C]
+void main() {
+ int x,result1=0,result2=0;
+ assume(x>=1 && x<= N);
+ result1 = 2*x; 
+ int state = 1;
+ int y=x;
+ while(state!=0) {
+  switch(state) {
+   case 0: state = 1; break;
+   case 1: state = 2; break;
+   case 2: x--; y++;
+   if(x == 0) state = 0;
+   else state = 1; break;
+  } 
+ }
+ result2 = y;
+ assert(result1 == result2);
+}
+\end{lstlisting}
+&
+\begin{minipage}{4.40cm}
+\centering
+\vspace*{0.3cm}
+\scalebox{.45}{\import{figures/}{example.pspdftex}}
+%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
+\end{minipage}
+&
+\begin{lstlisting}[mathescape=true,language=C]
+                 SAT solver  ACDCL  
+restarts           1          0
+conflicts  :       5          2
+decisions  :       36         2 
+propagations :     1365       240  
+conflict literals  65         2   
+\end{lstlisting}
+\\
+\hline
+\end{tabular}
+\caption{C Program, and statistics for decision procedure}
+\label{example}
+\end{figure}
+
 \begin{figure}[htp]
@@
 
+\Omit {
 \lstdefinestyle{mystyle}{
@@
 \lstset{style=mystyle}
+}
 Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.

### 2016-10-10 example.tex
example description

@@
 \hline
-C program & CFG & Solver Statistics\\
+C program & Control-Flow Graph & Decision Procedure Statistics\\
 \hline
@@
                  SAT solver  ACDCL 
+                 ----------- ----- 
 restarts           1          0
@@
 propagations :     1365       240  
-conflict literals  65         2   
+conflict literals  65         4   
 \end{lstlisting}
@@
 \end{tabular}
-\caption{C Program, and statistics for decision procedure}
+\caption{C Program, CFG and statistics for decision procedure}
 \label{example}
 \end{figure}
+Let us consider a simple program in Figure~\ref{example} that checks 
+the equivalence of a state-machine based interger multiplication against  
+a specification that uses multiplication operator (*).  The assertion 
+simply compares the result of two multiplication for an input $x$.  To 
+avoid overflow error, we assume $x$ has an upper bound $N$.  The 
+control flow graph is shown in the middle column of Figure~\ref{example}.
+
+We analyze the program using three different analyzers -- {\em bounded 
+model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
+For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop, the 
+right column shows the statistics from MiniSAT~\cite{minisat} solver 
+for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
+constraint propagations and performs 5 conflict analysis with 65 conflict 
+literals.  Moreover, the solver performs a restart, presumably to recover 
+from bad decisions.
+
+On the other hand, abstract interpretation using forward analysis 
+is too imprecise to verify the safety of the above program using 
+interval domain.  The imprecision is due to control-flow join at 
+node $n1$.  State-of-the-art abstract interpretation tool, Astr{\'e}e, 
+requires external hints, provided by manually annotating the code 
+with partition directives at $n1$, to prove safety.  The partition 
+directive helps the tool to analyze each program execution paths separately.  
+To do so, Astr{\'e}e requires a trace partitioning domain along with interval 
+to determine safety.  Consequently, the analysis becomes very expensive 
+for a large value of $N$.  In general, the imprecision is either intended 
+  by the tool because such high precision analysis is normally not required for runtime error analysis or the imprecision is unavoidable due to the complexity 
+of the application under analysis.
+
+Figure~\ref{} shows an analysis using ACDCL with interval domain.  The 
+analysis associates an interval for each control location and variable, 
+as shown in the abstract conflict graph.
+
+The result of the analysis is shown in right side of Figure~\ref{example}.  
+Compared to a SAT solver, ACDCL makes only a single decisions which lead 
+to 240 interval constraint propagations, and learns a single conflict clause 
+involving only one literal.
 

### 2016-10-11 example.tex
figure added

@@
 &
-\begin{lstlisting}[mathescape=true,language=C]
-                 SAT solver  ACDCL 
-                 ----------- ----- 
-restarts           1          0
-conflicts  :       5          2
-decisions  :       36         2 
-propagations :     1365       240  
-conflict literals  65         4   
-\end{lstlisting}
+\begin{minipage}{4.40cm}
+\centering
+\vspace*{0.3cm}
+\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
+%\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
+\end{minipage}
 \\
@@
 to determine safety.  Consequently, the analysis becomes very expensive 
-for a large value of $N$.  In general, the imprecision is either intended 
-  by the tool because such high precision analysis is normally not required for runtime error analysis or the imprecision is unavoidable due to the complexity 
+for a large value of $N$.  In general, the imprecision is either intended by 
+the tool because such high precision analysis is normally not required for runtime error analysis or the imprecision is unavoidable due to the complexity 
 of the application under analysis.
+
+\Omit{
+\begin{tabular}{l|l|l|l|l|l}
+\hline
+Solver & Decisions & propagations & conflict & conflict literals & restarts \\
+SAT & 1 & 1365 & 5 & 65 & 1 \\
+ACDCL & 1 & 23 & 1 & 1 & 0 \\
+\hline
+\end{tabular}
+%\caption{Solver-statistics}
+\label{solver}
+\end{figure}
+}
+\begin{lstlisting}[mathescape=true,language=C]
+
+                 SAT solver  ACDCL 
+                 ----------- ----- 
+restarts           1          0
+conflicts  :       5          2
+decisions  :       36         2 
+propagations :     1365       240  
+conflict literals  65         4   
+\end{lstlisting}
 

### 2016-10-11 contributions.tex
minor

--- 2016-10-07 contributions.tex
@@
 \begin{enumerate}
-\item We present a novel program analysis technique based on 
-abstract conflict driven clause learning algorithm~\cite{dhk2013-popl}.
-To this end, we present a procedure for lifting CDCL algorithm to 
-richer lattice structure such as {\em interval} and {\em octagon} 
-abstract domains.
-
-\item To enable lifting of CDCL algorithm, we implement 
-our own template-based numerical abstract domains that 
-supports complementation property.  Our proposed technique 
-is embodied in our tool called, {\em Titlis}.  To the best 
-of our knowledge, this is the first instantiation of CDCL 
+\item We present a novel program analysis technique that 
+lift CDCL algorithm to relational numerical abstract domains.   
+Our proposed technique is embodied in our tool called, {\em Titlis}.  
+To the best of our knowledge, this is the first instantiation of CDCL 
 algorithm with relational abstract domains.
@@
 the efficiency of abstract domains while being path-sensitive and 
-bit-precise at the same time.  This is attributed to the intelligent 
-decision heuristics which exploits the high-level structure of the 
-problem combined with stronger deductions and clause learning 
-mechanism aided by the richer abstract domains.
+bit-precise at the same time.  
 \end{enumerate}

### 2016-10-11 introduction.tex
minor

@@
 Clause learning SAT solvers requires a weak complementation property 
-of the partial assignments domain.  However,  the most commonly 
-used APRON~\cite{apron} library that implements abstract domains, is specialized 
-for abstract interpretation framework which does not require complementation 
+of the partial assignments domain.  However,  abstract domain 
+library, such as APRON~\cite{apron}, is specialized for abstract 
+interpretation framework which does not require complementation 
 operator.  Thus, we implement our own template-based numerical abstract 
-domains that support complementation property.  Our proposed technique 
-is embodied in our tool called, {\em Titlis}.  To the best of our 
-knowledge, this is the first instantiation of CDCL 
-algorithm with relational abstract domains.
+domains that support complementation property, required for clause learning 
+in the domain.

### 2016-10-11 contributions.tex
example added

@@
 
+\item We present an abstract model search algorithm that strictly 
+generalizes BCP procedure to compute fix-point over template polyhedra 
+domain.  Our abstract conflict analysis algorithm present a non-trivial 
+lifting of the Unique Implication Point (UIP) procedure to abstract 
+lattices in order to learn a generalized reason for conflicts.
+
 \item  Experiments on a set of benchmarks drawn from bit-vector 

### 2016-10-11 example.tex
example added

--- 2016-10-10 example.tex
@@
 \hline
-C program & Control-Flow Graph & Decision Procedure Statistics\\
+C program & Control-Flow Graph & Abstract Conflict Graph \\
 \hline
@@
 void main() {
- int x,result1=0,result2=0;
- assume(x>=1 && x<= N);
- result1 = 2*x; 
+int x,sepc=0,impl=0;
+assume(x>=1 && x<=N);
+spec = 2*x; 
 int state = 1;
@@
  switch(state) {
-   case 0: state = 1; break;
-   case 1: state = 2; break;
+  case 0: state = 1; 
+  break;
+  case 1: state = 2; 
+  break;
   case 2: x--; y++;
-   if(x == 0) state = 0;
-   else state = 1; break;
+  if(x == 0) state=0;
+  else state=1;break;
  } 
 }
- result2 = y;
- assert(result1 == result2);
+impl = y;
+assert(spec == impl);
 }
@@
 \end{tabular}
-\caption{C Program, CFG and statistics for decision procedure}
+\caption{C Program, CFG and Abstract Conflict Graph}
 \label{example}
@@
 a specification that uses multiplication operator (*).  The assertion 
-simply compares the result of two multiplication for an input $x$.  To 
-avoid overflow error, we assume $x$ has an upper bound $N$.  The 
+simply compares the result of two multiplication for an integer input $x$.  
+To avoid overflow error, we assume $x$ has an upper bound $N$.  The 
 control flow graph is shown in the middle column of Figure~\ref{example}.
@@
 model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
-For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop, the 
-right column shows the statistics from MiniSAT~\cite{minisat} solver 
+For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop,
+table~\ref{solver} shows the statistics from MiniSAT~\cite{minisat} solver 
 for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
@@
 with partition directives at $n1$, to prove safety.  The partition 
-directive helps the tool to analyze each program execution paths separately.  
-To do so, Astr{\'e}e requires a trace partitioning domain along with interval 
-to determine safety.  Consequently, the analysis becomes very expensive 
-for a large value of $N$.  In general, the imprecision is either intended 
-  by the tool because such high precision analysis is normally not required for runtime error analysis or the imprecision is unavoidable due to the complexity 
-of the application under analysis.
+directive helps the tool to analyze each program execution paths 
+separately.  To do so, Astr{\'e}e requires a product of trace 
+partitioning domain and an interval domain to determine safety.  
+Consequently, the analysis becomes very expensive for a large value 
+of $N$.  In general, the imprecision is either intended by the tool 
+because such high precision analysis is normally not required for 
+runtime error analysis or the imprecision is unavoidable due to the 
+complexity of the application under analysis.  
+%
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l|l|l|l|l}
+\hline
+Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
+SAT & 36 & 1365 & 5 & 65 & 1 \\ \hline
+ACDCL & 1 & 23 & 1 & 1 & 0 \\
+\hline
+\end{tabular}
+}
+\end{center}
+\caption{Solver-statistics}
+\label{solver}
+\end{table}
+%
+The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
+interval domain.  The analysis associates an interval for each control 
+location and variable, as shown in the abstract conflict graph.  Interval 
+constraints generated from forward analysis in the initial deduction phase 
+infers that $\langle x:[-\infty,2], spec:[2,4], impl:[-1,\infty],
+state:[1,\infty], y:[1,+\infty] \rangle$.  Clearly, ACDCL can not prove safety.
 
-Figure~\ref{} shows an analysis using ACDCL with interval domain.  The 
-analysis associates an interval for each control location and variable, 
-as shown in the abstract conflict graph.
+Hence, ACDCL makes an arbitrary decision $x:[-\infty,1]$ to refine the analysis.  
+The decision constrains the interval of $x$.  The deductions made during fixed point
+iteration are represented by abstract conflict graph in right-side of 
+Figure~\ref{fig:example}.   Nodes in the graph denotes the corresponding 
+updates to the program variables in the cfg.  Interval analysis concludes 
+that the $(n12:\bot)$, that is the decision $x:[-\infty,1]$ leads to a 
+{\em conflict}. Intuitively, it means that the program is {\em safe} for 
+$x:[-\infty,1]$.
 
-The result of the analysis is shown in right side of Figure~\ref{example}.  
-Compared to a SAT solver, ACDCL makes only a single decisions which lead 
+A clause learning SAT solver would learn the reason for conflict at this 
+point and then backtrack to a level such that the learnt clause is UNIT.  
+Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
+(\(I: x = [2,+\infty]\)), that is all error traces must satisfy $(x \geq 2)$ 
+at node $I$.  The analysis discards all interval constraints that
+lead to the conflict and backtrack to decision level 0.  ACDCL then 
+performs interval analysis with the learnt constraint $(x>=2)$ which again 
+leads to a conflict.  The analysis cannot backtrack further and therefore 
+it terminates proving that the program is safe.  A DPLL-style solver would 
+perform proof by cases.  However, decision and clause learning are used 
+to avoid case based reasoning and prevents enumeration behavior.
+
+The result of the analysis using ACDCL is shown in Figure~\ref{example}.  
+Compared to a SAT solver, ACDCL makes only a single decision which lead 
 to 240 interval constraint propagations, and learns a single conflict clause 
-involving only one literal.
-
-\begin{figure}[htp]
-\centering
-\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{cfg-conflict.pspdftex}}
-\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
-\end{figure}
-
-\Omit {
-\lstdefinestyle{mystyle}{
-    numbers=left,                    
-    numbersep=5pt,                  
-    tabsize=2,
-    mathescape=true,
-    language=C
-}
-\lstset{style=mystyle}
-}
-Let us consider a simple safe program $P$ as shown in Figure~\ref{fig:filter}A.
-Astr{\'e}e fails to verify the safety using interval and trace-partition domain
-due to control-flow join at location $n5$.  Astr{\'e}e requires external hints, provided by 
-manually annotating the code with partition directives at $n1$, to prove safety.  
-In general, the imprecision is either intended by the tool because such high precision analysis is 
-normally not required for runtime error analysis or the imprecision is unavoidable 
-due to the complexity of the application under analysis.
-
-On the other hand, ACDCL uses decision and learning to generate the proof 
-using simple interval domain. The analysis associates an interval with each location and variable.  
-Assuming $bool: c = [0,1]$, ACDCL performs backward propagation 
-starting from $n6$ by computing the weakest precondition for every statement 
-in $P$ and immediately infers that (\(n4: x = \bot)\), thus proving safety as shown in 
-red color in figure~\ref{fig:filter}A.  Now, let us consider program $P'$ in
-Figure~\ref{fig:filter}B. Assuming $char: c$, ACDCL performs both forward and
-backward propagations (shown in red color in Figure~\ref{fig:filter}B).  
-The analysis then perform sequence of decisions starting with $c = [0,255]$ and reaches a 
-conflict when the decision is (\(n1:c == [0,0]\)) which is connected to $(n6: \bot)$ 
-and suffices to prove safety.  The deductions made during fixed point
-iteration are represented by abstract conflict graph shown in Figure~\ref{fig:filter}C. 
-Similar to conflict analysis phase in SAT solvers, ACDCL learns (\(n1: c = [1,255]\)), 
-that is all error traces must satisfy (\(c \neq 0\)) at n1.  The analysis backtracks 
-discarding all assumptions. Interval analysis is run with the learnt constraint and 
-can prove safety.  Thus, decisions and clause learning are used to avoid case based reasoning. 
-The advantage of ACDCL over propositional SAT solver is that the decision heuristics in 
-ACDCL can exploit the program structure by making decisions on interesting program 
-variables (for example variables in conditional branches or loop variables). Experimental 
-evidence shows that this leads to significantly less number of decisions compared 
-to the propositional solver. Compared to classical abstract interpretation 
-based tools, ACDCL automatically performs program and property driven trace partitioning 
-using decision and clause learning to generate proofs using simpler domain -- thus, it is more 
-precise than classical abstract interpretation.
+involving only one literal.  Compared to abstract interpretation, ACDCL does 
+not require external hints to proof the program, thus automatically performing 
+program and property driven trace partitioning to generate proofs using decision 
+and clause learning.

### 2016-10-11 example.tex
minor

@@
 iteration are represented by abstract conflict graph in right-side of 
-Figure~\ref{fig:example}.   Nodes in the graph denotes the corresponding 
+figure~\ref{example}.   Nodes in the graph denotes the corresponding 
 updates to the program variables in the cfg.  Interval analysis concludes 
@@
 
-The result of the analysis using ACDCL is shown in Figure~\ref{example}.  
+The result of the analysis using ACDCL is shown in table~\ref{solver}.  
 Compared to a SAT solver, ACDCL makes only a single decision which lead 
-to 240 interval constraint propagations, and learns a single conflict clause 
+to 23 interval constraint propagations, and learns a single conflict clause 
 involving only one literal.  Compared to abstract interpretation, ACDCL does 

### 2016-10-11 experiments.tex
experiment

@@
-\section{Experimental Results}\label{experiments}
+\section{Experimental Results}
+We classify our benchmarks into two groups: {\em bounds checking}, 
+{\em equivalence checking}.  We use interval, octagon, and equality 
+domain for our verification task.  We analyze programs with bounded 
+loops.  All loops in the program are completely unrolled before analysis.

### 2016-10-12 example.tex
Related work

@@
 \end{tabular}
-\caption{C Program, CFG and Abstract Conflict Graph}
+\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
 \label{example}
@@
 
-Hence, ACDCL makes an arbitrary decision $x:[-\infty,1]$ to refine the analysis.  
-The decision constrains the interval of $x$.  The deductions made during fixed point
-iteration are represented by abstract conflict graph in right-side of 
-figure~\ref{example}.   Nodes in the graph denotes the corresponding 
-updates to the program variables in the cfg.  Interval analysis concludes 
-that the $(n12:\bot)$, that is the decision $x:[-\infty,1]$ leads to a 
-{\em conflict}. Intuitively, it means that the program is {\em safe} for 
-$x:[-\infty,1]$.
+Hence, ACDCL makes a decision $x:[-\infty,1]$ to refine the analysis.  
+The branching variable $x$ is chosen by an {\em ordered} decision 
+heuristics that choose variables appearing in conditional branches 
+(assume statement here) first before choosing other program variables. 
+The decision constrains the interval of $x$.  The deductions made 
+during fixed point iteration are represented by abstract conflict 
+graph in right-side of figure~\ref{example}.  Nodes in the graph 
+denotes the corresponding updates to the program variables in 
+the CFG.  Interval analysis concludes that the $(n12:\bot)$, that 
+is the decision $x:[-\infty,1]$ leads to a {\em conflict}.  Thus, 
+the program is {\em safe} for $x:[-\infty,1]$.
 

### 2016-10-12 introduction.tex
example added
Related work

@@
 standard static analyses~\cite{se2011} for program safety aims to compute 
-invariants by applying transformers that refines an abstract 
-value.  Most lattice structure used in static analysis have 
-meet and join operations but lack negations.  The meet operation 
+invariants as fixed-points of abstract transformers.  Most lattice structure 
+used in static analysis have meet and join operations.  The meet operation 
 precisely model conjunction, and the join operation over-approximate 
@@
 
+\rmcmt{cite Natural-domain SMT}
+\rmcmt{relational domains}
+
 The instantiation of ACDCL has so far been limited to interval domains only.  
 In this paper, we identify specific properties of the CDCL algorithm that 
-are necessary to lift the propositional skeleton of CDCL to other richer 
-lattice structures and relational abstract domains such as octagons, 
-polyhedra, zones and equalities.  To this end, we present a 
-characterization of abstract model search procedure and conflict 
-analysis procedure that lift the abstract deduction transformer and the 
-first-UIP algorithm to relational lattice structures.
+are necessary to lift the propositional skeleton of CDCL to relational 
+abstract domains such as template polyhedra domains~\cite{sriram} (octagons, zones and 
+equalities).  To this end, we present a characterization of abstract 
+model search procedure and conflict analysis procedure that lift 
+the abstract deduction transformer and the first-UIP algorithm to 
+relational lattice structures.
 

### 2016-10-12 related_work.tex
Related work

\section{Related Work}
The natural domain SMT framework proposed in~\cite{ndsmt} operates on
a fixed first-order partial assignment lattice structure, where first-order 
variables are mapped to domain values, similar to constants lattice 
in program analysis.  However, ACDCL can be instantiated on different 
abstract lattices.  The abstract lattice in natural domain SMT does 
not have complementable meet irreducibles, and therefore does not 
support generalized clause learning~\cite{sas}.  On the other hand, 
most abstract lattice have complementation property, thus enabling 
ACDCL to perform generalized clause learning.

### 2016-10-12 algorithm.tex
minor
conflict analysis

--- 2016-10-09 algorithm.tex
@@
 %\rmcmt{safety as satisfiability}
-\subsection{Static Analyses Equations for Safety}
-Currently, ACDCL handles loops in the program by unrolling the bounded loops. 
+\subsection{Static Analysis Equations for Safety}
+Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+perform safety analysis by computing fixed point to infer invariants 
+over program variables.  However, bounded model checking (BMC) tries to search 
+for a counterexample in a bounded execution trace, by symbolically executing a 
+program up to a finite bound.  
+Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+For a set $N$ of program analysis constraints defined over a set 
+of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+and the \rmcmt{final valuation} of constraint variables.
+\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
 %
@@
 %
-\Omit {
-Static program analysis based on abstract
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
-verify certain classes of properties for safety-critical systems.  In
-abstract interpretation, a given program is analysed with respect to a set
-of given abstract domains.  However, ACDCL is a novel program analysis 
-technique that embeds an abstract domain inside the CDCL algorithm.
-From the abstract interpretation point of view, ACDCL is an abstract 
-interpreter that uses decision and learning to increase transformer precision. 
-From the decision procedure perspective, ACDCL is a SAT solver for program 
-analysis constraints and is thus a strict generalisation of propositional CDCL solvers.
-Constraint propagation in ACDCL uses fixed point iteration, decisions restrict the 
-range of intervals and learning generates program analysis constraints (not assumptions) 
-that preserve the error reachability.  Moreover, ACDCL implicitly 
-perform program and property driven trace partitioning to increase the precision
-of the analysis.  
-}
+\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
@@
 decision level 0 after a conflict in which case it returns {\em SAFE}.
+
+\paragraph \em{Solver State:}  The state of a ACDCL solver is a tuple 
+of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
+is a sequence of labelled information of the form $(m,s)$ where 
+$m$ is a meet irreducible and $s = decision$ if $m$ is a decision, 
+or $s = deduction$ if $m$ is inferred by a deduction. And $S$ is 
+a set of abstract deduction transformer.  A trail in SAT solver 
+stores variable assignments of the form $(p, t)$, where $p$ is 
+a propositional variable that appears at most once in the trail 
+and $t$ is a truth assignment (true or false).  Whereas, a trail 
+$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
+inferred by deduction or decision phase where a variable in trail 
+can be assigned multiple times, each time with increasingly precise 
+bounds.  A current valuation $v$ is a meet of all elements of trail
+$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
+empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
+$m_i \in \mathcal{T}$.  A solver is in {\em conflict} if some clauses in
+$\mathcal{S}$ is not abstractly satisfied by $v$.

### 2016-10-12 decision.tex
minor changes
introduction
conflict analysis

@@
-\section{Decision}
+\subsection{Decision}
 A decision $\mathcal{D}$ is a meet irreducible that refines the 
-current partial assignment.  A decision can be of the form 
+current abstract valuation.  A decision can be of the form 
 $\langle x \preceq c \rangle$ or $\langle x \succeq c \rangle$, 
@@
 trail $\mathcal{T}$, that is, it must not contradict with 
-the elements in the trail.  Each variable in propositional CDCL 
-is assigned at most once at a particular decision level.  However, 
-a variable can be assigned several times in our lifting of CDCL to 
-numerical domains, each time with increasingly precise bounds.
+the elements in the trail.
 

### 2016-10-12 domains.tex
conflict analysis

@@
 \section{Program and Abstract Domains}
-In this section, we define a program and the notion of concrete domain 
-and abstract domain. 
+In this section, we define a notion of program analysis constraints, concrete domain and abstract domain. 
+\Omit {
 \subsection{Programs:}
@@
 assume that all the procedures are inlined.  
-
-\textbf{Program Analysis Constraints:}  A common approach to analyse 
-a program $\mathcal{P}$ is to represent $\mathcal{P}$ as set of 
-equations called, {\em program analysis constraints}.  The 
+}
+\textbf{Program Analysis Constraints:}  A program $\mathcal{P}$ is represented as set of equations called, {\em program analysis constraints}.  The 
 program analysis constraints, as shown below, is an assignment, 

### 2016-10-12 learning.tex
conflict analysis

@@
-\section{Clause Learning in Abstract Lattice}
+\section{Abstract Conflict Analysis}
+A conflict analysis in a SAT solver can be seen as abductive 
+reasoning where a singleton assignment $c$ is replaced by a partial 
+assignment that is sufficient to infer $c$.  Thus, a conflict 
+analysis is a generalization of an underapproximation of a set of 
+countermodels~\cite{sas,dhk2013-popl}, where a countermodel is a 
+set of structures that do not satisfy a formula.  Since relational 
+abstract domains have more complex structure than a partial 
+assignments domain, so the lifting of {\em first-UIP} algorithm~\cite{uip} 
+and clause learning to any arbitraty lattice structure is a non-trivial task.
+
+The first-UIP algorithm in SAT solver works on a data structure called 
+{\em implication graph} that computes a cut that suffices to produce a 
+conflict.  The trail in ACDCL implicitly encodes a graph structure called 
+{\em abstract conflict graph}, which records the dependencies between 
+deductions made during abstract model search phase.  For example, consider a
+formula $\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
+abstract valuation $v:= (x=[5,5])$.  
+Figure~\ref{conflict} shows a snapshot of abstract implication graph that 
+stores the deductions obtained from abstract model search using octagon domain.  
+An arrow in the graph indicate that the predecessors of each node $n$ are 
+sufficient to deduce $n$.  The last deduction conflict with the constraint 
+$(y \leq 0)$.  Every cut in the graph, marked as {\em cut0, cut1}, provides a 
+reason for conflict that can be used in learning.  An UIP is a special node such 
+that any path from the last decision node to the conflict node must pass 
+through it.  Unlike a SAT solver, a UIP in ACDCL may involve meet irreducibles 
+that involves the same variables.  The first-UIP is a node closest to the 
+conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
+asserting cuts, that is, it yields clauses that generates new deduction after 
+backtracking.  For our example, there exists multiple incomaparable reasons 
+for conflict.  Here, cut0 is the first UIP.  Choosing cut0 yields a learnt 
+clause $L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee 
+-y+10<0)$.  A learnt clause $\mathcal{L}$ is obtained by negating the reason 
+for conflict.  
+%
+\begin{figure}
+%\caption{An example of Octagon}\label{octagon}
+\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
+\label{conflict}
+\end{figure} 
+%    
+\paragraph {\em Backjumping:}
+A backjumping procedure undo all the assignments in the trail up to 
+a decision level that restores the solver to a consistent state 
+(non-conflicting).  The backjumping level is defined by the literal 
+of the conflict clause that is closest to the root (decision level 0) 
+where the conflict clause is still unit. If a conflict clause is 
+globally unit, then the backjumping level is the root of the search tree.
+
+The abstract clause learning and backjumping procedures in the abstract 
+conflict graph is stated in terms of the state of ACDCL solver as follows. 
+\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
+\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
+\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
+\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
+(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is
+consistent} \]
+
+\subsection{Clause Learning in Abstract Lattice}
+\Omit {
 Conflict graph for Intervals
@@
 
-\textit{Backjumping:}
-The backjumping level is defined by the literal of the conflict clause assigned
-at the level that is the closest to the conflict one. In other words, the
-backjumping level is the level closest to the root (decision level 0)  where the
-conflict clause is still unit. If a conflict clause is "globally"  unit, then
-the backjumping level is the root of the search tree.
 1.DPLL style -- chronological backtracking \\
@@
 
-\Omit{
 **********************************************

### 2016-10-13 paper.tex
domain text
domain text updated
CDCL
octagon figure
definitions added
definition of deduction transformer
clauses and literals
decision
decision text
static analysis equation
experiment
abstract

@@
 \usepackage{times}
-\usepackage{graphicx}
+\usepackage{array}
+\usepackage{graphicx,wrapfig}
 \usepackage{cite}
@@
 \usepackage{wasysym} 
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{paralist}
@@
 \newcommand{\Omit}[1]{}
+\newcommand{\mydef}[1]{\begin{definition}#1\end{definition}}
+\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}
+\newcommand{\pscmt}[1]{{\color{blue}{#1}}}
 
@@
 
-\title{Lifting CDCL to Numerical Domains}
+\title{Lifting CDCL to Relational Domains}
 
-\author{Rajdeep Mukherjee\inst{1} Peter Schrammel   Daniel Kroening}
+\author{Rajdeep Mukherjee\inst{1}\hskip 1em  Peter Schrammel\inst{2}\hskip 1em  Daniel Kroening\inst{1}}
 
@@
 \email{p.schrammel@sussex.ac.uk}}
-\Omit {
-\author{Rajdeep Mukherjee\inst{1} \and
-Peter Schrammel\inst{2} \and
-Daniel Kroening\inst{1}}
-
-\institute{University of Oxford, UK \and University of Sussex, UK \\
-\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
-\email{p.schrammel@sussex.ac.uk}}
-}
 
@@
 %
-The Conflict Driven Clause Learning (CDCL) algorithm can be seen as an 
-alternation between over-approximate deduction with boolean 
-constraint propagation, and under-approximate abduction, with 
-conflict analysis.  {\em Abstract Conflict Driven Clause Learning} 
-(ACDCL) is a novel program analysis that embeds an abstract domain 
-inside the CDCL algorithm. From decision procedure perspective, 
-ACDCL is a SAT solver for program analysis constraints and is a 
-strict generalisation of propositional CDCL solvers. From Abstract 
-Interpretation point of view, ACDCL can be viewed as an abstract 
-interpreter that uses decision and learning to increase the precision 
-of abstract transformers.  In this paper, we instantiate ACDCL with 
-a relational and a non-relational numerical abstract domain and 
-evaluate the performance of our analyzer with program verifiers that 
+  \emph{Abstract Conflict-Driven Learning} (ACDL) is a combination of
+  abstract interpretation and the well-known Conflict-Driven Clause
+  Learning (CDCL) algorithm that is used in SAT solvers. ACDL has been
+  successfully used with non-relational abstract domains such as
+  intervals to improve the performance of a solver for floating point
+  arithmetic, for instance. 
+%
+Whereas the
+algorithms typically used in SAT solvers can be rather easily
+lifted from the Boolean lattice to other non-relational domains,
+lifting them to relational domains is challenging.
+%because of inter-"literal" dependencies
+%because of the weaker properties of "literals"
+%
+In this paper we present the adaptations of algorithms and heuristics
+for propagation, decision making and conflict learning necessary in
+order to enable the use with relational domains.
+%
+We have implemented these algorithms inside the 2LS verification tool
+for the example of the template polyhedra domain.
+%
+We evaluate the performance of our analyzer in comparison with program verifiers that 
 uses propositional-level solvers.  Experimental evidence on a set of 
@@
 %===============================================================================
-\input{background}
+%\input{background}
 %===============================================================================
@@
 %===============================================================================
-\input{definition}
+\input{example}
 %===============================================================================
@@
 %===============================================================================
-\input{example}
+%\input{cdcl}
+%===============================================================================
+
+%===============================================================================
+\input{domains}
 %===============================================================================
@@
 %===============================================================================
-\input{domains}
+\input{deductions}
 %===============================================================================
@@
 %===============================================================================
-\input{deductions}
+%\input{definition}
 %===============================================================================

### 2016-10-13 contributions.tex
contributions

--- 2016-10-11 contributions.tex
@@
-\section{Contributions}
+\paragraph{Contributions}
 In this paper, we make the following contributions.
-\begin{enumerate}
+\begin{compactenum}
 \item We present a novel program analysis technique that 
-lift CDCL algorithm to relational numerical abstract domains.   
-Our proposed technique is embodied in our tool called, {\em Titlis}.  
+lifts the CDCL algorithm to relational numerical abstract domains.   
+Our proposed technique is embodied in our tool called, {\em
+  Titlis}. \pscmt{It's better not to create a zoo of tool names}  
 To the best of our knowledge, this is the first instantiation of CDCL 
@@
 
-\item  Experiments on a set of benchmarks drawn from bit-vector 
-regression suite in sv-comp, numerical programs and few 
-control-intensive benchmarks from hardware domain shows that 
+\item We present an abstract model search algorithm that strictly 
+generalizes the Boolean Constraint Propagation (BCP) procedure to
+compute a fixed point over a relational domain.  Our abstract conflict analysis algorithm present a non-trivial 
+lifting of the Unique Implication Point (UIP) procedure to relational abstract 
+domains in order to learn a generalized reason for conflicts.
+
+\item  Experiments on a set of benchmarks drawn from the bitvector 
+category in SV-COMP as well as numerical programs and a few 
+control-intensive benchmarks from hardware domain show that 
 the strict generalization of CDCL algorithm effectively combines 
@@
 bit-precise at the same time.  
-\end{enumerate}
+\end{compactenum}

### 2016-10-13 domains.tex
minor
some comments in 2 and 3

@@
 \section{Program and Abstract Domains}
-In this section, we define a notion of program analysis constraints, concrete domain and abstract domain. 
+In this section, we define a notion of program analysis constraints, concrete
+domain and abstract domain. \\
 \Omit {
@@
 
+\pscmt{we should give the (simplified) SSA for an example here}
+
 \subsection{Concrete Semantics}
+\pscmt{not sure that adds anything for TACAS}
 The {\em concrete} domain is a lattice 
@@
 
-\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
-in a complete lattice structure $A$ is a minimum complementable element 
-$M \in A$ that has the following property.
-\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
-\end{definition}
+\subsection{Template Polyhedra Abstract Domain}
 
-In case of partial assignments, meet irreducibles are the singleton
-assignments, for example $\{x \mapsto true \}$, which represents the set 
-of all propositional assignments where the literal $x$ is mapped to true.  
-An important property of meet irreducibles in case of partial assignment is that 
-they have precise complements.  For example, the complement of $\{x \mapsto true \}$
-is a singleton element,$\{x \mapsto false \}$, in the partial assignment domain.
+\pscmt{TODO}
 
-\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
-irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
-irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
-is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
-complementable, then $A$ is said to be have complementable meet irreducibles.  
-\end{definition}
-
-\subsection{Interval Abstract Domain}
+\paragraph{Interval Abstract Domain} \pscmt{TODO: simplify}
 The {\em Interval} abstract domain is a lattice 
@@
 
-\subsection{Octagon Abstract Domain}
+\paragraph{Octagon Abstract Domain} \pscmt{TODO: simplify}
 The {\em Octagon} abstract domain is a lattice 
@@
 
+\subsection{Properties of Domain Elements for Decision and Learning}
+
+\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
+in a complete lattice structure $A$ is a minimum complementable element 
+$M \in A$ that has the following property.
+\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
+\end{definition}
+
+In case of partial assignments, meet irreducibles are the singleton
+assignments, for example $\{x \mapsto true \}$, which represents the set 
+of all propositional assignments where the literal $x$ is mapped to true.  
+An important property of meet irreducibles in case of partial assignment is that 
+they have precise complements.  For example, the complement of $\{x \mapsto true \}$
+is a singleton element,$\{x \mapsto false \}$, in the partial assignment domain.
+
+\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
+irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
+irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
+is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
+complementable, then $A$ is said to be have complementable meet irreducibles.  
+\end{definition}
+
 %========================
@@
 
-\subsection{Properties of Domain Elements for Decision and Learning}
 One of the requirements for a clause learning SAT solver 
@@
 search space.
-
+\rmcmt{define precise complement}
 Thus, clause learning SAT solvers requires a weak complementation property
@@
 %For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
-%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2) \land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
+%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
+  meet-irreducible. It's trivial to do that with APRON. The reason was
+a different one: APRON does not support all C operators, e.g. the bitwise operators.}
 For example, the complementation of the octagon in Figure~\ref{octagon} can be written as disjunction of:

### 2016-10-13 domains.tex
complement property

--- 2016-10-12 domains.tex
@@
 }
-\textbf{Program Analysis Constraints:}  A program $\mathcal{P}$ is represented as set of equations called, {\em program analysis constraints}.  The 
+\textbf{Program Analysis Constraints:}  A program $\mathcal{P}$ is represented 
+as set of equations called, {\em program analysis constraints}.  The 
 program analysis constraints, as shown below, is an assignment, 
@@
 \subsection{Properties of Domain Elements for Decision and Learning}
-One of the requirements for a clause learning SAT solver 
+An important property for a clause learning SAT solver 
 is that each element of the partial assignments domain 
-needs to have a decomposition into precisely complementable 
-elements.  This property of the underlying domain helps to 
+are decomposable into precisely complementable 
+elements.  This property of the domain helps to 
 guide the search away from the conflicting region of the 
 search space.  
-
+\rmcmt{define precise complement}
 Thus, clause learning SAT solvers requires a weak complementation property
-of the abstraction.  However, numerical abstract domain,  such as 
-intervals and octagons are intersections of complementable half-spaces. 
+of the abstraction.  Most numerical abstract domain, such as intervals, 
+octagons, polyhedra have complementation property, that is, the domain 
+elements are intersections of complementable half-spaces. 
+\Omit {
 The commonly used library for numerical abstract domains  
@@
 program~\cite{se2011}. 
-
-APRON library is specialized for the abstract interpretation 
-framework which does not require complementation property.  
-To this end, we implement our own template-based 
-interval, octagon and equality domains which supports 
-complementation operator.  
+}
+Abstract Domain library, such as APRON C library~\cite{apron}, 
+is specialized for the abstract interpretation framework which 
+does not require complementation property.  To this end, we 
+implement our own template-based polyhedra domain and interval 
+domain which supports complementation operator.  
 %For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:

### 2016-10-13 domains.tex

--- 2016-10-13 domains.tex
@@
 
-One of the requirements for a clause learning SAT solver 
+\subsection{Properties of Domain Elements for Decision and Learning}
+An important property for a clause learning SAT solver 
 is that each element of the partial assignments domain 
-needs to have a decomposition into precisely complementable 
-elements.  This property of the underlying domain helps to 
+are decomposable into precisely complementable 
+elements.  This property of the domain helps to 
 guide the search away from the conflicting region of the 
@@
 Thus, clause learning SAT solvers requires a weak complementation property
-of the abstraction.  However, numerical abstract domain,  such as 
-intervals and octagons are intersections of complementable half-spaces. 
+of the abstraction.  Most numerical abstract domain, such as intervals, 
+octagons, polyhedra have complementation property, that is, the domain 
+elements are intersections of complementable half-spaces. 
+\Omit {
 The commonly used library for numerical abstract domains  

### 2016-10-13 experiments.tex
experiment description

--- 2016-09-22 experiments.tex
@@
-\section{Experimental Results}\label{experiments}
+\section{Experimental Results}
+We have implemented the instantiation of ACDCL on relational domains 
+inside 2LS verification tool~\cite{}.  The abstract domains used for 
+our experiments are intervals, template polyhedra, octagons and equalities.  
+We verified a total of 70 benchmarks in ANSI-C, which are derived from 
+(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
+cycle accurate software models of hardware circuits automatically 
+generated using v2c~\cite{tacas16}, (c) controller code generated 
+from Simulink (d) several hand-crafted benchmarks for equivalence 
+checking and bounded loop analysis.  All the benchmarks containing 
+bounded loops are completely unrolled before analysis.  We compare 
+our tool with a state-of-the-art bounded model checker CBMC~\cite{cbmc} 
+and a commercial abstract interpretation tool, Astr{\'e}e.  CBMC uses 
+MiniSAT solver in the backend.  Astr\{'e}e was configured with interval 
+domain, trace-partition domain, and relational domains.  To enable precise 
+analysis using Astr{\'e}e, all our benchmarks are manually instrumented with 
+partition directive which provides external hint to guide the trace partitioning 
+heuristics inside the tool.  Usually, such high-precision makes the analysis
+very expensive, and is not needed for static analysis of the code.  Without trace 
+partitioning, the analysis using Astr{\'e}e shows high degree of imprecision.

### 2016-10-13 contributions.tex
abstract

@@
 lifts the CDCL algorithm to relational numerical abstract domains.   
-Our proposed technique is embodied in our tool called, {\em
-  Titlis}. \pscmt{It's better not to create a zoo of tool names}  
-To the best of our knowledge, this is the first instantiation of CDCL 
+We have implemented our technique in 2LS verification tool.  To 
+the best of our knowledge, this is the first instantiation of CDCL 
 algorithm with relational abstract domains.
@@
 generalizes the Boolean Constraint Propagation (BCP) procedure to
-compute a fixed point over a relational domain.  Our abstract conflict analysis algorithm present a non-trivial 
-lifting of the Unique Implication Point (UIP) procedure to relational abstract 
+compute a fixed point over a relational domain.  Our abstract 
+conflict analysis algorithm present a non-trivial lifting of the 
+Unique Implication Point (UIP) procedure to relational abstract 
 domains in order to learn a generalized reason for conflicts.

### 2016-10-13 domains.tex
abstract

@@
 of the abstraction.  Most numerical abstract domain, such as intervals, 
-octagons, polyhedra have complementation property, that is, the domain 
-elements are intersections of complementable half-spaces. 
+octagons, polyhedra have complementation property, that is, they are
+representable as intersections of complementable half-spaces. 
 \Omit {
@@
 %\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+\Omit {
 \pscmt{That's no valid motivation. We never complement a whole octagon, but just a
   meet-irreducible. It's trivial to do that with APRON. The reason was
-a different one: APRON does not support all C operators, e.g. the bitwise operators.}
-For example, the complementation of the octagon in Figure~\ref{octagon} can be written as disjunction of:
+a different one: APRON does not support all C operators, e.g. the bitwise
+operators.} } For example, the complementation of the octagon in Figure~\ref{octagon} 
+can be written as disjunction of:
 \[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]

### 2016-10-13 paper.tex
abstract

@@
 %
-  \emph{Abstract Conflict-Driven Learning} (ACDL) is a combination of
-  abstract interpretation and the well-known Conflict-Driven Clause
-  Learning (CDCL) algorithm that is used in SAT solvers. ACDL has been
-  successfully used with non-relational abstract domains such as
-  intervals to improve the performance of a solver for floating point
-  arithmetic, for instance. 
+\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a combination of
+abstract interpretation and the well-known Conflict-Driven Clause Learning 
+(CDCL) algorithm that is used in SAT solvers.  ACDCL alternates between a 
+model search that performs overapproximate deduction with constraint 
+propagation and a conflict analysis that performs under-approximate 
+abduction with heuristic choice.  
 %
-Whereas the
-algorithms typically used in SAT solvers can be rather easily
-lifted from the Boolean lattice to other non-relational domains,
-lifting them to relational domains is challenging.
-%because of inter-"literal" dependencies
-%because of the weaker properties of "literals"
+ACDCL has been successfully used with non-relational abstract domains 
+such as intervals to improve the performance of a floating-point 
+decision procedure.  Whereas the propositional level algorithms used 
+in SAT solvers can be rather easily lifted from the boolean lattice to 
+other non-relational domains, lifting them to relational domains is 
+challenging because of relational properties of deductions and clause learning.
 %
-In this paper we present the adaptations of algorithms and heuristics
-for propagation, decision making and conflict learning necessary in
-order to enable the use with relational domains.
+In this paper, we present the non-trivial adaptations of the model search and 
+conflict analysis algorithms in ACDCL in order to enable the use of relational 
+abstract domains.  We have implemented these algorithms inside the 2LS 
+verification tool and instantiated our analysis over a template polyhedra 
+domain.  We evaluate the performance of our analyzer in comparison with 
+program verifiers that uses propositional-level solvers and a static 
+analysis tool.  
 %
-We have implemented these algorithms inside the 2LS verification tool
-for the example of the template polyhedra domain.
-%
-We evaluate the performance of our analyzer in comparison with program verifiers that 
-uses propositional-level solvers.  Experimental evidence on a set of 
-\rmcmt{N} benchmarks \rmcmt{drawn from X} shows significant reduction 
-in the number of decisions and backjumping iterations along with stronger 
-deductions and learned clause aided by the richer abstract domains.
+Experimental evidence on a set of benchmarks drawn from bit-vector 
+regression in SV-COMP'16, bit-precise software models of hardware 
+circuits auto-generated from v2c tool, and several bounds checking benchmarks 
+shows significant reduction in the number of decisions and backjumping iterations 
+along with stronger deductions and learned clauses aided by the richer abstract 
+domains.
 %
@@
 %===============================================================================
-\input{experiments}
+%\input{experiments}
 %===============================================================================

### 2016-10-13 experiments.tex
abstract
experiment reference

--- 2016-09-22 experiments.tex
@@
-\section{Experimental Results}\label{experiments}
+\section{Experimental Results}
+We have implemented the instantiation of ACDCL on relational domains 
+inside 2LS verification tool~\cite{2ls}.  The abstract domains used for 
+our experiments are intervals, template polyhedra, octagons and equalities.  
+We verified a total of 70 benchmarks in ANSI-C, which are derived from 
+(a) bit-vector regression category in SV-COMP\'16, (c) bit-precise and 
+cycle accurate software models of hardware circuits automatically 
+generated using v2c~\cite{mtk2016}, (c) controller code generated 
+from Simulink (d) several hand-crafted benchmarks for equivalence 
+checking and bounded loop analysis.  All the benchmarks containing 
+bounded loops are completely unrolled before analysis.  We compare 
+our tool with a state-of-the-art bounded model checker CBMC and a 
+commercial static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT 
+solver in the backend.  Astr{\'e}e was configured with interval 
+domain, trace-partition domain, and relational domains.  To enable precise 
+analysis using Astr{\'e}e, all our benchmarks are manually instrumented with 
+partition directive which provides external hint to guide the trace partitioning 
+heuristics inside the tool.  Usually, such high-precision makes the analysis
+very expensive, and is not needed for static analysis of the code.  Without trace 
+partitioning, the analysis using Astr{\'e}e shows high degree of imprecision.

### 2016-10-13 paper.tex
experiment reference

@@
 
-\author{Rajdeep Mukherjee\inst{1}\hskip 1em  Peter Schrammel\inst{2}\hskip 1em  Daniel Kroening\inst{1}}
+\author{Rajdeep Mukherjee\inst{1}\hskip 1em  Peter Schrammel\inst{2}\hskip 1em
+Daniel Kroening\inst{1}\hskip 1em Tom Melham\inst{1}}
 
 \institute{University of Oxford, UK \and University of Sussex, UK \\
-\email{\{rajdeep.mukherjee,kroening\}@cs.ox.ac.uk},
+\email{\{rajdeep.mukherjee,kroening,tom.melham\}@cs.ox.ac.uk},
 \email{p.schrammel@sussex.ac.uk}}
@@
 other non-relational domains, lifting them to relational domains is 
-challenging because of relational properties of deductions and clause learning.
+challenging because of relational properties of deductions and learnt clauses.
 %
@@
 shows significant reduction in the number of decisions and backjumping iterations 
-along with stronger deductions and learned clauses aided by the richer abstract 
+along with stronger deductions and learnt clauses aided by the richer abstract 
 domains.
@@
 %===============================================================================
-%\input{experiments}
+\input{experiments}
 %===============================================================================

### 2016-10-13 conclusion.tex
conclusion
git update
effect of decision heuristics

@@
 \section{Conclusions}
-In this paper, we instantiate ACDCL solver with different numerical domains -- 
-intervals and octagons.  Experimental evaluation on different benchmarks shows that
+In this paper, we instantiate ACDCL solver with relational domains.  We explained specific properties of model search and conflict analysis procedure necessary
+for such lifting.  Experimental evaluation on different benchmarks shows that
 ACDCL is able to suitably exploit the expressivity of richer
@@
 mechanism aided by the richer abstract domains.
+
+Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check.

### 2016-10-13 experiments.tex
effect of decision heuristics

--- 2016-10-11 experiments.tex
@@
 \section{Experimental Results}
-We classify our benchmarks into two groups: {\em bounds checking}, 
-{\em equivalence checking}.  We use interval, octagon, and equality 
-domain for our verification task.  We analyze programs with bounded 
-loops.  All loops in the program are completely unrolled before analysis.
+We have implemented the instantiation of ACDCL on relational domains 
+in 2LS verification tool~\cite{2ls}.  The abstract domains used for 
+our experiments are intervals, template polyhedra, octagons and equalities.  
+We verified a total of 70 benchmarks in ANSI-C, which are derived from 
+(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
+cycle accurate software models of hardware circuits automatically 
+generated using v2c~\cite{mtk2016}, (c) controller code generated 
+from Simulink (d) several hand-crafted benchmarks for equivalence 
+checking and bounded loop analysis.  All benchmarks containing 
+bounded loops are completely unrolled before analysis.  We compare 
+our tool with a state-of-the-art bounded model checker CBMC and a 
+commercial static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT 
+solver in the backend.  Astr{\'e}e was configured with interval 
+domain, trace-partition domain, and relational domains.  To enable precise 
+analysis using Astr{\'e}e, all our benchmarks are manually instrumented with 
+partition directives which provides external hint to the tool to guide 
+the trace partitioning heuristics.  Usually, such high-precision is not 
+needed for static analysis, since it makes the analysis very expensive.  
+Without trace partitioning, the analysis using Astr{\'e}e shows high 
+degree of imprecision.
+
+\paragraph {\em Observation of Analysis}
+
+\paragraph {\em Decision Heuristics} We compare the performance of 
+different decision heuristic on our benchmarks.  We observe that 
+the ordered heuristic outperforms other heuristics on control-intensive 
+benchmarks due to its ability to prioritize decisions on variables that 
+appears in conditional branches.  For straight-line code and most of 
+the bit-vector regression suite, random heuristics performs the best.  
+Whereas, the activity based heuristics such as Berkmin heuristic which 
+works well in propositional cases performs best for safe benchmarks 
+that encountered maximum number of conflicts to prove safety.
+
+\paragraph {\em Clause Learning}

### 2016-10-13 paper.tex

@@
 %
-\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a combination of
-abstract interpretation and the well-known Conflict-Driven Clause Learning 
-(CDCL) algorithm that is used in SAT solvers.  ACDCL alternates between a 
-model search that performs overapproximate deduction with constraint 
-propagation and a conflict analysis that performs under-approximate 
-abduction with heuristic choice.  
+\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalization
+of the well-known Conflict-Driven Clause Learning (CDCL) algorithm that is
+used in SAT solvers.  ACDCL alternates between a model search, which
+performs overapproximate deduction with constraint propagation, and a
+conflict analysis, which performs under-approximate abduction with heuristic
+choice.
 %
-ACDCL has been successfully used with non-relational abstract domains 
-such as intervals to improve the performance of a floating-point 
-decision procedure.  Whereas the propositional level algorithms used 
-in SAT solvers can be rather easily lifted from the boolean lattice to 
-other non-relational domains, lifting them to relational domains is 
-challenging because of relational properties of deductions and learnt clauses.
+ACDCL has been successfully used with non-relational abstract domains such
+as intervals to improve the performance of a floating-point decision
+procedure.  Whereas the propositional reasoning used in SAT solvers can be
+rather easily lifted from the boolean lattice to other non-relational
+domains, lifting them to relational domains is challenging because of
+relational properties of deductions and learnt clauses.
 %
-In this paper, we present the non-trivial adaptations of the model search and 
-conflict analysis algorithms in ACDCL in order to enable the use of relational 
-abstract domains.  We have implemented these algorithms inside the 2LS 
-verification tool and instantiated our analysis over a template polyhedra 
-domain.  We evaluate the performance of our analyzer in comparison with 
-program verifiers that uses propositional-level solvers and a static 
+In this paper, we present the non-trivial adaptations of the model search
+and conflict analysis algorithms in ACDCL in order to enable the use of
+relational abstract domains.  We have implemented these algorithms inside
+the 2LS verification tool and instantiated our analysis over a template
+polyhedra domain.  We evaluate the performance of our analyzer in comparison
+with program verifiers that uses propositional-level solvers and a static
 analysis tool.
@@
 Experimental evidence on a set of benchmarks drawn from bit-vector
-regression in SV-COMP'16, bit-precise software models of hardware 
-circuits auto-generated from v2c tool, and several bounds checking benchmarks 
-shows significant reduction in the number of decisions and backjumping iterations 
-along with stronger deductions and learnt clauses aided by the richer abstract 
-domains.
+regression in SV-COMP'16, bit-precise software models of hardware circuits
+auto-generated from v2c tool, and several bounds checking benchmarks shows
+significant reduction in the number of decisions and backjumping iterations
+along with stronger deductions and learnt clauses aided by the richer
+abstract domains.
 %

### 2016-10-13 experiments.tex
minor

@@
 
-\paragraph {\em Decision Heuristics} We compare the performance of 
+\paragraph {\em \textbf{Decision Heuristics}} We compare the performance of 
 different decision heuristic on our benchmarks.  We observe that 
@@
 Whereas, the activity based heuristics such as Berkmin heuristic which 
-works well in propositional cases performs best for safe benchmarks 
-that encountered maximum number of conflicts to prove safety.
+works well in propositional cases performs best for benchmarks that 
+encountered the maximum number of conflicts to prove safety, thus allowing 
+the heuristics to choose the decison variable from the set of learnt clauses.
 
-\paragraph {\em Clause Learning}
+\paragraph {\em \textbf{Clause Learning}}

### 2016-10-13 learning.tex
minor

@@
 formula $\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
-abstract valuation $v:= (x=[5,5])$.  
-Figure~\ref{conflict} shows a snapshot of abstract implication graph that 
+abstract valuation $v:= (x=5)$.  
+Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
 stores the deductions obtained from abstract model search using octagon domain.  

### 2016-10-14 example.tex
experiment reference
example finalised

--- 2016-10-11 example.tex
@@
 void main() {
-int x,sepc=0,impl=0;
-assume(x>=1 && x<=N);
-spec = 2*x; 
-int state = 1;
-int y=x;
-while(state!=0) {
- switch(state) {
-  case 0: state = 1; 
-  break;
-  case 1: state = 2; 
-  break;
-  case 2: x--; y++;
-  if(x == 0) state=0;
-  else state=1;break;
- } 
-}
-impl = y;
-assert(spec == impl);
+ int x,z;
+ int val;
+ assume(val>=0 && val<=N);
+ _Bool c;
+ if(c)
+  x = val;
+ else 
+  x = -val;
+ z = x * x;
+ assert(z>=0);
 }
@@
 \end{tabular}
-\caption{C Program, CFG and Abstract Conflict Graph}
+\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
 \label{example}
 \end{figure}
-Let us consider a simple program in Figure~\ref{example} that checks 
-the equivalence of a state-machine based interger multiplication against  
-a specification that uses multiplication operator (*).  The assertion 
-simply compares the result of two multiplication for an integer input $x$.  
-To avoid overflow error, we assume $x$ has an upper bound $N$.  The 
-control flow graph is shown in the middle column of Figure~\ref{example}.
+Let us consider a simple program in Figure~\ref{example} that performs 
+multiplication of two machine integers and checks that the result of the 
+multiplication is always positive for an integer input $x$.  To avoid 
+overflow error, we assume $x$ has an upper bound $N$.  The control 
+flow graph is shown in the middle column of Figure~\ref{example}.
 
@@
 model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
-For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop,
-table~\ref{solver} shows the statistics from MiniSAT~\cite{minisat} solver 
-for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
-constraint propagations and performs 5 conflict analysis with 65 conflict 
-literals.  Moreover, the solver performs a restart, presumably to recover 
-from bad decisions.
-
-On the other hand, abstract interpretation using forward analysis 
-is too imprecise to verify the safety of the above program using 
-interval domain.  The imprecision is due to control-flow join at 
-node $n1$.  State-of-the-art abstract interpretation tool, Astr{\'e}e, 
-requires external hints, provided by manually annotating the code 
-with partition directives at $n1$, to prove safety.  The partition 
-directive helps the tool to analyze each program execution paths 
-separately.  To do so, Astr{\'e}e requires a product of trace 
-partitioning domain and an interval domain to determine safety.  
-Consequently, the analysis becomes very expensive for a large value 
-of $N$.  In general, the imprecision is either intended by the tool 
-because such high precision analysis is normally not required for 
-runtime error analysis or the imprecision is unavoidable due to the 
-complexity of the application under analysis.  
+\Omit {
+The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
+and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
+the solver performs a restart, presumably to recover from bad decisions.  
+}
+Standard BMC tools convert the program to a bit-vector equation and pass 
+them to a SAT solver.  However, standard forward interval analysis is 
+too imprecise to verify the safety of the above program.  The 
+imprecision is due to control-flow join at node $n4$.  State-of-the-art 
+abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
+by manually annotating the code with partition directives at $n1$, 
+to prove safety.  The partition directive helps the tool to analyze 
+each program execution paths separately.  To do so, Astr{\'e}e 
+requires a product of trace partitioning domain and an interval domain 
+to determine safety.  However, a relational domain such as octagons is 
+able to prove the program without any partition.  In general, the 
+imprecision is either intended by the tool because such high 
+precision analysis is normally not required for runtime error 
+analysis or the imprecision is unavoidable due to the complexity 
+of the application under analysis.  
 %
@@
 Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
-SAT & 36 & 1365 & 5 & 65 & 1 \\ \hline
-ACDCL & 1 & 23 & 1 & 1 & 0 \\
+SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
+ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
+ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
 \hline
@@
 \end{center}
-\caption{Solver-statistics}
+\caption{Solver statistics}
 \label{solver}
@@
 The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
-interval domain.  The analysis associates an interval for each control 
-location and variable, as shown in the abstract conflict graph.  Interval 
-constraints generated from forward analysis in the initial deduction phase 
-infers that $\langle x:[-\infty,2], spec:[2,4], impl:[-1,\infty],
-state:[1,\infty], y:[1,+\infty] \rangle$.  Clearly, ACDCL can not prove safety.
+octagon domain (top) and an interval domain (bottom).  The analysis 
+associates an abstract element for each control location and variable, 
+as shown in the abstract conflict graph in Firgure~\ref{example}.  
+Octagonal constraints deduced by ACDCL solver are sufficient to prove 
+safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
+an automatic interval analysis is also sufficient to prove safety with 
+only a single decision.  Interval constraints generated from forward 
+analysis in the initial deduction phase infers that 
+$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL can not prove safety.
 
-Hence, ACDCL makes an arbitrary decision $x:[-\infty,1]$ to refine the analysis.  
-The decision constrains the interval of $x$.  The deductions made during fixed point
-iteration are represented by abstract conflict graph in right-side of 
-figure~\ref{example}.   Nodes in the graph denotes the corresponding 
-updates to the program variables in the cfg.  Interval analysis concludes 
-that the $(n12:\bot)$, that is the decision $x:[-\infty,1]$ leads to a 
-{\em conflict}. Intuitively, it means that the program is {\em safe} for 
-$x:[-\infty,1]$.
+Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
+ACDCL implements several decision heuritics.  In this case, the 
+branching variable $c$ is chosen by an {\em ordered} decision 
+heuristics that select variables appearing in conditional branches 
+first before choosing other program variables. 
+The decision constrains the interval of $x:=[0,5]$.  The deductions made 
+during fixed point iteration are represented by abstract conflict 
+graph in right-side of figure~\ref{example}.  Nodes in the abstract 
+conflict graph denotes the corresponding updates to the program 
+variables in the CFG.  Interval analysis concludes that the $(ERROR:\bot)$, that 
+is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
+the program is {\em safe} for $c:=[1,1]$.
 
@@
 Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
-(\(I: x = [2,+\infty]\)), that is all error traces must satisfy $(x \geq 2)$ 
+(\(I: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
 at node $I$.  The analysis discards all interval constraints that
 lead to the conflict and backtrack to decision level 0.  ACDCL then 
-performs interval analysis with the learnt constraint $(x>=2)$ which again 
+performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
 leads to a conflict.  The analysis cannot backtrack further and therefore 
@@
 
-The result of the analysis using ACDCL is shown in table~\ref{solver}.  
-Compared to a SAT solver, ACDCL makes only a single decision which lead 
-to 23 interval constraint propagations, and learns a single conflict clause 
-involving only one literal.  Compared to abstract interpretation, ACDCL does 
-not require external hints to proof the program, thus automatically performing 
-program and property driven trace partitioning to generate proofs using decision 
-and clause learning.
+For $N=46000$, table~\ref{solver} shows the statistics from 
+MiniSAT~\cite{minisat} solver for analysis using BMC, and 
+analysis using ACDCL with interval and octagon domains.  
+Compared to a SAT solver, there is a significant reduction 
+in the number of decisions, propagations, learnt clauses and 
+restrarts.  Compared to abstract interpretation, ACDCL does 
+not require external hints to proof the program, thus 
+automatically performing program and property driven trace 
+partitioning to generate proofs using decision and clause learning.

### 2016-10-14 background.tex
background.tex

--- 2016-10-12 related_work.tex
@@
-\section{Related Work}
+\section{Background}
 The natural domain SMT framework proposed in~\cite{ndsmt} operates on

### 2016-10-14 domains.tex
introduction

@@
 
-\subsection{Properties of Domain Elements for Decision and Learning}
-
 \begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
@@
 \subsection{Properties of Domain Elements for Decision and Learning}
+%
 An important property for a clause learning SAT solver 

### 2016-10-14 introduction.tex
introduction

@@
 
-\rmcmt{cite Natural-domain SMT}
 \rmcmt{relational domains}
@@
 are necessary to lift the propositional skeleton of CDCL to relational 
-abstract domains such as template polyhedra domains~\cite{sriram} (octagons, zones and 
-equalities).  To this end, we present a characterization of abstract 
-model search procedure and conflict analysis procedure that lift 
+abstract domains such as template polyhedra domains~\cite{sriram} 
+(octagons, zones and equalities).  To this end, we present a characterization 
+of abstract model search procedure and conflict analysis procedure that lift 
 the abstract deduction transformer and the first-UIP algorithm to 
@@
 
-Clause learning SAT solvers requires a weak complementation property 
-of the partial assignments domain.  However,  abstract domain 
-library, such as APRON~\cite{apron}, is specialized for abstract 
-interpretation framework which does not require complementation 
-operator.  Thus, we implement our own template-based numerical abstract 
-domains that support complementation property, required for clause learning 
-in the domain.
+Clause learning SAT solvers require a precise complementation property 
+of the partial assignments domain.  However, standard static analysis 
+based on abstract interpretation do not require complementation operator.   
+In this paper, we implement our own template-based numerical abstract 
+domains that precisely complements minimum complementable elements in 
+the abstract domain.

### 2016-10-14 domains.tex
section 3.4

@@
 
-In case of partial assignments, meet irreducibles are the singleton
+A meet irreducible in partial assignments domain are the singleton 
 assignments, for example $\{x \mapsto true \}$, which represents the set 
 of all propositional assignments where the literal $x$ is mapped to true.  
-An important property of meet irreducibles in case of partial assignment is that 
-they have precise complements.  For example, the complement of $\{x \mapsto true \}$
-is a singleton element,$\{x \mapsto false \}$, in the partial assignment domain.
+A meet irreducible in interval domain is $\langle x \leq n \rangle$ or 
+$\langle x \geq n \rangle$.
 
+%
 \begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
@@
 
+\Omit {
+An important property of meet irreducibles in case of partial assignments 
+domain is that they have precise complements.  
+For example, the complement of $\{x \mapsto true \}$ is a 
+singleton element,$\{x \mapsto false \}$, in the partial assignments domain. 
+}
+
+%
 %========================
@@
 
-\subsection{Properties of Domain Elements for Decision and Learning}
+\subsection{Properties of Domain Elements for Learning}
 %
-An important property for a clause learning SAT solver 
+\pscmt{TODO: Review by Peter}
+An important property of a clause learning SAT solver 
 is that each element of the partial assignments domain 
 are decomposable into precisely complementable 
-elements.  This property of the domain helps to 
-guide the search away from the conflicting region of the 
+elements.  This property of the domain helps to guide 
+the model search away from the conflicting region of the 
 search space.
-\rmcmt{define precise complement}
-Thus, clause learning SAT solvers requires a weak complementation property
-of the abstraction.  Most numerical abstract domain, such as intervals, 
-octagons, polyhedra have complementation property, that is, they are
-representable as intersections of complementable half-spaces. 
+
+Most numerical abstract domains, such as intervals, octagons, 
+polyhedra lacks precise complements, but they can be 
+represented as intersections of complementable half-spaces, 
+each of which have precise complements.  For example, intervals 
+lack complements, but the decomposition of intervals into 
+meet-irreducibles have complements.  
+\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
+\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
+Similarly, the complementation of the octagon in Figure~\ref{octagon} 
+can be written as disjunction of:
+\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
+
+Standard abstract interpretation does not require complementation property.  
+Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
+for abstract interpretation.  We implement a template-based polyhedra domain 
+that returns precise complements by decomposing it into half-spaces.
+
 \Omit {
@@
 }
-Abstract Domain library, such as APRON C library~\cite{apron}, 
-is specialized for the abstract interpretation framework which 
-does not require complementation property.  To this end, we 
-implement our own template-based polyhedra domain and interval 
+\Omit {
+To this end, we implement our own template-based polyhedra domain and interval 
 domain which supports complementation operator.  
@@
 %\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
-\Omit {
 \pscmt{That's no valid motivation. We never complement a whole octagon, but just a
@@
 a different one: APRON does not support all C operators, e.g. the bitwise
-operators.} } For example, the complementation of the octagon in Figure~\ref{octagon} 
-can be written as disjunction of:
-\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
+operators.} 
+}

### 2016-10-14 algorithm.tex
moved part of 4 up into 3

@@
 \section{Abstract Conflict Driven Clause Learning}
-In a typical abstract interpretation based galois-connection settings over 
-an over-approximate domain, every concrete elements have a unique over-approximate 
-representation in the abstract.  Likewise, every concrete transformer is
-over-approximated by a unique abstract transformer.  We now define a abstract 
-deduction transformer.
+In a typical abstract interpretation based Galois-connection setting
+over an over-approximate domain, every concrete element has a unique
+over-approximate representation in the abstract.  Likewise, every
+concrete transformer is over-approximated by a unique abstract
+transformer.  We now define an abstract deduction transformer.
 
@@
 %\rmcmt{safety as satisfiability}
-\subsection{Static Analysis Equations for Safety}
-Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
-perform safety analysis by computing fixed point to infer invariants 
-over program variables.  However, bounded model checking (BMC) tries to search 
-for a counterexample in a bounded execution trace, by symbolically executing a 
-program up to a finite bound.  
-Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
-For a set $N$ of program analysis constraints defined over a set 
-of constraint variables $Var = \{X_i| i \in N\}$, representing a program
-$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
-for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
-and the \rmcmt{final valuation} of constraint variables.
-\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
-{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
+
 %
@@
 %
-\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
+%\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
@@
 reason trail maps elements of propagation trail to transformers that were 
-used to drive them.
+used to derive them.
 

### 2016-10-14 domains.tex
moved part of 4 up into 3

--- 2016-10-14 domains.tex
@@
 \[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
+
+\subsection{Static Analysis Equations for Safety}
+Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+perform safety analysis by computing fixed point to infer invariants 
+over program variables.  However, bounded model checking (BMC) tries to search 
+for a counterexample in a bounded execution trace, by symbolically executing a 
+program up to a finite bound.  
+Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+For a set $N$ of program analysis constraints defined over a set 
+of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+and the \rmcmt{final valuation} of constraint variables.
+\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]

### 2016-10-14 domains.tex

--- 2016-10-14 domains.tex
@@
 }
+
+\subsection{Static Analysis Equations for Safety}
+Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+perform safety analysis by computing fixed point to infer invariants 
+over program variables.  However, bounded model checking (BMC) tries to search 
+for a counterexample in a bounded execution trace, by symbolically executing a 
+program up to a finite bound.  
+Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+For a set $N$ of program analysis constraints defined over a set 
+of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+and the \rmcmt{final valuation} of constraint variables.
+\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]

### 2016-10-14 algorithm.tex
moved more stuff from 4 to 3

@@
 \section{Abstract Conflict Driven Clause Learning}
-In a typical abstract interpretation based Galois-connection setting
-over an over-approximate domain, every concrete element has a unique
-over-approximate representation in the abstract.  Likewise, every
-concrete transformer is over-approximated by a unique abstract
-transformer.  We now define an abstract deduction transformer.
-
-\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
-transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
-domain $A$ is a sound approximation of a concrete model transformer
-$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
-$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
-$m \in A$ is a meet irreducible.   
-\end{definition}
-
-Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
-an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
-\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
-\rangle$.  An abstract deduction transformer is typically computed in the form 
-of strongest post-condition or a weakest pre-condition of a formula in the 
-abstract domain.
-
-A meet decomposition of the outcome of abstract deduction transformer 
-is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
-\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
-$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
-\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
-precisely complementable.
-
-\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
-element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
-\end{definition}
-
-For a program with $N$ variables, let $L$ be the total number of 
-meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
-maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
-of all octagonal inequalities in the set $L$.
-%\rmcmt{safety as satisfiability}
-
 %

### 2016-10-14 domains.tex
moved more stuff from 4 to 3

@@
 
+\subsection{Abstract Transformers}
+
+In a typical abstract interpretation based Galois-connection setting
+over an over-approximate domain, every concrete element has a unique
+over-approximate representation in the abstract.  Likewise, every
+concrete transformer is over-approximated by a unique abstract
+transformer.  We now define an abstract deduction transformer.
+
+\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+domain $A$ is a sound approximation of a concrete model transformer
+$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+$m \in A$ is a meet irreducible.   
+\end{definition}
+
+Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
+\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+\rangle$.  An abstract deduction transformer is typically computed in the form 
+of strongest post-condition or a weakest pre-condition of a formula in the 
+abstract domain.
+
+A meet decomposition of the outcome of abstract deduction transformer 
+is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
+\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+precisely complementable.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
+
+For a program with $N$ variables, let $L$ be the total number of 
+meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
+maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
+$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
+of all octagonal inequalities in the set $L$.
 %

### 2016-10-14 experiments.tex
update section 6
graph added

@@
 solver in the backend.  Astr{\'e}e was configured with interval 
-domain, trace-partition domain, and relational domains.  To enable precise 
-analysis using Astr{\'e}e, all our benchmarks are manually instrumented with 
-partition directives which provides external hint to the tool to guide 
-the trace partitioning heuristics.  Usually, such high-precision is not 
-needed for static analysis, since it makes the analysis very expensive.  
-Without trace partitioning, the analysis using Astr{\'e}e shows high 
-degree of imprecision.
+domain, trace-partition domain, and relational domains.  All the 
+bounded loops are completely unrolled in Astr{\'e}e.
 
+\Omit {
+To enable precise analysis using Astr{\'e}e, all our benchmarks are 
+manually instrumented with partition directives which provides external 
+hint to the tool to guide the trace partitioning heuristics.  Usually, 
+such high-precision is not needed for static analysis, since it makes 
+the analysis very expensive.  Without trace partitioning, the 
+analysis using Astr{\'e}e shows high degree of imprecision. 
+}
 \paragraph {\em Observation of Analysis}
@@
 
+\paragraph {\em \textbf{Propagation Strategy}}
+
 \paragraph {\em \textbf{Clause Learning}}

### 2016-10-14 graphs.tex
graph added

\section{Experiments}
\label{section:experiments}
\Omit {
We performed a number of experiments to demonstrate the utility and
applicability of {\algorithmName}.  All experiments were performed
on an Intel Xeon X5667 at 3\,GHz running Fedora 20 with 64-bit binaries.
Each individual run was limited to 13\,GB 
of memory and 900 seconds of
CPU time, % (with one exception noted below), 
enforced by the operating
system kernel.  We took the \emph{loops} meta-category (143 benchmarks) from the SV-COMP'15 benchmark
set.%
\footnote{\url{http://sv-comp.sosy-lab.org/2015/benchmarks.php}}
%

\subsection{\algorithmName\ Verifies More Programs Than the Algorithms it Simulates}

Table~\ref{tab:results} gives a comparison between \systemName\ running
\algorithmName\ (column 6) and \emph{the same system} running as an incremental
bounded model checker (IBMC) (column 2), incremental $k$-induction (i.e. without invariant inference, column 3)
and as an abstract interpreter (AI) (column 4).
%
 \algorithmName\ is more complete than each of the restricted modes. 
This is not self-evident since it could be much less efficient and,
thus, fail to solve the problems within the given time or memory limits.
%
$k$-induction can solve 60.8\% of the benchmarks, 13 more than IBMC. 
%
32\% of the benchmarks can be solved by abstract interpretation (bugs
are only exposed if they are reachable with 0 loop unwindings).  
%
\algorithmName\ solves 62.9\% of the benchmarks, 
proving 3 more properties than $k$-induction.

%%%%%%%%%%%%%%%%%%%%%%%%%% results table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{table}
\centering
\begin{tabular}{|l|ccc|c|c|c|c|}
\hline
& IBMC & $k$-induction & AI & portfolio & algorithmName &
                                                               CPAchecker & ESBMC \\
\hline
counterexamples & \bf 38 & \bf 38 &  17 & \bf 38 & \bf 38 & 36 &  35 \\
proofs          &    36 &  49 &  30 &  51 &  52 & 59 & \bf 91 \\
false proofs    &     0 &   0 &   0 &   0 &   0 &  2 &  12\\
false alarms    &     2 &   2 &   0 &   2 &   2 &  2 &   0\\
inconclusive    &     0 &   0 &  93 &   0 &   0 &  4 &   2\\
timeout         &    65 &  53 &   3 &  50 &  51 & 38 &   2 \\
memory out      &     2 &   1 &   0 &   2 &   0 &  2 &   1 \\
total runtime   & 17.1h & 13.8h & 0.89h & 13.3h & 13.2h & 10.9h & 0.54h \\
\hline
\end{tabular}~\\[1ex]
\caption{Comparison between algorithmName, the algorithms it subsumes,
the portfolio, and CPAchecker. The rows false alarms and false proofs 
indicate soundness bugs of the tool implementations.
\label{tab:results}}
\end{table}

\subsection{\algorithmName\ is at Least as Good as Their Na\"ive Portfolio}

To show
that \algorithmName\ is more than a mixture of three techniques and
that they strengthen each other, consider column 5 of
Table~\ref{tab:results}.  This gives the results of an ideal portfolio
in which the three restricted techniques are run in parallel on and
the portfolio terminates when the first returns a conclusive result.
Thus the CPU time taken is three times the time taken by the fastest
technique for each benchmark (in practice these could be run in
parallel, giving a lower \emph{wall clock} time).
%
In our setup, \algorithmName\ had a disadvantage as each component of
virtual portfolio had the same %time and 
memory restriction as
{\algorithmName}, thus effectively giving the portfolio three times as much memory.

Still, {\algorithmName} is slightly faster and more
accurate than the portfolio as can be seen in Table \ref{tab:results}.
%
The scatter plot in Figure~\ref{fig:results}a shows the results for
each benchmark:
%
one can observe that {\algorithmName} is up to one order of magnitude
slower on many unsafe benchmarks, which is obviously due to the
additional work of invariant inference that
{\algorithmName} has to perform in contrast to IBMC.
%
However, note that {\algorithmName} is faster than the 
portfolio on some safe and even one unsafe benchmarks.
This suggests that {\algorithmName} is more than the sum of its parts.

\subsection{\algorithmName\ is Comparable with State-of-the-Art Approaches}

We compared our implementation of \algorithmName\ with
CPAchecker%
\footnote{SVCOMP'15 version, http://cpachecker.sosy-lab.org/}%
%
and ESBMC%
\footnote{SVCOMP'15 version, http://www.esbmc.org/}%
, which uses $k$-induction.
%
The results are shown in the last three columns in Table \ref{tab:results}
and in the scatter plot in Figure~\ref{fig:results}b. Additional
results are given in \pponly{the extended version
  \cite{extended-version}}\rronly{Appendix \ref{sec:further}}.
%
In comparison to CPAchecker, the winner of SVCOMP'15,
our prototype of {\algorithmName} is overall a bit slower and proves 
fewer properties (due to more timeouts), but as
Figure~\ref{fig:results}b shows, it significantly outperforms
CPAchecker on most benchmarks.
%
ESBMC exposes fewer bugs, but proves many more properties and is 
significantly faster. However, it has 6 times
more soundness bugs than our implementation.%
\footnote{The two false alarms in our current implementation are due to 
limited support for dynamic memory allocation.}
%
These results show that our prototype implementation of
\algorithmName\ can keep up with state-of-the-art verification tools.

}
%%%%%%%%%%%%%%%%%%%%%%%%%% kiki vs vbs vs CPAchecker %%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{tabular}{@{\hspace{-1.5em}}c@{\hspace{1em}}c}
\begin{tikzpicture}[scale=0.75]
	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=kIkI (time in seconds),
			ylabel=portfolio (time in seconds), 
			legend style={at={(0.8,0.15)},
			anchor=north,legend columns=-1 },
			]
\addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
	table [meta=label] {scatter-kiki-vbs-sas.dat};
	\legend{Safe,Unsafe}
\addplot [domain=.1:1500] {x};
\addplot [red,sharp plot, domain=.1:1500] {900}
          node [below] at (axis cs:10,850) {timeout};
\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
          node [left,rotate=90] at (axis cs:700,10) {timeout}
 node [right,black] at (axis cs:10,3) {portfolio faster}
 node [right,black] at (axis cs:1,55) {kIkI faster};
%\addplot [red,sharp plot, update limits=false] coordinates{(900,.1) (900,1500)}
%	node [left] at {axis cs:700,200} {timeout};
\end{loglogaxis}
\end{tikzpicture}
 &
\begin{tikzpicture}[scale=0.75]
	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=kIkI (time in seconds),
			ylabel=CPAchecker (time in seconds),
			legend style={at={(0.8,0.15)},
			anchor=north,legend columns=-1 },
			]
\addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
	table [meta=label] {scatter-kiki-vbs-sas.dat};
	\legend{Safe,Unsafe}
\addplot [domain=.1:1500] {x};
\addplot [red,sharp plot, domain=.1:1500] {900}
          node [below] at (axis cs:10,850) {timeout};
\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
          node [left,rotate=90] at (axis cs:700,150) {timeout}
 node [right,black] at (axis cs:10,3) {CPAchecker faster}
 node [right,black] at (axis cs:1,55) {kIkI faster};
\end{loglogaxis}
\end{tikzpicture} \\
(a) & (b)
\end{tabular}
\caption{\label{fig:results}
Runtime Comparison
}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-10-14 example.tex

@@
 \end{tabular}
-\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
-\label{example}
+\caption{\label{example}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
+\pscmt{change \&\& in the CFG to $\wedge$; the constraint towards to error does not show up correctly; in the second oct analysis box put $\wedge$ between the constraints; add an arrow from the interval Error box with three dots to indicate that the analysis continues (even though you explain it in the text)}
+}
 \end{figure}

### 2016-10-14 example.tex
more comments on Fig1

--- 2016-10-14 example.tex
@@
 \end{tabular}
-\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
-\label{example}
+\caption{\label{example}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
+\pscmt{change \&\& in the CFG to $\wedge$; the constraint towards to error does not show up correctly; in the second oct analysis box put $\wedge$ between the constraints; add an arrow from the interval Error box with three dots to indicate that the analysis continues (even though you explain it in the text); shorten ``val'' to ``v''}
+}
 \end{figure}

### 2016-10-14 domains.tex
ssa fig, domain edits

@@
-\section{Program and Abstract Domains}
-In this section, we define a notion of program analysis constraints, concrete
-domain and abstract domain. \\
-\Omit {
-\subsection{Programs:}
-%
-A {\em program} $\mathcal{P}$ is formally defined as follows.
+\section{Preliminaries}
+This section explains the program model and the abstract domains that
+we are going to use.
+% \subsection{Programs:}
+% %
+% A {\em program} $\mathcal{P}$ is formally defined as follows.
+% \[
+% \begin{array}[t]{@{}lll}
+% \mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
+% \mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
+% \mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
+% \mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
+% \end{array}
+% \]
+% Consider sets of expressions $Exp$ and Boolean expressions $BExp$
+% over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
+% can take numeric values in $Val$.  A procedure is denoted as 
+% $\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
+% is an assignment, conditional, sequential concatenation or a loop. \\
+% \textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
+% where $Loc$ is the set of locations with an unique start 
+% location $(init)$ and error location $(err)$, $E$ is the set 
+% of control-flow edges that are labelled with the set 
+% $lbl \in Statement$.  For the purpose of illustration, we 
+% assume that all the procedures are inlined.
+
+\begin{figure}
 \[
-\begin{array}[t]{@{}lll}
-\mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
-\mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
-\mathit{Statement} & {:}{:}{=} & CStatement \mid \mathit{function(Var_1,\dots,Var_n)} \\
-\mathit{CStatement} & {:}{:}{=} & x {:}{=} exp \mid \mathit{ITE}(b, s_1,s_2) \mid \mathit{s_1;s_2} \mid loop\{s\} \\
+\begin{array}{rcl}
+g_0 &=& \true \\
+\multicolumn{3}{l}{0 \leq v\leq N} \\
+g_1 &=& g_0\wedge c\\
+x_0 &=& v \\
+x_1 &=& -v \\
+x_2 &=& g_1?x_0:x_1 \\
+g_2 &=& g_1 \vee g_0\wedge \neg c\\
+z &=& x_2 * x_2 \\
+\multicolumn{3}{l}{z<0 \wedge g_2}
 \end{array}
 \]
-Consider sets of expressions $Exp$ and Boolean expressions $BExp$
-over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
-can take numeric values in $Val$.  A procedure is denoted as 
-$\mathit{function(Var_1,\dots,Var_n)}$.  A $\mathit{CStatement}$ 
-is an assignment, conditional, sequential concatenation or a loop. \\
+\caption{\label{fig:ssa}
+SSA for the example in Figure~\ref{fig:example}
+}
+\end{figure}
 
-\textbf{Control-flow Graph:} A CFG is a triple $(Loc, E, lbl)$, 
-where $Loc$ is the set of locations with an unique start 
-location $(init)$ and error location $(err)$, $E$ is the set 
-of control-flow edges that are labelled with the set 
-$lbl \in Statement$.  For the purpose of illustration, we 
-assume that all the procedures are inlined.  
-}
-\textbf{Program Analysis Constraints:}  A program $\mathcal{P}$ is represented 
-as set of equations called, {\em program analysis constraints}.  The 
-program analysis constraints, as shown below, is an assignment, 
-or boolean expressions $BExp$ over program variables $Var$ or
-a non-deterministic choice (ternary experssion) between statements $s_1$ and
-$s_2$.  All the bounded loops are completely unrolled and represented using 
-following constraints. Similarly, an assertion is of type $BExp$.  
-\[c::= x:= exp | BExp | (exp) ? s_1, s_2\]
-For example, figure~\ref{swssa} presents the program analysis constraints for 
-a C program in Static Single Assignment (SSA) form.
+%-------------------------------------------------------------------------------
+\subsection{Program Representation}  
+%
+We consider loop-free programs (respectively bounded unwindings of
+programs with loops or recursion) with safety properties given as
+assertions in the code.
+%
+Such a program is represented as set $\mathcal{P}$ of
+\emph{statements} $s$.  These are constraints as typically obtained by
+translating the program into single static assignment (SSA) form via a
+data flow analysis.
+%
+Figure~\ref{fig:ssa} shows the constraints for the example program in
+Figure~\ref{fig:example}.
+%
+Assignments such as \texttt{x=v;} become equalities $x_1=v$ where the
+left-hand side variable gets a fresh name (subscript).
+%
+Control flow is encoded into guard variables such as $g_1=g_0\wedge c$.
+%
+Data flow joins become conditional expressions, e.g.\ $x_3=g_1?x_1:x_2$.
+%
+Assertions $\mathcal{A}$ are constraints $a$ of the form $g_3
+\Rightarrow z\geq 0$, meaning that if $g_3$ holds (i.e.\ the assertion
+is reachable) then the assertion must hold.
 
-\pscmt{we should give the (simplified) SSA for an example here}
+%For example, figure~\ref{swssa} presents the program analysis constraints for 
+%a C program in Static Single Assignment (SSA) form.
 
-\subsection{Concrete Semantics}
-\pscmt{not sure that adds anything for TACAS}
-The {\em concrete} domain is a lattice 
-of concrete environments, $Env = Var \rightarrow Val$, and is 
-defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
-A transformer, $post_{stmt}$, of concrete domain defines 
-the effect of a statement $stmt$ on the concrete domain, 
-$post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.
+% \subsection{Concrete Semantics}
+% \pscmt{not sure that adds anything for TACAS}
+% The {\em concrete} domain is a lattice 
+% of concrete environments, $Env = Var \rightarrow Val$, and is 
+% defined by $CDom = (Env, \sle_{C}, \join_{C}, \meet_{C})$.
+% A transformer, $post_{stmt}$, of concrete domain defines 
+% the effect of a statement $stmt$ on the concrete domain, 
+% $post_{stmt} : \powerset(Env) \rightarrow \powerset(Env)$.  
+% 
+% A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
+% where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
+% of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
+% $0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
+% if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.
 
-A state in concrete domain is a tuple $\langle l, \sigma \rangle$, 
-where $l$ is a location and $\sigma \in Env$.  A trace is a sequence 
-of states $(l_0, \sigma_0), \ldots (l_n, \sigma_n)$ such that for all 
-$0 \leq i \leq n$, there exists a cfg egde $(l_{i}, l_{i+1})$ 
-if $\sigma_{i+1} \in post_{i}(\sigma_{i})$.
+%-------------------------------------------------------------------------------
+\paragraph{Safety problem}  
+%
+Based on above program representation, the safety problem that our
+ACDL solver will solve is the following:
+%
+\begin{equation}\label{eq:safety}
+\bigwedge_{s\in\mathcal{P}} s \wedge \neg \bigwedge_{a\in\mathcal{A}} a
+\text{ is unsatisfiable if and only if the program is safe.}
+\end{equation}
 
-\subsection{Template Polyhedra Abstract Domain}
+%-------------------------------------------------------------------------------
+\subsection{Abstract Domain}
+%-------------------------------------------------------------------------------
+The results that we present on lifting ACDL to relational domains hold
+for any relational domain.
+% 
+In our prototype implementation and for illustrative purposes in this
+paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
+$\mathcal{B}$ is the Boolean domain
+$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$,
+$b$ the number of Boolean variables in the program, and $\mathcal{R}$
+is a \emph{relational} domain over the numerical (bitvector) variables
+in their respective data types (currently, signed and unsigned
+integers are supported).
+%
+We instantiate $\mathcal{R}$ with the template polyhedra domain.
 
-\pscmt{TODO}
+For notational convenience we will denote elements of
+$\mathcal{B}^b$ by their concretisations to propositional formulae,
+e.g.\ for the Boolean variable vector $(g_1,c)$ with abstract value
+$(\false,\top)$ we will write $\neg g_1$.
 
-\paragraph{Interval Abstract Domain} \pscmt{TODO: simplify}
-The {\em Interval} abstract domain is a lattice 
-$ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
-$ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
-the set of intervals of type $[l,u]$ over numeric data 
-type with $l \leq u$. The least element is $\bot$ and the 
-greatest element is $\top$ which maps all variables to their
-minimum ($min$) and maximum ($max$) values.  An interval 
-$\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
-partial order $\sle_{I}$ over elements in the set $Itv$ is 
-given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
-A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
-x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
-$\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
-A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
-x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
-$\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
-The galois connection between the concrete environment and intervals is 
-given as follows.
-\[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
-\[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
-For a program with $N$ variables, there are a total of 
-$2*N$ intervals. The intervals restrict the set of values 
-taken by program variables at each program point, starting 
-from $[-\infty, \infty]$.
+%-------------------------------------------------------------------------------
+\paragraph{Template Polyhedra Abstract Domain}
+%
+The template polyhedra domain~\cite{sriram} represents sets $X$ of
+valuations of a vector of variables $\vec{x}$ such that
+$\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
+and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
+by a special element to denote the minimal element $\bot$ of the lattice.
+%
+The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
+\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is
+applied component-wise. There are several optimisation-based
+techniques for computing the domain
+operations~\cite{sriram,GS07b,BJKS15}. In our implementation we use
+the strategy iteration approach available in 2LS~\cite{BJKS15}.
+%
+For notational convenience we will use conjunctions of linear
+inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
+of the domain. $\true$ is $\top$ and $\false$ is $\bot$.
 
-\paragraph{Octagon Abstract Domain} \pscmt{TODO: simplify}
-The {\em Octagon} abstract domain is a lattice 
-$OctDom = (OctElm, \sle_{O}, \join_{O}, \meet_{O})$, where
-$OctElm: (Var \times Var \rightarrow (\mathbb{R} \cup \{\infty\})) \cup \bot$. 
-The least element is $\bot$ that contains all unsatisfiable 
-set of inequalities and the greatest element is $\top$ which 
-maps the bounds of all octagonal inequalities to $\infty$. 
-The partial order $\sle_{O}$ over elements in the set $OctElm$ is 
-given by $O_1 \sle_{O} O_2$ iff the bounds of each inequalities in $0_1$ 
-is included (by $\leq$ order) in the bounds of corresponding inequalities 
-in $O_2$, that is, the octagons are ordered by the inclusion relations.
+% \paragraph{Interval Abstract Domain} 
+% The {\em Interval} abstract domain is a lattice 
+% $ItvDom = (ItvElm, \sle_{I}, \join_{I}, \meet_{I})$, where
+% $ItvElm: (Var \rightarrow Itv) \cup \{\bot\}$, and $Itv$ is 
+% the set of intervals of type $[l,u]$ over numeric data 
+% type with $l \leq u$. The least element is $\bot$ and the 
+% greatest element is $\top$ which maps all variables to their
+% minimum ($min$) and maximum ($max$) values.  An interval 
+% $\langle x, [min, v] \rangle$ is written as $x \leq v$.  The 
+% partial order $\sle_{I}$ over elements in the set $Itv$ is 
+% given by $I_1 \sle_{I} I_2$ if $I_2$ contains $I_1$.
+% A join $(\join_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
+% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
+% $\langle x_1 \rightarrow [min(l_1, l_2), max(u_1, u_2)]$.
+% A meet $(\meet_{I})$ of two intervals $\langle x_1 \rightarrow [l_1, u_1], 
+% x_1 \rightarrow [l_2, u_2] \rangle$ is an interval 
+% $\langle x_1 \rightarrow [max(l_1, l_2), min(u_1, u_2)] \rangle$.
+% The galois connection between the concrete environment and intervals is 
+% given as follows.
+% \[\alpha(x) = \{[inf(x), sup(x)] | x \in Var\} \qquad   \alpha(\emptyset) = \bot \]
+% \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
+Template polyhedra subsume, for example, interval and octagon domains.
 
-The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
-(v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
-and is computed by taking piece-wise maximum of bounds of corresponding 
-octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
-However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
-(v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
-computed by taking piece-wise minimum of bounds of corresponding 
-octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
-octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
-the octagons that abstract the same concrete values.
+\begin{wrapfigure}{r}{4.5cm}
+\vspace*{-3ex}
+\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
+\caption{An example of Octagon}\label{octagon}
+\vspace*{-3ex}
+\end{wrapfigure}
 
-The octagon domain is a relational abstract domain that permits $2n^2$ 
-linear inequalities between $n$ program variables.  The octagonal 
-inequalities are of two types: binary or unary inequalities as shown below.
-\[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
-An element in octagon domain, $OctElm$, is a conjunction of such 
-inequalities and is called an octagon.  Figure~\ref{octagon} shows an 
+For a program with $N$ variables, the matrix $\mat{C}$ for intervals
+generates at most $2N$ inequality constraints, one for the upper and lower
+bounds of each variable.
+%
+For octagons, we have at most $2N^2$ inequalities, one for the upper
+and lower bounds of each variable and sums and differences for each
+pair of variables. Figure~\ref{octagon} shows an 
 example octagon and its associated inequalities.
+
+The interval domain is a non-relational domain because a single
+inequality only contains a single variable.
 %
-\begin{wrapfigure}{r}{5.5cm}
-\caption{An example of Octagon}\label{octagon}
-\scalebox{.65}{\import{figures/}{octagons.pspdftex}}
-\end{wrapfigure} 
+The octagon domain, however, is relational.  A fundamental difference
+between relational and non-relational domains is that relational
+domains require the computation of a \emph{closure} in order to obtain
+a normal form that is usually required for efficient domain operations.
+An example of a closure computation is shown below.
+\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
+The closure computes all implied domain constraints.  For octagons,
+the closure operation is very expensive (cubic complexity in the
+number of program variables).
 %
-For building program analyzers using octagon domains, the domain also 
-provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
-The widening operator is used to accelerate convergence for loops in the program
-and has a quadratic complexity in the number of variables.  If the bound of an 
-octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
-to $\infty$.  However, the closure operator is often used to reduce the degree 
-of over-approximation resulting from the join operation. Closure always produces 
-unique octagonal inequalities, but it is very expensive to compute because of 
-cubic complexity in the number of program variables. An example of closure
-operation is shown below.
-\[((x-y \leq 4) \wedge (y-z \leq 5))^* \implies ((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]
+We will see that our algorithm performs closure computations lazily,
+postponing them to the point where they are actually necessary.
+
+% \paragraph{Octagon Abstract Domain} \pscmt{TODO: simplify}
+% The {\em Octagon} abstract domain is a lattice 
+% $OctDom = (OctElm, \sle_{O}, \join_{O}, \meet_{O})$, where
+% $OctElm: (Var \times Var \rightarrow (\mathbb{R} \cup \{\infty\})) \cup \bot$. 
+% The least element is $\bot$ that contains all unsatisfiable 
+% set of inequalities and the greatest element is $\top$ which 
+% maps the bounds of all octagonal inequalities to $\infty$. 
+% The partial order $\sle_{O}$ over elements in the set $OctElm$ is 
+% given by $O_1 \sle_{O} O_2$ iff the bounds of each inequalities in $0_1$ 
+% is included (by $\leq$ order) in the bounds of corresponding inequalities 
+% in $O_2$, that is, the octagons are ordered by the inclusion relations.
+
+% The join $(\join_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
+% (v_i \times v_j \rightarrow N_2) \rangle$, is not necessarily an octagon 
+% and is computed by taking piece-wise maximum of bounds of corresponding 
+% octagonal inequalities, $\langle (v_i \times v_j \rightarrow max(N_1, N_2) \rangle$.
+% However, the meet $(\meet_{O})$ of two octagons, $\langle (v_i \times v_j \rightarrow N_1),
+% (v_i \times v_j \rightarrow N_2) \rangle$, is always an octagon and is 
+% computed by taking piece-wise minimum of bounds of corresponding 
+% octagonal inequalities, $\langle (v_i \times v_j \rightarrow min(N_1, N_2)$.  A closed 
+% octagon is the smallest octagon following the partial order $\sle_{O}$, among all 
+% the octagons that abstract the same concrete values.
+
+% The octagon domain is a relational abstract domain that permits $2n^2$ 
+% linear inequalities between $n$ program variables.  The octagonal 
+% inequalities are of two types: binary or unary inequalities as shown below.
+% \[Binary: \pm v_i \pm v_j \leq a, v_i \neq v_j \qquad Unary: v_i \leq b, \{a, b\} \in \mathbb{R} \cup \infty \]  
+% An element in octagon domain, $OctElm$, is a conjunction of such 
+% inequalities and is called an octagon.  
+%
+% For building program analyzers using octagon domains, the domain also 
+% provides few operators like {\em widening ($\nabla$)} and {\em closure ($*$)}.  
+% The widening operator is used to accelerate convergence for loops in the program
+% and has a quadratic complexity in the number of variables.  If the bound of an 
+% octagonal inequality increases every iteration, then $\nabla$ sets the bounds 
+% to $\infty$.  However, the closure operator is often used to reduce the degree 
+% of over-approximation resulting from the join operation.
+
+\pscmt{TODO from here}
 

### 2016-10-14 example.tex
ssa fig, domain edits

@@
 \end{tabular}
-\caption{\label{example}
+\caption{\label{fig:example}
 C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
@@
 \end{figure}
-Let us consider a simple program in Figure~\ref{example} that performs 
+Let us consider a simple program in Figure~\ref{fig:example} that performs 
 multiplication of two machine integers and checks that the result of the 
@@
 overflow error, we assume $x$ has an upper bound $N$.  The control 
-flow graph is shown in the middle column of Figure~\ref{example}.
+flow graph is shown in the middle column of Figure~\ref{fig:example}.
 
@@
 %
-The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
+The right-side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
 octagon domain (top) and an interval domain (bottom).  The analysis 
 associates an abstract element for each control location and variable, 
-as shown in the abstract conflict graph in Firgure~\ref{example}.  
+as shown in the abstract conflict graph in Firgure~\ref{fig:example}.  
 Octagonal constraints deduced by ACDCL solver are sufficient to prove 
@@
 during fixed point iteration are represented by abstract conflict 
-graph in right-side of figure~\ref{example}.  Nodes in the abstract 
+graph in right-side of figure~\ref{fig:example}.  Nodes in the abstract 
 conflict graph denotes the corresponding updates to the program 

### 2016-10-15 example.tex
example

@@
  int x,z;
- int val;
- assume(val>=0 && val<=N);
+ int v;
+ assume(v>=0 $\wedge$ v<=N);
  _Bool c;
  if(c)
-   x = val;
+   x = v;
  else 
-   x = -val;
+   x = -v;
  z = x * x;

### 2016-10-15 domains.tex
meet irreducibles

@@
 
-\begin{figure}
+%-------------------------------------------------------------------------------
+\subsection{Program Representation}
+
+\begin{wrapfigure}{r}{3cm}
 \[
@@
 }
-\end{figure}
+\end{wrapfigure}
 
-%-------------------------------------------------------------------------------
-\subsection{Program Representation}  
-%
 We consider loop-free programs (respectively bounded unwindings of
@@
 %-------------------------------------------------------------------------------
-\paragraph{Safety problem}  
+\paragraph{Safety equation}  
 %
-Based on above program representation, the safety problem that our
-ACDL solver will solve is the following:
+Based on above program representation, the safety equation solved by the ACDCL solver is the following:
 %
 \begin{equation}\label{eq:safety}
-\bigwedge_{s\in\mathcal{P}} s \wedge \neg \bigwedge_{a\in\mathcal{A}} a
+\varphi:= \bigwedge_{s\in\mathcal{P}} s \wedge \neg \bigwedge_{a\in\mathcal{A}} a
 \text{ is unsatisfiable if and only if the program is safe.}
@@
 $\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$,
-$b$ the number of Boolean variables in the program, and $\mathcal{R}$
+$b$ is the number of Boolean variables in the program, and $\mathcal{R}$
 is a \emph{relational} domain over the numerical (bitvector) variables
@@
 \scalebox{.65}{\import{figures/}{octagons.pspdftex}}
-\caption{An example of Octagon}\label{octagon}
+\caption{Example of an octagon}\label{octagon}
 \vspace*{-3ex}
@@
 
-\pscmt{TODO from here}
+% \begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
+% mapping of variables to an element of abstract domain, for example 
+% a mapping of variable $x$ to an interval environment is given by 
+% $\langle x \mapsto [2,5] \rangle$ or a mapping of $\{x,y\}$ to octagon 
+% environment is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
+% An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
+% value or to $\bot$.  
+% \end{definition}
 
-The abstract transformer, $apost_{stmt}$, captures the effect of different program 
-statements in the abstract domain. The transformer is precise for octagonal 
-assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
-as shown below.
-\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
-
-\begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
-mapping of variables to an element of abstract domain, for example 
-a mapping of variable $x$ to an interval environment is given by 
-$\langle x \mapsto [2,5] \rangle$ or a mapping of $\{x,y\}$ to octagon 
-environment is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
-An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
-value or to $\bot$.  
+\begin{definition}{(Meet Irreducible)} A \emph{meet irreducible} $m$ 
+in a complete lattice structure $A$ is an element with the following property.
+\[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
 \end{definition}
 
-\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
-in a complete lattice structure $A$ is a minimum complementable element 
-$M \in A$ that has the following property.
-\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
+The meet irreducibles in the Boolean domain $\mathcal{B}$ are $\true$
+and $\false$. The meet irreducibles in the interval domain are all
+intervals that can be represented by a single inequality, i.e.\ $x \leq d$ or $x \geq d$ for constants $d$.
+
+\begin{definition}{(Complementable Meet Irreducible)} 
+A meet irreducible $m\in A$ is called complementable iff $\exists
+\bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
+%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
+%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
+%is the set complement of $\bar{M}$.  
+%If every meet irreducible in $A$ is complementable, then $A$ is said to
+%be have complementable meet irreducibles.
 \end{definition}
 
-A meet irreducible in partial assignments domain are the singleton 
-assignments, for example $\{x \mapsto true \}$, which represents the set 
-of all propositional assignments where the literal $x$ is mapped to true.  
-A meet irreducible in interval domain is $\langle x \leq n \rangle$ or 
-$\langle x \geq n \rangle$.
+The meet irreducibles in the Boolean and template polyhedra domains
+are all complementable.
 
-%
-\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
-irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
-irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
-is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
-complementable, then $A$ is said to be have complementable meet irreducibles.  
-\end{definition}
-
-\Omit {
-An important property of meet irreducibles in case of partial assignments 
-domain is that they have precise complements.  
-For example, the complement of $\{x \mapsto true \}$ is a 
-singleton element,$\{x \mapsto false \}$, in the partial assignments domain. 
-}
+%An important property of meet irreducibles in case of partial assignments 
+%domain is that they have precise complements.  
+%For example, the complement of $\{x \mapsto true \}$ is a 
+%singleton element,$\{x \mapsto false \}$, in the partial assignments domain.
 
@@
 transformer.  We now define an abstract deduction transformer.
+
+The abstract transformer, $apost_{stmt}$, captures the effect of different program 
+statements in the abstract domain. The transformer is precise for octagonal 
+assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
+as shown below.
+\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
 

### 2016-10-15 domains.tex
meet irreducibles

@@
 \begin{definition}{(Complementable Meet Irreducible)} 
-A meet irreducible $m\in A$ is called complementable iff $\exists
+A meet irreducible $m\in A$ is called \emph{complementable} iff $\exists
 \bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
@@
 %singleton element,$\{x \mapsto false \}$, in the partial assignments domain.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$a=\bigscap_{m\in M} m$.
+%$forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
 
@@
 precisely complementable.
-
-\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
-element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
-\end{definition}
 

### 2016-10-15 domains.tex
meet decomposition

@@
 
-\begin{definition}{(Meet Irreducible)} A \emph{meet irreducible} $m$ 
+\begin{definition} 
+A \emph{meet irreducible} $m$ 
 in a complete lattice structure $A$ is an element with the following property.
@@
 
-\begin{definition}{(Complementable Meet Irreducible)} 
+\begin{definition} 
 A meet irreducible $m\in A$ is called \emph{complementable} iff $\exists
@@
 
-\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+\begin{definition}
+ A \emph{meet decomposition} of an abstract
 element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$a=\bigscap_{m\in M} m$.
+$a=\bigsqcap_{m\in M} m$.
 %$forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
 \end{definition}
+
+For example, the meet decomposition of the interval domain element
+% $(x,y) \in [2,4]\times[3, 5]$ 
+$2\leq x\leq 4 \wedge 3\leq y\leq 5$ is
+the set $\{x\geq 2, x\leq 4, y\geq 3, y\leq 5\}$.
+% \langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+% $\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+% \langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+% precisely complementable.
 
@@
 
-A meet decomposition of the outcome of abstract deduction transformer 
-is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
-\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
-$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
-\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
-precisely complementable.
-
-For a program with $N$ variables, let $L$ be the total number of 
-meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
-maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
-of all octagonal inequalities in the set $L$.
+% For a program with $N$ variables, let $L$ be the total number of 
+% meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
+% maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
+% $2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
+% of all octagonal inequalities in the set $L$.
 %

### 2016-10-15 domains.tex

--- 2016-10-15 domains.tex
@@
 
-%-------------------------------------------------------------------------------
-\subsection{Program Representation}
-
-\begin{wrapfigure}{r}{3cm}
+\begin{figure}
 \[
@@
 }
-\end{wrapfigure}
+\end{figure}
 
+%-------------------------------------------------------------------------------
+\subsection{Program Representation}  
+%
 We consider loop-free programs (respectively bounded unwindings of
@@
 \scalebox{.65}{\import{figures/}{octagons.pspdftex}}
-\caption{Example of an octagon}\label{octagon}
+\caption{An example of Octagon}\label{octagon}
 \vspace*{-3ex}
@@
 
-% \begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
-% mapping of variables to an element of abstract domain, for example 
-% a mapping of variable $x$ to an interval environment is given by 
-% $\langle x \mapsto [2,5] \rangle$ or a mapping of $\{x,y\}$ to octagon 
-% environment is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
-% An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
-% value or to $\bot$.  
-% \end{definition}
+\pscmt{TODO from here}
 
-\begin{definition}{(Meet Irreducible)} A \emph{meet irreducible} $m$ 
-in a complete lattice structure $A$ is an element with the following property.
-\[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
+The abstract transformer, $apost_{stmt}$, captures the effect of different program 
+statements in the abstract domain. The transformer is precise for octagonal 
+assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
+as shown below.
+\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
+
+\begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
+mapping of variables to an element of abstract domain, for example 
+a mapping of variable $x$ to an interval environment is given by 
+$\langle x \mapsto [2,5] \rangle$ or a mapping of $\{x,y\}$ to octagon 
+environment is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
+An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
+value or to $\bot$.  
 \end{definition}
 
-The meet irreducibles in the Boolean domain $\mathcal{B}$ are $\true$
-and $\false$. The meet irreducibles in the interval domain are all
-intervals that can be represented by a single inequality, i.e.\ $x \leq d$ or $x \geq d$ for constants $d$.
-
-\begin{definition}{(Complementable Meet Irreducible)} 
-A meet irreducible $m\in A$ is called complementable iff $\exists
-\bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
-%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
-%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
-%is the set complement of $\bar{M}$.  
-%If every meet irreducible in $A$ is complementable, then $A$ is said to
-%be have complementable meet irreducibles.
+\begin{definition}{(Meet Irreducibles)} A {\em meet irreducible} $(M)$ 
+in a complete lattice structure $A$ is a minimum complementable element 
+$M \in A$ that has the following property.
+\[\forall M_1, M_2 \in A, M_1 \meet M2 = M \implies (M = M_1 \lor M = M_2), M \neq \top \]  
 \end{definition}
 
-The meet irreducibles in the Boolean and template polyhedra domains
-are all complementable.
+A meet irreducible in partial assignments domain are the singleton 
+assignments, for example $\{x \mapsto true \}$, which represents the set 
+of all propositional assignments where the literal $x$ is mapped to true.  
+A meet irreducible in interval domain is $\langle x \leq n \rangle$ or 
+$\langle x \geq n \rangle$.
 
-%An important property of meet irreducibles in case of partial assignments 
-%domain is that they have precise complements.  
-%For example, the complement of $\{x \mapsto true \}$ is a 
-%singleton element,$\{x \mapsto false \}$, in the partial assignments domain.
+%
+\begin{definition}{(Complementable Meet Irreducibles)} A complementable meet
+irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
+irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
+is the set complement of $\bar{M}$.  If every meet irreducible in $A$ is
+complementable, then $A$ is said to be have complementable meet irreducibles.  
+\end{definition}
+
+\Omit {
+An important property of meet irreducibles in case of partial assignments 
+domain is that they have precise complements.  
+For example, the complement of $\{x \mapsto true \}$ is a 
+singleton element,$\{x \mapsto false \}$, in the partial assignments domain. 
+}
 
@@
 transformer.  We now define an abstract deduction transformer.
-
-The abstract transformer, $apost_{stmt}$, captures the effect of different program 
-statements in the abstract domain. The transformer is precise for octagonal 
-assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
-as shown below.
-\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
 

### 2016-10-15 experiments.tex
figure 1

@@
 encountered the maximum number of conflicts to prove safety, thus allowing 
-the heuristics to choose the decison variable from the set of learnt clauses.
+the heuristics to choose the decison variable among the set of learnt clauses.
 

### 2016-10-15 domains.tex
more restructuring in 3

--- 2016-10-15 domains.tex
@@
 %-------------------------------------------------------------------------------
-\paragraph{Safety equation}  
+\paragraph{Safety problem}  
 %
-Based on above program representation, the safety equation solved by the ACDCL solver is the following:
+Based on above program representation, the safety problem solved by the ACDCL solver is the following:
 %
@@
 \end{equation}
+
+% \subsection{Static Analysis Equations for Safety}
+% Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+% perform safety analysis by computing fixed point to infer invariants 
+% over program variables.  However, bounded model checking (BMC) tries to search 
+% for a counterexample in a bounded execution trace, by symbolically executing a 
+% program up to a finite bound.  
+% Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+% For a set $N$ of program analysis constraints defined over a set 
+% of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+% $\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+% for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+% and the \rmcmt{final valuation} of constraint variables.
+% \[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+% {\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]
 
@@
 paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
-$\mathcal{B}$ is the Boolean domain
-$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$,
-$b$ is the number of Boolean variables in the program, and $\mathcal{R}$
-is a \emph{relational} domain over the numerical (bitvector) variables
-in their respective data types (currently, signed and unsigned
-integers are supported).
+$\mathcal{B}$ is the partial Boolean assignments domain
+$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$ (as
+used in SAT solvers~\cite{sas12}), $b$ is the number of Boolean
+variables in the program, and $\mathcal{R}$ is a \emph{relational}
+domain over the numerical (bitvector) variables in their respective
+data types (currently, signed and unsigned integers are supported).
 %
@@
 
+%-------------------------------------------------------------------------------
+\subsection{Properties of Abstract Domains for Learning}
+%
+An important property of a clause learning SAT solver is that each
+element of the partial Boolean assignments domain is decomposable into
+complementable elements.  This property of the domain is required to
+learn elements that help to guide the model search away from the
+conflicting region of the search space.
+%
+Most numerical abstract domains, such as intervals and octagons lack
+complements, but they can be represented as intersections of
+complementable half-spaces, each of which has a complement. We
+formalise this in the sequel.
+
 \begin{definition} 
@@
 
-The meet irreducibles in the Boolean domain $\mathcal{B}$ are $\true$
-and $\false$. The meet irreducibles in the interval domain are all
-intervals that can be represented by a single inequality, i.e.\ $x \leq d$ or $x \geq d$ for constants $d$.
-
-\begin{definition} 
-A meet irreducible $m\in A$ is called \emph{complementable} iff $\exists
-\bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
-%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
-%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
-%is the set complement of $\bar{M}$.  
-%If every meet irreducible in $A$ is complementable, then $A$ is said to
-%be have complementable meet irreducibles.
-\end{definition}
-
-The meet irreducibles in the Boolean and template polyhedra domains
-are all complementable.
+The meet irreducibles in the partial Boolean assignments domain
+$\mathcal{B}$ are $\true$ and $\false$. The meet irreducibles in the
+template polyhedra domain are all elements that concretise to
+half-spaces, i.e.\ they can be represented by a single inequality. For
+the interval domain, these are \ $x \leq d$ or $x \geq d$ for
+constants $d$.
 
@@
 
+For polyhedra this intuitively means that each polyhedron can be
+written as an intersection of half-spaces.
+%
 For example, the meet decomposition of the interval domain element
@@
 
+\begin{definition} 
+A meet irreducible $m\in A$ is called \emph{complementable} iff $\exists
+\bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
+%A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
+%irreducible $M \in A$ such that $\bar{M} \in A$ and the concretisation of $M$ 
+%is the set complement of $\bar{M}$.  
+%If every meet irreducible in $A$ is complementable, then $A$ is said to
+%be have complementable meet irreducibles.
+\end{definition}
+
+The meet irreducibles in the partial Boolean assignments domain and
+template polyhedra domains are all complementable.
+%
+Note that for template polyhedra, all non-meet-irreducible elements $e$
+(except $\bot$ and $\top$) are not complementable within $A$.
+%
+However, we can complement each element in the meet decomposition of $e$ and
+re-interpret the obtained set as a disjunction. 
+%
+Thus the complement of the octagon in Figure~\ref{octagon} 
+can be written as a disjunction of meet irreducibles:
+\[(x\leq -3) \lor (x\geq 2) \lor (y\leq -2) \lor (y\geq 3) \lor (x+y\geq 3) \lor (x-y\geq 2) \lor (y-x\geq 4) \lor (-x-y\geq 3)\]
+
+Standard abstract interpretation does not require complementation property.  
+Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
+for abstract interpretation.  We implement a template-based polyhedra domain 
+that returns precise complements by decomposing it into half-spaces.
+
+%-------------------------------------------------------------------------------
 \subsection{Abstract Transformers}
-
+%
 In a typical abstract interpretation based Galois-connection setting
@@
 
-\subsection{Properties of Domain Elements for Learning}
-%
-\pscmt{TODO: Review by Peter}
-An important property of a clause learning SAT solver 
-is that each element of the partial assignments domain 
-are decomposable into precisely complementable 
-elements.  This property of the domain helps to guide 
-the model search away from the conflicting region of the 
-search space.
-
-Most numerical abstract domains, such as intervals, octagons, 
-polyhedra lacks precise complements, but they can be 
-represented as intersections of complementable half-spaces, 
-each of which have precise complements.  For example, intervals 
-lack complements, but the decomposition of intervals into 
-meet-irreducibles have complements.  
-\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
-\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
-Similarly, the complementation of the octagon in Figure~\ref{octagon} 
-can be written as disjunction of:
-\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
-
-Standard abstract interpretation does not require complementation property.  
-Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
-for abstract interpretation.  We implement a template-based polyhedra domain 
-that returns precise complements by decomposing it into half-spaces.
-
-\Omit {
-The commonly used library for numerical abstract domains  
-is the APRON C library~\cite{apron}.  This library is 
-used for the static analysis of the numerical variables 
-of a program by abstract interpretation. APRON provides a 
-C API interface to various abstract domains and libraries 
-such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
-{\em Linear Equalities} library.  The aim of such analysis is 
-to compute invariants over numerical variables in the 
-program~\cite{se2011}. 
-}
-\Omit {
-To this end, we implement our own template-based polyhedra domain and interval 
-domain which supports complementation operator.  
-%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
-%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
-%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
-\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
-  meet-irreducible. It's trivial to do that with APRON. The reason was
-a different one: APRON does not support all C operators, e.g. the bitwise
-operators.} 
-}
-
-\subsection{Static Analysis Equations for Safety}
-Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
-perform safety analysis by computing fixed point to infer invariants 
-over program variables.  However, bounded model checking (BMC) tries to search 
-for a counterexample in a bounded execution trace, by symbolically executing a 
-program up to a finite bound.  
-Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
-For a set $N$ of program analysis constraints defined over a set 
-of constraint variables $Var = \{X_i| i \in N\}$, representing a program
-$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
-for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
-and the \rmcmt{final valuation} of constraint variables.
-\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
-{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]
+% \Omit {
+% The commonly used library for numerical abstract domains  
+% is the APRON C library~\cite{apron}.  This library is 
+% used for the static analysis of the numerical variables 
+% of a program by abstract interpretation. APRON provides a 
+% C API interface to various abstract domains and libraries 
+% such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
+% {\em Linear Equalities} library.  The aim of such analysis is 
+% to compute invariants over numerical variables in the 
+% program~\cite{se2011}. 
+% }
+% \Omit {
+% To this end, we implement our own template-based polyhedra domain and interval 
+% domain which supports complementation operator.  
+% %For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
+% %\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
+% %\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+% \pscmt{That's no valid motivation. We never complement a whole octagon, but just a
+%  meet-irreducible. It's trivial to do that with APRON. The reason was
+% a different one: APRON does not support all C operators, e.g. the bitwise
+% operators.} 
+% }

### 2016-10-15 domains.tex
sync, some stuff on transformers

@@
 \begin{definition}
- A \emph{meet decomposition} of an abstract
+ A \emph{meet decomposition} $\mathit{decomp}(a)$ of an abstract
 element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
@@
 \subsection{Abstract Transformers}
-%
+%-------------------------------------------------------------------------------
 In a typical abstract interpretation based Galois-connection setting
@@
 transformer.  We now define an abstract deduction transformer.
+
+$\llbracket s \rrbracket^\sharp(a)=\min \{a'\mid a'\sqsubset a, a'\models s\}$
+
+\pscmt{define $a'\models s$}
+
+We define a specialised variant of the transformer to compute deductions w.r.t.\ a given subdomain $L\subseteq A$.
+
+$\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\sqsubset a, a'\in L, a'\models s\}$
+
+We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq a\meet\llbracket s \rrbracket^\sharp(a,L)$.
+
+For example,
+$a=(0\leq y \leq 1 \wedge 5\leq z)$
+$\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
+$\llbracket x=y+z\rrbracket^\sharp(a,\{y\})=\top$.
+
+%$a=(x-y \leq 1, y-x \leq 1)$
+%$\llbracket x=y+z\rrbracket^\sharp($
+
+Moreover, we want to know what the reasons for a specific deduction are.
+$ded(s,a,L)=\{R\rightarrow d\mid R=reason(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}$
+where 
+$reason(s,a,L,d)=\min \{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}$.
+
+\pscmt{TODO from here}
 
@@
 %========================
-\input{literals-clauses}
+%\input{literals-clauses}
 %========================

### 2016-10-15 example.tex

@@
 \caption{\label{fig:example}
-C Program, Control-flow Graph (CFG) and Abstract Conflict Graph
-\pscmt{change \&\& in the CFG to $\wedge$; the constraint towards to error does not show up correctly; in the second oct analysis box put $\wedge$ between the constraints; add an arrow from the interval Error box with three dots to indicate that the analysis continues (even though you explain it in the text); shorten ``val'' to ``v''}
-}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
 \end{figure}

### 2016-10-15 domains.tex
complement properties

--- 2016-10-15 domains.tex
@@
 \end{definition}
+For example, the meet decomposition of an interval is given as follows.
+\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
+\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
 
@@
 %
-\pscmt{TODO: Review by Peter}
 An important property of a clause learning SAT solver 
-is that each element of the partial assignments domain 
-are decomposable into precisely complementable 
-elements.  This property of the domain helps to guide 
-the model search away from the conflicting region of the 
-search space.
+is that a partial assignments domain can be decomposed 
+into a set of precisely complementable singleton assignments.  
+This property of the partial assignments domain helps to 
+guide the model search away from the conflicting region 
+of the search space and allows case-based analysis.
 
-Most numerical abstract domains, such as intervals, octagons, 
-polyhedra lacks precise complements, but they can be 
-represented as intersections of complementable half-spaces, 
-each of which have precise complements.  For example, intervals 
-lack complements, but the decomposition of intervals into 
-meet-irreducibles have complements.  
-\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
-\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
+The precise complementation property of partial assignment lattice 
+can be generalized to other lattice structures.  Most numerical 
+abstract domains, such as intervals, octagons, polyhedra, can be 
+decomposed into half-spaces, that admits precise complements.  
+For example, intervals lack complements, but the meet decomposition 
+of intervals into meet-irreducibles have precise complements.  
+Numerical abstract domains that admit complementable decomposition 
+are shown in Figure~\ref{complement}.  Our template-based domain 
+implementation precise complements a meet irreducible in an abstraction.
+% 
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l l}
+\hline
+Domain & Precise Complements
+Interval & $(x \leq n) \quad \longrightarrow (x > n)$ \\ \hline
+Octagon & $(x+y < 1) \quad \longrightarrow (-x-y < 0)$ \\ \hline
+Equality & $(x==y) \quad \longrightarrow (x \neq y)$ \\ \hline 
+\end{tabular}
+\label{fig:complement}
+%\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
+\end{figure}
+%
+\Omit{
 Similarly, the complementation of the octagon in Figure~\ref{octagon} 
@@
 \[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
-
-Standard abstract interpretation does not require complementation property.  
+Standard abstract interpretation do not require complementation property
+The application of set-complements on over-approximating 
+domains does not give necessary invariants required to reason about assertions
+in programs. 
 Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
-for abstract interpretation.  We implement a template-based polyhedra domain 
-that returns precise complements by decomposing it into half-spaces.
-
+for abstract interpretation.  
+}
 \Omit {
@@
 } 
-
+\Omit {
 \subsection{Static Analysis Equations for Safety}
@@
 {\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
+}

### 2016-10-15 introduction.tex
complement properties

@@
 relational lattice structures.
-
-Clause learning SAT solvers require a precise complementation property 
-of the partial assignments domain.  However, standard static analysis 
-based on abstract interpretation do not require complementation operator.   
-In this paper, we implement our own template-based numerical abstract 
-domains that precisely complements minimum complementable elements in 
-the abstract domain.

### 2016-10-15 domains.tex
domain section 3.5 added

--- 2016-10-15 domains.tex
@@
 
+A meet decomposition of the outcome of abstract deduction transformer 
+is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
+\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
+$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
+\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
+precisely complementable.
+
+\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
+element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
+$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
+\end{definition}
+For example, the meet decomposition of an interval is given as follows.
+\[x \mapsto [2,5] \mathrel{\mathop{\longrightarrow}^{\mathrm{decompose}}} \{\langle x
+\geq 2 \rangle \meet \langle x \leq 5 \rangle \} \]
+
 % For a program with $N$ variables, let $L$ be the total number of 
@@
 
+\subsection{Properties of Domain Elements for Learning}
+%
+An important property of a clause learning SAT solver 
+is that a partial assignments domain can be decomposed 
+into a set of precisely complementable singleton assignments.  
+This property of the partial assignments domain helps to 
+guide the model search away from the conflicting region 
+of the search space and allows case-based analysis.  
+%
+\begin{definition}{(Precise complement)} A precise complement of an abstract
+element $a \in A$ in abstract lattice $A$ is an element $\bar{a}$ such that 
+$\gamma(\bar{a} = \neg \gamma(a))$.
+\end{definition}
+%
+The precise complementation property of partial assignment lattice 
+can be generalized to other lattice structures.  Most numerical 
+abstract domains, such as intervals, octagons, polyhedra, can be 
+decomposed into half-spaces, that admits precise complements.  
+For example, intervals lack complements, but the meet decomposition 
+of intervals into meet-irreducibles have precise complements.  
+Numerical abstract domains that admit complementable decomposition 
+are shown in Figure~\ref{complement}.  Our template-based domain 
+implementation precisely complements a meet irreducible in an abstraction.
+% 
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l}
+\hline
+Domain & Precise Complements \\ \hline 
+Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
+Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
+Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
+\end{tabular}
+}
+\end{center}
+\label{fig:complement}
+%\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
+\end{table}
+%
+\Omit{
+Similarly, the complementation of the octagon in Figure~\ref{octagon} 
+can be written as disjunction of:
+\[(x<-2) \lor (x>1) \lor (y<-1) \lor (y>2) \lor (x+y>2) \lor (x-y>1) \lor (y-x>3) \lor (-x-y>2)\]
+Standard abstract interpretation do not require complementation property
+The application of set-complements on over-approximating 
+domains does not give necessary invariants required to reason about assertions
+in programs. 
+Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
+for abstract interpretation.  
+}
+\Omit {
+The commonly used library for numerical abstract domains  
+is the APRON C library~\cite{apron}.  This library is 
+used for the static analysis of the numerical variables 
+of a program by abstract interpretation. APRON provides a 
+C API interface to various abstract domains and libraries 
+such as {\em BOX}, {\em OCTAGON}, {\em Convex Polyhedra} and
+{\em Linear Equalities} library.  The aim of such analysis is 
+to compute invariants over numerical variables in the 
+program~\cite{se2011}. 
+}
+\Omit {
+To this end, we implement our own template-based polyhedra domain and interval 
+domain which supports complementation operator.  
+%For example, the octagon in Figure~\ref{octagon} can be written as a conjunction of:
+%\[(x>=-2) \land (x<=1) \land (y>=-1) \land (y<=2) \land (x+y<=2)
+%\land (x-y<=1) \land (y-x<=3) \land (-x-y<=2)\] 
+\pscmt{That's no valid motivation. We never complement a whole octagon, but just a
+  meet-irreducible. It's trivial to do that with APRON. The reason was
+a different one: APRON does not support all C operators, e.g. the bitwise
+operators.} 
+} 
+\Omit {
+\subsection{Static Analysis Equations for Safety}
+Static program analysis based on abstract interpretation~\cite{DBLP:conf/emsoft/Cousot07} 
+perform safety analysis by computing fixed point to infer invariants 
+over program variables.  However, bounded model checking (BMC) tries to search 
+for a counterexample in a bounded execution trace, by symbolically executing a 
+program up to a finite bound.  
+Similar to BMC, ACDCL searches for a counterexample by solving the formula shown below.
+For a set $N$ of program analysis constraints defined over a set 
+of constraint variables $Var = \{X_i| i \in N\}$, representing a program
+$\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+and the \rmcmt{final valuation} of constraint variables.
+\[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+{\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \] 
+}
 % \Omit {

### 2016-10-15 domains.tex
some more on transformers

--- 2016-10-15 domains.tex
@@
 the strategy iteration approach available in 2LS~\cite{BJKS15}.
+%
+The concretisation $\gamma(\vec{d})$ is the set $\{\vec{x}\mid
+\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$.
 %
@@
 %-------------------------------------------------------------------------------
-In a typical abstract interpretation based Galois-connection setting
-over an over-approximate domain, every concrete element has a unique
-over-approximate representation in the abstract.  Likewise, every
-concrete transformer is over-approximated by a unique abstract
-transformer.  We now define an abstract deduction transformer.
+%In a typical abstract interpretation based Galois-connection setting
+%over an over-approximate domain, every concrete element has a unique
+%over-approximate representation in the abstract.  Likewise, every
+%concrete transformer is over-approximated by a unique abstract
+%transformer.  We now define an abstract deduction transformer.
 
-$\llbracket s \rrbracket^\sharp(a)=\min \{a'\mid a'\sqsubset a, a'\models s\}$
+An abstract transformer transforms an abstract value $a$ through a
+statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
+The best abstract transformer is
+\[\llbracket s
+\rrbracket^\sharp(a)=\min \{a'\mid a'\models s\}\] 
+where we write
+$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
+\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
+all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
+of $a$. Any abstract transformer that computes a non-minimal element
+is still a sound over-approximating transformer and can be used in our
+algorithm.
 
-\pscmt{define $a'\models s$}
+To make our algorithm efficient, we have to focus abstract
+transformers on performing only the minimally necessary
+work. 
+%
+We thus define a specialised variant of the transformer to compute
+deductions w.r.t.\ a given subdomain $L\sqsubseteq A$, $\bot,\top \in L$.
+\[\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\in L, a'\models s\}\]
+We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
 
-We define a specialised variant of the transformer to compute deductions w.r.t.\ a given subdomain $L\subseteq A$.
-
-$\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\sqsubset a, a'\in L, a'\models s\}$
-
-We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq a\meet\llbracket s \rrbracket^\sharp(a,L)$.
-
-For example,
-$a=(0\leq y \leq 1 \wedge 5\leq z)$
+For example, for
+$a=(0\leq y \leq 1 \wedge 5\leq z)$ we have
 $\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
@@
 $reason(s,a,L,d)=\min \{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}$.
-
-\pscmt{TODO from here}
 

### 2016-10-15 deductions.tex
acdcl algorithm
moved abstract transformers into sec 5

--- 2016-10-09 deductions.tex
@@
 \section{Abstract Model Search}
-In a typical abstract interpretation based galois-connection settings over 
-an over-approximate domain, every concrete elements have a unique over-approximate 
-representation in the abstract.  Likewise, every concrete transformer is
-over-approximated by a unique abstract transformer.  We now define a abstract 
-deduction transformer.
-
-\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
-transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
-domain $A$ is a sound approximation of a concrete model transformer
-$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
-$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
-$m \in A$ is a meet irreducible.   
-\end{definition}
-
-Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
-an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
-\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
-\rangle$.  An abstract deduction transformer is typically computed in the form 
-of strongest post-condition or a weakest pre-condition of a formula in the 
-abstract domain.
-
-A meet decomposition of the outcome of abstract deduction transformer 
-is obtained by taking a meet of the element $\langle y:[3, 5] \rangle, 
-\langle x:[2, 4] \rangle \in ItvDom$ which gives set of meet irreducibles, 
-$\{ \langle y \succeq 3 \rangle, \langle y \preceq 5 \rangle, 
-\langle x \succeq 2 \rangle, \langle x \preceq 4 \rangle \}$, that are 
-precisely complementable.
-
-\begin{definition}{(Meet Decomposition)} A meet decomposition of an abstract
-element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$\forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
-\end{definition}
-
-For a program with $N$ variables, let $L$ be the total number of 
-meet irreducibles returned by a domain $D$.  For $D$ = {\em ItvDom}, the 
-maximum value of $L$ is $2*N$, whereas the maximum value of $L$ is 
-$2*(N^2)$ for $D$ = {\em OctDom}. Note that an octagon is the conjunction 
-of all octagonal inequalities in the set $L$.
 %
@@
 %    
-\begin{algorithm2e}[htbp]
+\begin{algorithm2e}[t]
 \DontPrintSemicolon
@@
 \SetKwData{sat}{SAT}
+\SetKwData{conflict}{CONFLICT}
 \SetKwData{unsat}{UNSAT}
@@
 \Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
-abstract value $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
-\Output{The status (\sat or \unsat or \unknown)}
+abstract valuation $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
+\Output{The status (\sat or \conflict or \unknown)}
 worklist $\leftarrow$ initialize\_worklist($S$) \;
@@
     worklist.delete() \;
-    \return \sat \;
+    \return \conflict \;
   }
@@
 \uIf{$v$ is $\gamma$-complete} {
-  \return \unsat;
+  \return \sat;
 }
-$d \leftarrow decide(v)$ \;
-\uIf{$v \sqsubseteq d$} {
+\uElse {
  \return \unknown \;
 }
-$\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
-$v \leftarrow v \meet d$ \; 
-$\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
-worklist.update(v, d) \; 
 }
 \end{small}
-\caption{Abstract Model Search\label{Alg:ms}}
+\caption{Abstract Model Search $\langle deduce(S,v,\mathcal{T},\mathcal{R}) \rangle$ \label{Alg:ms}}
 \end{algorithm2e}
 %  
-\subsection{Algorithm for Abstract Model Search}
-Algorithm~\ref{Alg:ms} presents the abstract model search procedure. 
-In propositional case, a model search involves two steps -- 
-{\em deductions} using unit rule refines current partial assignments 
-and {\em decisions} to heuristically guess a value for an unassigned 
+A model search procedure in a SAT solver involves two steps -- {\em deductions} 
+using unit rule refines current partial assignments and 
+{\em decisions} to heuristically guess a value for an unassigned 
 literal.  A unit rule overapproximates a model transformer and deduction 
 computes a greatest fixed point over the partial assignments
-domain~\cite{dhk2013-popl}.  Similarly, an abstract model search computes
-a greatest fixed point over meet irreducible deduction transformer in $S$.
+domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
+that computes a greatest fixed point over meet irreducible deduction 
+transformer in $S$.  
+%
 
+%-------------------------------------------------------------------------------
+\subsection{Abstract Transformers} \label{sec:abst}
+%-------------------------------------------------------------------------------
+
+To make our algorithm efficient, we have to focus abstract
+transformers on performing only the minimally necessary
+work. 
+%
+We thus define a specialised variant of the transformer to compute
+deductions w.r.t.\ a given subdomain $L\sqsubseteq A$; $\bot,\top \in L$.
+\[\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\in L, a'\models s\}\]
+We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
+%
+For example, for
+$a=(0\leq y \leq 1 \wedge 5\leq z)$ we have
+$\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
+$\llbracket x=y+z\rrbracket^\sharp(a,\{y,z\})=\top$.
+%
+If $L=A$ then intuitively $a$ is intersected with the constraint $s$.
+Yet, with the help of $L$ we can also control whether to perform
+forward (right-hand side to left-hand side, $L=\{x\}$ in the above
+example) or backward (lhs to rhs, $L=\{y,z\}$) propagation.
+
+Moreover, we want to know what the reasons for a specific deduction are.
+\[ded(s,a,L)=\{R\rightarrow d\mid R=reasons(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}\]
+where 
+$reasons(s,a,L,d)=\min \{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}$.
+
+\pscmt{give an example for $ded$}
+
+\pscmt{define $ded$ for all statements}
+
+%-------------------------------------------------------------------------------
+\subsection{Algorithm for Deduction Phase}
+%-------------------------------------------------------------------------------
+%
+\rmcmt{Talk about lazy closure operation}
+Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.
 A {\em worklist} is initialized following a heuristics choice for iteration 
@@
 $\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with UNSAT.  However,
+is inferred by the domain, the algorithm terminates with CONFLICT.  However,
 when a fixed-point is reached, the abstract valuation $v$ is checked for set of
@@
 abstract valuation $v$ is abstractly satisfying, the abstract model 
-search terminates with SAT.  Otherwise, a decision is made by calling a 
-procedure $decide(v)$ which returns a meet irreducible $d$.  If $d$ refines the 
-current abstract value $v$, then the worklist is updated with all transformers that are dependant on the current decision $d$ along with the update of the corresponding propagation and reason trail,
-$\mathcal{T}$ and $\mathcal{R}$. Else, the model search procedure 
-terminates with UNKNOWN.
+search terminates with SAT.  Otherwise, the algorithm returns UNKNOWN and the 
+model search procedure makes a new decision.

### 2016-10-15 domains.tex
moved abstract transformers into sec 5

--- 2016-10-15 domains.tex
@@
 %-------------------------------------------------------------------------------
+\paragraph{Abstract Transformers}
+%
+%In a typical abstract interpretation based Galois-connection setting
+%over an over-approximate domain, every concrete element has a unique
+%over-approximate representation in the abstract.  Likewise, every
+%concrete transformer is over-approximated by a unique abstract
+%transformer.  We now define an abstract deduction transformer.
+%
+An abstract transformer transforms an abstract value $a$ through a
+statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
+The best abstract transformer is
+\[\llbracket s
+\rrbracket^\sharp(a)=\min \{a'\mid a'\models s\}\] 
+where we write
+$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
+\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
+all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
+of $a$. Any abstract transformer that computes a non-minimal element
+is still a sound over-approximating transformer and can be used in our
+algorithm.
+
+In a forward analysis, we can deduce, for example, $\llbracket
+x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6)$ for $a=(0\leq y \leq 2 \wedge 1 \leq y-z \leq
+1)$.
+%
+We will give more details in Section~\ref{sec:abst}.
+
+%-------------------------------------------------------------------------------
 \subsection{Properties of Abstract Domains for Learning}
+%-------------------------------------------------------------------------------
 %
@@
 
-%-------------------------------------------------------------------------------
-\subsection{Abstract Transformers}
-%-------------------------------------------------------------------------------
-%In a typical abstract interpretation based Galois-connection setting
-%over an over-approximate domain, every concrete element has a unique
-%over-approximate representation in the abstract.  Likewise, every
-%concrete transformer is over-approximated by a unique abstract
-%transformer.  We now define an abstract deduction transformer.
+% The abstract transformer, $apost_{stmt}$, captures the effect of different program 
+% statements in the abstract domain. The transformer is precise for octagonal 
+% assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
+% as shown below.
+% \[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
 
-An abstract transformer transforms an abstract value $a$ through a
-statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
-The best abstract transformer is
-\[\llbracket s
-\rrbracket^\sharp(a)=\min \{a'\mid a'\models s\}\] 
-where we write
-$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
-\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
-all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
-of $a$. Any abstract transformer that computes a non-minimal element
-is still a sound over-approximating transformer and can be used in our
-algorithm.
+% \begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
+% transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+% domain $A$ is a sound approximation of a concrete model transformer
+% $ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
+% $\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+% $m \in A$ is a meet irreducible.   
+% \end{definition}
 
-To make our algorithm efficient, we have to focus abstract
-transformers on performing only the minimally necessary
-work. 
-%
-We thus define a specialised variant of the transformer to compute
-deductions w.r.t.\ a given subdomain $L\sqsubseteq A$, $\bot,\top \in L$.
-\[\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\in L, a'\models s\}\]
-We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
-
-For example, for
-$a=(0\leq y \leq 1 \wedge 5\leq z)$ we have
-$\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
-$\llbracket x=y+z\rrbracket^\sharp(a,\{y\})=\top$.
-
-%$a=(x-y \leq 1, y-x \leq 1)$
-%$\llbracket x=y+z\rrbracket^\sharp($
-
-Moreover, we want to know what the reasons for a specific deduction are.
-$ded(s,a,L)=\{R\rightarrow d\mid R=reason(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}$
-where 
-$reason(s,a,L,d)=\min \{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}$.
-
-The abstract transformer, $apost_{stmt}$, captures the effect of different program 
-statements in the abstract domain. The transformer is precise for octagonal 
-assignments $(x:=y+1)$ but imprecise for non-octagonal assignments $(x:=y+z)$, 
-as shown below.
-\[apost_{x:=y+1}(a) = b = \langle x-y \leq 1, y-x \leq 1 \rangle \qquad apost_{x:=y+z}(b) = \langle \top \rangle \]
-
-\begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
-transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
-domain $A$ is a sound approximation of a concrete model transformer
-$ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
-$\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
-$m \in A$ is a meet irreducible.   
-\end{definition}
-
-Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
-an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
-\rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
-\rangle$.  An abstract deduction transformer is typically computed in the form 
-of strongest post-condition or a weakest pre-condition of a formula in the 
-abstract domain.
+% Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+% an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
+% \rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+% \rangle$.  An abstract deduction transformer is typically computed in the form 
+% of strongest post-condition or a weakest pre-condition of a formula in the 
+% abstract domain.
 
@@
 
+\Omit{
+%
+An important property of a clause learning SAT solver 
+is that a partial assignments domain can be decomposed 
+into a set of precisely complementable singleton assignments.  
+This property of the partial assignments domain helps to 
+guide the model search away from the conflicting region 
+of the search space and allows case-based analysis.  
+%
+\begin{definition}{(Precise complement)} A precise complement of an abstract
+element $a \in A$ in abstract lattice $A$ is an element $\bar{a}$ such that 
+$\gamma(\bar{a} = \neg \gamma(a))$.
+\end{definition}
+%
+The precise complementation property of partial assignment lattice 
+can be generalized to other lattice structures.  Most numerical 
+abstract domains, such as intervals, octagons, polyhedra, can be 
+decomposed into half-spaces, that admits precise complements.  
+For example, intervals lack complements, but the meet decomposition 
+of intervals into meet-irreducibles have precise complements.  
+Numerical abstract domains that admit complementable decomposition 
+are shown in Figure~\ref{complement}.  Our template-based domain 
+implementation precisely complements a meet irreducible in an abstraction.
+% 
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l}
+\hline
+Domain & Precise Complements \\ \hline 
+Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
+Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
+Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
+\end{tabular}
+}
+\end{center}
+\label{fig:complement}
+%\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
+\end{table}
+}
 % \Omit {

### 2016-10-15 learning.tex
section 5,6 update

@@
 analysis is a generalization of an underapproximation of a set of 
-countermodels~\cite{sas,dhk2013-popl}, where a countermodel is a 
+countermodels~\cite{sas12,dhk2013-popl}, where a countermodel is a 
 set of structures that do not satisfy a formula.  Since relational 
@@
 
+\paragraph {\textbf{Abstract Conflict Graph}}
 The first-UIP algorithm in SAT solver works on a data structure called 
@@
 {\em abstract conflict graph}, which records the dependencies between 
-deductions made during abstract model search phase.  For example, consider a
-formula $\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
-abstract valuation $v:= (x=5)$.  
-Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
+deductions made during abstract model search phase.  Nodes in the graph 
+represent elements of the trail $\mathcal{T}$.  Nodes can be either 
+{\em decision} node or {\em deduction} node. Edges can be extracted from
+the transformers in the reason trail.  Incoming arrows to node $n$ 
+indicate that the predecessors of $n$ are sufficient to deduce $n$.  
+A decision node has no incoming edges.  
+For example, consider a formula 
+$\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
+abstract valuation $v:= (x=5)$.  Figure~\ref{conflict} shows a 
+snapshot of an abstract conflict graph that 
 stores the deductions obtained from abstract model search using octagon domain.  
-An arrow in the graph indicate that the predecessors of each node $n$ are 
-sufficient to deduce $n$.  The last deduction conflict with the constraint 
-$(y \leq 0)$.  Every cut in the graph, marked as {\em cut0, cut1}, provides a 
-reason for conflict that can be used in learning.  An UIP is a special node such 
-that any path from the last decision node to the conflict node must pass 
+The last deduction conflict with the constraint $(y \leq 0)$.
+
+\paragraph {\textbf{Lifting UIP to relational domains}}
+An UIP is a special node in the abstract conflict graph such that 
+any path from the last decision node to the conflict node must pass 
 through it.  Unlike a SAT solver, a UIP in ACDCL may involve meet irreducibles 
@@
 asserting cuts, that is, it yields clauses that generates new deduction after 
-backtracking.  For our example, there exists multiple incomaparable reasons 
-for conflict.  Here, cut0 is the first UIP.  Choosing cut0 yields a learnt 
-clause $L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee 
--y+10<0)$.  A learnt clause $\mathcal{L}$ is obtained by negating the reason 
-for conflict.  
+backtracking.  Every cut in the graph reason for conflict that can be used in learning.  
+A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
+conflict node and iteratively finds new nodes that contradict with the 
+transformer stored in $reasons[\bot]$.
+
+For our example, there exists multiple incomaparable reasons for conflict,
+marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
+Choosing cut0 yields a learnt clause 
+$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$.  
+A learnt clause $\mathcal{L}$ is obtained by negating the reason for conflict.   
+\rmcmt{shall we call learnt clause or learnt transformer ?}
 %
@@
 
+\Omit {
 \subsection{Clause Learning in Abstract Lattice}
-\Omit {
 Conflict graph for Intervals

### 2016-10-15 deductions.tex
refactoring domain.tex figure 2

@@
 $\begin{array}{l@{\,\,}c@{\,\,}l}
-SSA &\iff& ((guard0 == TRUE) \land \\
+SSA &\iff& ((g0 == TRUE) \land \\
     &    & (cond == (x == y)) \land \\
-    &    & (guard1 == (cond \&\& guard0)) \land \\
+    &    & (g1 == (cond \&\& guard0)) \land \\
     &    & (x' == 1u + x) \land \\
-    &    & (x' == 1u + y || !guard1))
+    &    & (x' == 1u + y || !1))
 \end{array}$
@@
 \caption{C Program, corresponding SSA and Octagonal Inequalities}
-\label{swssa}
+\label{ssa}
 \end{figure}
@@
 
-\pscmt{give an example for $ded$}
+\pscmt{give an example for $ded$, link it to the reason trail}
 
-\pscmt{define $ded$ for all statements}
+\pscmt{define $ded$ for all statements, connect this to worklist}
+
+\pscmt{link livevars to $L$ -- the subdomain spanned by livevars}
 

### 2016-10-15 domains.tex
refactoring domain.tex figure 2

@@
 \subsection{Program Representation}  
-
+\Omit {
 \begin{wrapfigure}{r}{3cm}
@@
 \end{wrapfigure}
-
+}
+\begin{figure}[t]
+\scriptsize
+\begin{tabular}{l|l}
+\hline
+ SSA & Octagon \\
+\hline
+\begin{minipage}{4.20cm}
+$\begin{array}{rcl}
+(g_0 = \true) \wedge (c = 0 \leq v\leq N) \\
+(g_1 = g_0 \wedge c) \wedge (x_0 = v) \wedge \\
+(x_1 = -v) \wedge (x_2 = g_1?x_0:x_1) \wedge \\
+(g_2 = g_1 \vee g_0\wedge \neg c) \wedge \\
+(z  = x_2 * x_2) \wedge (z<0 \wedge g_2)
+\end{array}$
+\end{minipage}
+&
+\begin{minipage}{4.50cm}
+$\begin{array}{l@{\,\,}c@{\,\,}l}
+C &\iff& ((x \geq -5) \land (x \leq 5) \land \\
+  &    & (z \geq 0) \land (z \leq 25) \land \\
+  &    & (x-z \leq 5) \land (x-z \geq -30) \land \\
+  &    & (x+z \geq -5) \land (x+z \leq 30))  
+\end{array}$
+\end{minipage}
+\\
+\hline
+\end{tabular}
+\caption{SSA and Octagonal Inequalities}
+\label{ssa}
+\end{figure}
+%
 We consider loop-free programs (respectively bounded unwindings of
@@
 %
-Figure~\ref{fig:ssa} shows the constraints for the example program in
-Figure~\ref{fig:example}.
+Figure~\ref{ssa} shows the SSA constraints for the program in figure~\ref{fig:example}. 
 %
@@
 is reachable) then the assertion must hold.
-
+%
+%
+%
 %For example, figure~\ref{swssa} presents the program analysis constraints for 
@@
 %-------------------------------------------------------------------------------
-The results that we present on lifting ACDL to relational domains hold
-for any relational domain.
+The work of~\cite{sas12,dhk2013-popl} shows that CDCL based SAT solvers 
+operates on the partial assignment domains. We present a generic instantiation 
+of CDCL architecture that can be extended to any relational and non-relational
+domains.  
 % 
@@
 paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
-$\mathcal{B}$ is the partial Boolean assignments domain
-$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$ (as
-used in SAT solvers~\cite{sas12}), $b$ is the number of Boolean
-variables in the program, and $\mathcal{R}$ is a \emph{relational}
-domain over the numerical (bitvector) variables in their respective
+$\mathcal{B}$ is the Boolean domain
+$\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$, 
+$b$ is the number of Boolean variables in the program, 
+and $\mathcal{R}$ is a \emph{relational} domain over the 
+numerical (bitvector) variables in their respective
 data types (currently, signed and unsigned integers are supported).
@@
 %-------------------------------------------------------------------------------
-\subsection{Properties of Abstract Domains for Learning}
+\subsection{Properties of Abstract Domains}
 %-------------------------------------------------------------------------------
 %
+\Omit {
 An important property of a clause learning SAT solver is that each
-element of the partial Boolean assignments domain is decomposable into
+element of the partial assignments domain is decomposable into
 complementable elements.  This property of the domain is required to
@@
 complements, but they can be represented as intersections of
-complementable half-spaces, each of which has a complement. We
-formalise this in the sequel.
+complementable half-spaces, each of which has a complement. 
+}
 
+We identify specific properties of domain elements necessary for 
+abstract model search and learning in abstractions.  We formalise 
+this in the sequel.
+%
 \begin{definition} 
-A \emph{meet irreducible} $m$ 
-in a complete lattice structure $A$ is an element with the following property.
+A \emph{meet irreducible} $m$ in a complete lattice 
+structure $A$ is an element with the following property.
 \[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
+\rmcmt{mention minimum complementable element}
 \end{definition}
-
-The meet irreducibles in the partial Boolean assignments domain
-$\mathcal{B}$ are $\true$ and $\false$. The meet irreducibles in the
-template polyhedra domain are all elements that concretise to
-half-spaces, i.e.\ they can be represented by a single inequality. For
-the interval domain, these are \ $x \leq d$ or $x \geq d$ for
-constants $d$.
+%
+The meet irreducibles in boolean domain $\mathcal{B}$ are 
+$\{x \mapsto \true\}$ and $\{x \mapsto \false\}$. The meet 
+irreducibles in the template polyhedra domain are all elements 
+that concretise to half-spaces, i.e.\ they can be represented 
+by a single inequality. For the interval domain, these are 
+\ $x \leq d$ or $x \geq d$ for constants $d$.
 
@@
 % $(x,y) \in [2,4]\times[3, 5]$ 
-$2\leq x\leq 4 \wedge 3\leq y\leq 5$ is
+$decomp(2\leq x\leq 4 \wedge 3\leq y\leq 5)$ is
 the set $\{x\geq 2, x\leq 4, y\geq 3, y\leq 5\}$.
@@
 
+An important property of a clause learning SAT solver 
+is that a partial assignments domain can be decomposed 
+into a set of precisely complementable singleton assignments.  
+This property of the partial assignments domain helps to 
+guide the model search away from the conflicting region 
+of the search space and allows case-based analysis.  Complementation 
+operator in abstract domains is different from the notion of precise
+complements. 
+%
+\begin{definition}{(Precise complement)} Let $A$ be an lattice. A precise 
+complement of an abstract element $a \in A$ is an element $\bar{a} \in A$ 
+such that the negation of concretization $(\gamma)$ of $a$ is equivalent to 
+concretization of $\bar{a}$, that is, $\gamma(\bar{a}) = \neg \gamma(a))$.
+\end{definition}
+%
+The precise complementation property of partial assignment lattice 
+can be generalized to other lattice structures.  Most numerical 
+abstract domains, such as intervals, octagons, polyhedra, can be 
+decomposed into half-spaces, that admits precise complements.  
+For example, intervals lack complements, but the meet decomposition 
+of intervals into meet-irreducibles have precise complements.  
+Numerical abstract domains that admit complementable decomposition 
+are shown in Figure~\ref{complement}.  Our template-based domain 
+implementation precisely complements a meet irreducible in an abstraction.
+% 
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l}
+\hline
+Domain & Precise Complements \\ \hline 
+Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
+Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
+Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
+\end{tabular}
+}
+\end{center}
+\label{fig:complement}
+\end{table}
+
+\Omit {
 The meet irreducibles in the partial Boolean assignments domain and
@@
 that returns precise complements by decomposing it into half-spaces.    
+}
 
@@
 %
-An important property of a clause learning SAT solver 
-is that a partial assignments domain can be decomposed 
-into a set of precisely complementable singleton assignments.  
-This property of the partial assignments domain helps to 
-guide the model search away from the conflicting region 
-of the search space and allows case-based analysis.  
-%
-\begin{definition}{(Precise complement)} A precise complement of an abstract
-element $a \in A$ in abstract lattice $A$ is an element $\bar{a}$ such that 
-$\gamma(\bar{a} = \neg \gamma(a))$.
-\end{definition}
-%
-The precise complementation property of partial assignment lattice 
-can be generalized to other lattice structures.  Most numerical 
-abstract domains, such as intervals, octagons, polyhedra, can be 
-decomposed into half-spaces, that admits precise complements.  
-For example, intervals lack complements, but the meet decomposition 
-of intervals into meet-irreducibles have precise complements.  
-Numerical abstract domains that admit complementable decomposition 
-are shown in Figure~\ref{complement}.  Our template-based domain 
-implementation precisely complements a meet irreducible in an abstraction.
-% 
-\begin{table}
-\begin{center}
-{
-\begin{tabular}{l|l}
-\hline
-Domain & Precise Complements \\ \hline 
-Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
-Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
-Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
-\end{tabular}
-}
-\end{center}
-\label{fig:complement}
-%\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
-\end{table}
 }

### 2016-10-15 domains.tex
various fixes

--- 2016-10-15 domains.tex
@@
 \[\llbracket s
-\rrbracket^\sharp(a)=\min \{a'\mid a'\models s\}\] 
-where we write
+\rrbracket^\sharp(a)=\bigsqcup \{a'\mid a'\models s\}\] where we write
 $a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
@@
 all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
-of $a$. Any abstract transformer that computes a non-minimal element
-is still a sound over-approximating transformer and can be used in our
-algorithm.
+of $a$. Any abstract transformer that over-approximates the least
+upper bound is a sound transformer and can be used in our algorithm.
 
 In a forward analysis, we can deduce, for example, $\llbracket
-x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6)$ for $a=(0\leq y \leq 2 \wedge 1 \leq y-z \leq
-1)$.
+x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6)$ for $a=(0\leq y \leq 2
+\wedge 1 \leq y-z \leq 1)$.
 %
@@
 
-Standard abstract interpretation does not require complementation property.  
-Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
-for abstract interpretation.  We implement a template-based polyhedra domain 
-that returns precise complements by decomposing it into half-spaces.
+Standard abstract interpretation does not require a complementation
+operator. Hence, abstract domain libraries, such as APRON~\cite{apron}
+do not provide it.  However, it can be implemented with the help of a
+meet decomposition as explained above.
+%
+Another feature that abstract domain libraries do not provide is to
+track the reasons why certain deductions have been made. We come back
+to this point in Section~\ref{sec:abst}.
 

### 2016-10-15 experiments.tex
refactoring domain.tex figure 2
various fixes

@@
 We have implemented the instantiation of ACDCL on relational domains
-in 2LS verification tool~\cite{2ls}.  The abstract domains used for 
-our experiments are intervals, template polyhedra, octagons and equalities.  
+in 2LS verification tool~\cite{2ls}.  We used the abstract domains
+provided by 2LS (intervals, octagons and equalities) for our
+experiments. Although the performance of these implementations is not
+competitive with dedicated implementations as found in
+APRON~\cite{apron}, for instance, they have the advantage that they
+handle all C operators (including bitwise operations) out of the box.
+
 We verified a total of 70 benchmarks in ANSI-C, which are derived from 

### 2016-10-15 learning.tex
abstract unit rule

@@
 $L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$.  
-A learnt clause $\mathcal{L}$ is obtained by negating the reason for conflict.   
-\rmcmt{shall we call learnt clause or learnt transformer ?}
 %
@@
 %    
-\paragraph {\em Backjumping:}
+\paragraph{\textbf{Clause Learning in Abstract Lattice}}
+A learnt clause $\mathcal{L}$ in propositional solver is obtained 
+by negating the reason for conflict.  An asserting clause guarantees 
+new dductions using unit rule~\cite{cdcl}.  However, we model a 
+learnt clause as abstract deduction transformer, called {\em abstract unit 
+transformer (AUnit)}.  For an abstract lattice $A$ and a set of 
+countermodels $C \subseteq A$ of $\varphi$ such that \meet{C} does not 
+satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is $\top$ if $\exists a \in A$ 
+such that $a \subseteq \meet{C}$.  However, if $t \in C$ and 
+$\forall t' \in C \ {t}$ such that $a \sqsubseteq t'$, $AUnit$ 
+infers a valid deduction $\bar{t}$, which directs the search 
+away from the conflict region. Else, $AUnit$ returns $\top$.
+
+\paragraph {\textbf{Backjumping}}
 A backjumping procedure undo all the assignments in the trail up to 

### 2016-10-15 domains.tex

--- 2016-10-15 domains.tex
@@
 %
-\Omit {
 An important property of a clause learning SAT solver is that each
-element of the partial assignments domain is decomposable into
+element of the boolean domain is decomposable into
 complementable elements.  This property of the domain is required to
@@
 complementable half-spaces, each of which has a complement. 
-}
-
-We identify specific properties of domain elements necessary for 
-abstract model search and learning in abstractions.  We formalise 
-this in the sequel.
+%
+% We identify specific properties of domain elements necessary for 
+% abstract model search and learning in abstractions.  
+We formalise this in the sequel.
 %
@@
 \[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
-\rmcmt{mention minimum complementable element}
+%\rmcmt{mention minimum complementable element}
 \end{definition}
 %
-The meet irreducibles in boolean domain $\mathcal{B}$ are 
-$\{x \mapsto \true\}$ and $\{x \mapsto \false\}$. The meet 
+The meet irreducibles in the boolean domain $\mathcal{B}$ for a variable $x$ are 
+$x$ and $\neg x$. The meet 
 irreducibles in the template polyhedra domain are all elements 
@@
 \begin{definition} 
-A meet irreducible $m\in A$ is called \emph{complementable} iff $\exists
-\bar{m} \in A: m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.  
+An element $a\in A$ is called \emph{precisely complementable}
+iff $\exists \bar{a} \in A: \neg\gamma(\bar{a})=\gamma(a)$, 
+i.e.\ there is an element whose negated concretisation equals
+the concretisation of $a$.
+%m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.
 %A complementable meet irreducible $\bar{M}$ of an abstract lattice $A$ is the complement of a meet 
@@
 
-An important property of a clause learning SAT solver 
-is that a partial assignments domain can be decomposed 
-into a set of precisely complementable singleton assignments.  
-This property of the partial assignments domain helps to 
-guide the model search away from the conflicting region 
-of the search space and allows case-based analysis.  Complementation 
-operator in abstract domains is different from the notion of precise
-complements. 
+% An important property of a clause learning SAT solver 
+% is that a partial assignments domain can be decomposed 
+% into a set of precisely complementable singleton assignments.  
+% This property of the partial assignments domain helps to 
+% guide the model search away from the conflicting region 
+% of the search space and allows case-based analysis.  Complementation 
+% operator in abstract domains is different from the notion of precise
+% complements. 
 %
-\begin{definition}{(Precise complement)} Let $A$ be an lattice. A precise 
-complement of an abstract element $a \in A$ is an element $\bar{a} \in A$ 
-such that the negation of concretization $(\gamma)$ of $a$ is equivalent to 
-concretization of $\bar{a}$, that is, $\gamma(\bar{a}) = \neg \gamma(a))$.
-\end{definition}
+% \begin{definition}{(Precise complement)} Let $A$ be an lattice. A precise 
+% complement of an abstract element $a \in A$ is an element $\bar{a} \in A$ 
+% such that the negation of concretization $(\gamma)$ of $a$ is equivalent to 
+% concretization of $\bar{a}$, that is, $\gamma(\bar{a}) = \neg \gamma(a))$.
+% \end{definition}
 %
-The precise complementation property of partial assignment lattice 
-can be generalized to other lattice structures.  Most numerical 
-abstract domains, such as intervals, octagons, polyhedra, can be 
-decomposed into half-spaces, that admits precise complements.  
-For example, intervals lack complements, but the meet decomposition 
-of intervals into meet-irreducibles have precise complements.  
-Numerical abstract domains that admit complementable decomposition 
-are shown in Figure~\ref{complement}.  Our template-based domain 
-implementation precisely complements a meet irreducible in an abstraction.
+% The precise complementation property of partial assignment lattice 
+% can be generalized to other lattice structures.  Most numerical 
+% abstract domains, such as intervals, octagons, polyhedra, can be 
+% decomposed into half-spaces, that admits precise complements.  
+% Numerical abstract domains that admit complementable decomposition 
+% are shown in Figure~\ref{complement}.  Our template-based domain 
+% implementation precisely complements a meet irreducible in an abstraction.
+% % 
+% \begin{table}
+% \begin{center}
+% {
+% \begin{tabular}{l|l}
+% \hline
+% Domain & Precise Complements \\ \hline 
+% Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
+% Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
+% Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
+% \end{tabular}
+% }
+% \end{center}
+% \label{fig:complement}
+% \end{table}
+% }
+
+Note that for many domains $A$ including template polyhedra
+most domain elements are not precisely complementable within $A$.
 %
-\begin{table}
-\begin{center}
-{
-\begin{tabular}{l|l}
-\hline
-Domain & Precise Complements \\ \hline 
-Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
-Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
-Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
-\end{tabular}
-}
-\end{center}
-\label{fig:complement}
-\end{table}
-
-\Omit {
-The meet irreducibles in the partial Boolean assignments domain and
-template polyhedra domains are all complementable.
+In fact, for template polyhedra all non-meet-irreducible elements $e$
+(except $\bot$ and $\top$) are not precisely complementable,
+whereas all meet irreducibles are precisely complementable.
 %
-Note that for template polyhedra, all non-meet-irreducible elements $e$
-(except $\bot$ and $\top$) are not complementable within $A$.
-%
-However, we can complement each element in the meet decomposition of $e$ and
+Hence, we can complement each element in the meet decomposition of $e$ and
 re-interpret the obtained set as a disjunction. 
@@
 can be written as a disjunction of meet irreducibles:
-\[(x\leq -3) \lor (x\geq 2) \lor (y\leq -2) \lor (y\geq 3) \lor (x+y\geq 3) \lor (x-y\geq 2) \lor (y-x\geq 4) \lor (-x-y\geq 3)\]
+\[(x{\leq} -3) \lor (x{\geq} 2) \lor (y{\leq} -2) \lor (y{\geq} 3) \lor (x+y{\geq} 3) \lor (x-y{\geq} 2) \lor (y-x{\geq} 4) \lor (-x-y{\geq} 3)\]
 
-Standard abstract interpretation does not require complementation property.  
-Abstract Domain library, such as APRON C library~\cite{apron} is specialized 
-for abstract interpretation.  We implement a template-based polyhedra domain 
-that returns precise complements by decomposing it into half-spaces.    
-}
+Standard abstract interpretation does not require a complementation
+operator. Hence, abstract domain libraries, such as APRON~\cite{apron}
+do not provide it.  However, it can be implemented with the help of a
+meet decomposition as explained above.
+%
+Another feature that abstract domain libraries do not provide is to
+track the reasons why certain deductions have been made. We come back
+to this point in Section~\ref{sec:abst}.
 

### 2016-10-15 graphs.tex

@@
-\section{Experiments}
-\label{section:experiments}
-\Omit {
-We performed a number of experiments to demonstrate the utility and
-applicability of {\algorithmName}.  All experiments were performed
-on an Intel Xeon X5667 at 3\,GHz running Fedora 20 with 64-bit binaries.
-Each individual run was limited to 13\,GB 
-of memory and 900 seconds of
-CPU time, % (with one exception noted below), 
-enforced by the operating
-system kernel.  We took the \emph{loops} meta-category (143 benchmarks) from the SV-COMP'15 benchmark
-set.%
-\footnote{\url{http://sv-comp.sosy-lab.org/2015/benchmarks.php}}
-%
-
-\subsection{\algorithmName\ Verifies More Programs Than the Algorithms it Simulates}
-
-Table~\ref{tab:results} gives a comparison between \systemName\ running
-\algorithmName\ (column 6) and \emph{the same system} running as an incremental
-bounded model checker (IBMC) (column 2), incremental $k$-induction (i.e. without invariant inference, column 3)
-and as an abstract interpreter (AI) (column 4).
-%
- \algorithmName\ is more complete than each of the restricted modes. 
-This is not self-evident since it could be much less efficient and,
-thus, fail to solve the problems within the given time or memory limits.
-%
-$k$-induction can solve 60.8\% of the benchmarks, 13 more than IBMC. 
-%
-32\% of the benchmarks can be solved by abstract interpretation (bugs
-are only exposed if they are reachable with 0 loop unwindings).  
-%
-\algorithmName\ solves 62.9\% of the benchmarks, 
-proving 3 more properties than $k$-induction.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%% results table %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{table}
-\centering
-\begin{tabular}{|l|ccc|c|c|c|c|}
-\hline
-& IBMC & $k$-induction & AI & portfolio & algorithmName &
-                                                               CPAchecker & ESBMC \\
-\hline
-counterexamples & \bf 38 & \bf 38 &  17 & \bf 38 & \bf 38 & 36 &  35 \\
-proofs          &    36 &  49 &  30 &  51 &  52 & 59 & \bf 91 \\
-false proofs    &     0 &   0 &   0 &   0 &   0 &  2 &  12\\
-false alarms    &     2 &   2 &   0 &   2 &   2 &  2 &   0\\
-inconclusive    &     0 &   0 &  93 &   0 &   0 &  4 &   2\\
-timeout         &    65 &  53 &   3 &  50 &  51 & 38 &   2 \\
-memory out      &     2 &   1 &   0 &   2 &   0 &  2 &   1 \\
-total runtime   & 17.1h & 13.8h & 0.89h & 13.3h & 13.2h & 10.9h & 0.54h \\
-\hline
-\end{tabular}~\\[1ex]
-\caption{Comparison between algorithmName, the algorithms it subsumes,
-the portfolio, and CPAchecker. The rows false alarms and false proofs 
-indicate soundness bugs of the tool implementations.
-\label{tab:results}}
-\end{table}
-
-\subsection{\algorithmName\ is at Least as Good as Their Na\"ive Portfolio}
-
-To show
-that \algorithmName\ is more than a mixture of three techniques and
-that they strengthen each other, consider column 5 of
-Table~\ref{tab:results}.  This gives the results of an ideal portfolio
-in which the three restricted techniques are run in parallel on and
-the portfolio terminates when the first returns a conclusive result.
-Thus the CPU time taken is three times the time taken by the fastest
-technique for each benchmark (in practice these could be run in
-parallel, giving a lower \emph{wall clock} time).
-%
-In our setup, \algorithmName\ had a disadvantage as each component of
-virtual portfolio had the same %time and 
-memory restriction as
-{\algorithmName}, thus effectively giving the portfolio three times as much memory.
-
-Still, {\algorithmName} is slightly faster and more
-accurate than the portfolio as can be seen in Table \ref{tab:results}.
-%
-The scatter plot in Figure~\ref{fig:results}a shows the results for
-each benchmark:
-%
-one can observe that {\algorithmName} is up to one order of magnitude
-slower on many unsafe benchmarks, which is obviously due to the
-additional work of invariant inference that
-{\algorithmName} has to perform in contrast to IBMC.
-%
-However, note that {\algorithmName} is faster than the 
-portfolio on some safe and even one unsafe benchmarks.
-This suggests that {\algorithmName} is more than the sum of its parts.
-
-\subsection{\algorithmName\ is Comparable with State-of-the-Art Approaches}
-
-We compared our implementation of \algorithmName\ with
-CPAchecker%
-\footnote{SVCOMP'15 version, http://cpachecker.sosy-lab.org/}%
-%
-and ESBMC%
-\footnote{SVCOMP'15 version, http://www.esbmc.org/}%
-, which uses $k$-induction.
-%
-The results are shown in the last three columns in Table \ref{tab:results}
-and in the scatter plot in Figure~\ref{fig:results}b. Additional
-results are given in \pponly{the extended version
-  \cite{extended-version}}\rronly{Appendix \ref{sec:further}}.
-%
-In comparison to CPAchecker, the winner of SVCOMP'15,
-our prototype of {\algorithmName} is overall a bit slower and proves 
-fewer properties (due to more timeouts), but as
-Figure~\ref{fig:results}b shows, it significantly outperforms
-CPAchecker on most benchmarks.
-%
-ESBMC exposes fewer bugs, but proves many more properties and is 
-significantly faster. However, it has 6 times
-more soundness bugs than our implementation.%
-\footnote{The two false alarms in our current implementation are due to 
-limited support for dynamic memory allocation.}
-%
-These results show that our prototype implementation of
-\algorithmName\ can keep up with state-of-the-art verification tools.
-
-}
-%%%%%%%%%%%%%%%%%%%%%%%%%% kiki vs vbs vs CPAchecker %%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%% scatter plots %%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[t]
@@
 	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
-	table [meta=label] {scatter-kiki-vbs-sas.dat};
+	table [meta=label] {plotdata/scatter-plot1.dat};
 	\legend{Safe,Unsafe}
@@
 \caption{\label{fig:results}
-Runtime Comparison
+Comparison
 }

### 2016-10-15 deductions.tex
various fixes
closure sketch

@@
 We thus define a specialised variant of the transformer to compute
-deductions w.r.t.\ a given subdomain $L\sqsubseteq A$; $\bot,\top \in L$.
-\[\llbracket s \rrbracket^\sharp(a,L)=\min \{a'\mid a'\in L, a'\models s\}\]
+deductions w.r.t.\ a given subdomain $L\subseteq A$; $\bot,\top \in L$.
+\[\llbracket s \rrbracket^\sharp(a,L)=\bigsqcup \{a'\mid a'\in L, a'\models s\}\]
 We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
@@
 where 
-$reasons(s,a,L,d)=\min \{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}$.
+$reasons(s,a,L,d)=\text{argmin}_{R\in\{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}} |R| $.
 
@@
 \rmcmt{Talk about lazy closure operation}
+
+\pscmt{
+Rough and sketchy:
+%
+$espan(V)=$all meet irreducibles $\in A$ that contain at most one variable that is not in $V$.
+%
+$select(S,PV)$ is the set of statements that contain variables in $PV$ where $PV$ is the set of variables in the previous deduction.
+%
+$LV(s)$ are the live variables of a statement $s$, i.e.\ the intersection of $PV$ with the variables in $s$.
+\begin{theorem}
+$$
+\gfp\lambda a. \bigsqcap_{s\in S} ded(s,a,A) = 
+\gfp\lambda a. \bigsqcap_{s\in select(S,PV)} ded(s,a,espan(LV(s)))
+$$
+\end{theorem}
+Proof sketch:
+When computing $ded$, we obtain a deduction that affects variables $V$.
+This has two consequences: (1) we have to compute the closure; (2) we have to propagate through all statements $s$ affected by the changes to $V$. 
+We do only (2): $espan$ augments the set of affected variables by one transitivity step that we would have to compute in the closure.
+$ded$ on this subdomain hence performs the deduction w.r.t.
+$a$ and $s$ plus one transitivity step of the closure.
+Thus, by induction, we eventually compute the entire closure just by doing (2).
+
+What are the consequences regarding complexity?
+Assuming we require $m$ iterations to reach the fixed point with eager closure; we have to perform $|A|^c|S|m$ (idealised) computation steps to do that (where $c=3$ for octagons).
+With the lazy closure we have to do $|L|^c|S|mk$ computations taking into account that we need $k$ times more iterations to converge.
+Hence, the approach pays off if $k<<\frac{|A|^c}{|L|^c}$, which seems likely for $c=3$.
+
+}
+
 Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.

### 2016-10-15 domains.tex
learning and domain update

@@
 An important property of a clause learning SAT solver is that each
-element of the boolean domain is decomposable into
+non-singleton element of the partial assignment domain is decomposable into
 complementable elements.  This property of the domain is required to
@@
 conflicting region of the search space.
+\rmcmt{Convention to be followed: In SAT world, partial assignment (PA) is a data structure. Previous ACDL paper 
+showed that PA is an abstract domain and called it partial assignment domain. Likewise, in this paper, 
+whenever we refer to SAT, we also call partial assignment domain. But When we refer 
+to our ACDCL product domain, we call boolean domain and polyhedra domain. We follow this 
+convention throughout. This would make a SAT reader happy as well as a Abstract Interpreter not sad !!}
 %
@@
 the concretisation of $a$.
+\rmcmt{$\gamma$ is not defined before and used here for the first time}
 %m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.

### 2016-10-15 learning.tex
learning and domain update

@@
 by negating the reason for conflict.  An asserting clause guarantees 
-new dductions using unit rule~\cite{cdcl}.  However, we model a 
-learnt clause as abstract deduction transformer, called {\em abstract unit 
-transformer (AUnit)}.  For an abstract lattice $A$ and a set of 
-countermodels $C \subseteq A$ of $\varphi$ such that \meet{C} does not 
+new deductions using unit rule~\cite{cdcl}.  However, we model a 
+learnt clause as meet irreducible deduction transformer, 
+called {\em abstract unit transformer (AUnit)}.  For an 
+abstract lattice $A$ with complementable meet irreducibles 
+and a set of countermodels $C \subseteq A$ of $\varphi$ such that \meet{C} does not 
 satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is $\top$ if $\exists a \in A$ 
 such that $a \subseteq \meet{C}$.  However, if $t \in C$ and 
-$\forall t' \in C \ {t}$ such that $a \sqsubseteq t'$, $AUnit$ 
+$\forall t' \in C \\ \{t\}$ such that $a \sqsubseteq t'$, $AUnit$ 
 infers a valid deduction $\bar{t}$, which directs the search 

### 2016-10-17 learning.tex
learning in relational lattice

--- 2016-10-15 learning.tex
@@
 Choosing cut0 yields a learnt clause 
-$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$.  
-A learnt clause $\mathcal{L}$ is obtained by negating the reason for conflict.   
-\rmcmt{shall we call learnt clause or learnt transformer ?}
+$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
+which is obtained by negating the reason for conflict.  
 %
@@
 %    
-\paragraph {\em Backjumping:}
+\paragraph{\textbf{Learning in Relational Lattice}}
+We model a learning as a meet irreducible deduction transformer, called 
+{\em abstract unit transformer (AUnit)}.  For an 
+abstract lattice $A$ with complementable meet irreducibles 
+and a set of conflicting meet irreducibles (countermodels) 
+$C \subseteq A$ of $\varphi$ such that \meet{C} does not 
+satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
+\[ AUnit(a) =
+ \begin{cases}
+  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
+  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
+  \setminus \{t\}. a  \sqsubseteq t' \\
+  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
+  \not\sqsubseteq t', \; \text{then} \\ 
+             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
+             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
+  \top & \quad \text{otherwise} \\
+ \end{cases}
+\]
+For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
+of meet irreducibles of $\varphi$.  Let $a= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15, k:[1,1] \rangle$, 
+then $AUnit_C(a) = (x+z < 10)$.  Applying the third rule, we get a decomposition 
+of $a$ into $a'$ and $a''$ where 
+$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
+$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  $AUnit_C(a)$ returns a new 
+deduction $x+z \leq 9$, which constrains the bound of the octagonal constraint.  
+Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
+and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
+the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
+then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
+and $a' \sqsubseteq (x \leq 5)$.
+
+\paragraph {\textbf{Backjumping}}
 A backjumping procedure undo all the assignments in the trail up to 

### 2016-10-17 domains.tex
minor changes

@@
 \section{Preliminaries}
-This section explains the program model and the abstract domains that
-we are going to use.
+This section introduces the program model and identifies properties of abstract
+domains that are necesary for learning in abstract lattice structures.
 % \subsection{Programs:}
@@
 \begin{figure}[t]
+\center
 \scriptsize
@@
 \hline
-\begin{minipage}{4.20cm}
+\begin{minipage}{4.50cm}
 $\begin{array}{rcl}
-(g_0 = \true) \wedge (c = 0 \leq v\leq N) \\
+\mathcal{P}:= (g_0 = \true) \wedge (c = 0 \leq v\leq N) \\
 (g_1 = g_0 \wedge c) \wedge (x_0 = v) \wedge \\
@@
 %
-We consider loop-free programs (respectively bounded unwindings of
-programs with loops or recursion) with safety properties given as
-assertions in the code.
+We consider programs with bounded loops 
+%(respectively bounded unwindings of programs with loops or recursion) with 
+containing safety properties given as assertions in the code.
 %
@@
 An important property of a clause learning SAT solver is that each
-non-singleton element of the partial assignment domain is decomposable into
-complementable elements.  This property of the domain is required to
-learn elements that help to guide the model search away from the
-conflicting region of the search space.
+non-singleton element of the partial assignment domain can be 
+decomposed into a set of precisely complementable singleton elements.  
+This property of the partial assignment domain is required to learn 
+elements that help to guide the model search away from the
+conflicting region of the search space.  Complementation operator 
+in abstract domains is different from the notion of precise complements. 
 \rmcmt{Convention to be followed: In SAT world, partial assignment (PA) is a data structure. Previous ACDL paper 
@@
 %
-% The precise complementation property of partial assignment lattice 
-% can be generalized to other lattice structures.  Most numerical 
-% abstract domains, such as intervals, octagons, polyhedra, can be 
-% decomposed into half-spaces, that admits precise complements.  
-% Numerical abstract domains that admit complementable decomposition 
-% are shown in Figure~\ref{complement}.  Our template-based domain 
-% implementation precisely complements a meet irreducible in an abstraction.
-% % 
-% \begin{table}
-% \begin{center}
-% {
-% \begin{tabular}{l|l}
-% \hline
-% Domain & Precise Complements \\ \hline 
-% Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
-% Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
-% Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
-% \end{tabular}
-% }
-% \end{center}
-% \label{fig:complement}
-% \end{table}
-% }
-
+The precise complementation property of partial assignment lattice 
+can be generalized to other lattice structures.  Most numerical 
+abstract domains, such as intervals, octagons, polyhedra, can be 
+decomposed into half-spaces, that admits precise complements.  
+Numerical abstract domains that admit complementable decomposition 
+are shown in Figure~\ref{fig:complement}.  Our template-based domain 
+implementation precisely complements a meet irreducible in an abstraction.
+% 
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l}
+\hline
+Domain & Precise Complements \\ \hline 
+Interval & \((x \leq n) \quad \longrightarrow (x > n)\) \\ \hline
+Octagon & \((x+y < 1) \quad \longrightarrow (-x-y < 0)\) \\ \hline
+Equality & \((x==y) \quad \longrightarrow (x \neq y)\) \\ \hline 
+\end{tabular}
+}
+\end{center}
+\label{fig:complement}
+\end{table}
+%
+\Omit {
 Note that for many domains $A$ including template polyhedra
@@
 %
-Thus the complement of the octagon in Figure~\ref{octagon} 
-can be written as a disjunction of meet irreducibles:
-\[(x{\leq} -3) \lor (x{\geq} 2) \lor (y{\leq} -2) \lor (y{\geq} 3) \lor (x+y{\geq} 3) \lor (x-y{\geq} 2) \lor (y-x{\geq} 4) \lor (-x-y{\geq} 3)\]
-
+%Thus the complement of the octagon in Figure~\ref{octagon} 
+%can be written as a disjunction of meet irreducibles:
+%\[(x{\leq} -3) \lor (x{\geq} 2) \lor (y{\leq} -2) \lor (y{\geq} 3) \lor (x+y{\geq} 3) \lor (x-y{\geq} 2) \lor (y-x{\geq} 4) \lor (-x-y{\geq} 3)\]
+}
 Standard abstract interpretation does not require a complementation

### 2016-10-17 contributions.tex
contribution

@@
 lifts the CDCL algorithm to relational numerical abstract domains.   
-We have implemented our technique in 2LS verification tool.  To 
-the best of our knowledge, this is the first instantiation of CDCL 
+We have implemented our technique in 2LS verification tool.  
+\Omit{
+To the best of our knowledge, this is the first instantiation of CDCL 
 algorithm with relational abstract domains. 
+}
 
@@
 
-\item  Experiments on a set of benchmarks drawn from the bitvector 
-category in SV-COMP as well as numerical programs and a few 
-control-intensive benchmarks from hardware domain show that 
-the strict generalization of CDCL algorithm effectively combines 
-the efficiency of abstract domains while being path-sensitive and 
-bit-precise at the same time.  
+\item  Experiments indicate a \rmcmt{N}X reduction in the number of 
+decisions, propagations, learning and backtracking iterations compared to a 
+SAT solver while being more precise compared to an abstract interpreter.  
+\Omit {
+set of benchmarks drawn from the bitvector regression category in 
+SV-COMP, bit-precise software models auto-generated from hardware 
+circuits domain and some hand crafted example show that 
+}
 \end{compactenum}

### 2016-10-17 experiments.tex
contribution

--- 2016-10-15 experiments.tex
@@
 We have implemented the instantiation of ACDCL on relational domains
-in 2LS verification tool~\cite{2ls}.  The abstract domains used for 
-our experiments are intervals, template polyhedra, octagons and equalities.  
+in 2LS verification tool~\cite{2ls}.  We have used the abstract domains
+provided by 2LS (intervals, octagons, polyhedra and equalities) for our
+experiments. Although the performance of the abstract domain implementations 
+in 2LS is not competitive with that of APRON~\cite{apron} library, but 
+the domain implementation in 2LS can handle all C operators 
+(including bitwise operations) out of the box and support precise
+complementation of meet irreducibles, which are necessary for our algorithm.
+
 We verified a total of 70 benchmarks in ANSI-C, which are derived from 
@@
 bounded loops are completely unrolled before analysis.  We compare 
-our tool with a state-of-the-art bounded model checker CBMC and a 
-commercial static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT 
-solver in the backend.  Astr{\'e}e was configured with interval 
-domain, trace-partition domain, and relational domains.  All the 
-bounded loops are completely unrolled in Astr{\'e}e.
+our tool with a state-of-the-art SAT-based bounded model checker 
+CBMC~\cite{cbmc} and a commercial static analysis tool, Astr{\'e}e.  
+CBMC uses MiniSAT solver in the backend.  Astr{\'e}e was configured with 
+default set of domains which includes interval, bit-field, congurence and 
+relational domains.  Additionally, to enable precise analysis using 
+Astr{\'e}e, we also used trace-partition domain.
 
@@
 }
-\paragraph {\em Observation of Analysis}
+\paragraph {\em \textbf{Observation of Analysis}}
 
@@
 \paragraph {\em \textbf{Clause Learning}}
+
+%================  
+\input{graphs}
+%================

### 2016-10-17 learning.tex
learning

@@
 \paragraph{\textbf{Learning in Relational Lattice}}
-We model a learning as a meet irreducible deduction transformer, called 
+We model learning as a meet irreducible deduction transformer, called 
 {\em abstract unit transformer (AUnit)}.  For an 
@@
 For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
-of meet irreducibles of $\varphi$.  Let $a= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15, k:[1,1] \rangle$, 
-then $AUnit_C(a) = (x+z < 10)$.  Applying the third rule, we get a decomposition 
-of $a$ into $a'$ and $a''$ where 
+of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
+x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
+decomposition of $a$ into $a'$ and $a''$ where 
 $a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
-$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  $AUnit_C(a)$ returns a new 
-deduction $x+z \leq 9$, which constrains the bound of the octagonal constraint.  
+$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
+which constrains the bound of the octagonal constraint.  
 Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$

### 2016-10-18 graphs.tex
decision plot

@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=kIkI (time in seconds),
-			ylabel=portfolio (time in seconds), 
+	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel= ACDCL (decisions),
+			ylabel=SAT (decisions), 
 			legend style={at={(0.8,0.15)},
@@
 	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
-	table [meta=label] {plotdata/scatter-plot1.dat};
+	table [meta=label] {plotdata/scatter-decision.dat};
 	\legend{Safe,Unsafe}
 \addplot [domain=.1:1500] {x};
-\addplot [red,sharp plot, domain=.1:1500] {900}
-          node [below] at (axis cs:10,850) {timeout};
-\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
-          node [left,rotate=90] at (axis cs:700,10) {timeout}
- node [right,black] at (axis cs:10,3) {portfolio faster}
- node [right,black] at (axis cs:1,55) {kIkI faster};
+%\addplot [red,sharp plot, domain=.1:1500] {900}
+%          node [below] at (axis cs:10,850) {timeout};
+%\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
+%          node [left,rotate=90] at (axis cs:700,10) {timeout}
+ %node [right,black] at (axis cs:10,3) {portfolio faster}
+ %node [right,black] at (axis cs:1,55) {kIkI faster};
 %\addplot [red,sharp plot, update limits=false] coordinates{(900,.1) (900,1500)}

### 2016-10-18 graphs.tex
decision plots
conflict statistics

@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel= ACDCL (decisions),
-			ylabel=SAT (decisions), 
+	\begin{loglogaxis} [xmin=0.1,xmax=4000, ymin=0.1, ymax=4000, xlabel= SAT (decision),
+			ylabel=ACDCL (decision), 
 			legend style={at={(0.8,0.15)},
@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=kIkI (time in seconds),
-			ylabel=CPAchecker (time in seconds),
+	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=SAT (Propagation),
+			ylabel=ACDCL (Propagation),
 			legend style={at={(0.8,0.15)},
@@
 	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
-	table [meta=label] {plotdata/scatter-plot1.dat};
+	table [meta=label] {plotdata/scatter-propagation.dat};
 	\legend{Safe,Unsafe}
 \addplot [domain=.1:1500] {x};
-\addplot [red,sharp plot, domain=.1:1500] {900}
-          node [below] at (axis cs:10,850) {timeout};
-\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
-          node [left,rotate=90] at (axis cs:700,150) {timeout}
- node [right,black] at (axis cs:10,3) {CPAchecker faster}
- node [right,black] at (axis cs:1,55) {kIkI faster};
+%\addplot [red,sharp plot, domain=.1:1500] {900}
+%          node [below] at (axis cs:10,850) {timeout};
+%\addplot [red,sharp plot, domain=.1:1500] coordinates{(900,.1) (900,1500)}
+%          node [left,rotate=90] at (axis cs:700,150) {timeout}
+ %node [right,black] at (axis cs:10,3) {CPAchecker faster}
+ %node [right,black] at (axis cs:1,55) {kIkI faster};
 \end{loglogaxis}

### 2016-10-18 domains.tex
experiments

@@
 in abstract domains is different from the notion of precise complements. 
-\rmcmt{Convention to be followed: In SAT world, partial assignment (PA) is a data structure. Previous ACDL paper 
-showed that PA is an abstract domain and called it partial assignment domain. Likewise, in this paper, 
-whenever we refer to SAT, we also call partial assignment domain. But When we refer 
-to our ACDCL product domain, we call boolean domain and polyhedra domain. We follow this 
-convention throughout. This would make a SAT reader happy as well as a Abstract Interpreter not sad !!}
 %

### 2016-10-18 experiments.tex
experiments

@@
 (including bitwise operations) out of the box and support precise
-complementation of meet irreducibles, which are necessary for our algorithm.
+complementation of meet irreducibles, which are necessary for conflict-driven 
+learning.
 
-We verified a total of 70 benchmarks in ANSI-C, which are derived from 
+We verified a total of 66 benchmarks in ANSI-C, which are derived from 
 (a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
@@
 from Simulink (d) several hand-crafted benchmarks for equivalence 
-checking and bounded loop analysis.  All benchmarks containing 
-bounded loops are completely unrolled before analysis.  We compare 
-our tool with a state-of-the-art SAT-based bounded model checker 
-CBMC~\cite{cbmc} and a commercial static analysis tool, Astr{\'e}e.  
-CBMC uses MiniSAT solver in the backend.  Astr{\'e}e was configured with 
-default set of domains which includes interval, bit-field, congurence and 
-relational domains.  Additionally, to enable precise analysis using 
-Astr{\'e}e, we also used trace-partition domain.
+checking and bounded loop analysis.  All bounded loops are completely 
+unrolled before analysis.  We compare our tool with a state-of-the-art 
+SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
+static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
+Astr{\'e}e was configured with default set of domains which includes 
+interval, bit-field, congurence and relational domains (octagons, polyhedra, 
+zones, equalities, filter).  Additionally, to enable precise analysis 
+using Astr{\'e}e, we also used trace-partition domain.
 
@@
 \paragraph {\em \textbf{Observation of Analysis}}
+Figure~\ref{fig:results} presents a comparison of the solver statistics
+(decision and learning) for the analysis using CBMC and ACDCL.  The cumulative 
+statistics corresponding to each phase of the algorithm for all 66 benchmarks 
+is presented in Table~\ref{result}.  The reasoning using ACDCL is atleast 
+\rmcmt{1.5X} effecient compared to SAT in terms of number of decisions, propagations 
+and learning.    
+%
+\begin{table}
+\begin{center}
+{
+\begin{tabular}{l|l|l|l|l|l}
+\hline
+Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
+SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
+ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
+\end{tabular}
+}
+\end{center}
+\caption{Solver statistics}
+\label{result}
+\end{table}
+%
+Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
+without any restarts.  The solver is restarted in 50 cases to avoid spending 
+too much time in branches that does not easily lead to a satisfying assignment or 
+stronger clause learning.  In contrast, the analysis using ACDCL solved all
+66 benchmarks without any restarts.  This is attributed to intelligent decision 
+heuristics which exploits the high-level structure of the program combined 
+with stronger deductions and clause learning mechanism aided by the richer 
+abstract domains.
+
+On the other hand, Ast{\'e}e is often faster than ACDCL, but the analysis using 
+Astr{\'e}e shows high degree of imprecision.  Astr{\'e}e reported 15 false alarms 
+out of 41 safe benchmarks.
 
 \paragraph {\em \textbf{Decision Heuristics}} We compare the performance of 
-different decision heuristic on our benchmarks.  We observe that 
-the ordered heuristic outperforms other heuristics on control-intensive 
-benchmarks due to its ability to prioritize decisions on variables that 
-appears in conditional branches.  For straight-line code and most of 
-the bit-vector regression suite, random heuristics performs the best.  
-Whereas, the activity based heuristics such as Berkmin heuristic which 
-works well in propositional cases performs best for benchmarks that 
-encountered the maximum number of conflicts to prove safety, thus allowing 
-the heuristics to choose the decison variable among the set of learnt clauses.
+different decision heuristics in ACDCL.  We observe that the ordered heuristic 
+outperforms other heuristics on control-intensive benchmarks due to its 
+ability to prioritize decisions on variables that appears in conditional 
+branches.  For straight-line code and most of the bit-vector regression 
+suite, random heuristics performs the best.  Whereas, the activity based 
+heuristics such as Berkmin heuristic which works well in propositional 
+cases performs best for benchmarks that encountered the maximum number 
+of conflicts to prove safety, thus allowing the heuristics to choose 
+the decison variable among the set of learnt clauses.
 

### 2016-10-18 graphs.tex
experiments

@@
 	\legend{Safe,Unsafe}
-\addplot [domain=.1:1500] {x};
+\addplot [domain=.1:4000] {x};
 %\addplot [red,sharp plot, domain=.1:1500] {900}
@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=.1,xmax=1500, ymin=.1, ymax=1500, xlabel=SAT (Propagation),
+	\begin{loglogaxis} [xmin=.1,xmax=83000, ymin=.1, ymax=83000, xlabel=SAT (Propagation),
 			ylabel=ACDCL (Propagation),
@@
 	\legend{Safe,Unsafe}
-\addplot [domain=.1:1500] {x};
+\addplot [domain=.1:83000] {x};
 %\addplot [red,sharp plot, domain=.1:1500] {900}
@@
 \caption{\label{fig:results}
-Comparison
+Comparison of SAT and ACDCL 
 }

### 2016-10-19 domains.tex
figure 2
minor
domains

@@
 \hline
- SSA & Octagon \\
+ SSA & Abstract Conflict Graph \\
 \hline
@@
 &
-\begin{minipage}{4.50cm}
-$\begin{array}{l@{\,\,}c@{\,\,}l}
-C &\iff& ((x \geq -5) \land (x \leq 5) \land \\
-  &    & (z \geq 0) \land (z \leq 25) \land \\
-  &    & (x-z \leq 5) \land (x-z \geq -30) \land \\
-  &    & (x+z \geq -5) \land (x+z \leq 30))  
-\end{array}$
+\begin{minipage}{5.2cm}
+\centering
+\scalebox{.45}{\import{figures/}{learning.pspdftex}}
 \end{minipage}
@@
 \end{tabular}
-\caption{SSA and Octagonal Inequalities}
+\caption{SSA and abstract conflict graph}
 \label{ssa}
@@
 the concretisation of $a$.
-\rmcmt{$\gamma$ is not defined before and used here for the first time}
 %m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.
@@
 decomposed into half-spaces, that admits precise complements.  
+For example, a precise complement of a meet irreducible $(x \leq n)$ 
+in interval domain is $(x > n)$, or a meet irreducible $(x+y < 1)$ in 
+octagon domain is $(-x-y < 0)$.  Our template-based domain implementation 
+precisely complements a meet irreducible in an abstraction.
+\Omit{
 Numerical abstract domains that admit complementable decomposition 
-are shown in Figure~\ref{fig:complement}.  Our template-based domain 
-implementation precisely complements a meet irreducible in an abstraction.
+are shown in Figure~\ref{fig:complement}.  
+}
 % 
+\Omit{
 \begin{table}
@@
 \end{table}
+}
 %

### 2016-10-19 example.tex
refactoring domain.tex figure 2
figure 2
sync

@@
 %
+\pscmt{you should give an example that requires a relational domain to
+solve...}
 \begin{figure}[t]
@@
 void main() {
- int x,z;
- int v;
- assume(v>=0 $\wedge$ v<=N);
- _Bool c;
- if(c)
-   x = v;
- else 
-   x = -v;
- z = x * x;
- assert(z>=0);
+int x,sepc=0,impl=0;
+assume(x>=1 && x<=N);
+spec = 2*x; 
+int state = 1;
+int y=x;
+while(state!=0) {
+ switch(state) {
+  case 0: state = 1; 
+  break;
+  case 1: state = 2; 
+  break;
+  case 2: x--; y++;
+  if(x == 0) state=0;
+  else state=1;break;
+ } 
+}
+impl = y;
+assert(spec == impl);
 }
@@
 \end{tabular}
-\caption{\label{fig:example}
-C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
+\caption{C Program, Control-flow Graph (CFG) and Abstract Conflict
+  Graph \pscmt{where is the relational domain?}}
+\label{example}
 \end{figure}
-Let us consider a simple program in Figure~\ref{fig:example} that performs 
-multiplication of two machine integers and checks that the result of the 
-multiplication is always positive for an integer input $x$.  To avoid 
-overflow error, we assume $x$ has an upper bound $N$.  The control 
-flow graph is shown in the middle column of Figure~\ref{fig:example}.
+Let us consider a simple program in Figure~\ref{example} that checks 
+the equivalence of a state-machine based interger multiplication against  
+a specification that uses multiplication operator (*).  The assertion 
+simply compares the result of two multiplication for an integer input $x$.  
+To avoid overflow error, we assume $x$ has an upper bound $N$.  The 
+control flow graph is shown in the middle column of Figure~\ref{example}.
 
@@
 model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
-\Omit {
-The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
-and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
-the solver performs a restart, presumably to recover from bad decisions.  
-}
-Standard BMC tools convert the program to a bit-vector equation and pass 
-them to a SAT solver.  However, standard forward interval analysis is 
-too imprecise to verify the safety of the above program.  The 
-imprecision is due to control-flow join at node $n4$.  State-of-the-art 
-abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
-by manually annotating the code with partition directives at $n1$, 
-to prove safety.  The partition directive helps the tool to analyze 
-each program execution paths separately.  To do so, Astr{\'e}e 
-requires a product of trace partitioning domain and an interval domain 
-to determine safety.  However, a relational domain such as octagons is 
-able to prove the program without any partition.  In general, the 
-imprecision is either intended by the tool because such high 
-precision analysis is normally not required for runtime error 
-analysis or the imprecision is unavoidable due to the complexity 
-of the application under analysis.  
+For $N=2$ and unwind limit of $bound=4$ for the {\em while} loop,
+table~\ref{solver} shows the statistics from MiniSAT~\cite{minisat} solver 
+for analysis using BMC.  The SAT solver makes 36 decisions, 1365 boolean 
+constraint propagations and performs 5 conflict analysis with 65 conflict 
+literals.  Moreover, the solver performs a restart, presumably to recover 
+from bad decisions.
+
+On the other hand, abstract interpretation using forward analysis 
+is too imprecise to verify the safety of the above program using 
+interval domain.  The imprecision is due to control-flow join at 
+node $n1$.  State-of-the-art abstract interpretation tool, Astr{\'e}e, 
+requires external hints, provided by manually annotating the code 
+with partition directives at $n1$, to prove safety.  The partition 
+directive helps the tool to analyze each program execution paths 
+separately.  To do so, Astr{\'e}e requires a product of trace 
+partitioning domain and an interval domain to determine safety.  
+Consequently, the analysis becomes very expensive for a large value 
+of $N$.  In general, the imprecision is either intended by the tool 
+because such high precision analysis is normally not required for 
+runtime error analysis or the imprecision is unavoidable due to the 
+complexity of the application under analysis.  
 %
@@
 Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
-SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
-ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline
-ACDCL (Octagons) & 0 & 7 & 0 & 0 & 0 \\ 
+SAT & 36 & 1365 & 5 & 65 & 1 \\ \hline
+ACDCL & 1 & 23 & 1 & 1 & 0 \\
 \hline
@@
 \end{center}
-\caption{Solver statistics}
+\caption{Solver-statistics}
 \label{solver}
@@
 %
-The right-side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
-octagon domain (top) and an interval domain (bottom).  The analysis 
-associates an abstract element for each control location and variable, 
-as shown in the abstract conflict graph in Firgure~\ref{fig:example}.  
-Octagonal constraints deduced by ACDCL solver are sufficient to prove 
-safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
-an automatic interval analysis is also sufficient to prove safety with 
-only a single decision.  Interval constraints generated from forward 
-analysis in the initial deduction phase infers that 
-$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL can not prove safety.
+The right-side of Figure~\ref{example} shows an analysis using ACDCL with 
+interval domain.  The analysis associates an interval for each control 
+location and variable, as shown in the abstract conflict graph.  Interval 
+constraints generated from forward analysis in the initial deduction phase 
+infers that $\langle x:[-\infty,2], spec:[2,4], impl:[-1,\infty],
+state:[1,\infty], y:[1,+\infty] \rangle$.  Clearly, ACDCL can not prove safety.
 
-Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
-ACDCL implements several decision heuritics.  In this case, the 
-branching variable $c$ is chosen by an {\em ordered} decision 
-heuristics that select variables appearing in conditional branches 
-first before choosing other program variables. 
-The decision constrains the interval of $x:=[0,5]$.  The deductions made 
+Hence, ACDCL makes a decision $x:[-\infty,1]$ to refine the analysis.  
+The branching variable $x$ is chosen by an {\em ordered} decision 
+heuristics that choose variables appearing in conditional branches 
+(assume statement here) first before choosing other program variables. 
+The decision constrains the interval of $x$.  The deductions made 
 during fixed point iteration are represented by abstract conflict 
-graph in right-side of figure~\ref{fig:example}.  Nodes in the abstract 
-conflict graph denotes the corresponding updates to the program 
-variables in the CFG.  Interval analysis concludes that the $(ERROR:\bot)$, that 
-is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
-the program is {\em safe} for $c:=[1,1]$.
+graph in right-side of figure~\ref{example}.  Nodes in the graph 
+denotes the corresponding updates to the program variables in 
+the CFG.  Interval analysis concludes that the $(n12:\bot)$, that 
+is the decision $x:[-\infty,1]$ leads to a {\em conflict}.  Thus, 
+the program is {\em safe} for $x:[-\infty,1]$.
 
@@
 Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
-(\(I: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
+(\(I: x = [2,+\infty]\)), that is all error traces must satisfy $(x \geq 2)$ 
 at node $I$.  The analysis discards all interval constraints that
 lead to the conflict and backtrack to decision level 0.  ACDCL then 
-performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
+performs interval analysis with the learnt constraint $(x>=2)$ which again 
 leads to a conflict.  The analysis cannot backtrack further and therefore 
@@
 
-For $N=46000$, table~\ref{solver} shows the statistics from 
-MiniSAT~\cite{minisat} solver for analysis using BMC, and 
-analysis using ACDCL with interval and octagon domains.  
-Compared to a SAT solver, there is a significant reduction 
-in the number of decisions, propagations, learnt clauses and 
-restrarts.  Compared to abstract interpretation, ACDCL does 
-not require external hints to proof the program, thus 
-automatically performing program and property driven trace 
-partitioning to generate proofs using decision and clause learning.
+The result of the analysis using ACDCL is shown in table~\ref{solver}.  
+Compared to a SAT solver, ACDCL makes only a single decision which lead 
+to 23 interval constraint propagations, and learns a single conflict clause 
+involving only one literal.  Compared to abstract interpretation, ACDCL does 
+not require external hints to proof the program, thus automatically performing 
+program and property driven trace partitioning to generate proofs using decision 
+and clause learning.
+
+\pscmt{clearly say that everything that follows is for finite
+  unwindings of programs}

### 2016-10-19 paper.tex
background.tex
graph added
ssa fig, domain edits
meet decomposition
refactoring domain.tex figure 2
learning in relational lattice
sync

--- 2016-10-13 paper.tex
@@
 \usepackage{hyperref}
+\usepackage{paralist}
 \makeatother
@@
 %
-\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a combination of
-abstract interpretation and the well-known Conflict-Driven Clause Learning 
-(CDCL) algorithm that is used in SAT solvers.  ACDCL alternates between a 
-model search that performs overapproximate deduction with constraint 
-propagation and a conflict analysis that performs under-approximate 
-abduction with heuristic choice.  
+  \emph{Abstract Conflict-Driven Learning} (ACDL) is a combination of
+  abstract interpretation and the well-known Conflict-Driven Clause
+  Learning (CDCL) algorithm that is used in SAT solvers. ACDL has been
+  successfully used with non-relational abstract domains such as
+  intervals to improve the performance of a solver for floating point
+  arithmetic, for instance. 
 %
-ACDCL has been successfully used with non-relational abstract domains 
-such as intervals to improve the performance of a floating-point 
-decision procedure.  Whereas the propositional level algorithms used 
-in SAT solvers can be rather easily lifted from the boolean lattice to 
-other non-relational domains, lifting them to relational domains is 
-challenging because of relational properties of deductions and clause learning.
+Whereas the
+algorithms typically used in SAT solvers can be rather easily
+lifted from the Boolean lattice to other non-relational domains,
+lifting them to relational domains is challenging.
+%because of inter-"literal" dependencies
+%because of the weaker properties of "literals"
 %
-In this paper, we present the non-trivial adaptations of the model search and 
-conflict analysis algorithms in ACDCL in order to enable the use of relational 
-abstract domains.  We have implemented these algorithms inside the 2LS 
-verification tool and instantiated our analysis over a template polyhedra 
-domain.  We evaluate the performance of our analyzer in comparison with 
-program verifiers that uses propositional-level solvers and a static 
-analysis tool.  
+In this paper we present the adaptations of algorithms and heuristics
+for propagation, decision making and conflict learning necessary in
+order to enable the use with relational domains.
 %
-Experimental evidence on a set of benchmarks drawn from bit-vector 
-regression in SV-COMP'16, bit-precise software models of hardware 
-circuits auto-generated from v2c tool, and several bounds checking benchmarks 
-shows significant reduction in the number of decisions and backjumping iterations 
-along with stronger deductions and learned clauses aided by the richer abstract 
-domains.
+We have implemented these algorithms inside the 2LS verification tool
+for the example of the template polyhedra domain.
+%
+We evaluate the performance of our analyzer in comparison with program verifiers that 
+uses propositional-level solvers.  Experimental evidence on a set of 
+\rmcmt{N} benchmarks \rmcmt{drawn from X} shows significant reduction 
+in the number of decisions and backjumping iterations along with stronger 
+deductions and learned clause aided by the richer abstract domains.
 %
 \end{abstract}
+
 %===============================================================================

### 2016-10-19 experiments.tex
learning

@@
 complementation of meet irreducibles, which are necessary for conflict-driven 
-learning.
+learning. Our tool and benchmarks are available online\footnote{http://www.cprover.org/acdcl/}.
 
@@
 Astr{\'e}e was configured with default set of domains which includes 
-interval, bit-field, congurence and relational domains (octagons, polyhedra, 
-zones, equalities, filter).  Additionally, to enable precise analysis 
-using Astr{\'e}e, we also used trace-partition domain.
+interval, bit-field, congurence, trace partitioning and relational 
+domains (octagons, polyhedra, zones, equalities, filter).
 
@@
 }
-\paragraph {\em \textbf{Observation of Analysis}}
+\paragraph {\textbf{Precision and Efficiency of Analysis}}
 Figure~\ref{fig:results} presents a comparison of the solver statistics
@@
 with stronger deductions and clause learning mechanism aided by the richer 
-abstract domains.
+abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
+but the analysis using Astr{\'e}e shows high degree of imprecision.  
+Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.
 
-On the other hand, Ast{\'e}e is often faster than ACDCL, but the analysis using 
-Astr{\'e}e shows high degree of imprecision.  Astr{\'e}e reported 15 false alarms 
-out of 41 safe benchmarks.
-
-\paragraph {\em \textbf{Decision Heuristics}} We compare the performance of 
+\paragraph {\textbf{Decision Heuristics}} We compare the performance of 
 different decision heuristics in ACDCL.  We observe that the ordered heuristic 
@@
 
-\paragraph {\em \textbf{Propagation Strategy}}
+\paragraph {\textbf{Propagation Strategy}}      
+Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis and
+{\em chaotic} analysis in ACDCL.  The choice of analysis influences the 
+total number of decisions and clause learning iterations.  Hence, the 
+propagation strategy has a significant influence on the run time which 
+can be visualised in Figure~\ref{prop-dec}.  Unlike the forward analysis, 
+the chaotic or multi-way strategy performs both forward and backward 
+propagations.  This may take several iterations to reach the fixed-point, but 
+significantly reduces the total number of decisions and learning.  
+\Omit {
+We observe that the deductions made during chaotic analysis helps the decision 
+heuristics to make better decisions compared to forward analysis.  This is 
+attributed to the precision of deductions inferred by the chaotic analysis.         
+}
 
-\paragraph {\em \textbf{Clause Learning}}
+\paragraph {\textbf{Learning}} Learning has a significant influence on
+the runtime of ACDCL.  We compare the first-UIP learning technique with 
+a DPLL-style conflict analysis without learning.  The effect of UIP computation 
+allows ACDCL to backtrack non-chronologically and guide the model search with 
+a learnt transformer.  However, the DPLL-style analysis exhibits
+case-enumeration behavior and thus immediately runs out of capacity for programs 
+which require relational domains.
 

### 2016-10-19 decision.tex
effect of decision heuristics

@@
 variable $v \in \mathcal{V}$, a bound $c$, and the polarity ($\succeq$ or 
-$\preceq$).
+$\preceq$).  A decision changes the sate of the solver by adding new element 
+$m$ and labelling information $s=decision$ to the trail, which is described below. 
+\[decide: \quad (\mathcal{E},S) \rightarrow (\mathcal{E}(m,s),S) \]
 
-{\em Titlis} support several decision heuristics namely, {\em ordered} 
-{\em longest-range}, {\em random}, {\em Berkmin}~\cite{} decision heuristics.  
+ACDL supports several decision heuristics namely, {\em ordered} 
+{\em longest-range}, {\em random}, {\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
 The {\em ordered} decision heuristics creates an ordering among non-singleton 

### 2016-10-19 example.tex

--- 2016-10-15 example.tex
@@
 &
-\begin{minipage}{4.40cm}
+\begin{minipage}{3.7cm}
 \centering
-\vspace*{0.3cm}
+%\vspace*{0.3cm}
 \scalebox{.45}{\import{figures/}{example.pspdftex}}
@@
 &
-\begin{minipage}{4.40cm}
+\begin{minipage}{4.50cm}
 \centering
@@
 performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
-leads to a conflict.  The analysis cannot backtrack further and therefore 
-it terminates proving that the program is safe.  A DPLL-style solver would 
-perform proof by cases.  However, decision and clause learning are used 
-to avoid case based reasoning and prevents enumeration behavior.
+leads to a conflict, as shown in right side of figure~\ref{ssa}.  The 
+analysis cannot backtrack further and therefore it terminates proving 
+that the program is safe.  A DPLL-style solver would perform proof by 
+cases.  However, decision and clause learning are used to avoid case 
+based reasoning and prevents enumeration behavior.
 
@@
 partitioning to generate proofs using decision and clause learning.
+
+\pscmt{clearly say that everything that follows is for finite
+  unwindings of programs}

### 2016-10-19 experiments.tex

--- 2016-10-18 experiments.tex
@@
 \paragraph {\em \textbf{Propagation Strategy}}      
+Figure~\ref{} presents a comparison between {\em forward} iteration strategy and
+{\em chaotic} iteration strategy in ACDCL.  The iteration strategy influences the total 
+number of decisions and clause learning iterations.  Forward iteration strategy
+is similar to forward analysis in abstract interpretation.  However, the chaotic 
+or multi-way strategy performs both-ways propagations which may take several 
+iterations to reach the fixed-point, but significantly reduces the total number of
+decisions and learning.  We observe that the deductions made during chaotic iteration 
+helps the decision heuristics to make better decisions compared to forward analysis.
 

### 2016-10-19 graphs.tex

@@
 \caption{\label{fig:results}
-Comparison of SAT and ACDCL 
+Comparison between SAT and ACDCL 
 }
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[t]
+\begin{tabular}{@{\hspace{-1.5em}}c@{\hspace{1em}}c}
+\begin{tikzpicture}[scale=0.75]
+	\begin{loglogaxis} [xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, xlabel=Forward
+  Propagation, ylabel=Chaotic Propagation, 
+			legend style={at={(0.8,0.15)},
+			anchor=north,legend columns=-1 },
+			]
+\addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
+	scatter/classes={a={mark=square,draw=black}}] 
+	table [meta=label] {plotdata/scatter-chaotic-forward.dat};
+	%\legend{Safe,Unsafe}
+  \addplot [domain=.1:2500] {x};
+\end{loglogaxis}
+\end{tikzpicture}
+ &
+\begin{tikzpicture}[scale=0.75]
+	\begin{loglogaxis} [xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, xlabel=Forward 
+  Propagation,ylabel=Chaotic Propagation, 
+			legend style={at={(0.8,0.15)},
+			anchor=north,legend columns=-1 },
+			]
+\addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
+	scatter/classes={a={mark=square,draw=black}}] 
+	table [meta=label] {plotdata/scatter-chaotic-forward.dat};
+	%\legend{Safe,Unsafe}
+  \addplot [domain=.1:2500] {x};
+\end{loglogaxis}
+\end{tikzpicture} \\
+(a) & (b)
+\end{tabular}
+\caption{\label{fig:results}
+Comparison between Chaotic and Forward Propagation Strategies 
+}
+\end{figure}

### 2016-10-19 background.tex
background

@@
-\section{Background}
-The natural domain SMT framework proposed in~\cite{ndsmt} operates on
-a fixed first-order partial assignment lattice structure, where first-order 
-variables are mapped to domain values, similar to constants lattice 
-in program analysis.  However, ACDCL can be instantiated on different 
-abstract lattices.  The abstract lattice in natural domain SMT does 
+\section{Related Work}
+The work of~\cite{sas13} present an instantiation of ACDCL as a decision 
+procedure for floating-point arithmetic that uses interval constraint 
+propagation.  The use of ACDCL as program analysis is proposed
+in~\cite{sas12,tacas12}.  The analysis operate over the abstract domain of 
+floating-point and machine-integer.  Thus, it is restricted to bounds 
+checking for numerical programs.  A similar technique that lifts DPLL(T) 
+to programs is Satisfiability Modulo Path Programs (SMPP). SMPP enumerates 
+path programs using a SAT formula which are then verified using abstract 
+interpretation.  The lifting of CDCL to first-order theory is proposed 
+in~\cite{dpll,cp09} and natural domain SMT framework~\cite{ndsmt}.  
+\Omit{
+operates on a fixed first-order partial assignment lattice structure, 
+where first-order variables are mapped to domain values, similar to constants lattice 
+in program analysis.  
+}
+However, unlike previous works that operate on a fixed first-order 
+lattice structure, ACDCL can be instantiated with different 
+abstract domains.  This involves model search and learning in 
+abstract lattices.     
+\Omit {
+The abstract lattice in natural domain SMT does 
 not have complementable meet irreducibles, and therefore does not 
@@
 ACDCL to perform generalized clause learning.
+}

### 2016-10-19 related_work.tex
background

@@
 \section{Related Work}
-The natural domain SMT framework proposed in~\cite{ndsmt} operates on
-a fixed first-order partial assignment lattice structure, where first-order 
-variables are mapped to domain values, similar to constants lattice 
-in program analysis.  However, ACDCL can be instantiated on different 
-abstract lattices.  The abstract lattice in natural domain SMT does 
+The work of~\cite{} present an instantiation of ACDCL as a decision 
+procedure for floating-point arithmetic that uses interval constraint 
+propagation.  The use of ACDCL as program analysis is proposed
+in~\cite{sas12,tacas12}.  The analysis operate over the abstract domain of 
+floating-point and machine-integer.  Thus, it is restricted to bounds 
+checking for numerical programs.  A similar technique that lifts DPLL(T) 
+to programs is Satisfiability Modulo Path Programs (SMPP). SMPP enumerates 
+path programs using a SAT formula which are then verified using abstract 
+interpretation.  The lifting of CDCL to first-order theory is proposed 
+in~\cite{dpll,cp09} and natural domain SMT framework~\cite{ndsmt}.  
+\Omit{
+operates on a fixed first-order partial assignment lattice structure, 
+where first-order variables are mapped to domain values, similar to constants lattice 
+in program analysis.  
+}
+However, unlike previous works that operate on a fixed first-order 
+lattice structure, ACDCL can be instantiated with different 
+abstract domains.  This involves model search and learning in 
+abstract lattices.     
+\Omit {
+The abstract lattice in natural domain SMT does 
 not have complementable meet irreducibles, and therefore does not 

### 2016-10-19 algorithm.tex
algorithm description

@@
 set of meet irreducibles for the current decision.  The model search phase 
-alternates between the decision and deduction until the result of deduction 
-is either $\gamma$-complete~\cite{gamma} or the deduction leads to a {\em CONFLICT}.  
+alternates between the decision and deduction until the abstract value $v$ 
+precisely represents the set of models or the deduction leads to a {\em CONFLICT}.  
+Checking whether $v$ precisely represents set of models corresponds to 
+$\gamma$-completeness~\cite{dhk2013-popl} check in abstract interpretation.  
 If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 

### 2016-10-19 experiments.tex
comments in section 7

--- 2016-10-19 experiments.tex
@@
 Astr{\'e}e was configured with default set of domains which includes 
-interval, bit-field, congurence, trace partitioning and relational 
+interval, bit-field, congruence, trace partitioning and relational 
 domains (octagons, polyhedra, zones, equalities, filter).
@@
 statistics corresponding to each phase of the algorithm for all 66 benchmarks 
-is presented in Table~\ref{result}.  The reasoning using ACDCL is atleast 
-\rmcmt{1.5X} effecient compared to SAT in terms of number of decisions, propagations 
-and learning.    
+is presented in Table~\ref{result}.  The reasoning using ACDCL is at least 
+\rmcmt{1.5X} more efficient compared to propositional SAT in terms of number of decisions, propagations 
+and learning \pscmt{what learning???}.    
 %
@@
 outperforms other heuristics on control-intensive benchmarks due to its 
-ability to prioritize decisions on variables that appears in conditional 
+ability to prioritize decisions on variables that appear in conditional 
 branches.  For straight-line code and most of the bit-vector regression 
-suite, random heuristics performs the best.  Whereas, the activity based 
+suite, random heuristics performs the best.  Whereas activity-based 
 heuristics such as Berkmin heuristic which works well in propositional 
-cases performs best for benchmarks that encountered the maximum number 
-of conflicts to prove safety, thus allowing the heuristics to choose 
+cases performs best for benchmarks that encountered \pscmt{???} the maximum number 
+of conflicts to prove safety, thus allowing the heuristics \pscmt{???} to choose  
 the decison variable among the set of learnt clauses.
@@
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis and
-{\em chaotic} analysis in ACDCL.  The choice of analysis influences the 
+Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis \pscmt{iteration/propagation strategy} and
+{\em chaotic} analysis in ACDCL \pscmt{what about backward?}.  The choice of analysis influences the 
 total number of decisions and clause learning iterations.  Hence, the 
 propagation strategy has a significant influence on the run time which 
-can be visualised in Figure~\ref{prop-dec}.  Unlike the forward analysis, 
+can be visualised in Figure~\ref{prop-dec}. \pscmt{Are the graphs up-to-date?} Unlike the forward analysis, 
 the chaotic or multi-way strategy performs both forward and backward 
 propagations.  This may take several iterations to reach the fixed-point, but 
-significantly reduces the total number of decisions and learning.  
+significantly reduces the total number of decisions and learning. \pscmt{Why?} 
 \Omit {
@@
 the runtime of ACDCL.  We compare the first-UIP learning technique with 
-a DPLL-style conflict analysis without learning.  The effect of UIP computation 
+a DPLL-style conflict analysis without learning \pscmt{you learn the negation of the last decision}.  The effect of UIP computation 
 allows ACDCL to backtrack non-chronologically and guide the model search with 
@@
 case-enumeration behavior and thus immediately runs out of capacity for programs 
-which require relational domains.
+which require relational domains. \pscmt{Is there a significant difference between intervals and octagons that indicates that for relational domains DPLL is even worse than in the propositional case?}
 

### 2016-10-19 decision.tex
decision

@@
 current abstract valuation.  A decision can be of the form 
-$\langle x \preceq c \rangle$ or $\langle x \succeq c \rangle$, 
-where $c$ is the bound and $x$ is a program variable.  A 
-decision must always be consistent with respect to the 
-trail $\mathcal{T}$, that is, it must not contradict with 
-the elements in the trail.
+$\langle x R c \rangle$ or $\langle x R y \rangle$, where 
+$R \in \{\preceq, \succeq\}$, $c$ is the bound and 
+$\{x, y\}$ are program variables.  A decision must always 
+be consistent with respect to the trail $\mathcal{T}$, that 
+is, it must not contradict with the elements in the trail.
 
@@
 
-ACDL supports several decision heuristics namely, {\em ordered} 
-{\em longest-range}, {\em random}, {\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
+ACDCL supports several decision heuristics namely, {\em ordered} 
+{\em longest-range}, {\em random}, {\em relational} and 
+{\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
 The {\em ordered} decision heuristics creates an ordering among non-singleton 
@@
 restricted.  A {\em random} decision heuristics arbitrarily picks a variable 
-for making decision.  Whereas, a {\em berkmin} decision heuristics is inspired 
+for making decision.  A {\em relational} decision heuristics is of the form 
+$x R y$ for $R \in \{preceq, succceq\}$ and is only relevant for relational 
+abstract domains.  Whereas, a {\em berkmin} decision heuristics is inspired 
 from decision heuristics used in Berkmin~\cite{eugoldberg07} SAT solver, which 

### 2016-10-19 conclusion.tex
background

@@
 In this paper, we instantiate ACDCL solver with relational domains.  We explained specific properties of model search and conflict analysis procedure necessary
-for such lifting.  Experimental evaluation on different benchmarks shows that
+for such lifting.  Experimental evaluation on different benchmark sets shows that
 ACDCL is able to suitably exploit the expressivity of richer
@@
 
-Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check.
+\rmcmt{Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check. Implement dedicated abstract domains to exploit full performance.
+generalization for relational domain, instantiate acdl with trace-based domains.}

### 2016-10-19 background.tex
more comments
reference

@@
 \section{Related Work}
-The work of~\cite{sas13} present an instantiation of ACDCL as a decision 
-procedure for floating-point arithmetic that uses interval constraint 
+The work of~\cite{sas13,DBLP:journals/fmsd/BrainDGHK14} present 
+an instantiation of ACDCL as a decision procedure for 
+floating-point arithmetic that uses interval constraint 
 propagation.  The use of ACDCL as program analysis is proposed
@@
 }
+\pscmt{\cite{DBLP:journals/fmsd/BrainDGHK14} and \cite{DBLP:conf/esop/BrainDKS14} are also ACDL instantiations (the latter is even a relational one, but without backtracking)}

### 2016-10-19 learning.tex
comments in section 6

@@
 reasoning where a singleton assignment $c$ is replaced by a partial 
-assignment that is sufficient to infer $c$.  Thus, a conflict 
-analysis is a generalization of an underapproximation of a set of 
+assignment that is sufficient to infer $c$ \pscmt{???}.  Thus, a conflict 
+analysis is a generalization of an under-approximation of a set of 
 countermodels~\cite{sas12,dhk2013-popl}, where a countermodel is a 
-set of structures that do not satisfy a formula.  Since relational 
+set of structures that do not satisfy a formula \pscmt{this explanation does not help the reader, who will mix up counterexamples and countermodels}.  Since relational 
 abstract domains have more complex structure than a partial 
 assignments domain, so the lifting of {\em first-UIP} algorithm~\cite{uip} 
-and clause learning to any arbitraty lattice structure is a non-trivial task.
+and clause learning to any arbitrary lattice structure is a non-trivial task.
 
@@
 any path from the last decision node to the conflict node must pass 
-through it.  Unlike a SAT solver, a UIP in ACDCL may involve meet irreducibles 
-that involves the same variables.  The first-UIP is a node closest to the 
+through it.  Unlike a SAT solver, a UIP in ACDCL may have meet irreducibles 
+that involve the same variables.  The first-UIP is a node closest \pscmt{unique?} to the 
 conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
@@
 A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
-conflict node and iteratively finds new nodes that contradict with the 
+conflict node and iteratively finds new nodes that contradict \pscmt{we have to define what that means in the preliminaries} with the 
 transformer stored in $reasons[\bot]$.
@@
 marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
-Choosing cut0 yields a learnt clause 
+Choosing cut0 yields a learnt clause \pscmt{use $\leq$ instead of $<$} 
 $L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
@@
 %
-\begin{figure}
+\begin{figure}[t]
 %\caption{An example of Octagon}\label{octagon}
 \scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
-\label{conflict}
+\caption{\label{conflict}
+\pscmt{push constants to the lhs of the inequalities; labels cut0, cut1 not fully visible}
+}
 \end{figure} 
@@
 \paragraph{\textbf{Learning in Relational Lattice}}
-We model learning as a meet irreducible deduction transformer, called 
-{\em abstract unit transformer (AUnit)}.  For an 
+We model learning as a meet irreducible deduction transformer \pscmt{???}, called 
+{\em abstract unit transformer (AUnit)} \pscmt{why the name? seems confusing as we are talking about learning and not about deduction}.  For an 
 abstract lattice $A$ with complementable meet irreducibles 
-and a set of conflicting meet irreducibles (countermodels) 
-$C \subseteq A$ of $\varphi$ such that \meet{C} does not 
+and a set of conflicting meet irreducibles (countermodels) \pscmt{The term countermodels for $C$ is misleading since $\bigsqcap C$ is a set of countermodels, but the elements of $C$ individually are no countermodels.}
+$C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
 satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
+\pscmt{explain which cases are different in non-relational domains}
 \[ AUnit(a) =
@@
 \]
-For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ be the conflicting set 
+For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ \pscmt{why suddenly a different notation for $k$?
+The variable name $k$ might be confusing because it is usually used for counting or indices} be the conflicting set
 of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
@@
 consistent} \]   
+\pscmt{What is $\mathcal{E}$? What is $m$? What is the notation $\mathcal{E}_1(m,s)\mathcal{E}_2$? What does consistent mean?}
+\pscmt{Is there any difference between non-relational and relational?}
 

### 2016-10-19 domains.tex
more comments

@@
 \end{tabular}
-\caption{SSA and abstract conflict graph}
+\caption{SSA and abstract conflict graph \pscmt{Where do you refer to this graph?}}
 \label{ssa}
@@
 %
-Figure~\ref{ssa} shows the SSA constraints for the program in figure~\ref{fig:example}. 
+Figure~\ref{ssa} shows the SSA constraints for the program in Figure~\ref{fig:example}. 
 %
@@
 \[\llbracket s
-\rrbracket^\sharp(a)=\bigsqcup \{a'\mid a'\models s\}\] where we write
+\rrbracket^\sharp(a)=\bigsqcup \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write
 $a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
@@
 In a forward analysis, we can deduce, for example, $\llbracket
-x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6)$ for $a=(0\leq y \leq 2
+x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6) \sqcap a$ for $a=(0\leq y \leq 2
 \wedge 1 \leq y-z \leq 1)$.
@@
 %
-The precise complementation property of partial assignment lattice 
-can be generalized to other lattice structures.  Most numerical 
-abstract domains, such as intervals, octagons, polyhedra, can be 
-decomposed into half-spaces, that admits precise complements.  
-For example, a precise complement of a meet irreducible $(x \leq n)$ 
-in interval domain is $(x > n)$, or a meet irreducible $(x+y < 1)$ in 
-octagon domain is $(-x-y < 0)$.  Our template-based domain implementation 
-precisely complements a meet irreducible in an abstraction.
-\Omit{
+The precise complementation property of partial assignment lattice can
+be generalized to other lattice structures.  Most numerical abstract
+domains, such as intervals, octagons, polyhedra, can be decomposed
+into half-spaces, that admits precise complements.  For example, a
+precise complement of a meet irreducible $(x \leq 2)$ in the interval
+domain over integers is $(x \geq 3)$, or the precise complement of
+the meet irreducible $(x+y \leq 1)$ in the octagon domain over
+integers is $(x+y \geq 2)$.  Our template-based domain implementation
+precisely complements a meet irreducible in an abstraction.  \Omit{
   Numerical abstract domains that admit complementable decomposition
-are shown in Figure~\ref{fig:complement}.  
-}
+  are shown in Figure~\ref{fig:complement}.  }
 % 
@@
 %
-Another feature that abstract domain libraries do not provide is to
-track the reasons why certain deductions have been made. We come back
-to this point in Section~\ref{sec:abst}.
+%Another feature that abstract domain libraries do not provide is to
+%track the reasons why certain deductions have been made. We come back
+%to this point in Section~\ref{sec:abst}.
 
@@
 
-\Omit{
-%
-}
 % \Omit {

### 2016-10-19 example.tex
more comments

@@
 \caption{\label{fig:example}
-C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph \pscmt{I do not understand the arrow from n2 to n3 on the right hand side}}
 \end{figure}
@@
 analysis in the initial deduction phase infers that 
-$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL can not prove safety.
+$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.
 
-Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
+Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis. \pscmt{Do we want to use the product domain right away to be consistent?}  
 ACDCL implements several decision heuritics.  In this case, the 
@@
 during fixed point iteration are represented by abstract conflict 
-graph in right-side of figure~\ref{fig:example}.  Nodes in the abstract 
+graph in right-side of Figure~\ref{fig:example}.  Nodes in the abstract 
 conflict graph denotes the corresponding updates to the program 
-variables in the CFG.  Interval analysis concludes that the $(ERROR:\bot)$, that 
+variables in the CFG.  Interval analysis concludes that $(\text{Error}:\bot)$, that 
 is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
@@
 performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
-leads to a conflict, as shown in right side of figure~\ref{ssa}.  The 
+leads to a conflict, as shown in right side of Figure~\ref{ssa}.  The 
 analysis cannot backtrack further and therefore it terminates proving 

### 2016-10-19 decision.tex
comments in sec 5

@@
 A decision $\mathcal{D}$ is a meet irreducible that refines the 
-current abstract valuation.  A decision can be of the form 
+current abstract valuation.  A decision can be of the form \pscmt{that's too restrictive?}
 $\langle x R c \rangle$ or $\langle x R y \rangle$, where 

### 2016-10-19 deductions.tex
minor changes
comments in sec 5

@@
 %
+\Omit {
 \begin{figure}[t]
@@
 \end{figure}
+}
 %
@@
 A model search procedure in a SAT solver involves two steps -- {\em deductions} 
-using unit rule refines current partial assignments and 
+using the unit rule refines current partial assignments and 
 {\em decisions} to heuristically guess a value for an unassigned 
-literal.  A unit rule overapproximates a model transformer and deduction 
+literal.  The unit rule overapproximates a model transformer and deduction 
 computes a greatest fixed point over the partial assignments
@@
 that computes a greatest fixed point over meet irreducible deduction 
-transformer in $S$.  
+transformer in $S$ \pscmt{???}.  
 %
@@
 deductions w.r.t.\ a given subdomain $L\subseteq A$; $\bot,\top \in L$.
-\[\llbracket s \rrbracket^\sharp(a,L)=\bigsqcup \{a'\mid a'\in L, a'\models s\}\]
+\[\llbracket s \rrbracket^\sharp(a,L)=\bigsqcup \{a'\mid a'\in L, a' \sqsubseteq a, a'\models s\}\]
 We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
@@
 
-Moreover, we want to know what the reasons for a specific deduction are.
-\[ded(s,a,L)=\{R\rightarrow d\mid R=reasons(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}\]
-where 
-$reasons(s,a,L,d)=\text{argmin}_{R\in\{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}} |R| $.
+%Moreover, we want to know what the reasons for a specific deduction are.
+%\[ded(s,a,L)=\{R\rightarrow d\mid R=reasons(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}\]
+%where 
+%$reasons(s,a,L,d)=\text{argmin}_{R\in\{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}} |R| $.
 
@@
 iteration strategy starts with a set of assertions. On the other hand, a 
-{\em chaotic} iteration strategy initializes the worklist with set of all 
-transformers.  For every meet irreducible deduction transformer $s \in S$ 
+{\em chaotic} \pscmt{that's not a good name because it is not chaotic in the meaning of chaotic iterations} iteration strategy initializes the worklist with set of all 
+transformers.  For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
 passed to the abstract domain, if the domain infers a new meet irreducible 
-$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$.  Additionally, the reason 
-trail $\mathcal{R}$ is updated with the transformer $s$ as the reason for inferring
+$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$ \pscmt{It usually infers several ones.}.  Additionally, the reason 
+trail $\mathcal{R}$ \pscmt{define} is updated with the transformer $s$ as the reason for inferring
 $\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
@@
 models.  This is determined by checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{gamma}.  This can be achieved by checking if there is a 
+$\gamma$-complete~\cite{dhk2013-popl}.  This can be achieved by checking if there is a 
 concrete solution $c$ such that the inequality $c(x) \subseteq \gamma \circ
-v(x)$ holds for every meet irreducible $x$.  If $v$ is $\gamma$-complete, that is, the 
+v(x)$ holds for every meet irreducible $x$ \pscmt{No, you have to consider $\bigsqcap x$}.  If $v$ is $\gamma$-complete, that is, the 
 abstract valuation $v$ is abstractly satisfying, the abstract model 

### 2016-10-19 algorithm.tex
comments in sec 5

@@
 The algorithm begins with a set $S$ of abstract deduction transformer 
-$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+$\widehat{ded_{\varphi}}$ \pscmt{???} for a formula $\varphi$ over an abstract 
 domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
@@
 respectively.  The propagation trail stores all meet irreducibles inferred 
-by the abstact model search phase (deductions and decisions), while the 
+by the abstract model search phase (deductions and decisions), while the 
 reason trail maps elements of propagation trail to transformers that were 
@@
 The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
-fixed point over the transformers in F that refines the abstract valuation 
-$v$, similar to BCP step in SAT solvers.  If the result of $deduce$ 
-returns {\em CONFLICT} ($\perp$), the algorithm terminates with {\em safe}. 
+fixed point over the transformers in F \pscmt{what is F?} that refines the abstract valuation 
+$v$, similar to the BCP \pscmt{mentioned before?} step in SAT solvers.  If the result of $deduce$ 
+returns {\em CONFLICT} \pscmt{don't use capitals} ($\perp$), the algorithm terminates with {\em safe}. 
 Else, the analysis enters into the while($true$) loop and makes a new decision 
 by a call to $decide(v)$ which returns a new meet irreducible $d$ that is
-consistent to the state of the solver.  If $d$ refines the current abstract 
+consistent to the state of the solver. \pscmt{why do we have to store $v$ and $\mathcal{T}$?} If $d$ refines the current abstract 
 value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 
@@
 
-For example, a decision in interval domain restrict the range of 
+For example, a decision in interval domain restricts the range of 
 intervals for variables, so the analysis jumps under a 
-greatest fixed-point. Note that the widening operation in abstract interpretation 
+greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
 jumps above a least fixed-point, so decisions can be viewed as 
-{\em dual widening}.  The procedure $deduce$ is called next to infer new 
+{\em dual widening} \pscmt{what do we learn?}.  The procedure $deduce$ is called next to infer new 
 set of meet irreducibles for the current decision.  The model search phase 
@@
 If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
-algorithm returns the error trace $v$ and terminates with UNSAFE or the 
+algorithm returns the error trace $v$ \pscmt{this is not a trace} and terminates with UNSAFE or the 
 current abstraction is insufficient to determine the satisfiability of $\varphi$.  
-However, in case of {\em CONFLICT}, the algorithm enters in to 
+However, in case of {\em CONFLICT}, the algorithm enters in the 
 $conflict\_analysis()$ phase to learn the reason for {\em conflict}.  
@@
 the choice of Unique Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically 
-choose one reason.  A learnt clause must include asserting cuts which 
+chooses one reason.  A learnt clause must include asserting cuts which 
 guarantees derivation of new meet irreducibles after backtracking. The clause 
@@
 is conflicting ($\perp$), that is, the abstract valuation $v$ does not
-abstractly satisfy the formula.  If no further backtrack is possible, 
+abstractly satisfy the formula \pscmt{define}.  If no further backtrack is possible, 
 then the algorithm terminates and $\varphi$ is {\em SAFE}. Else, the 
 algorithm makes a new decision and the above process is repeated 
-until a real counterexample is obtained or the algorithm backtracks to 
+until a real counterexample \pscmt{is this a model?} is obtained or the algorithm backtracks to 
 decision level 0 after a conflict in which case it returns {\em SAFE}.
@@
 or $s = deduction$ if $m$ is inferred by a deduction. And $S$ is 
-a set of abstract deduction transformer.  A trail in SAT solver 
+a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
 stores variable assignments of the form $(p, t)$, where $p$ is 
@@
 inferred by deduction or decision phase where a variable in trail 
-can be assigned multiple times, each time with increasingly precise 
-bounds.  A current valuation $v$ is a meet of all elements of trail
+can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
+bounds \pscmt{that's specific for a particular domain}.  A current valuation $v$ is a meet of all elements of trail
 $\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
@@
 $\mathcal{S}$ is not abstractly satisfied by $v$.
+
+\pscmt{define ``consistent with trail''}

### 2016-10-19 decision.tex
more comments in section 5

@@
 singleton value, for example, $\langle x:[1,1] \rangle$.  Otherwise, 
-the variable is non-singleton.  Given a set of variables $\mathcal{V}$, 
+the variable is non-singleton \pscmt{this definition only works for non-relational domains}.  Given a set of variables $\mathcal{V}$, 
 a decision phase heuristically chooses a non-singleton branching 
@@
 
-ACDCL supports several decision heuristics namely, {\em ordered} 
+ACDCL supports several decision heuristics namely, {\em ordered}, 
 {\em longest-range}, {\em random}, {\em relational} and 
@@
 This variable ordering is path-sensitive as it gives an effect of trace
-partitioning.  A {\em longest-range} heuristics simply keeps track of the
+partitioning.
+
+The {\em longest-range} heuristics simply keeps track of the
 interval $[l,u]$ range ($r=u-l$) of a variable and chooses a variable with 
-the longest range value.  This ensures a fairness policy in selecting a 
+the longest range value \pscmt{why is this restricted to variables?}.  This ensures a fairness policy in selecting a 
 variable since it guarantees that the intervals of variables are uniformly 
-restricted.  A {\em random} decision heuristics arbitrarily picks a variable 
-for making decision.  A {\em relational} decision heuristics is of the form 
+restricted.
+
+The {\em random} decision heuristics arbitrarily picks a variable 
+for making decision. \pscmt{uniformly distributed?}
+
+The {\em relational} decision heuristics is of the form 
 $x R y$ for $R \in \{preceq, succceq\}$ and is only relevant for relational 
-abstract domains.  Whereas, a {\em berkmin} decision heuristics is inspired 
+abstract domains.
+
+Whereas, a {\em berkmin} decision heuristics is inspired 
 from decision heuristics used in Berkmin~\cite{eugoldberg07} SAT solver, which 
@@
 of a variable as the number of occurrences of that variable in conflict 
-clauses only.
+clauses only. \pscmt{how do you adapt this to relational domains?}
 
-A bound of a variable is heuristically chosen to be an approximation of the 
+A bound of a variable \pscmt{why restricted to variables?} is heuristically chosen to be an approximation of the 
 arithmetic average of the current bounds.  However, the polarity ($\preceq$ or

### 2016-10-19 graphs.tex

@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, xlabel=Forward 
-  Propagation,ylabel=Chaotic Propagation, 
-			legend style={at={(0.8,0.15)},
-			anchor=north,legend columns=-1 },
+	\begin{axis} [
+  title={Performance of Decision Heuristics},
+  xmin=0, xmax=70, 
+  ymin=0, ymax=1500, 
+  xlabel={Benchmarks},
+  ylabel={time (in seconds)},
+  xtick={},
+  ytick={}, %{0, 0.5, 1.0, 10, 50, 100, 200, 500, 700, 1000, 1500},
+	legend pos=north west,
  ]
-\addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
-	scatter/classes={a={mark=square,draw=black}}] 
-	table [meta=label] {plotdata/scatter-chaotic-forward.dat};
-	%\legend{Safe,Unsafe}
-  \addplot [domain=.1:2500] {x};
-\end{loglogaxis}
+ \addplot table[x=benchmark,y=time] {plotdata/berkmin.dat};
+ %\addplot table[x=benchmark,y=time] {plotdata/random.dat};
+ \legend{Berkmin}
+ %\legend{Random}
+ %\addplot [domain=.1:2500] {x};
+ \end{axis}
 \end{tikzpicture} \\
@@
 \end{tabular}
-\caption{\label{fig:results}
-Comparison between Chaotic and Forward Propagation Strategies 
+\caption{\label{prop-dec}
+Effect of Propagation Strategies and Decision Heuristics 
 }
 \end{figure}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-10-19 algorithm.tex
notes from today's call

@@
 The algorithm begins with a set $S$ of abstract deduction transformer 
-$\widehat{ded_{\varphi}}$ \pscmt{???} for a formula $\varphi$ over an abstract 
+$\widehat{ded_{\varphi}}$ \pscmt{define in terms of P and A} for a formula $\varphi$ over an abstract 
 domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
@@
 reason trail maps elements of propagation trail to transformers that were 
-used to derive them.
+used to derive them. \pscmt{define operator to turn trail into abstract value $\bigsqcap\mathcal{T}=\bigsqcap_{m\in\mathcal{T}}m$}
 
 The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
-fixed point over the transformers in F \pscmt{what is F?} that refines the abstract valuation 
-$v$, similar to the BCP \pscmt{mentioned before?} step in SAT solvers.  If the result of $deduce$ 
+fixed point over the transformers in $\varphi$ that refines the abstract valuation 
+$v$, similar to the BCP step in SAT solvers.  If the result of $deduce$ 
 returns {\em CONFLICT} \pscmt{don't use capitals} ($\perp$), the algorithm terminates with {\em safe}. 
@@
 
-For example, a decision in interval domain restricts the range of 
-intervals for variables, so the analysis jumps under a 
-greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
-jumps above a least fixed-point, so decisions can be viewed as 
-{\em dual widening} \pscmt{what do we learn?}.  The procedure $deduce$ is called next to infer new 
+For example, a decision in the interval domain restricts the range of 
+intervals for variables.
+% so the analysis jumps under a 
+%greatest fixed-point \pscmt{???}. Note that the widening operation in abstract interpretation 
+%jumps above a least fixed-point, so decisions can be viewed as 
+%{\em dual widening} \pscmt{what do we learn?}.  
+The procedure $deduce$ is called next to infer a new 
 set of meet irreducibles for the current decision.  The model search phase 
 alternates between the decision and deduction until the abstract value $v$ 
-precisely represents the set of models or the deduction leads to a {\em CONFLICT}.  
-Checking whether $v$ precisely represents set of models corresponds to 
+precisely \pscmt{not true} represents the set of models or the deduction leads to a {\em CONFLICT}.  
+Checking whether the abstract model $v$ concretises to a set of models of $\varphi$; this corresponds to a 
 $\gamma$-completeness~\cite{dhk2013-popl} check in abstract interpretation.  
 If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
-algorithm returns the error trace $v$ \pscmt{this is not a trace} and terminates with UNSAFE or the 
+algorithm returns the abstract model $v$, which is a set of concrete models, and terminates with UNSAFE or the 
 current abstraction is insufficient to determine the satisfiability of $\varphi$.  

### 2016-10-19 deductions.tex
notes from today's call

@@
 iteration strategy starts with a set of assertions. On the other hand, a 
-{\em chaotic} \pscmt{that's not a good name because it is not chaotic in the meaning of chaotic iterations} iteration strategy initializes the worklist with set of all 
+{\em multi-way} iteration strategy initializes the worklist with set of all 
 transformers.  For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
@@
 models.  This is determined by checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{dhk2013-popl}.  This can be achieved by checking if there is a 
-concrete solution $c$ such that the inequality $c(x) \subseteq \gamma \circ
-v(x)$ holds for every meet irreducible $x$ \pscmt{No, you have to consider $\bigsqcap x$}.  If $v$ is $\gamma$-complete, that is, the 
+$\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
 abstract valuation $v$ is abstractly satisfying, the abstract model 

### 2016-10-19 graphs.tex
notes from today's call

@@
 	\begin{axis} [
-  title={Performance of Decision Heuristics},
+  title={Performance of Decision Heuristics}, ymode = log,
   xmin=0, xmax=70, 
@@
  \addplot table[x=benchmark,y=time] {plotdata/berkmin.dat};
- %\addplot table[x=benchmark,y=time] {plotdata/random.dat};
- \legend{Berkmin}
- %\legend{Random}
- %\addplot [domain=.1:2500] {x};
+ \addplot table[x=benchmark,y=time] {plotdata/random.dat};
+ \addlegendentry{Berkmin}
+ \addlegendentry{Random}
+ %\addplot [domain=.1:2500] {};
  \end{axis}

### 2016-10-19 learning.tex
notes from today's call

@@
 conflict node and iteratively finds new nodes that contradict \pscmt{we have to define what that means in the preliminaries} with the 
-transformer stored in $reasons[\bot]$.
+transformer stored in $\mathcal{R}[\bot]$.
 

### 2016-10-19 algorithm.tex
some fixes in algs

@@
 \SetKwData{safe}{safe}
+\SetKwData{sat}{SAT}
 \SetKwData{unsafe}{unsafe}
@@
 \Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
-$v \leftarrow \top$ \;
-$\mathcal{T} \leftarrow \epsilon$ \;
+$\mathcal{T} \leftarrow \langle\rangle$ \;
 $\mathcal{R} \leftarrow \emptyset$ \;
-$result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;
-\uIf{$result$ == \conflict} {
-  \return \safe \;
-}
+$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
+\lIf{$\mathit{result}$ = \conflict} {
+  \return \safe}
 \While{$true$} 
 {
-  \uIf{$v$ is $\gamma$-complete} {
-    \return \unsafe;
+\lIf{$\mathit{result}$ = \sat} {
+  \return \unsafe}
+  $d \leftarrow decide(\mathit{abs}(\mathcal{T}))$ \;
+%  $v \leftarrow v \meet d$ \; 
+  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
+  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{nil} \;
+  $\mathit{result} \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
+  \Do{$\mathit{result} = \conflict$} {
+%    learn() \;
+    \lIf{$\neg \mathit{conflict\_analysis}(S,\mathcal{T},\mathcal{R})$} {
+      \return \safe
     }
-  $d \leftarrow decide(v)$ \;
-  $v \leftarrow v \meet d$ \; 
-  $\mathcal{T} \leftarrow \mathcal{T} . d$ \; 
-  $\mathcal{R}[|\mathcal{T}|] \leftarrow nil$ \;
-  $result \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
-  \Do{$result == \conflict$} {
-    learn() \;
-    \uIf{$\neg$ conflict\_analysis()} {
-      return \safe \;  
-    }
-    result $\leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$ \;
+    $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
   }
@@
 reason trail maps elements of propagation trail to transformers that were 
-used to derive them. \pscmt{define operator to turn trail into abstract value $\bigsqcap\mathcal{T}=\bigsqcap_{m\in\mathcal{T}}m$}
+used to derive them.
+We convert the propagation trail into an abstract value as follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m\text{ in }\mathcal{T}}m$ with $\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.
 
@@
 $v$, similar to the BCP step in SAT solvers.  If the result of $deduce$ 
-returns {\em CONFLICT} \pscmt{don't use capitals} ($\perp$), the algorithm terminates with {\em safe}. 
+returns \textsf{CONFLICT} ($\perp$), the algorithm terminates with \textsf{safe}. 
 Else, the analysis enters into the while($true$) loop and makes a new decision 
@@
 alternates between the decision and deduction until the abstract value $v$ 
-precisely \pscmt{not true} represents the set of models or the deduction leads to a {\em CONFLICT}.  
+precisely \pscmt{not true} represents the set of models or the deduction leads to a \textsf{CONFLICT}.  
 Checking whether the abstract model $v$ concretises to a set of models of $\varphi$; this corresponds to a 
@@
 current abstraction is insufficient to determine the satisfiability of $\varphi$.  
-However, in case of {\em CONFLICT}, the algorithm enters in the 
-$conflict\_analysis()$ phase to learn the reason for {\em conflict}.  
+However, in case of \textsf{CONFLICT}, the algorithm enters in the 
+$conflict\_analysis()$ phase to learn the reason for the conflict.  
 There can be multiple incomparable reasons for conflict -- based on 
@@
 abstractly satisfy the formula \pscmt{define}.  If no further backtrack is possible, 
-then the algorithm terminates and $\varphi$ is {\em SAFE}. Else, the 
+then the algorithm terminates and $\varphi$ is \textsf{safe}. Else, the 
 algorithm makes a new decision and the above process is repeated 
 until a real counterexample \pscmt{is this a model?} is obtained or the algorithm backtracks to 
-decision level 0 after a conflict in which case it returns {\em SAFE}.
+decision level 0 after a conflict in which case it returns \textsf{safe}.
 
@@
 empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
-$m_i \in \mathcal{T}$.  A solver is in {\em conflict} if some clauses in
+$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
 $\mathcal{S}$ is not abstractly satisfied by $v$.

### 2016-10-19 deductions.tex
some fixes in algs

@@
 \begin{small}
-\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
-abstract valuation $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
-\Output{The status (\sat or \conflict or \unknown)}
+\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
+\Output{\sat or \conflict or \unknown}
 worklist $\leftarrow$ initialize\_worklist($S$) \;
-\Loop{}
-{
-// Greatest fixed point \;
 \While{$!worklist.empty()$} 
 {
-  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
-  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
-  \uIf{$meet\_irrd == \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \;
-    worklist.delete() \;
+  $(s,\mathit{live\_vars}) \leftarrow$ worklist.pop() \;
+  $v \leftarrow ded(s, \mathit{espan}(\mathit{live\_vars}), \mathit{abs}(\mathcal{T}))$\;
+  \uIf{$v = \bot$} {
+    $\mathcal{R}[\bot] \leftarrow s$ \pscmt{how is this used in the conflict analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}\;
+    worklist.clear() \pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...} \;
     \return \conflict \;
@@
   {
-    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
-    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
+    $\mathcal{T} \leftarrow \mathcal{T} \cdot \mathit{decomp}(v)$ \; 
+%    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
     $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    worklist.update(v, s) \; 
+    worklist.update($v$) \; 
   }
 }
-\uIf{$v$ is $\gamma$-complete} {
+\lIf{$\mathit{abs}(\mathcal{T})$ is $\gamma$-complete} {
   \return \sat;
 }
-\uElse {
+\lElse {
  \return \unknown \;
-}
 }

### 2016-10-19 decision.tex
decision

@@
 A decision $\mathcal{D}$ is a meet irreducible that refines the 
-current abstract valuation.  A decision can be of the form \pscmt{that's too restrictive?}
-$\langle x R c \rangle$ or $\langle x R y \rangle$, where 
-$R \in \{\preceq, \succeq\}$, $c$ is the bound and 
-$\{x, y\}$ are program variables.  A decision must always 
-be consistent with respect to the trail $\mathcal{T}$, that 
-is, it must not contradict with the elements in the trail.
+current abstract valuation.  For example, a decision in interval 
+domain can be of the form $\langle x R c \rangle$ where 
+$R \in \{\leq, \geq\}$, $c$ is the bound.  A decision in 
+relational domain such as octagon is of the form $ax - by \leq c$, 
+where $x$ and $y$ are variables, $a,b \in \{-1,0,1\}$ are co-efficients, 
+and $c \in \mathbb{R}\cup\{\infty\}$ is the bound of the inequality.    
+A decision must always be consistent with respect to the trail 
+$\mathcal{T}$, that is, it must not contradict with the elements 
+in the trail.
 

### 2016-10-19 learning.tex
conflict analysis

@@
 A conflict analysis in a SAT solver can be seen as abductive 
-reasoning where a singleton assignment $c$ is replaced by a partial 
-assignment that is sufficient to infer $c$ \pscmt{???}.  Thus, a conflict 
+reasoning which underapproximate a set of models that do not satisfy a 
+formula~\cite{sas12,dhk2013-popl}.  
+\Omit {
+where a singleton assignment $c$ is replaced by a partial 
+assignment that is sufficient to infer $c$.  Thus, a conflict 
 analysis is a generalization of an under-approximation of a set of 
-countermodels~\cite{sas12,dhk2013-popl}, where a countermodel is a 
-set of structures that do not satisfy a formula \pscmt{this explanation does not help the reader, who will mix up counterexamples and countermodels}.  Since relational 
-abstract domains have more complex structure than a partial 
-assignments domain, so the lifting of {\em first-UIP} algorithm~\cite{uip} 
-and clause learning to any arbitrary lattice structure is a non-trivial task.
+countermodels, where a countermodel is a 
+set of structures that do not satisfy a formula 
+}
+Since relational abstract domains have more complex structure than a partial 
+assignments domain or interval domain, so the generalization of {\em first-UIP} algorithm~\cite{uip} and clause learning in a relational lattice structure is non-trivial.
 

### 2016-10-20 example.tex
redundant comment removed
comments on example

@@
 \caption{\label{fig:example}
-C Program, Control-flow Graph (CFG) and Abstract Conflict Graph \pscmt{I do not understand the arrow from n2 to n3 on the right hand side}}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph \pscmt{val $\rightarrow$ v; better rename $I$ to $n_0$ (also in the text); arrow from $n_2$ to $n_4$ on the rhs}}
 \end{figure}
@@
 
-Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis. \pscmt{Do we want to use the product domain right away to be consistent?}  
+Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
 ACDCL implements several decision heuritics.  In this case, the 
@@
 partitioning to generate proofs using decision and clause learning.
-
-\pscmt{clearly say that everything that follows is for finite
-  unwindings of programs}

### 2016-10-20 domains.tex
some fixes in algs
preps for harmonising 4

@@
 \end{tabular}
-\caption{SSA and abstract conflict graph \pscmt{Where do you refer to this graph?}}
+\caption{SSA and abstract conflict graph}
 \label{ssa}
@@
 %
-We consider programs with bounded loops 
-%(respectively bounded unwindings of programs with loops or recursion) with 
-containing safety properties given as assertions in the code.
+In this paper we consider safety checking of loop-free programs, including
+bounded unwindings of programs with loops or recursion.  We assume
+that safety properties given as assertions $\mathcal{A}$ in the code.
 %
-Such a program is represented as set $\mathcal{P}$ of
-\emph{statements} $s$.  These are constraints as typically obtained by
-translating the program into single static assignment (SSA) form via a
-data flow analysis.
+Such a program is represented as set
+$\mathcal{S}=\mathcal{P}\cup\{\neg \bigwedge_{a\in\mathcal{A}} a\}$
+where $\mathcal{P}$ are the statements of the program in the form of
+constraints as typically obtained by translating the program into
+single static assignment (SSA) form via a data flow analysis.
 %
@@
 %
-Assertions $\mathcal{A}$ are constraints $a$ of the form $g_3
-\Rightarrow z\geq 0$, meaning that if $g_3$ holds (i.e.\ the assertion
-is reachable) then the assertion must hold.
+The assertions $\mathcal{A}$ are constraints of the form $g_3
+\Rightarrow z\geq 0$, for instance, meaning that if $g_3$ holds
+(i.e.\ the assertion is reachable) then the assertion must hold.
 %
@@
 %-------------------------------------------------------------------------------
-\paragraph{Safety problem}  
+%\paragraph{Safety problem}  
 %
-Based on above program representation, the safety problem solved by the ACDCL solver is the following:
-%
-\begin{equation}\label{eq:safety}
-\varphi:= \bigwedge_{s\in\mathcal{P}} s \wedge \neg \bigwedge_{a\in\mathcal{A}} a
-\text{ is unsatisfiable if and only if the program is safe.}
-\end{equation}
+Based on the above program representation, we define $\varphi:=
+\bigwedge_{s\in\mathcal{S}} s$. The formula $\varphi$ is unsatisfiable
+if and only if the program is safe.
 

### 2016-10-20 example.tex
fixed code in Fig. 1

@@
 \begin{lstlisting}[mathescape=true,language=C]
-void main() {
- int x,z;
- int v;
+void foo(int v, 
+         _Bool c) 
+{
   assume(v>=0 $\wedge$ v<=N);
- _Bool c;
+  int x; 
   if(c)
@@
     x = -v;
- z = x * x;
+  int z = x * x;
   assert(z>=0);
@@
 multiplication of two machine integers and checks that the result of the 
-multiplication is always positive for an integer input $x$.  To avoid 
-overflow error, we assume $x$ has an upper bound $N$.  The control 
+multiplication is always positive for an integer input $v$.  To avoid 
+overflow error, we assume $v$ has an upper bound $N$.  The control 
 flow graph is shown in the middle column of Figure~\ref{fig:example}.

### 2016-10-20 contributions.tex
example resolve

@@
 
-\item  Experiments indicate a \rmcmt{N}X reduction in the number of 
-decisions, propagations, learning and backtracking iterations compared to a 
-SAT solver while being more precise compared to an abstract interpreter.  
+\item  Experiments indicate two order of magnitude reduction 
+in the number of decisions, propagations, and conflicts compared 
+to a SAT solver while being more precise compared to an abstract interpreter.  
 \Omit {

### 2016-10-20 example.tex
example resolve

--- 2016-10-20 example.tex
@@
 \caption{\label{fig:example}
-C Program, Control-flow Graph (CFG) and Abstract Conflict Graph \pscmt{val $\rightarrow$ v; better rename $I$ to $n_0$ (also in the text); arrow from $n_2$ to $n_4$ on the rhs}}
+C Program, Control-flow Graph (CFG) and Abstract Conflict Graph}
 \end{figure}
@@
 to determine safety.  However, a relational domain such as octagons is 
-able to prove the program without any partition.  In general, the 
-imprecision is either intended by the tool because such high 
+able to prove the program without any partition.  
+\Omit {
+In general, the imprecision is either intended by the tool because such high 
 precision analysis is normally not required for runtime error 
@@
 of the application under analysis.  
+}
 %
@@
 Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
-(\(I: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
-at node $I$.  The analysis discards all interval constraints that
+(\(n0: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
+at node $n0$.  The analysis discards all interval constraints that
 lead to the conflict and backtrack to decision level 0.  ACDCL then 

### 2016-10-20 learning.tex
section 6

@@
 Since relational abstract domains have more complex structure than a partial 
-assignments domain or interval domain, so the generalization of {\em first-UIP} algorithm~\cite{uip} and clause learning in a relational lattice structure is non-trivial.
+assignments domain or interval domain, so finding a generalized reason for 
+conflict through UIP computation~\cite{uip} is non-trivial.
 

### 2016-10-20 contributions.tex
learning text

@@
 generalizes the Boolean Constraint Propagation (BCP) procedure to
-compute a fixed point over a relational domain.  Our abstract 
-conflict analysis algorithm present a non-trivial lifting of the 
-Unique Implication Point (UIP) procedure to relational abstract 
-domains in order to learn a generalized reason for conflicts.
+compute a fixed point over a relational domain.
+
+\item We propose several domain specific decision heuristics, that 
+exploits the expressivity of the underlying abstract domain to make 
+effective decisions.
+
+\item We present a non-trivial lifting of the conflict analysis procedure 
+to relational abstract domains in order to learn a generalized reason 
+for conflicts.
 

### 2016-10-20 learning.tex
figure 4
learning text

@@
 For example, consider a formula 
-$\varphi:= (x=5 \wedge y=x \wedge y=y+x \wedge y \leq 0)$ with current 
+$\varphi:= (x=5 \; \wedge \; y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
 abstract valuation $v:= (x=5)$.  Figure~\ref{conflict} shows a 
@@
 any path from the last decision node to the conflict node must pass 
-through it.  Unlike a SAT solver, a UIP in ACDCL may have meet irreducibles 
-that involve the same variables.  The first-UIP is a node closest \pscmt{unique?} to the 
-conflict, and the last UIP is the decision node itself.  Computing UIPs ensures 
-asserting cuts, that is, it yields clauses that generates new deduction after 
-backtracking.  Every cut in the graph reason for conflict that can be used in learning.  
-A first-UIP algorithm traverses the trail $\mathcal{T}$ starting from the 
-conflict node and iteratively finds new nodes that contradict \pscmt{we have to define what that means in the preliminaries} with the 
-transformer stored in $\mathcal{R}[\bot]$.
-
+through it.  Unlike a SAT solver, a UIP for relational domain may 
+involve meet irreducibles that contains the same variables.  The 
+first-UIP is a unique node closest to the conflict, and the last UIP 
+is the decision node itself.  Computing UIPs ensures asserting
+cuts~\cite{cdcl,DBLP:journals/fmsd/BrainDGHK14}, that is, it 
+yields clauses that generates new deduction after backtracking.  
+Every cut in the graph is a reason for conflict that can be 
+used in learning.  An abstract first-UIP algorithm~\cite{DBLP:journals/fmsd/BrainDGHK14} 
+traverses the trail $\mathcal{T}$ starting from the conflict node and 
+computes a cut that suffices to produce a conflict. 
+\Omit {
+contradict with the transformer stored in $\mathcal{R}[\bot]$.  
+}
 For our example, there exists multiple incomaparable reasons for conflict,
 marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
-Choosing cut0 yields a learnt clause \pscmt{use $\leq$ instead of $<$} 
-$L_0: (x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
+Choosing cut0 yields a learnt clause 
+$(x+y<15 \vee -x-y<-15 \vee -x+y<5 \vee x-y<-5 \vee y-10<0 \vee -y+10<0)$, 
 which is obtained by negating the reason for conflict.  
@@
 \scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
-\caption{\label{conflict}
-\pscmt{push constants to the lhs of the inequalities; labels cut0, cut1 not fully visible}
-}
+\caption{\label{conflict} Abstract Conflict Graph}
 \end{figure} 
@@
 \paragraph{\textbf{Learning in Relational Lattice}}
-We model learning as a meet irreducible deduction transformer \pscmt{???}, called 
-{\em abstract unit transformer (AUnit)} \pscmt{why the name? seems confusing as we are talking about learning and not about deduction}.  For an 
-abstract lattice $A$ with complementable meet irreducibles 
-and a set of conflicting meet irreducibles (countermodels) \pscmt{The term countermodels for $C$ is misleading since $\bigsqcap C$ is a set of countermodels, but the elements of $C$ individually are no countermodels.}
-$C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
-satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is defined as follows. 
-\pscmt{explain which cases are different in non-relational domains}
+Learning in propositional solver yields a asserting clause~\cite{cdcl} 
+that express the negation of the conflict reasons.  However, we model 
+learning in relational domains as learning a transformer which infers 
+new meet irreducibles using the abstract unit rule called 
+{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 
+transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
+a generalization of propositional unit rule for numerical domains.  For 
+an abstract lattice $A$ with complementable meet irreducibles and a set 
+of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
+satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 
 \[ AUnit(a) =
  \begin{cases}
-  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \\
+  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \qquad \qquad \qquad
+  \qquad \qquad \qquad \quad (1)\\
   \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
-  \setminus \{t\}. a  \sqsubseteq t' \\
+  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad (2) \\
   \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
@@
              & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
-             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a' \not\sqsubseteq t  \\
-  \top & \quad \text{otherwise} \\
+             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
+             \not\sqsubseteq t   \qquad \qquad \qquad \qquad (3) \\
+  \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
  \end{cases}
 \]
-For example, let $C = \langle x+z \geq 10,  k \geq 0 \rangle$ \pscmt{why suddenly a different notation for $k$?
-The variable name $k$ might be confusing because it is usually used for counting or indices} be the conflicting set
-of meet irreducibles of $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
-x+z \leq 15, k:[1,1] \rangle$.  Applying the third rule of $AUnit_C$, we get a 
+
+\pscmt{explain which cases are different in non-relational domains}
+For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
+be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
+x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying the rule (3) of $AUnit_C$, we get a 
 decomposition of $a$ into $a'$ and $a''$ where 
 $a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
-$a''= \langle k:[1,1] \rangle \sqsubseteq (k \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
+$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
 which constrains the bound of the octagonal constraint.  
 Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
-and let $a = \langle x:[3,4], y:[5,6] \rangle$, then $AUnit_C(a) = \bot$ using
-the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x:[3,4] \rangle$, 
+and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
+the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
 then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$

### 2016-10-20 algorithm.tex
comments on example
learning

@@
   $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{nil} \;
-  $\mathit{result} \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
+  $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$\;
   \Do{$\mathit{result} = \conflict$} {
@@
 reason trail maps elements of propagation trail to transformers that were 
-used to derive them.
-We convert the propagation trail into an abstract value as follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m\text{ in }\mathcal{T}}m$ with $\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.
+used to derive them. A propagation trail is mapped to an abstract value as 
+follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with 
+$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.
 
@@
 
-\pscmt{define ``consistent with trail''}
+%\pscmt{define ``consistent with trail''}

### 2016-10-20 learning.tex
learning

@@
   \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
-  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad (2) \\
+  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad \; \; (2) \\
   \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
@@
              & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
-             \not\sqsubseteq t   \qquad \qquad \qquad \qquad (3) \\
+             \not\sqsubseteq t   \qquad \qquad \qquad \qquad \; \; (3) \\
   \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
@@
 \]
-
-\pscmt{explain which cases are different in non-relational domains}
-For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
+$AUnit$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
+(3) of $AUnit$ infers a valid meet irreducible, which implies that $C$ is unit. 
+However, rule (3) is domain-specific and only applies to numerical abstract
+domains.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
 be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
-x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying the rule (3) of $AUnit_C$, we get a 
+x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $AUnit_C$, we get a 
 decomposition of $a$ into $a'$ and $a''$ where 
@@
 and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
-the first rule, since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
-then $AUnit_C(a') = (y > 7)$ using second rule, since $a' \sqsubseteq (x \geq 2)$
-and $a' \sqsubseteq (x \leq 5)$.
+the rule (1), since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
+then $AUnit_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
+and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $AUnit$ 
+returns $\top$.
 
@@
 \[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
-(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is
-consistent} \]   
-\pscmt{What is $\mathcal{E}$? What is $m$? What is the notation $\mathcal{E}_1(m,s)\mathcal{E}_2$? What does consistent mean?}
-\pscmt{Is there any difference between non-relational and relational?}
+(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict}} \]
 

### 2016-10-20 algorithm.tex
minor

--- 2016-10-19 algorithm.tex
@@
 \begin{small}
-\Input{Set $S$ of abstract deduction transformer, 
-$\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract domain $A$}
-\Output{The status (\safe or \unsafe) and a counterexample if \unsafe}
+\Input{%An abstract domain $A$, 
+A program $\mathcal{S}$, a
+  propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
+and a conflict analysis heuristics $\mathit{conflict\_analysis}$.}
+\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
 $\mathcal{T} \leftarrow \langle\rangle$ \;
 $\mathcal{R} \leftarrow \emptyset$ \;
-$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
+$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R}, H)$ \;
 \lIf{$\mathit{result}$ = \conflict} {
@@
   $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
-  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{nil} \;
-  $\mathit{result} \leftarrow \deduce(S,v,\mathcal{T},\mathcal{R})$\;
+  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{decision} \;
+  $\mathit{result} \leftarrow \deduce(\mathcal{S},\mathcal{T},\mathcal{R})$\;
   \Do{$\mathit{result} = \conflict$} {
 %    learn() \;
-    \lIf{$\neg \mathit{conflict\_analysis}(S,\mathcal{T},\mathcal{R})$} {
+    \lIf{$\neg \mathit{conflict\_analysis}(\mathcal{S},\mathcal{T},\mathcal{R})$} {
       \return \safe
     }
-    $\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R})$ \;
+    $\mathit{result} \leftarrow \deduce(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
   }
@@
 \end{small}
-\caption{Abstract Conflict Driven Clause Learning $\langle ACDCL(S) \rangle$ \label{Alg:acdcl}}
+\caption{Abstract Conflict Driven Clause Learning $ACDCL(A)$ \label{Alg:acdcl}}
 \end{algorithm2e}

### 2016-10-20 example.tex
minor

@@
 {
-  assume(v>=0 $\wedge$ v<=N);
+  assume(v>=0 && v<=N);
   int x; 

### 2016-10-20 domains.tex
fixed code in Fig. 1
domain

@@
-\section{Preliminaries}
+\section{Program Model and Abstract Domain}
 This section introduces the program model and identifies properties of abstract
@@
 $\begin{array}{rcl}
-\mathcal{P}:= (g_0 = \true) \wedge (c = 0 \leq v\leq N) \\
-(g_1 = g_0 \wedge c) \wedge (x_0 = v) \wedge \\
+\mathcal{P}:= (g_0 = (0 \leq v\leq N)) \\
+(g_1 = (g_0 \wedge c)) \wedge (x_0 = v) \wedge \\
 (x_1 = -v) \wedge (x_2 = g_1?x_0:x_1) \wedge \\
-(g_2 = g_1 \vee g_0\wedge \neg c) \wedge \\
+(g_2 = (g_1 \vee g_0\wedge \neg c)) \wedge \\
 (z  = x_2 * x_2) \wedge (z<0 \wedge g_2)
@@
 %
-In this paper we consider safety checking of loop-free programs, including
+In this paper, we consider safety checking of loop-free programs, including
 bounded unwindings of programs with loops or recursion.  We assume
@@
 %
-Based on the above program representation, we define $\varphi:=
-\bigwedge_{s\in\mathcal{S}} s$. The formula $\varphi$ is unsatisfiable
-if and only if the program is safe.
+Based on the above program representation, we define a safety equation as, 
+$\varphi:= \bigwedge_{s\in\mathcal{S}} s$. The formula $\varphi$ is 
+unsatisfiable if and only if the program is safe.
 
@@
 meet decomposition as explained above.
+
+\begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
+mapping of variables to an element of abstract domain. For example, 
+\Omit {
+a mapping of variable $x$ to an interval environment is given by 
+$\langle x \mapsto [2,5] \rangle$ or a 
+}
+a mapping of constraint variables $\{x,y\}$ to octagon environment 
+is given by $\langle x-y \mapsto 0, y-x \mapsto 0 \rangle$.  
+An abstract valuation is {\em atomic} if each variable is mapped to a singleton 
+value or to $\bot$.  An abstract valuation $(v)$ abstractly satisfies a formula 
+$\varphi$ if for every variable $\mathcal{x}$ in $\varphi$, there is a concrete solution 
+$(c)$ such that $c(\mathcal{x}) \subseteq \gamma \circ v(\mathcal{x})$ holds. 
+\end{definition}
+
 %

### 2016-10-20 learning.tex
minor
domain

@@
 \[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
-(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict}} \]
+(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict} \]
 

### 2016-10-20 deductions.tex
comments on deduction algorithm

@@
   \uIf{$v = \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \pscmt{how is this used in the conflict analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}\;
-    worklist.clear() \pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...} \;
+    $\mathcal{R}[\bot] \leftarrow s$ \pscmt{how is this used in the conflict
+    analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}
+    \rmcmt{UIP uses the transformer stored at $\mathcal{R}[\bot]$ to derive the
+    cut. No, we dont want to make the trail inconsistent by adding $\bot$}\;
+    worklist.clear() \pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...}
+    \rmcmt{When a bottom is deduced, the direction of search
+    changes--so the worklist is populated with new set of statement and must be
+    flushed out} \;
     \return \conflict \;

### 2016-10-20 domains.tex
comments on deduction algorithm

@@
 value or to $\bot$.  An abstract valuation $(v)$ abstractly satisfies a formula 
-$\varphi$ if for every variable $\mathcal{x}$ in $\varphi$, there is a concrete solution 
-$(c)$ such that $c(\mathcal{x}) \subseteq \gamma \circ v(\mathcal{x})$ holds. 
+$\varphi$ if for every variable $x$ in $\varphi$, there is a concrete solution 
+$(c)$ such that $c(x) \subseteq \gamma \circ v(x)$ holds. 
 \end{definition}

### 2016-10-20 appendix.tex
appendix

--- 2016-10-19 deductions.tex
@@
-\section{Abstract Model Search}
-%
-\Omit {
-\begin{figure}[t]
-\scriptsize
-\begin{tabular}{l|l|l}
-\hline
-C program & SSA & Octagon \\
-\hline
-\begin{lstlisting}[mathescape=true,language=C]
-int main() {
- unsigned x, y;
- __CPROVER_assume(x==y);
- x++;
- assert(x==y+1);
-}
-\end{lstlisting}
-&
-\begin{minipage}{4.40cm}
-$\begin{array}{l@{\,\,}c@{\,\,}l}
-SSA &\iff& ((g0 == TRUE) \land \\
-    &    & (cond == (x == y)) \land \\
-    &    & (g1 == (cond \&\& guard0)) \land \\
-    &    & (x' == 1u + x) \land \\
-    &    & (x' == 1u + y || !1))
-\end{array}$
-\end{minipage}
-&
-\begin{minipage}{3.75cm}
-$\begin{array}{l@{\,\,}c@{\,\,}l}
-C &\iff& ((x' > 1) \land (-x'-y < -2) \land \\
-  &    & (-x-x' < -2) \land (y-x' < 0) \land \\                                                                
-  &    & (x-x' < 0) \land (y > 0) \land \\
-  &    & (x > 0) \land (-x'-y < 0) \land \\
-  &    & (x+y > 1) \land (y-x < 1) \land \\
-  &    & (x'-y < 2) \land (x-y < 1) \land \\
-  &    & (x+y > 0) \land (x+x' > 0) \land \\
-  &    & (x'-x < 2))
-\end{array}$
-\end{minipage}
-\\
-\hline
-\end{tabular}
-\caption{C Program, corresponding SSA and Octagonal Inequalities}
-\label{ssa}
-\end{figure}
-}
-%
-%    
-\begin{algorithm2e}[t]
-\DontPrintSemicolon
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
-\SetKwData{sat}{SAT}
-\SetKwData{conflict}{CONFLICT}
-\SetKwData{unsat}{UNSAT}
-\SetKwData{unknown}{UNKNOWN}
-\SetKwData{true}{true}
-\SetKwInOut{Input}{input}
-\SetKwInOut{Output}{output}
-\SetKwFor{Loop}{Loop}{}{}
-\SetKw{KwNot}{not}
-\begin{small}
-\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$,
-abstract valuation $v$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
-\Output{The status (\sat or \conflict or \unknown)}
-worklist $\leftarrow$ initialize\_worklist($S$) \;
-\Loop{}
-{
-// Greatest fixed point \;
-\While{$!worklist.empty()$} 
-{
-  $\langle live\_vars, s \rangle \leftarrow$ worklist.pop() \;
-  $\mathit{meet\_irrd} \leftarrow$ domain(s, live\_vars, $v$)\;
-  \uIf{$meet\_irrd == \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \;
-    worklist.delete() \;
-    \return \conflict \;
-  }
-  \uElse
-  {
-    $\mathcal{T} \leftarrow \mathcal{T} . \mathit{meet\_irrd}$ \; 
-    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
-    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    worklist.update(v, s) \; 
-  }
-}
-\uIf{$v$ is $\gamma$-complete} {
-  \return \sat;
-}
-\uElse {
- \return \unknown \;
-}
-}
-\end{small}
-\caption{Abstract Model Search $\langle deduce(S,v,\mathcal{T},\mathcal{R}) \rangle$ \label{Alg:ms}}
-\end{algorithm2e}
-%  
-A model search procedure in a SAT solver involves two steps -- {\em deductions} 
-using the unit rule refines current partial assignments and 
-{\em decisions} to heuristically guess a value for an unassigned 
-literal.  The unit rule overapproximates a model transformer and deduction 
-computes a greatest fixed point over the partial assignments
-domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
-that computes a greatest fixed point over meet irreducible deduction 
-transformer in $S$ \pscmt{???}.  
-%
-
-%-------------------------------------------------------------------------------
-\subsection{Abstract Transformers} \label{sec:abst}
-%-------------------------------------------------------------------------------
-
-To make our algorithm efficient, we have to focus abstract
-transformers on performing only the minimally necessary
-work. 
-%
-We thus define a specialised variant of the transformer to compute
-deductions w.r.t.\ a given subdomain $L\subseteq A$; $\bot,\top \in L$.
-\[\llbracket s \rrbracket^\sharp(a,L)=\bigsqcup \{a'\mid a'\in L, a' \sqsubseteq a, a'\models s\}\]
-We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
-%
-For example, for
-$a=(0\leq y \leq 1 \wedge 5\leq z)$ we have
-$\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
-$\llbracket x=y+z\rrbracket^\sharp(a,\{y,z\})=\top$.
-%
-If $L=A$ then intuitively $a$ is intersected with the constraint $s$.
-Yet, with the help of $L$ we can also control whether to perform
-forward (right-hand side to left-hand side, $L=\{x\}$ in the above
-example) or backward (lhs to rhs, $L=\{y,z\}$) propagation.
-
-%Moreover, we want to know what the reasons for a specific deduction are.
-%\[ded(s,a,L)=\{R\rightarrow d\mid R=reasons(s,a,L,d), d \in decomp(\llbracket s \rrbracket^\sharp(a,L))\}\]
-%where 
-%$reasons(s,a,L,d)=\text{argmin}_{R\in\{R'\mid R'\subseteq decomp(a), \llbracket s\rrbracket^\sharp(R,L)=d\}} |R| $.
-
-\pscmt{give an example for $ded$, link it to the reason trail}
-
-\pscmt{define $ded$ for all statements, connect this to worklist}
-
-\pscmt{link livevars to $L$ -- the subdomain spanned by livevars}
-
-%-------------------------------------------------------------------------------
-\subsection{Algorithm for Deduction Phase}
-%-------------------------------------------------------------------------------
-%
-\rmcmt{Talk about lazy closure operation}
+\appendix
+\section{Omitted proofs}
 
@@
 
-Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.
-A {\em worklist} is initialized following a heuristics choice for iteration 
-strategies. A {\em forward} iteration strategy initializes the worklist with 
-transformers that has constants in the right-hand side. Whereas, a {\em backward} 
-iteration strategy starts with a set of assertions. On the other hand, a 
-{\em multi-way} iteration strategy initializes the worklist with set of all 
-transformers.  For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
-passed to the abstract domain, if the domain infers a new meet irreducible 
-$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$ \pscmt{It usually infers several ones.}.  Additionally, the reason 
-trail $\mathcal{R}$ \pscmt{define} is updated with the transformer $s$ as the reason for inferring
-$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with CONFLICT.  However,
-when a fixed-point is reached, the abstract valuation $v$ is checked for set of
-models.  This is determined by checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
-abstract valuation $v$ is abstractly satisfying, the abstract model 
-search terminates with SAT.  Otherwise, the algorithm returns UNKNOWN and the 
-model search procedure makes a new decision.
+\section{Detailed experimental results}

### 2016-10-20 deductions.tex
appendix

--- 2016-10-19 deductions.tex
@@
 \end{theorem}
-Proof sketch:
-When computing $ded$, we obtain a deduction that affects variables $V$.
-This has two consequences: (1) we have to compute the closure; (2) we have to propagate through all statements $s$ affected by the changes to $V$. 
-We do only (2): $espan$ augments the set of affected variables by one transitivity step that we would have to compute in the closure.
-$ded$ on this subdomain hence performs the deduction w.r.t.
-$a$ and $s$ plus one transitivity step of the closure.
-Thus, by induction, we eventually compute the entire closure just by doing (2).
-
-What are the consequences regarding complexity?
-Assuming we require $m$ iterations to reach the fixed point with eager closure; we have to perform $|A|^c|S|m$ (idealised) computation steps to do that (where $c=3$ for octagons).
-With the lazy closure we have to do $|L|^c|S|mk$ computations taking into account that we need $k$ times more iterations to converge.
-Hence, the approach pays off if $k<<\frac{|A|^c}{|L|^c}$, which seems likely for $c=3$.
-
+Proof sketch: see Appendix
 }

### 2016-10-20 decision.tex
decision phase

@@
 and $c \in \mathbb{R}\cup\{\infty\}$ is the bound of the inequality.  
-A decision must always be consistent with respect to the trail 
-$\mathcal{T}$, that is, it must not contradict with the elements 
+That is, a decision can specify relation between variables, for example 
+$\{x \leq y)$.  A decision must always be consistent with respect to 
+the trail $\mathcal{T}$, that is, it must not contradict with the elements 
 in the trail.
 
-Let $\mathcal{V}$ be a set of all singletons and non-singletons 
-variables.  A singleton variable is one that has been assigned a 
-singleton value, for example, $\langle x:[1,1] \rangle$.  Otherwise, 
-the variable is non-singleton \pscmt{this definition only works for non-relational domains}.  Given a set of variables $\mathcal{V}$, 
-a decision phase heuristically chooses a non-singleton branching 
-variable $v \in \mathcal{V}$, a bound $c$, and the polarity ($\succeq$ or 
-$\preceq$).  A decision changes the sate of the solver by adding new element 
+A singleton meet irreducible $s$ in an abstract domain $A$ is an element 
+whose lower bound and upper bound are the same.  For example, 
+$\langle x:[1,1] \rangle$ is a singleton meet irreducible in interval domain.  
+For an octagon domain, $\langle 1 \geq x-y \leq 1 \rangle$ is a singleton meet 
+irreducible for the template $(x-y)$, though the concretisation of individual 
+template variables ($x$, $y$) may not be singletons.
+
+Given a set of variables $\mathcal{V}$ and a current abstract value $v$, 
+a decision phase heuristically returns a non-singleton meet irreducible 
+by choosing a set of branching variables $\{B\} \subseteq \mathcal{V}$, 
+a bound $c$, and a polarity ($\leq$ or $\geq$).  For interval domain, $|B|=1$.  
+A decision changes the sate of the solver by adding a new meet irreducible 
 $m$ and labelling information $s=decision$ to the trail, which is described below. 
 \[decide: \quad (\mathcal{E},S) \rightarrow (\mathcal{E}(m,s),S) \]
+
+%Let $\mathcal{V}$ be a set of all singletons and non-singletons.  
+%Otherwise, the variable is non-singleton
 
@@
 The {\em ordered} decision heuristics creates an ordering among non-singleton 
-variables, thereby making decisions on conditional variables (variables that 
-appear in conditional branches) first before choosing numerical variables.  
-This variable ordering is path-sensitive as it gives an effect of trace
-partitioning.
+meet irreducibles, thereby making decisions on meet irreducibles that involve 
+conditional variables (variables that appear in conditional branches) first 
+before choosing meet irreducibles with numerical variables.  
+The ordering is path-sensitive as it gives an effect of trace partitioning.
 
 The {\em longest-range} heuristics simply keeps track of the
-interval $[l,u]$ range ($r=u-l$) of a variable and chooses a variable with 
-the longest range value \pscmt{why is this restricted to variables?}.  This ensures a fairness policy in selecting a 
-variable since it guarantees that the intervals of variables are uniformly 
-restricted.
-
-The {\em random} decision heuristics arbitrarily picks a variable 
-for making decision. \pscmt{uniformly distributed?}
-
-The {\em relational} decision heuristics is of the form 
-$x R y$ for $R \in \{preceq, succceq\}$ and is only relevant for relational 
+interval $[l,u]$ range ($r=u-l$) of a meet irreducible and chooses 
+a meet irreducible with the longest range value.  This ensures a 
+fairness policy in selecting a variable since it guarantees that 
+the intervals of meet irreducibles are uniformly restricted.  
+%
+The {\em random} decision heuristics arbitrarily picks a meet irreducible  
+for making decision. 
+%
+The {\em relational} decision heuristics is only relevant for relational 
 abstract domains.  
-
+%
 Whereas, a {\em berkmin} decision heuristics is inspired 
-from decision heuristics used in Berkmin~\cite{eugoldberg07} SAT solver, which 
-keeps track of the activity of a variable that participate in conflict clauses 
+from decision heuristic used in Berkmin~\cite{eugoldberg07} SAT solver.  
+The berkmin heuristic is currently implemented for interval constraints only.  
+The heuristic keeps track of the activity of an interval meet irreducible 
+that participate in conflict clauses. 
+\Omit {
 as well as variables that actively contribute to conflicts but do not explicitly 
-appear in conflict clauses.  The set of conflict clauses is organized 
-chronologically with the top clause as the one deduced in the last.  A 
-branching variable is chosen among the free variables whose literals are 
-in the top unsatisfied conflict clause.  A similar decision heuristics is 
-also implemented in Chaff~\cite{chaff} SAT solver, that computes the activity 
-of a variable as the number of occurrences of that variable in conflict 
-clauses only. \pscmt{how do you adapt this to relational domains?}
-
-A bound of a variable \pscmt{why restricted to variables?} is heuristically chosen to be an approximation of the 
-arithmetic average of the current bounds.  However, the polarity ($\preceq$ or
-$\succeq$) of a variable is chosen randomly.
+appear in conflict clauses.  The set of conflict clauses is 
+organized chronologically with the top clause 
+as the one deduced in the last.  A branching variable is chosen among the 
+free variables whose literals are in the top unsatisfied conflict clause.  
+A similar decision heuristics is also implemented in Chaff~\cite{chaff} SAT 
+solver, that computes the activity of a variable as the number of occurrences 
+of that variable in conflict clauses only. 
+}
+%
+A bound of a meet irreducible is heuristically chosen to be an approximation of the 
+arithmetic average of the current bounds.  However, the polarity ($\leq$ or $\geq$) 
+of a meet irreducible is chosen randomly.

### 2016-10-20 domains.tex
decision phase

@@
 In this paper, we consider safety checking of loop-free programs, including
-bounded unwindings of programs with loops or recursion.  We assume
-that safety properties given as assertions $\mathcal{A}$ in the code.
+bounded unwindings of programs with loops or recursion\rmcmt{why can't we simply 
+say we consider programs with bounded loops}.  We assume
+that safety properties are given as assertions $\mathcal{A}$ in the program.
 %

### 2016-10-20 deductions.tex
consistent naming and fonts in alg1 and 2

--- 2016-10-20 deductions.tex
@@
 \SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
+\SetKw{return}{return}
 \SetKwFunction{continue}{continue}
 \SetKwFunction{assign}{assign}
-\SetKwData{sat}{SAT}
-\SetKwData{conflict}{CONFLICT}
-\SetKwData{unsat}{UNSAT}
-\SetKwData{unknown}{UNKNOWN}
+\SetKwData{sat}{sat}
+\SetKwData{conflict}{conflict}
+\SetKwData{unsat}{unsat}
+\SetKwData{unknown}{unknown}
 \SetKwData{true}{true}
@@
 \begin{small}
-\Input{Set $S$ of $\widehat{ded_{\varphi}}$ for $\varphi$ over Abstract domain $A$, propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$}
+\Input{%Abstract domain $A$
+Program $\mathcal{S}$,
+propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$,
+propagation heuristics $H$}
 \Output{\sat or \conflict or \unknown}
-worklist $\leftarrow$ initialize\_worklist($S$) \;
-\While{$!worklist.empty()$} 
+$\mathit{worklist} \leftarrow H.\mathit{initialize\_worklist}(\mathcal{S})$ \;
+\While{$!\mathit{worklist.empty}()$} 
 {
-  $(s,\mathit{live\_vars}) \leftarrow$ worklist.pop() \;
+  $(s,\mathit{live\_vars}) \leftarrow \mathit{worklist.pop}()$ \;
   $v \leftarrow ded(s, \mathit{espan}(\mathit{live\_vars}), \mathit{abs}(\mathcal{T}))$\;
   \uIf{$v = \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \pscmt{how is this used in the conflict analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}\;
-    worklist.clear() \pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...} \;
+    $\mathcal{R}[\bot] \leftarrow s$ \;
+%\pscmt{how is this used in the conflict
+%    analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}
+%    \rmcmt{UIP uses the transformer stored at $\mathcal{R}[\bot]$ to derive the
+%    cut. No, we dont want to make the trail inconsistent by adding $\bot$}
+    $\mathit{worklist.clear}()$ \;
+%\pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...}
+%    \rmcmt{When a bottom is deduced, the direction of search
+%    changes--so the worklist is populated with new set of statement and must be
+%    flushed out} 
     \return \conflict \;
@@
     $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    worklist.update($v$) \; 
+    $H.\mathit{update\_worklist}(v, \mathit{worklist})$ \; 
   }
@@
 }
-\lElse {
  \return \unknown \;
-}
+
 \end{small}
-\caption{Abstract Model Search $\langle deduce(S,v,\mathcal{T},\mathcal{R}) \rangle$ \label{Alg:ms}}
+\caption{Abstract Model Search $\mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R},H)$ \label{Alg:ms}}
 \end{algorithm2e}
@@
 
-Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model search procedure.
+Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model
+search procedure. \pscmt{The propagation heuristics $H$ provides two
+  functions $\mathit{initialize\_worklist}$ and
+  $\mathit{update\_worklist}$. Describe them for the different
+  heuristics (already done for initialize, TBD for update...}
 A {\em worklist} is initialized following a heuristics choice for iteration 
@@
 $\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with CONFLICT.  However,
+is inferred by the domain, the algorithm terminates with \textsf{conflict}.  However,
 when a fixed-point is reached, the abstract valuation $v$ is checked for set of
@@
 abstract valuation $v$ is abstractly satisfying, the abstract model 
-search terminates with SAT.  Otherwise, the algorithm returns UNKNOWN and the 
+search terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
 model search procedure makes a new decision.

### 2016-10-20 experiments.tex
deduction

@@
 \paragraph {\textbf{Precision and Efficiency of Analysis}}
-Figure~\ref{fig:results} presents a comparison of the solver statistics
-(decision and learning) for the analysis using CBMC and ACDCL.  The cumulative 
-statistics corresponding to each phase of the algorithm for all 66 benchmarks 
-is presented in Table~\ref{result}.  The reasoning using ACDCL is at least 
-\rmcmt{1.5X} more efficient compared to propositional SAT in terms of number of decisions, propagations 
-and learning \pscmt{what learning???}.    
+Figure~\ref{fig:results} presents a comparison of the analysis 
+using CBMC and ACDCL.  The cumulative statistics corresponding to 
+each phase of the algorithm for all 66 benchmarks is presented in 
+Table~\ref{result}.  Experimental result shows a reduction of at least 
+two orders of magnitude in the number of decisions, propagations and conflicts 
+compared to analyis using CBMC.    
 %
@@
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis \pscmt{iteration/propagation strategy} and
-{\em chaotic} analysis in ACDCL \pscmt{what about backward?}.  The choice of analysis influences the 
-total number of decisions and clause learning iterations.  Hence, the 
-propagation strategy has a significant influence on the run time which 
-can be visualised in Figure~\ref{prop-dec}. \pscmt{Are the graphs up-to-date?} Unlike the forward analysis, 
-the chaotic or multi-way strategy performs both forward and backward 
-propagations.  This may take several iterations to reach the fixed-point, but 
-significantly reduces the total number of decisions and learning. \pscmt{Why?} 
+Figure~\ref{prop-dec} presents a comparison between {\em forward} propagation 
+strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of 
+propagation strategy influences the total number of decisions and clause 
+learning iterations.  Hence, the propagation strategy has a significant 
+influence on the run time which can be visualised in Figure~\ref{prop-dec}.  
+Unlike the forward propagation, the multi-way strategy may take several 
+iterations to reach the fixed-point, but it significantly reduces the 
+total number of decisions and conflicts.  This is attributed to the higher 
+precision of the meet irreducibles inferred by the multi-way strategy which 
+subsequently aids the decision heuristics to make better decisions. 
 \Omit {
@@
 \paragraph {\textbf{Learning}} Learning has a significant influence on
-the runtime of ACDCL.  We compare the first-UIP learning technique with 
-a DPLL-style conflict analysis without learning \pscmt{you learn the negation of the last decision}.  The effect of UIP computation 
-allows ACDCL to backtrack non-chronologically and guide the model search with 
-a learnt transformer.  However, the DPLL-style analysis exhibits
-case-enumeration behavior and thus immediately runs out of capacity for programs 
-which require relational domains. \pscmt{Is there a significant difference between intervals and octagons that indicates that for relational domains DPLL is even worse than in the propositional case?}
+the runtime of ACDCL.  We compare the UIP-based learning technique with 
+an analysis that performs chronoligical backtracking without learning. 
+The effect of UIP computation allows ACDCL to backtrack non-chronologically 
+and guide the model search with a learnt transformer.  However, the analysis 
+without learning exhibits case-enumeration behavior and thus immediately 
+runs out of capacity for most benchmarks.
 

### 2016-10-20 graphs.tex
deduction

@@
 			ylabel=ACDCL (decision), 
-			legend style={at={(0.8,0.15)},
-			anchor=north,legend columns=-1 },
+			legend pos = north west,
+			%legend style={at={(0.8,0.15)},
+			%anchor=north,legend columns=-1 },
 			]
@@
 			ylabel=ACDCL (Propagation),
-			legend style={at={(0.8,0.15)},
-			anchor=north,legend columns=-1 },
+			legend pos = north west,
+      %legend style={at={(0.8,0.15)},
+			%anchor=north,legend columns=-1 },
 			]

### 2016-10-20 algorithm.tex

@@
 \SetKwFunction{print}{print}
-\SetKwFunction{return}{return}
+\SetKw{return}{return}
 \SetKwFunction{nondet}{nondet}
@@
 %\SetKwFunction{isf}{isFeasible}
-\SetKwData{conflict}{CONFLICT}
+\SetKwData{conflict}{conflict}
 \SetKwData{safe}{safe}
-\SetKwData{sat}{SAT}
+\SetKwData{sat}{sat}
 \SetKwData{unsafe}{unsafe}
@@
 $\mathcal{R} \leftarrow \emptyset$ \;
-$\mathit{result} \leftarrow \deduce(S,\mathcal{T},\mathcal{R}, H)$ \;
+$\mathit{result} \leftarrow \mathit{deduce}(S,\mathcal{T},\mathcal{R}, H)$ \;
 \lIf{$\mathit{result}$ = \conflict} {
@@
   \return \unsafe}
-  $d \leftarrow decide(\mathit{abs}(\mathcal{T}))$ \;
+  $d \leftarrow \mathit{decide}(\mathit{abs}(\mathcal{T}))$ \;
 %  $v \leftarrow v \meet d$ \; 
@@
   $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{decision} \;
-  $\mathit{result} \leftarrow \deduce(\mathcal{S},\mathcal{T},\mathcal{R})$\;
+  $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$\;
   \Do{$\mathit{result} = \conflict$} {
@@
     }
-    $\mathit{result} \leftarrow \deduce(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
+    $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
   }
@@
 $v$, similar to the BCP step in SAT solvers.  If the result of $deduce$ 
-returns \textsf{CONFLICT} ($\perp$), the algorithm terminates with \textsf{safe}. 
+returns \textsf{conflict} ($\bot$), the algorithm terminates with \textsf{safe}. 
 Else, the analysis enters into the while($true$) loop and makes a new decision 
@@
 alternates between the decision and deduction until the abstract value $v$ 
-precisely \pscmt{not true} represents the set of models or the deduction leads to a \textsf{CONFLICT}.  
+precisely \pscmt{not true} represents the set of models or the deduction leads to a \textsf{conflict}.  
 Checking whether the abstract model $v$ concretises to a set of models of $\varphi$; this corresponds to a 
@@
 If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
-algorithm returns the abstract model $v$, which is a set of concrete models, and terminates with UNSAFE or the 
+algorithm returns the abstract model $v$, which is a set of concrete models, and terminates with \textsf{unsafe} or the 
 current abstraction is insufficient to determine the satisfiability of $\varphi$.  
-However, in case of \textsf{CONFLICT}, the algorithm enters in the 
-$conflict\_analysis()$ phase to learn the reason for the conflict.  
+However, in case of \textsf{conflict}, the algorithm enters in the 
+$\mathit{conflict\_analysis}$ phase to learn the reason for the conflict.  
 There can be multiple incomparable reasons for conflict -- based on 
@@
 learning and backtracking continues as long as the result of deduction 
-is conflicting ($\perp$), that is, the abstract valuation $v$ does not
+is conflicting ($\bot$), that is, the abstract valuation $v$ does not
 abstractly satisfy the formula \pscmt{define}.  If no further backtrack is possible, 
@@
 
-\paragraph \em{Solver State:}  The state of a ACDCL solver is a tuple 
+\paragraph{Solver State.}  The state of a ACDCL solver is a tuple 
 of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
 is a sequence of labelled information of the form $(m,s)$ where 
-$m$ is a meet irreducible and $s = decision$ if $m$ is a decision, 
-or $s = deduction$ if $m$ is inferred by a deduction. And $S$ is 
+$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
+or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
 a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 

### 2016-10-20 conclusion.tex
conclusion

--- 2016-10-13 conclusion.tex
@@
 \section{Conclusions}
-In this paper, we instantiate ACDCL solver with relational domains.  We explained specific properties of model search and conflict analysis procedure necessary
-for such lifting.  Experimental evaluation on different benchmarks shows that
+In this paper, we present a procedure for lifting the model search 
+in satisfiability solvers to relational abstract domains.  We explain 
+specific properties of abstract domain elements that are necessary to infer 
+domain-specific deductions.  We present decision heuristics that takes 
+into account the high-level structure of programs and the underlying 
+expressivity of the abstract domains.  Our conflict analysis procedure 
+learns abstract transformer following UIP computation over relational 
+elements.  We instantiate our algorithms as a program analyzer to determine
+safety of C programs.  Experimental evaluation shows that ACDCL is able 
+to suitably exploit the expressivity of richer abstract domains as well 
+as precision of CDCL architecture.
+
+Generalizations of conflict reasons for relational domains 
+and non-relational domain such as trace-based abstrac domains is a 
+possible direction of future work.  We believe that domain refinement 
+in ACDCL will also help to dynamically adjust the precision of the 
+deduction in an effective manner.
+
+\Omit {
+Experimental evaluation on different benchmark sets shows that
 ACDCL is able to suitably exploit the expressivity of richer
@@
 mechanism aided by the richer abstract domains.
-
-Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check.
+Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check. Implement dedicated abstract domains to exploit full performance.
+generalization for relational domain, instantiate acdl with trace-based domains.}

### 2016-10-20 conclusion.tex
compile

@@
 
-\Omit {
+\Omit{
 Experimental evaluation on different benchmark sets shows that
@@
 mechanism aided by the richer abstract domains.
-Future work -- domain refinement, incremental solving, invariant inference, intelligent $\gamma$-completeness check. Implement dedicated abstract domains to exploit full performance.
-generalization for relational domain, instantiate acdl with trace-based domains.}
+Future work -- domain refinement, incremental solving, 
+invariant inference, intelligent $\gamma$-completeness check. 
+Implement dedicated abstract domains to exploit full performance.
+generalization for relational domain, instantiate acdl with trace-based domains.
+}

### 2016-10-20 experiments.tex
decision heuristics

@@
 
-\paragraph {\textbf{Decision Heuristics}} We compare the performance of 
-different decision heuristics in ACDCL.  We observe that the ordered heuristic 
-outperforms other heuristics on control-intensive benchmarks due to its 
-ability to prioritize decisions on variables that appear in conditional 
-branches.  For straight-line code and most of the bit-vector regression 
-suite, random heuristics performs the best.  Whereas activity-based 
-heuristics such as Berkmin heuristic which works well in propositional 
-cases performs best for benchmarks that encountered \pscmt{???} the maximum number 
-of conflicts to prove safety, thus allowing the heuristics \pscmt{???} to choose  
-the decison variable among the set of learnt clauses.
-
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec} presents a comparison between {\em forward} propagation 
+Figure~\ref{prop-dec}(a) presents a comparison between {\em forward} propagation 
 strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of 
@@
 
+\paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
+visualization of different decision heuristics in ACDCL.  We observe that 
+the ordered heuristic outperforms other heuristics due to its ability to 
+prioritize decisions on meet irreducibles that appear in conditional 
+branches.  For straight-line code and most of the bit-vector regression 
+suite, random heuristics performs the best.
+
+\Omit{
+Whereas activity-based heuristics such as Berkmin heuristic which 
+works well in propositional cases performs best for benchmarks 
+that encountered the maximum number of conflicts to prove safety, 
+thus allowing the heuristics to choose the decison variable among the set of learnt clauses.   
+}
+
 \paragraph {\textbf{Learning}} Learning has a significant influence on

### 2016-10-20 learning.tex
deduction
figure 4

@@
 For example, consider a formula 
-$\varphi:= (x=5 \; \wedge \; y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
-abstract valuation $v:= (x=5)$.  Figure~\ref{conflict} shows a 
-snapshot of an abstract conflict graph that 
+$\varphi:= (y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
+abstract valuation $v:= (x-5 \geq 0 \wedge -x+5 \geq 0)$.  
+Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
 stores the deductions obtained from abstract model search using octagon domain.  

### 2016-10-20 conclusion.tex
introduction

@@
 Generalizations of conflict reasons for relational domains 
-and non-relational domain such as trace-based abstrac domains is a 
+and non-relational domain such as trace-based abstract domains is a 
 possible direction of future work.  We believe that domain refinement 

### 2016-10-20 introduction.tex
sync
introduction

@@
 %
+%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
 Static program analysis based on abstract 
@@
 behavior. 
-
+%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
 On the other hand, the combination of precise conjunction in partial assignments 
@@
 
-\rmcmt{relational domains}
-
-The instantiation of ACDCL has so far been limited to interval domains only.  
-In this paper, we identify specific properties of the CDCL algorithm that 
-are necessary to lift the propositional skeleton of CDCL to relational 
-abstract domains such as template polyhedra domains~\cite{sriram} 
-(octagons, zones and equalities).  To this end, we present a characterization 
-of abstract model search procedure and conflict analysis procedure that lift 
-the abstract deduction transformer and the first-UIP algorithm to 
-relational lattice structures.
+The instantiation of ACDCL has so far been instantiated as floating-point
+decision procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and program analyzer 
+that only uses interval constraints~\cite{tacas12}.  Hence, the analysis is 
+limited to bounds checking and numercial programs.  However, verification of 
+certain class of programs require more expressive domains like octagons or 
+equalities or polyhedra.  These domains capture interesting properties of
+programs which intervals can not specify.  In this paper, we identify specific 
+properties of the CDCL algorithm that are necessary to lift the propositional 
+skeleton of CDCL to relational abstract domains such as template polyhedra 
+domains~\cite{sriram} (octagons, zones and equalities).  To this end, we 
+present a characterization of abstract model search procedure and conflict 
+analysis procedure that lift the abstract deduction transformer and the conflict 
+analysis procedure to relational lattice structures.

### 2016-10-20 graphs.tex
decision plot

@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, xlabel=Forward
-  Propagation, ylabel=Chaotic Propagation, 
+	\begin{loglogaxis} [
+      xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, 
+      xlabel=Forward Propagation, ylabel=Chaotic Propagation, 
+      title={Performance of Propagation Heuristics},
 			legend style={at={(0.8,0.15)},

### 2016-10-20 experiments.tex
experiment discussion

@@
 \paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
-visualization of different decision heuristics in ACDCL.  We observe that 
-the ordered heuristic outperforms other heuristics due to its ability to 
-prioritize decisions on meet irreducibles that appear in conditional 
-branches.  For straight-line code and most of the bit-vector regression 
-suite, random heuristics performs the best.
-
+visualization of different decision heuristics in ACDCL.  Random heuristic
+performs the best for most safe benchmarks and all bit-vector category 
+benchmarks.  Whereas, the ordered heuristic performs better for programs 
+with conditional branches since it prioritizes decisions on meet 
+irreducibles that appear in conditional branches over meet irreducibles 
+that involve numerical variables.  Whereas, the runtimes for berkmin 
+heuristic is slightly greater than the random and ordered heuristic 
+for our benchmarks. 
 \Omit{

### 2016-10-20 experiments.tex
description of result

@@
 static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
-Astr{\'e}e was configured with default set of domains which includes 
+Astr{\'e}e was configured with range of abstract domains which includes 
 interval, bit-field, congruence, trace partitioning and relational 
-domains (octagons, polyhedra, zones, equalities, filter).
-
+domains (octagons, polyhedra, zones, equalities, filter).  Whereas, 
+ACDCL was instantiated with a product domain of booleans and template 
+polyhedra.   
+%
 \Omit {
@@
 Figure~\ref{fig:results} presents a comparison of the analysis 
-using CBMC and ACDCL.  The cumulative statistics corresponding to 
-each phase of the algorithm for all 66 benchmarks is presented in 
-Table~\ref{result}.  Experimental result shows a reduction of at least 
-two orders of magnitude in the number of decisions, propagations and conflicts 
+using CBMC and ACDCL.  Figure~\ref{fig:results}(a) clearly shows that 
+the SAT solver made significantly more decisions compared to ACDCL 
+for all the benchmarks.  Whereas, the extreme right points below the 
+diagonal in Figure~\ref{fig:results}(b) shows the propagations in SAT solver  
+is maximum for the benchmarks that exhibit relational behavior.  The 
+cumulative statistics corresponding to each phase of the algorithm 
+for all 66 benchmarks is presented in Table~\ref{result}.  
+Experimental result shows a reduction of at least two orders of 
+magnitude in the number of decisions, propagations and conflicts 
 compared to analyis using CBMC.    

### 2016-10-21 paper.tex
paper title
decision graphs
appendix

--- 2016-10-13 paper.tex
@@
 \usepackage{cite}
+\usepackage{tikz}
+\usetikzlibrary{plotmarks}
+\usepackage{pgfplotstable}
+\usepackage{filecontents}
+\usepackage{pgfplots}
 \usepackage{amsfonts}
@@
 \usepackage{amsmath}
-\usepackage{mathptm}
+\usepackage{stmaryrd}
+%\usepackage{mathptm}
 \usepackage{color}
@@
 \usepackage{hyperref}
+\usepackage[subnum]{cases}
 \makeatother
+
+\newcommand{\extendedonly}[1]{#1}
+\newcommand{\paperonly}[1]{}
 
@@
 
-\title{Lifting CDCL to Relational Domains}
+\title{Model Search in Relational Domains}
 
@@
 as intervals to improve the performance of a floating-point decision
-procedure.  Whereas the propositional reasoning used in SAT solvers can be
-rather easily lifted from the boolean lattice to other non-relational
-domains, lifting them to relational domains is challenging because of
-relational properties of deductions and learnt clauses.
+procedure.  Whereas the propositional reasoning used in satisfiability 
+solvers can be rather easily lifted from the boolean lattice to other 
+non-relational domains, lifting them to relational domains is 
+challenging because of relational properties of deductions.
 %
@@
 \end{abstract}
+
 %===============================================================================
 \input{introduction}
-%===============================================================================
-
-%===============================================================================
-%\input{background}
 %===============================================================================
@@
 \input{example}
-%===============================================================================
-
-%===============================================================================
-%\input{cdcl}
 %===============================================================================
@@
 %===============================================================================
-%\input{definition}
-%===============================================================================
-
-%===============================================================================
 \input{decision}
@@
 %===============================================================================
+\input{background}
+%===============================================================================
+
+%===============================================================================
 \input{conclusion}
@@
 
+\extendedonly{
+\input{appendix}
+}
+
 \end{document}

### 2016-10-21 introduction.tex
introduction fix

@@
 domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl}
-solvers to reason about disjunction without enumerating cases.  
-Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
+based satisfiability (SAT) solvers to reason about disjunction without 
+enumerating cases.  Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
 a formal recipe for applying abstract interpretation~\cite{se2011} to solve the 
@@
 
-The instantiation of ACDCL has so far been instantiated as floating-point
-decision procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and program analyzer 
+ACDCL has so far been instantiated as a floating-point
+decision procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyzer 
 that only uses interval constraints~\cite{tacas12}.  Hence, the analysis is 
-limited to bounds checking and numercial programs.  However, verification of 
+limited to bounds checking in simple numercial programs.  However, verification of 
 certain class of programs require more expressive domains like octagons or 

### 2016-10-21 learning.tex

@@
 \section{Abstract Conflict Analysis}
-A conflict analysis in a SAT solver can be seen as abductive 
-reasoning which underapproximate a set of models that do not satisfy a 
+Conflict analysis in a SAT solver can be seen as abductive 
+reasoning which underapproximates a set of models that do not satisfy a 
 formula~\cite{sas12,dhk2013-popl}.  
@@
 }
-Since relational abstract domains have more complex structure than a partial 
+Since relational abstract domains have more complex structure than the partial 
 assignments domain or interval domain, so finding a generalized reason for 
@@
 \paragraph {\textbf{Abstract Conflict Graph}}
-The first-UIP algorithm in SAT solver works on a data structure called 
-{\em implication graph} that computes a cut that suffices to produce a 
+The first-UIP algorithm in SAT solvers works on a data structure called 
+{\em implication graph} and computes a cut that suffices to produce a 
 conflict.  The trail in ACDCL implicitly encodes a graph structure called 
 {\em abstract conflict graph}, which records the dependencies between 
-deductions made during abstract model search phase.  Nodes in the graph 
-represent elements of the trail $\mathcal{T}$.  Nodes can be either 
-{\em decision} node or {\em deduction} node. Edges can be extracted from
+deductions made during the abstract model search phase.  Nodes in the graph 
+represent elements of the trail $\mathcal{T}$.  A node can be either a
+{\em decision} node or a {\em deduction} node. Edges can be extracted from
 the transformers in the reason trail.  Incoming arrows to node $n$ 
@@
 Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
-stores the deductions obtained from abstract model search using octagon domain.  
-The last deduction conflict with the constraint $(y \leq 0)$.
+stores the deductions obtained from abstract model search using the octagon domain.  
+The last deduction conflicts with the constraint $(y \leq 0)$.
 
@@
 through it.  Unlike a SAT solver, a UIP for relational domain may 
-involve meet irreducibles that contains the same variables.  The 
+involve meet irreducibles that contain the same variables.  The 
 first-UIP is a unique node closest to the conflict, and the last UIP 
@@
 cuts~\cite{cdcl,DBLP:journals/fmsd/BrainDGHK14}, that is, it 
-yields clauses that generates new deduction after backtracking.  
+yields clauses that generate new deductions after backtracking.  
 Every cut in the graph is a reason for conflict that can be 
@@
 }
-For our example, there exists multiple incomaparable reasons for conflict,
-marked as {\em cut0, cut1}, in figure~\ref{conflict}.  Here, cut0 is the first UIP.  
+For our example, there exist multiple incomaparable reasons for conflict,
+marked as {\em cut0, cut1}, in Figure~\ref{conflict}.  Here, cut0 is the first UIP.  
 Choosing cut0 yields a learnt clause 
@@
 %    
-\paragraph{\textbf{Learning in Relational Lattice}}
-Learning in propositional solver yields a asserting clause~\cite{cdcl} 
-that express the negation of the conflict reasons.  However, we model 
+\paragraph{\textbf{Learning in relational domains}}
+Learning in a propositional solver yields an asserting clause~\cite{cdcl} 
+that expresses the negation of the conflict reasons.  However, we model 
 learning in relational domains as learning a transformer which infers 
 new meet irreducibles using the abstract unit rule called 
-{\em abstract unit transformer (AUnit)}.  We add $AUnit$ to the set of 
+{\em abstract unit transformer (AUnit)}.  We add $\mathit{AUnit}$ to the set of 
 transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
@@
 an abstract lattice $A$ with complementable meet irreducibles and a set 
-of meet irreducibles $C \subseteq A$ of $\varphi$ such that $\bigsqcap C$ does not 
-satisfy $\varphi$, $AUnit_C: A \rightarrow A$ is formally defined as follows. 
+of meet irreducibles $C \subseteq A$ of $\varphi$ \pscmt{???} such that $\bigsqcap C$ does not 
+satisfy $\varphi$, $\mathit{AUnit}_C: A \rightarrow A$ is formally defined as follows. 
 \[ AUnit(a) =
@@
 \]
-$AUnit$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
-(3) of $AUnit$ infers a valid meet irreducible, which implies that $C$ is unit. 
+$\mathit{AUnit}$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
+(3) of $\mathit{AUnit}$ infers a valid meet irreducible, which implies that $C$ is unit. 
 However, rule (3) is domain-specific and only applies to numerical abstract
-domains.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
+domains \pscmt{???}.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
 be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
-x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $AUnit_C$, we get a 
+x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $\mathit{AUnit}_C$, we get a 
 decomposition of $a$ into $a'$ and $a''$ where 
 $a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
-$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $AUnit_C(a) = (x+z < 10)$, 
+$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $\mathit{AUnit}_C(a) = (x+z < 10)$, 
 which constrains the bound of the octagonal constraint.  
 Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
-and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $AUnit_C(a) = \bot$ using
+and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $\mathit{AUnit}_C(a) = \bot$ using
 the rule (1), since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
-then $AUnit_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
-and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $AUnit$ 
+then $\mathit{AUnit}_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
+and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $\mathit{AUnit}$ 
 returns $\top$.
@@
 \paragraph {\textbf{Backjumping}}
-A backjumping procedure undo all the assignments in the trail up to 
-a decision level that restores the solver to a consistent state 
+A backjumping procedure undoes all the assignments in the trail up to 
+a decision level \pscmt{explained?} that restores the solver to a consistent state 
 (non-conflicting).  The backjumping level is defined by the literal 

### 2016-10-21 experiments.tex
comments in exp

@@
 %
-\begin{table}
+\begin{table}[t]
 \begin{center}
@@
 without any restarts.  The solver is restarted in 50 cases to avoid spending 
-too much time in branches that does not easily lead to a satisfying assignment or 
+too much time in branches that do not easily lead to a satisfying assignment or 
 stronger clause learning.  In contrast, the analysis using ACDCL solved all
-66 benchmarks without any restarts.  This is attributed to intelligent decision 
+66 benchmarks without any restarts.  This is attributed to intelligent \pscmt{hm...} decision 
 heuristics which exploits the high-level structure of the program combined 
-with stronger deductions and clause learning mechanism aided by the richer 
+with the stronger deduction and clause learning mechanisms aided by the richer 
 abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
-but the analysis using Astr{\'e}e shows high degree of imprecision.  
-Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.
+but the analysis using Astr{\'e}e shows a high degree of imprecision.  
+Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks. \pscmt{even with trace partitioning?}
 
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec}(a) presents a comparison between {\em forward} propagation 
-strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of 
+Figure~\ref{prop-dec}(a) \pscmt{in Fig 5 you distinguish between safe and unsafe, why not here too?} presents a comparison between {\em forward} propagation 
+strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of the
 propagation strategy influences the total number of decisions and clause 
 learning iterations.  Hence, the propagation strategy has a significant 
-influence on the run time which can be visualised in Figure~\ref{prop-dec}.  
+influence on the run time which can be visualised in Figure~\ref{prop-dec}(a).  
 Unlike the forward propagation, the multi-way strategy may take several 
-iterations to reach the fixed-point, but it significantly reduces the 
+iterations \pscmt{forward does not need several? - reformulate} to reach the fixed-point, but it significantly reduces the 
 total number of decisions and conflicts.  This is attributed to the higher 
 precision of the meet irreducibles inferred by the multi-way strategy which 
-subsequently aids the decision heuristics to make better decisions. 
+subsequently aids the decision heuristics to make better decisions. \pscmt{maybe repeat that it is a simultaneous forward/backward iteration...}
 \Omit {
@@
 benchmarks.  Whereas, the ordered heuristic performs better for programs 
-with conditional branches since it prioritizes decisions on meet 
+with conditional branches since it prioritises decisions on meet 
 irreducibles that appear in conditional branches over meet irreducibles 
@@
 heuristic is slightly greater than the random and ordered heuristic 
-for our benchmarks. 
+for our benchmarks. \pscmt{I don't see ordered in the graph. The differences between Berkmin and random are insignificant.}
 \Omit{
@@
 and guide the model search with a learnt transformer.  However, the analysis 
-without learning exhibits case-enumeration behavior and thus immediately 
-runs out of capacity for most benchmarks.
+without learning \pscmt{reformulate} exhibits case-enumeration behavior and thus immediately 
+runs out of capacity for most benchmarks. \pscmt{how many, what are the bounds on time and memory?}
 

### 2016-10-21 domains.tex
compile
domain comments

@@
 %-------------------------------------------------------------------------------
-The work of~\cite{sas12,dhk2013-popl} shows that CDCL based SAT solvers 
-operates on the partial assignment domains. We present a generic instantiation 
-of CDCL architecture that can be extended to any relational and non-relational
+SAT solvers operates on the partial assignment domains~\cite{sas12,dhk2013-popl}.  
+We present an instantiation of ACDCL as program analyzer that uses relational
 domains.  
 % 
-In our prototype implementation and for illustrative purposes in this
-paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
+%In our prototype implementation and for illustrative purposes 
+In this paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
 $\mathcal{B}$ is the Boolean domain
 $\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$, 
-$b$ is the number of Boolean variables in the program, 
+$b$ is the number of boolean variables in the program, 
 and $\mathcal{R}$ is a \emph{relational} domain over the 
-numerical (bitvector) variables in their respective
+numerical (bitvector) variables of respective
 data types (currently, signed and unsigned integers are supported).
@@
 $\mathcal{B}^b$ by their concretisations to propositional formulae,
-e.g.\ for the Boolean variable vector $(g_1,c)$ with abstract value
-$(\false,\top)$ we will write $\neg g_1$.
+e.g.\ for the boolean variable vector $(g_1,c)$ \rmcmt{what is g1, what is c?} 
+with abstract value $(\false,\top)$ we will write $\neg g_1$.
 
@@
 The template polyhedra domain~\cite{sriram} represents sets $X$ of
-valuations of a vector of variables $\vec{x}$ such that
+valuations of a vector of variables $\vec{x}$ \rmcmt{not clear what is a
+valuation?} such that
 $\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
@@
 techniques for computing the domain
-operations~\cite{sriram,GS07b,BJKS15}. In our implementation we use
+operations~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
 the strategy iteration approach available in 2LS~\cite{BJKS15}.
@@
 inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
-of the domain. $\true$ is $\top$ and $\false$ is $\bot$.
+of the \rmcmt{template polyhedra ?} domain. $\true$ is $\top$ and $\false$ is $\bot$.
 
@@
 Template polyhedra subsume, for example, interval and octagon domains.
+\rmcmt{can we rephrase this statement to say that -- In this paper, we used
+interval, octagon and equality ?}
 
@@
 domains require the computation of a \emph{closure} in order to obtain
-a normal form that is usually required for efficient domain operations.
-An example of a closure computation is shown below.
+a normal form that is usually required for efficient \rmcmt{or precise?} 
+domain operations.  An example of a closure computation for octagonal 
+inequalities is shown below.
 \[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
@@
 %
-We will see that our algorithm performs closure computations lazily,
-postponing them to the point where they are actually necessary.
+We will see that our model search algorithm performs closure computations lazily,
+postponing them to the point where they are actually necessary.\rmcmt{are we
+showing this ?}
 
@@
 statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
-The best abstract transformer is
+The best abstract transformer is \rmcmt{is this equation updated ?}
 \[\llbracket s
@@
 %
-We will give more details in Section~\ref{sec:abst}.
+A detailed description of abstract transformer is presented in Section~\ref{sec:abst}.
 

### 2016-10-21 algorithm.tex
sec 4

@@
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-The algorithm begins with a set $S$ of abstract deduction transformer 
-$\widehat{ded_{\varphi}}$ \pscmt{define in terms of P and A} for a formula $\varphi$ over an abstract 
-domain $A$, an abstract valuation $\langle v \leftarrow \top \rangle$, 
-and an empty propagation trail and reason trail, $\mathcal{T}$, $\mathcal{R}$, 
-respectively.  The propagation trail stores all meet irreducibles inferred 
-by the abstract model search phase (deductions and decisions), while the 
-reason trail maps elements of propagation trail to transformers that were 
-used to derive them. A propagation trail is mapped to an abstract value as 
-follows; $\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with 
-$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty sequence $\langle\rangle$.
+The algorithm takes a program $\mathcal{S}$ and propagation, decision,
+and conflict analysis heuristics as inputs.  As internal state, it
+maintains a propagation trail $\mathcal{T}$ and a reason trail
+\pscmt{map?} $\mathcal{R}$.  The propagation trail is an initially
+empty sequence of meet irreducibles inferred by the abstract model
+search phase (deductions and decisions). The reason trail maps
+\pscmt{certain?} elements of the propagation trail to transformers
+$s\in\mathcal{S}$ that were used to derive them. A propagation trail
+is converted into an abstract value as follows;
+$\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with
+$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty
+sequence $\langle\rangle$.
 
@@
 Else, the analysis enters into the while($true$) loop and makes a new decision 
-by a call to $decide(v)$ which returns a new meet irreducible $d$ that is
-consistent to the state of the solver. \pscmt{why do we have to store $v$ and $\mathcal{T}$?} If $d$ refines the current abstract 
+by a call to $\mathit{decide}$ which returns a new meet irreducible $d$ that is
+consistent with the state of the solver.
+\pscmt{what happens if we cannot make any decision any more?
+-- then we should either have deduced $\bot$ or $\mathit{abs}(\mathcal{T})$ must be $\gamma$-complete} If $d$ refines the current abstract
 value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 

### 2016-10-21 learning.tex
decision level

@@
 (3) of $\mathit{AUnit}$ infers a valid meet irreducible, which implies that $C$ is unit. 
-However, rule (3) is domain-specific and only applies to numerical abstract
-domains \pscmt{???}.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ 
-be a conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
+However, rule (3) is domain-specific and only applies to interval and octagon
+domain.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ be a 
+conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
 x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $\mathit{AUnit}_C$, we get a 
@@
 A backjumping procedure undoes all the assignments in the trail up to 
-a decision level \pscmt{explained?} that restores the solver to a consistent state 
-(non-conflicting).  The backjumping level is defined by the literal 
-of the conflict clause that is closest to the root (decision level 0) 
-where the conflict clause is still unit. If a conflict clause is 
-globally unit, then the backjumping level is the root of the search tree.
+a decision level that restores the solver to a non-conflicting state.  
+The backjumping level is defined by the literal of the conflict clause 
+that is closest to the root (decision level 0) where the conflict 
+clause is still unit.  If a conflict clause is globally unit, then the 
+backjumping level is the root of the search tree.
 

### 2016-10-21 algorithm.tex
more in sec4

@@
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-The algorithm takes a program $\mathcal{S}$ and propagation, decision,
-and conflict analysis heuristics as inputs.  As internal state, it
-maintains a propagation trail $\mathcal{T}$ and a reason trail
-\pscmt{map?} $\mathcal{R}$.  The propagation trail is an initially
-empty sequence of meet irreducibles inferred by the abstract model
-search phase (deductions and decisions). The reason trail maps
-\pscmt{certain?} elements of the propagation trail to transformers
-$s\in\mathcal{S}$ that were used to derive them. A propagation trail
-is converted into an abstract value as follows;
+The algorithm takes a program \pscmt{make this a set of transformers
+  (parametrised by the propagation heuristics)}
+$\mathcal{S}$ and propagation, decision, and conflict analysis
+heuristics as inputs.  As internal state, it maintains a propagation
+trail $\mathcal{T}$ and a reason trail \pscmt{map?} $\mathcal{R}$.
+The propagation trail is an initially empty sequence of meet
+irreducibles inferred by the abstract model search phase (deductions
+and decisions). The reason trail maps \pscmt{certain?} elements of the
+propagation trail to transformers $s\in\mathcal{S}$ that were used to
+derive them. The \emph{abstract value} $\mathit{abs}(\mathcal{T})$
+corresponding to the propagation trail $\mathcal{T}$ is the
+conjunction of the meet irreducibles on the trail:
 $\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with
@@
 
-The procedure $deduce(S,v,\mathcal{T},\mathcal{R})$ computes a greatest 
-fixed point over the transformers in $\varphi$ that refines the abstract valuation 
-$v$, similar to the BCP step in SAT solvers.  If the result of $deduce$ 
-returns \textsf{conflict} ($\bot$), the algorithm terminates with \textsf{safe}. 
-Else, the analysis enters into the while($true$) loop and makes a new decision 
-by a call to $\mathit{decide}$ which returns a new meet irreducible $d$ that is
-consistent with the state of the solver.
-\pscmt{what happens if we cannot make any decision any more?
--- then we should either have deduced $\bot$ or $\mathit{abs}(\mathcal{T})$ must be $\gamma$-complete} If $d$ refines the current abstract
-value $v$, the corresponding propagation and reason trail, $\mathcal{T}$ 
-and $\mathcal{R}$ are updated.
-
+The procedure $deduce$ computes a greatest fixed point over the
+transformers in $\mathcal{S}$ that refines the abstract value,
+similar to the BCP step in SAT solvers.  If the result of $deduce$
+returns \textsf{conflict} ($\bot$), the algorithm terminates with
+\textsf{safe}.  Else, the analysis enters into the while($true$) loop
+and makes a new decision by a call to $\mathit{decide}$ which returns
+a new meet irreducible $d$ that is consistent with the state of the
+solver. \pscmt{what happens if we cannot make any decision any more?
+  -- then we should either have deduced $\bot$ or
+  $\mathit{abs}(\mathcal{T})$ must be $\gamma$-complete} If $d$
+refines the current abstract value, the corresponding propagation
+and reason trail, $\mathcal{T}$ and $\mathcal{R}$ are updated.
+%
 For example, a decision in the interval domain restricts the range of 
@@
 %{\em dual widening} \pscmt{what do we learn?}.
-The procedure $deduce$ is called next to infer a new 
-set of meet irreducibles for the current decision.  The model search phase 
-alternates between the decision and deduction until the abstract value $v$ 
-precisely \pscmt{not true} represents the set of models or the deduction leads to a \textsf{conflict}.  
-Checking whether the abstract model $v$ concretises to a set of models of $\varphi$; this corresponds to a 
-$\gamma$-completeness~\cite{dhk2013-popl} check in abstract interpretation.  
-If $v$ is $\gamma$-complete, then it cannot be refined further.  Thus, the 
-algorithm returns the abstract model $v$, which is a set of concrete models, and terminates with \textsf{unsafe} or the 
-current abstraction is insufficient to determine the satisfiability of $\varphi$.  
+
+The procedure $deduce$ is called next to infer a new set of meet
+irreducibles for the current decision.  The model search phase
+alternates between the decision and deduction until the abstract value
+$v$ precisely \pscmt{not true} represents the set of models or the
+deduction leads to a \textsf{conflict}.  Checking whether the abstract
+model $v$ concretises to a set of models of $\varphi$; this
+corresponds to a $\gamma$-completeness~\cite{dhk2013-popl} check in
+abstract interpretation.  If $v$ is $\gamma$-complete, then it cannot
+be refined further.  Thus, the algorithm returns the abstract model
+$v$, which is a set of concrete models, and terminates with
+\textsf{unsafe} or the current abstraction is insufficient to
+determine the satisfiability of $\varphi$.
+
 However, in case of \textsf{conflict}, the algorithm enters in the
-$\mathit{conflict\_analysis}$ phase to learn the reason for the conflict.  
-There can be multiple incomparable reasons for conflict -- based on 
-the choice of Unique Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically 
-chooses one reason.  A learnt clause must include asserting cuts which 
-guarantees derivation of new meet irreducibles after backtracking. The clause 
-learning and backtracking continues as long as the result of deduction 
-is conflicting ($\bot$), that is, the abstract valuation $v$ does not
-abstractly satisfy the formula \pscmt{define}.  If no further backtrack is possible, 
-then the algorithm terminates and $\varphi$ is \textsf{safe}. Else, the 
-algorithm makes a new decision and the above process is repeated 
-until a real counterexample \pscmt{is this a model?} is obtained or the algorithm backtracks to 
-decision level 0 after a conflict in which case it returns \textsf{safe}.
+$\mathit{conflict\_analysis}$ phase to learn the reason for the
+conflict.  There can be multiple incomparable reasons for conflict --
+based on the choice of Unique Implication Point (UIP)~\cite{cdcl}.
+ACDCL heuristically chooses one reason.  A learnt clause must include
+asserting cuts which guarantees derivation of new meet irreducibles
+after backtracking. The clause learning and backtracking continues as
+long as the result of deduction is conflicting ($\bot$), that is, the
+abstract valuation $v$ does not abstractly satisfy the formula
+\pscmt{define}.  If no further backtrack is possible, then the
+algorithm terminates and $\varphi$ is \textsf{safe}. Else, the
+algorithm makes a new decision and the above process is repeated until
+a real counterexample \pscmt{is this a model?} is obtained or the
+algorithm backtracks to decision level 0 after a conflict in which
+case it returns \textsf{safe}.
 
-\paragraph{Solver State.}  The state of a ACDCL solver is a tuple 
+\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
 of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 

### 2016-10-21 algorithm.tex
comment

--- 2016-10-21 algorithm.tex
@@
   propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
-and a conflict analysis heuristics $\mathit{conflict\_analysis}$.}
+and a conflict analysis heuristics $\mathit{conflict\_analysis}$.\rmcmt{calling
+them heuristics is not a good idea since these are key phases of the cdcl
+algorithm. Each phase may have a heuristics for some sub phase but the algorithm
+of each phase still remians a procedure and not a heuristic}}
 \Output{The status \safe or \unsafe, and a counterexample if \unsafe.}

### 2016-10-21 algorithm.tex
more comment

@@
 can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
-bounds \pscmt{that's specific for a particular domain}.  A current valuation $v$ is a meet of all elements of trail
+bounds \pscmt{that's specific for a particular domain}.  
+\rmcmt{define trail refinement}
+\Omit {
+A current valuation $v$ is a meet of all elements of trail
 $\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
@@
 $\mathcal{S}$ is not abstractly satisfied by $v$.
-
+}
 %\pscmt{define ``consistent with trail''}

### 2016-10-21 algorithm.tex
more sec4

--- 2016-10-21 algorithm.tex
@@
   propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
-and a conflict analysis heuristics $\mathit{conflict\_analysis}$.\rmcmt{calling
-them heuristics is not a good idea since these are key phases of the cdcl
-algorithm. Each phase may have a heuristics for some sub phase but the algorithm
-of each phase still remians a procedure and not a heuristic}}
+and a conflict analysis heuristics $\mathit{conflict\_analysis}$.}
 \Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
@@
 
-The procedure $deduce$ is called next to infer a new set of meet
-irreducibles for the current decision.  The model search phase
-alternates between the decision and deduction until the abstract value
-$v$ precisely \pscmt{not true} represents the set of models or the
-deduction leads to a \textsf{conflict}.  Checking whether the abstract
-model $v$ concretises to a set of models of $\varphi$; this
-corresponds to a $\gamma$-completeness~\cite{dhk2013-popl} check in
-abstract interpretation.  If $v$ is $\gamma$-complete, then it cannot
-be refined further.  Thus, the algorithm returns the abstract model
-$v$, which is a set of concrete models, and terminates with
-\textsf{unsafe} or the current abstraction is insufficient to
-determine the satisfiability of $\varphi$.
+The procedure $\mathit{deduce}$ is called next to infer new meet
+irreducibles taking into account the current decision.  The model search phase
+alternates between the decision and deduction until $\mathit{deduce}$
+returns either \textsf{sat} or \textsf{conflict}.
+In the former case, we have found an abstract value that represents 
+models of $\varphi$ and thus we return \textsf{unsafe}.
+%
+% has found an abstract value that represents a set of models of $\varphi$ or the
+% deduction leads to a \textsf{conflict}.  Checking whether the abstract
+% model concretises to a set of models of $\varphi$; this
+% corresponds to a $\gamma$-completeness~\cite{dhk2013-popl} check in
+% abstract interpretation.  If $v$ is $\gamma$-complete, then it cannot
+% be refined further.  Thus, the algorithm returns the abstract model
+% $v$, which is a set of concrete models, and terminates with
+% \textsf{unsafe} or the current abstraction is insufficient to
+% determine the satisfiability of $\varphi$.
 
-However, in case of \textsf{conflict}, the algorithm enters in the
+In the latter case the algorithm enters in the
 $\mathit{conflict\_analysis}$ phase to learn the reason for the

### 2016-10-21 algorithm.tex
section 4 updates

@@
 $\mathcal{S}$ and propagation, decision, and conflict analysis
-heuristics as inputs.  As internal state, it maintains a propagation
+heuristics as inputs.  
+\rmcmt{Approximation of the concrete transformers in 
+$\mathcal{S}$ are typically available in abstract domain in the 
+form of strongest-post condition or weakest pre-condition. } 
+As internal state, it maintains a propagation
 trail $\mathcal{T}$ and a reason trail \pscmt{map?} $\mathcal{R}$.

### 2016-10-21 deductions.tex
deduction
section 4 updates

@@
   heuristics (already done for initialize, TBD for update...}
-A {\em worklist} is initialized following a heuristics choice for iteration 
-strategies. A {\em forward} iteration strategy initializes the worklist with 
-transformers that has constants in the right-hand side. Whereas, a {\em backward} 
+A {\em worklist} is a queue which contains \rmcmt{abstract deduction
+transformers}.  A {\em worklist} is initialized following a heuristics 
+choice for iteration strategies. A {\em forward} iteration 
+strategy initializes the worklist with \rmcmt{abstract} transformers that 
+has constants in the right-hand side. Whereas, a {\em backward} 
 iteration strategy starts with a set of assertions. On the other hand, a 
 {\em multi-way} iteration strategy initializes the worklist with set of all 
-transformers.  For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
+transformers and performs both forward and backward propagations in tandem.  
+For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
 passed to the abstract domain, if the domain infers a new meet irreducible 

### 2016-10-21 decision.tex
decision level
decision

@@
 the trail $\mathcal{T}$, that is, it must not contradict with the elements 
-in the trail.
+in the trail.  A new decision always increases the decision level by one.
 
@@
 a decision phase heuristically returns a non-singleton meet irreducible 
-by choosing a set of branching variables $\{B\} \subseteq \mathcal{V}$, 
-a bound $c$, and a polarity ($\leq$ or $\geq$).  For interval domain, $|B|=1$.  
-A decision changes the sate of the solver by adding a new meet irreducible 
-$m$ and labelling information $s=decision$ to the trail, which is described below. 
+over a set of branching variables $\{B\} \subseteq \mathcal{V}$, a 
+bound $c$, and a polarity ($\leq$ or $\geq$).  For interval domain, $|B|=1$ 
+since an interval meet irreducible is defined over a single variable.  For 
+octagons, $|B|=2$.  A decision adds a new meet irreducible $m$ to the trail.  
+Subsequently, the new state of the solver is defined over $m$ and corresponding 
+label information $s=decision$, which is described below. 
 \[decide: \quad (\mathcal{E},S) \rightarrow (\mathcal{E}(m,s),S) \]
@@
 before choosing meet irreducibles with numerical variables.  
-The ordering is path-sensitive as it gives an effect of trace partitioning.
+The ordering heuristic is similar to trace partitioning~\cite{toplas07}.
 

### 2016-10-21 algorithm.tex
sync

--- 2016-10-21 algorithm.tex
@@
   $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
-  $\mathcal{R}[|\mathcal{T}|] \leftarrow $\textsf{decision} \;
+  $\mathcal{R}[|\mathcal{T}|] \leftarrow \top$ \;
   $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$\;
@@
 sequence $\langle\rangle$.
+\pscmt{trail refinement}
 
@@
 
-\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
-of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
-is a sequence of labelled information of the form $(m,s)$ where 
-$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
-or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
-a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
-stores variable assignments of the form $(p, t)$, where $p$ is 
-a propositional variable that appears at most once in the trail 
-and $t$ is a truth assignment (true or false).  Whereas, a trail 
-$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
-inferred by deduction or decision phase where a variable in trail 
-can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
-bounds \pscmt{that's specific for a particular domain}.  
-\rmcmt{define trail refinement}
-\Omit {
-A current valuation $v$ is a meet of all elements of trail
-$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
-empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
-$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
-$\mathcal{S}$ is not abstractly satisfied by $v$.
-}
+%\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
+%of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
+%is a sequence of labelled information of the form $(m,s)$ where 
+%$m$ is a meet irreducible and $s = \mathsf{decision}$ if $m$ is a decision, 
+%or $s = \mathsf{deduction}$ if $m$ is inferred by a deduction. And $S$ is 
+%a set of abstract deduction transformer \pscmt{???}.  A trail in a SAT solver 
+%stores variable assignments of the form $(p, t)$, where $p$ is 
+%a propositional variable that appears at most once in the trail 
+%and $t$ is a truth assignment (true or false).  Whereas, a trail 
+%$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
+%inferred by deduction or decision phase where a variable in trail 
+%can be assigned \pscmt{constrained?} multiple times, each time with increasingly precise 
+%bounds \pscmt{that's specific for a particular domain}.  
+%\rmcmt{define trail refinement}
+%\Omit {
+%A current valuation $v$ is a meet of all elements of trail
+%$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
+%empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
+%$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
+%$\mathcal{S}$ is not abstractly satisfied by $v$.
+%}
 %\pscmt{define ``consistent with trail''}

### 2016-10-21 domains.tex
sync

@@
 
+\pscmt{In Figure~\ref{}}
 For notational convenience we will denote elements of
@@
 %
-The template polyhedra domain~\cite{sriram} represents sets $X$ of
+The template polyhedra domain~\cite{sriram} represents sets $X$ \pscmt{split this up into two sentences} of
 valuations of a vector of variables $\vec{x}$ \rmcmt{not clear what is a
@@
 techniques for computing the domain
-operations~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
+operations, such as $\sqcap$, $\sqcup$, etc~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
 the strategy iteration approach available in 2LS~\cite{BJKS15}.
@@
 inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
-of the \rmcmt{template polyhedra ?} domain. $\true$ is $\top$ and $\false$ is $\bot$.
+of the template polyhedra domain. $\true$ is $\top$ and $\false$ is $\bot$.
+\pscmt{give $\mat{C}$ and $\vec{d}$ for this example}
 
@@
 domains require the computation of a \emph{closure} in order to obtain
-a normal form that is usually required for efficient \rmcmt{or precise?} 
+a normal form that is usually required for precise and efficient 
 domain operations.  An example of a closure computation for octagonal 
@@
 \[\llbracket s
-\rrbracket^\sharp(a)=\bigsqcup \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write
+\rrbracket^\sharp(a)=\bigsqcap \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write \pscmt{fix this}
 $a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]

### 2016-10-21 experiments.tex
experiment graph

@@
 static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
-Astr{\'e}e was configured with range of abstract domains which includes 
+Astr{\'e}e is configured with range of abstract domains which includes 
 interval, bit-field, congruence, trace partitioning and relational 
 domains (octagons, polyhedra, zones, equalities, filter).  Whereas, 
-ACDCL was instantiated with a product domain of booleans and template 
-polyhedra.   
+ACDCL is instantiated with a product domain of booleans and template 
+polyhedra.  Additionally, the analysis using ACDCL is configured with 
+a heuristic choice of decision (ordered, random, berkmin), 
+propagation (forward and multi-way) and conflict-analysis (learning UIP,
+DPLL-style).  The timeout for our experiments is set to ~1 hour. 
 %
@@
 stronger clause learning.  In contrast, the analysis using ACDCL solved all
-66 benchmarks without any restarts.  This is attributed to intelligent \pscmt{hm...} decision 
+66 benchmarks without any restarts.  This is attributed to decision 
 heuristics which exploits the high-level structure of the program combined 
@@
 but the analysis using Astr{\'e}e shows a high degree of imprecision.  
-Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks. \pscmt{even with trace partitioning?}
+Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.
 
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec}(a) \pscmt{in Fig 5 you distinguish between safe and unsafe, why not here too?} presents a comparison between {\em forward} propagation 
+Figure~\ref{prop-dec}(a) presents a comparison between {\em forward} propagation 
 strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of the
@@
 influence on the run time which can be visualised in Figure~\ref{prop-dec}(a).  
-Unlike the forward propagation, the multi-way strategy may take several 
-iterations \pscmt{forward does not need several? - reformulate} to reach the fixed-point, but it significantly reduces the 
-total number of decisions and conflicts.  This is attributed to the higher 
+Compared to forward propagation, the multi-way strategy may take more iterations 
+to reach the fixed-point, but it significantly reduces the total number of 
+decisions and conflicts.  This is attributed to the higher 
 precision of the meet irreducibles inferred by the multi-way strategy which 
-subsequently aids the decision heuristics to make better decisions. \pscmt{maybe repeat that it is a simultaneous forward/backward iteration...}
+subsequently aids the decision heuristics to make better decisions.
+
 \Omit {
@@
 heuristic is slightly greater than the random and ordered heuristic 
-for our benchmarks. \pscmt{I don't see ordered in the graph. The differences between Berkmin and random are insignificant.}
+for our benchmarks.
+
 \Omit{
@@
 The effect of UIP computation allows ACDCL to backtrack non-chronologically 
-and guide the model search with a learnt transformer.  However, the analysis 
-without learning \pscmt{reformulate} exhibits case-enumeration behavior and thus immediately 
-runs out of capacity for most benchmarks. \pscmt{how many, what are the bounds on time and memory?}
+and guide the model search with a learnt transformer.  However, the DPLL-style 
+analysis exhibits case-enumeration behavior and could not finish within the time 
+bound for 20\% of our benchmarks.
 

### 2016-10-21 graphs.tex
experiment graph

--- 2016-10-20 graphs.tex
@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, xlabel=Forward
-  Propagation, ylabel=Chaotic Propagation, 
-			legend style={at={(0.8,0.15)},
-			anchor=north,legend columns=-1 },
+	\begin{loglogaxis} [
+      xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, 
+      xlabel=Forward Propagation, ylabel=Chaotic Propagation, 
+      title={Performance of Propagation Heuristics},
+			legend pos = north west,
+			%legend style={at={(0.8,0.15)},
+			%anchor=north,legend columns=-1 },
 			]
 \addplot [mark size=1pt,only marks,scatter,point meta=explicit symbolic,
-	scatter/classes={a={mark=square,draw=black}}] 
+	scatter/classes={s={mark=square},u={mark=triangle*,blue}},] 
 	table [meta=label] {plotdata/scatter-chaotic-forward.dat};
-	%\legend{Safe,Unsafe}
+	\legend{Safe,Unsafe}
   \addplot [domain=.1:2500] {x};

### 2016-10-21 paper.tex
abstract

@@
 In this paper, we present the non-trivial adaptations of the model search
-and conflict analysis algorithms in ACDCL in order to enable the use of
+and conflict analysis algorithms in order to enable the use of
 relational abstract domains.  We have implemented these algorithms inside
-the 2LS verification tool and instantiated our analysis over a template
-polyhedra domain.  We evaluate the performance of our analyzer in comparison
-with program verifiers that uses propositional-level solvers and a static
-analysis tool.
+the 2LS verification tool and instantiated our analyzer over a template 
+polyhedra domain to check safety properties of C programs.  We evaluate 
+the performance of our analyzer in comparison with program verifiers that 
+uses propositional-level solvers and a static analysis tool.
 %
-Experimental evidence on a set of benchmarks drawn from bit-vector
-regression in SV-COMP'16, bit-precise software models of hardware circuits
-auto-generated from v2c tool, and several bounds checking benchmarks shows
-significant reduction in the number of decisions and backjumping iterations
+Experimental evidence suggest 
+\Omit {
+on a set of benchmarks drawn from bit-vector regression 
+in SV-COMP'16, bit-precise software models of hardware 
+circuits auto-generated from v2c tool, and several 
+bounds checking benchmarks 
+}
+two orders of magnitude reduction in the number of 
+decisions and conflict analysis compared to a satisfiability 
+solver while being more precise compared to an abstract interpreter. 
+\Omit{
 along with stronger deductions and learnt clauses aided by the richer
 abstract domains.
+}
 %

### 2016-10-21 domains.tex
fixes in sec3

@@
 This section introduces the program model and identifies properties of abstract
-domains that are necesary for learning in abstract lattice structures.
+domains that are necessary for learning in abstract lattice structures.
 % \subsection{Programs:}
@@
 $\begin{array}{rcl}
-\mathcal{P}:= (g_0 = (0 \leq v\leq N)) \\
+\mathcal{S}:= (g_0 = (0 \leq v\leq N)) \\
 (g_1 = (g_0 \wedge c)) \wedge (x_0 = v) \wedge \\
@@
 (g_2 = (g_1 \vee g_0\wedge \neg c)) \wedge \\
-(z  = x_2 * x_2) \wedge (z<0 \wedge g_2)
+(z  = x_2 * x_2) \wedge (g_2 \wedge z<0)
 \end{array}$
@@
 %
-In this paper, we consider safety checking of loop-free programs, including
-bounded unwindings of programs with loops or recursion\rmcmt{why can't we simply 
-say we consider programs with bounded loops}.  We assume
-that safety properties are given as assertions $\mathcal{A}$ in the program.
+In this paper, we consider \emph{bounded programs} with safety
+properties given as assertions $\mathcal{A}$ in the program.
 %
-Such a program is represented as set
+A bounded program is obtained by a program transformation that unfolds
+loops and recursions a finite number of times. The result of this
+transformation is represented as set
 $\mathcal{S}=\mathcal{P}\cup\{\neg \bigwedge_{a\in\mathcal{A}} a\}$
-where $\mathcal{P}$ are the statements of the program in the form of
-constraints as typically obtained by translating the program into
-single static assignment (SSA) form via a data flow analysis.
+where $\mathcal{P}$ are the statements of the transformed program in the
+form of constraints as typically obtained by translating the program
+into single static assignment (SSA) form via a data flow analysis.
 %
-Figure~\ref{ssa} shows the SSA constraints for the program in Figure~\ref{fig:example}. 
+Figure~\ref{ssa} shows the SSA constraints for the program in
+Figure~\ref{fig:example}.
 %
@@
 %
-Based on the above program representation, we define a safety equation as, 
+Based on the above program representation, we define a safety equation \pscmt{why ``equation''?} as, 
 $\varphi:= \bigwedge_{s\in\mathcal{S}} s$. The formula $\varphi$ is 
@@
 %-------------------------------------------------------------------------------
-SAT solvers operates on the partial assignment domains~\cite{sas12,dhk2013-popl}.  
-We present an instantiation of ACDCL as program analyzer that uses relational
+SAT solvers operate on partial assignment domains~\cite{sas12,dhk2013-popl}.  
+We present an instantiation of ACDCL as program analyser that uses relational
 domains.  
@@
 In this paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
-$\mathcal{B}$ is the Boolean domain
+$\mathcal{B}$ is the boolean domain
 $\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$, 
@@
 
-\pscmt{In Figure~\ref{}}
 For notational convenience we will denote elements of
 $\mathcal{B}^b$ by their concretisations to propositional formulae,
-e.g.\ for the boolean variable vector $(g_1,c)$ \rmcmt{what is g1, what is c?} 
-with abstract value $(\false,\top)$ we will write $\neg g_1$.
+For example, the program in Figure~\ref{ssa} has four boolean variables
+written as the vector $(g_0,g_1,c,g_2)$; then we will denote an abstract value $(\top, \false, \true, \top) \in \mathcal{B}^4$ as $\neg g_1 \wedge c$.
 
@@
 %
-The template polyhedra domain~\cite{sriram} represents sets $X$ \pscmt{split this up into two sentences} of
-valuations of a vector of variables $\vec{x}$ \rmcmt{not clear what is a
-valuation?} such that
+An abstract value of the template polyhedra domain~\cite{sriram}
+represents a set $X$ of values of numerical variables $\vec{x}$. For
+example, in the program in Figure~\ref{ssa}, we have four numerical
+variables, written as the vector $\vec{x}:=(x_0,x_1,x_2,z)$.  We can
+represent sets of values for $\vec{x}$ such that
 $\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
 and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
-by a special element to denote the minimal element $\bot$ of the lattice.
+by a special element $\bot$ to denote the minimal element of the
+lattice.
 %
 The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
-\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is
-applied component-wise. There are several optimisation-based
-techniques for computing the domain
-operations, such as $\sqcap$, $\sqcup$, etc~\cite{sriram,GS07b,BJKS15}\rmcmt{like what operations?}. In our implementation we use
-the strategy iteration approach available in 2LS~\cite{BJKS15}.
+\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is applied
+component-wise. There are several optimisation-based
+techniques~\cite{sriram,GS07b,BJKS15} for computing the domain
+operations, such as $\sqcap$, $\sqcup$, etc. In our implementation we
+use the strategy iteration approach available in 2LS~\cite{BJKS15}.
 %
 The concretisation $\gamma(\vec{d})$ is the set $\{\vec{x}\mid
-\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$.
+\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$, i.e.\ the empty polyhedron.
 %
 For notational convenience we will use conjunctions of linear
-inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote elements
-of the template polyhedra domain. $\true$ is $\top$ and $\false$ is $\bot$.
-\pscmt{give $\mat{C}$ and $\vec{d}$ for this example}
+inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote the
+abstract value $\vec{d}=\vecv{0}{30}$ of the template polyhedra domain
+with $\mat{C}=\qmat{-1}{0}{1}{-1}$ and $\vec{x}=\vecv{x_1}{z}$;
+$\true$ is $\top$ and $\false$ is $\bot$.  Template polyhedra can
+express other domains, such as interval and octagon domains.
 
@@
 % \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
-Template polyhedra subsume, for example, interval and octagon domains.
-\rmcmt{can we rephrase this statement to say that -- In this paper, we used
-interval, octagon and equality ?}
 
@@
 \scalebox{.65}{\import{figures/}{octagons.pspdftex}}
-\caption{Example of an octagon}\label{octagon}
+\caption{Example of an octagon \pscmt{BUG: should be $x\geq -2$, $y\geq -1$; please rename x to $x_1$ and $y$ to $z$}}\label{octagon}
 \vspace*{-3ex}
@@
 For a program with $N$ variables, the matrix $\mat{C}$ for intervals
-generates at most $2N$ inequality constraints, one for the upper and lower
-bounds of each variable.
+has $2N$ rows; and hence it generates at most $2N$ inequalities, one
+for the upper and lower bounds of each variable.
 %
@@
 statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
-The best abstract transformer is \rmcmt{is this equation updated ?}
+The best abstract transformer is 
 \[\llbracket s
-\rrbracket^\sharp(a)=\bigsqcap \{a'\mid a'\sqsubseteq a, a'\models s\}\] where we write \pscmt{fix this}
-$a\models s$ if $\forall \vec{y}\in\gamma(a): (s[\vec{y}/\vec{x}]
-\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
-all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
-of $a$. Any abstract transformer that over-approximates the least
-upper bound is a sound transformer and can be used in our algorithm.
-
-In a forward analysis, we can deduce, for example, $\llbracket
-x=2(y+z)\rrbracket^\sharp(a)=(-2\leq x\leq 6) \sqcap a$ for $a=(0\leq y \leq 2
-\wedge 1 \leq y-z \leq 1)$.
+\rrbracket^\sharp(a)=a\sqcap \alpha(\{c\mid c\models s\})\] where we write 
+$c\models s$ if the concrete value $c$ satisfies the constraint $s$.
+%$\forall \vec{y}\in c: (s[\vec{y}/\vec{x}]
+%\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
+%all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
+%of $a$. 
+Any abstract transformer that over-approximates the best abstract
+transformer is a sound transformer and can be used in our algorithm.
 %
-A detailed description of abstract transformer is presented in Section~\ref{sec:abst}.
+For example, we can deduce $\llbracket
+x=2(y+z)\rrbracket^\sharp(a)=a\sqcap(-2\leq x\leq 6)$ for the abstract value $a=(0\leq y \leq 2 \wedge 1 \leq y-z \leq 1)$.
+%
+%A detailed description of abstract transformer is presented in Section~\ref{sec:abst}.
 

### 2016-10-21 decision.tex

@@
 ACDCL supports several decision heuristics namely, {\em ordered}, 
-{\em longest-range}, {\em random}, {\em relational} and 
+{\em longest-range}, {\em random} and 
 {\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
@@
 %
-The {\em relational} decision heuristics is only relevant for relational 
-abstract domains.  
+%The {\em relational} decision heuristics is only relevant for relational 
+%abstract domains.  
 %

### 2016-10-21 experiments.tex

@@
 \paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
-visualization of different decision heuristics in ACDCL.  Random heuristic
-performs the best for most safe benchmarks and all bit-vector category 
-benchmarks.  Whereas, the ordered heuristic performs better for programs 
-with conditional branches since it prioritises decisions on meet 
-irreducibles that appear in conditional branches over meet irreducibles 
-that involve numerical variables.  Whereas, the runtimes for berkmin 
-heuristic is slightly greater than the random and ordered heuristic 
-for our benchmarks.
+visualization of the performance of different decision heuristics in ACDCL.  
+Note that the runtimes for all decision heuristics are obtained with multi-way 
+propagtion strategy.  The runtimes are very close, but we observe some key
+characteristics of these heuristics.  Berkmin heuristic performs consistently 
+well for most safe benchmarks and all bit-vector category benchmarks.  Whereas, 
+the ordered heuristic performs better for programs with conditional branches 
+since it prioritises decisions on meet irreducibles that appear in conditional 
+branches over meet irreducibles that involve numerical variables.  Whereas, 
+the runtimes for random heuristic is marginally higher than the other two.  This 
+suggest that domain-specific decision heuristics are important for ACDCL.
 

### 2016-10-21 deductions.tex
sync
section 5

@@
 transformers}.  A {\em worklist} is initialized following a heuristics 
-choice for iteration strategies. A {\em forward} iteration 
-strategy initializes the worklist with \rmcmt{abstract} transformers that 
+choice for propagation strategies. A {\em forward} propagation strategy 
+initializes the worklist with \rmcmt{abstract} transformers that 
 has constants in the right-hand side. Whereas, a {\em backward} 

### 2016-10-21 domains.tex
checking consistency of notations, take 1

@@
 % %
-% A {\em program} $\mathcal{P}$ is formally defined as follows.
+% A {\em program} $\program$ is formally defined as follows.
 % \[
 % \begin{array}[t]{@{}lll}
-% \mathcal{P} & {:}{:}{=} & \mathit{Procedure} \\
+% \program & {:}{:}{=} & \mathit{Procedure} \\
 % \mathit{Procedure} & {:}{:}{=} & \mathit{Statement} \mid \mathit{Procedure} \\
@@
 % Consider sets of expressions $Exp$ and Boolean expressions $BExp$
-% over variables $Var$ of $\mathcal{P}$.  The variables in $Var$ 
+% over variables $Var$ of $\program$.  The variables in $Var$ 
 % can take numeric values in $Val$.  A procedure is denoted as 
@@
 $\begin{array}{rcl}
-\mathcal{S}:= (g_0 = (0 \leq v\leq N)) \\
+\conctransset:= (g_0 = (0 \leq v\leq N)) \\
 (g_1 = (g_0 \wedge c)) \wedge (x_0 = v) \wedge \\
@@
 In this paper, we consider \emph{bounded programs} with safety
-properties given as assertions $\mathcal{A}$ in the program.
+properties given as assertions $\assertions$ in the program.
 %
@@
 transformation is represented as set
-$\mathcal{S}=\mathcal{P}\cup\{\neg \bigwedge_{a\in\mathcal{A}} a\}$
-where $\mathcal{P}$ are the statements of the transformed program in the
+$\constraints=\program\cup\{\neg \bigwedge_{\assertion\in\assertions} \assertion\}$
+where $\program$ are the statements of the transformed program in the
 form of constraints as typically obtained by translating the program
@@
 %
-The assertions $\mathcal{A}$ are constraints of the form $g_3
+The assertions $\assertions$ are constraints of the form $g_3
 \Rightarrow z\geq 0$, for instance, meaning that if $g_3$ holds
@@
 Based on the above program representation, we define a safety equation \pscmt{why ``equation''?} as, 
-$\varphi:= \bigwedge_{s\in\mathcal{S}} s$. The formula $\varphi$ is 
+$\formula:= \bigwedge_{\constraint\in\constraints} \constraint$. The formula $\formula$ is 
 unsatisfiable if and only if the program is safe.
@@
 % of constraint variables $Var = \{X_i| i \in N\}$, representing a program
-% $\mathcal{P}$ and a concrete domain $\powerset(Env)$, the static analysis equation 
-% for safety of $\mathcal{P}$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
+% $\program$ and a concrete domain $\powerset(Env)$, the static analysis equation 
+% for safety of $\program$ is given as follows.  Here, $X_{init}, X_{Err}$ denote the initial valuation 
 % and the \rmcmt{final valuation} of constraint variables.
-% \[\varphi = X_{init} \subseteq Env \wedge \underset{p,q \in |\mathcal{P}|}
+% \[\formula = X_{init} \subseteq Env \wedge \underset{p,q \in |\program|}
 % {\bigwedge} \{ X_p \subseteq post_{(p,q)}(X_q) \} \wedge X_{err} \supset \emptyset \]
@@
 %In our prototype implementation and for illustrative purposes 
-In this paper we use a product domain $\mathcal{B}^b\times\mathcal{R}$ where
-$\mathcal{B}$ is the boolean domain
+In this paper we use a product domain $\domain=\booldomain^\booldim\times\reldomain$ where
+$\booldomain$ is the boolean domain
 $\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$, 
-$b$ is the number of boolean variables in the program, 
-and $\mathcal{R}$ is a \emph{relational} domain over the 
+$\booldim$ is the number of boolean variables in the program, 
+and $\reldomain$ is a \emph{relational} domain over the 
 numerical (bitvector) variables of respective
@@
 %
-We instantiate $\mathcal{R}$ with the template polyhedra domain.
+We instantiate $\reldomain$ with the template polyhedra domain.
 
 For notational convenience we will denote elements of
-$\mathcal{B}^b$ by their concretisations to propositional formulae,
+$\booldomain^\booldim$ by their concretisations to propositional formulae,
 For example, the program in Figure~\ref{ssa} has four boolean variables
-written as the vector $(g_0,g_1,c,g_2)$; then we will denote an abstract value $(\top, \false, \true, \top) \in \mathcal{B}^4$ as $\neg g_1 \wedge c$.
+written as the vector $(g_0,g_1,c,g_2)$; then we will denote an abstract value $(\top, \false, \true, \top) \in \booldomain^4$ as $\neg g_1 \wedge c$.
 
@@
 An abstract value of the template polyhedra domain~\cite{sriram}
-represents a set $X$ of values of numerical variables $\vec{x}$. For
+represents a set $\numconcval$ of values of numerical variables $\vec{\numvar}$. For
 example, in the program in Figure~\ref{ssa}, we have four numerical
-variables, written as the vector $\vec{x}:=(x_0,x_1,x_2,z)$.  We can
-represent sets of values for $\vec{x}$ such that
-$\mat{C}\vec{x}\leq\vec{d}$ for a fixed coefficient matrix $\mat{C}$
-and a constant vector $\vec{d}$. The domain of $\vec{d}$ is augmented
+variables, written as the vector $\vec{\numvar}:=(x_0,x_1,x_2,z)$.  We can
+represent sets of values for $\vec{\numvar}$ such that
+$\mat{C}\vec{\numvar}\leq\vec{\numabsval}$ for a fixed coefficient matrix $\mat{C}$
+and a constant vector $\vec{\numabsval}$. The domain of $\vec{\numabsval}$ is augmented
 by a special element $\bot$ to denote the minimal element of the
@@
 %
-The abstraction function is defined as $\alpha(X) = \min \{\vec{d}\mid
-\mat{C}\vec{x}\leq\vec{d}, \vec{x}\in X\}$ where $\min$ is applied
+The abstraction function is defined as $\alpha(\numconcval) = \min \{\vec{\numabsval}\mid
+\mat{C}\vec{\numvar}\leq\vec{\numabsval}, \vec{\numvar}\in \numconcval\}$ where $\min$ is applied
 component-wise. There are several optimisation-based
 techniques~\cite{sriram,GS07b,BJKS15} for computing the domain
-operations, such as $\sqcap$, $\sqcup$, etc. In our implementation we
+operations, such as $\meet$, $\sqcup$, etc. In our implementation we
 use the strategy iteration approach available in 2LS~\cite{BJKS15}.
 %
-The concretisation $\gamma(\vec{d})$ is the set $\{\vec{x}\mid
-\mat{C}\vec{x}\leq\vec{d}\}$; $\gamma(\bot)=\emptyset$, i.e.\ the empty polyhedron.
+The concretisation $\gamma(\vec{\numabsval})$ is the set $\{\vec{\numvar}\mid
+\mat{C}\vec{\numvar}\leq\vec{\numabsval}\}$; $\gamma(\bot)=\emptyset$, i.e.\ the empty polyhedron.
 %
@@
 inequalities, e.g.\ $x_1\geq 0 \wedge x_1-z\leq 30$ to denote the
-abstract value $\vec{d}=\vecv{0}{30}$ of the template polyhedra domain
-with $\mat{C}=\qmat{-1}{0}{1}{-1}$ and $\vec{x}=\vecv{x_1}{z}$;
+abstract value $\vec{\numabsval}=\vecv{0}{30}$ of the template polyhedra domain
+with $\mat{C}=\qmat{-1}{0}{1}{-1}$ and $\vec{\numvar}=\vecv{x_1}{z}$;
 $\true$ is $\top$ and $\false$ is $\bot$.  Template polyhedra can
@@
 %
-An abstract transformer transforms an abstract value $a$ through a
-statement $s$, i.e.\ it \emph{deduces} information from $a$ and $s$.
+An abstract transformer $\abstrans{\domain}{\constraint}$ transforms an abstract value $\absval$ through a
+constraint $\constraint$, i.e.\ it \emph{deduces} information from $\absval$ and $\constraint$.
 The best abstract transformer is 
-\[\llbracket s
-\rrbracket^\sharp(a)=a\sqcap \alpha(\{c\mid c\models s\})\] where we write 
-$c\models s$ if the concrete value $c$ satisfies the constraint $s$.
-%$\forall \vec{y}\in c: (s[\vec{y}/\vec{x}]
+\[\abstrans{\domain}{\constraint}(\absval)=\absval\meet \alpha(\{\val\mid \val\models \constraint\})\] where we write 
+$\val\models \constraint$ if the concrete value $\val$ satisfies the constraint $\constraint$.
+%$\forall \vec{y}\in c: (s[\vec{y}/\vec{\numvar}]
 %\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
-%all values $\vec{y}$ of its variables $\vec{x}$ in the concretisation
+%all values $\vec{y}$ of its variables $\vec{\numvar}$ in the concretisation
 %of $a$. 
@@
 %
-For example, we can deduce $\llbracket
-x=2(y+z)\rrbracket^\sharp(a)=a\sqcap(-2\leq x\leq 6)$ for the abstract value $a=(0\leq y \leq 2 \wedge 1 \leq y-z \leq 1)$.
+For example, we can deduce $\abstrans{\domain}{x=2(y+z)}(\absval)=\absval\meet(-2\leq x\leq 6)$ for the abstract value $\absval=(0\leq y \leq 2 \wedge 1 \leq y-z \leq 1)$.
 %
@@
 %
-The meet irreducibles in the boolean domain $\mathcal{B}$ for a variable $x$ are 
+The meet irreducibles in the boolean domain $\booldomain$ for a variable $x$ are 
 $x$ and $\neg x$. The meet 
@@
 \begin{definition}
-A \emph{meet decomposition} $\mathit{decomp}(a)$ of an abstract
-element $a \in A$ is a set of meet irreducibles $M \subseteq A$ such that 
-$a=\bigsqcap_{m\in M} m$.
+A \emph{meet decomposition} $\decomp(\absval)$ of an abstract
+element $\absval \in \domain$ is a set of meet irreducibles $M \subseteq \domain$ such that 
+$\absval=\bigsqcap_{m\in M} m$.
 %$forall m_i \in M, \meet(m_i) = a$, where $max(i) = |M|$.
@@
 \begin{definition} 
-An element $a\in A$ is called \emph{precisely complementable}
-iff $\exists \bar{a} \in A: \neg\gamma(\bar{a})=\gamma(a)$, 
+An element $\absval\in \domain$ is called \emph{precisely complementable}
+iff $\exists \bar{\absval} \in \domain: \neg\gamma(\bar{\absval})=\gamma(\absval)$, 
 i.e.\ there is an element whose negated concretisation equals
-the concretisation of $a$.
+the concretisation of $\absval$.
 %m\meet\bar{m}=\bot \wedge m\join\bar{m}=\top$.
@@
 \Omit {
-Note that for many domains $A$ including template polyhedra
+Note that for many domains $A$, including template polyhedra,
 most domain elements are not precisely complementable within $A$.
@@
 
+\pscmt{check}
 \begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
@@
 value or to $\bot$.  An abstract valuation $(v)$ abstractly satisfies a formula 
-$\varphi$ if for every variable $x$ in $\varphi$, there is a concrete solution 
+$\formula$ if for every variable $x$ in $\formula$, there is a concrete solution 
 $(c)$ such that $c(x) \subseteq \gamma \circ v(x)$ holds. 
@@
 % \begin{definition}{(Abstract Deduction Transformer)} An abstract deduction
-% transformer, $\widehat{ded_{\varphi}}$ for a formula $\varphi$ over an abstract 
+% transformer, $\widehat{ded_{\formula}}$ for a formula $\formula$ over an abstract 
 % domain $A$ is a sound approximation of a concrete model transformer
-% $ded_{\varphi}$, given by $\widehat{ded_{\varphi}} : A \rightarrow A$, such that 
-% $\forall a \in A: \widehat{ded_{\varphi}}(a) \in \{\top, \bot, m\}$, where 
+% $ded_{\formula}$, given by $\widehat{ded_{\formula}} : A \rightarrow A$, such that 
+% $\forall a \in A: \widehat{ded_{\formula}}(a) \in \{\top, \bot, m\}$, where 
 % $m \in A$ is a meet irreducible.   
@@
 
-% Let us consider a formula $\varphi = (x:=y-1)$ to be analyzed over 
+% Let us consider a formula $\formula = (x:=y-1)$ to be analyzed over 
 % an interval abstract domain, $A = ItvDom$, and let $a = \langle y:[3, 5]
-% \rangle \in ItvDom$, then $\widehat{ded_{\varphi}}(a) = a \meet \langle x:[2, 4]
+% \rangle \in ItvDom$, then $\widehat{ded_{\formula}}(a) = a \meet \langle x:[2, 4]
 % \rangle$.  An abstract deduction transformer is typically computed in the form 

### 2016-10-21 experiments.tex

@@
 ACDCL is instantiated with a product domain of booleans and template 
-polyhedra.  Additionally, the analysis using ACDCL is configured with 
+polyhedra.  \rmcmt{We use an interval and octagon templates for our
+experiments.} Additionally, the analysis using ACDCL is configured with 
 a heuristic choice of decision (ordered, random, berkmin), 

### 2016-10-21 algorithm.tex
checking consistency of notations, take 2

--- 2016-10-21 algorithm.tex
@@
 \DontPrintSemicolon
-\SetKwFunction{decide}{Decide}
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{backtrack}{backtrack}
-\SetKwFunction{print}{print}
 \SetKw{return}{return}
-\SetKwFunction{nondet}{nondet}
-\SetKwFunction{run}{run}
-\SetKwFunction{learn}{learn}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
 \SetKwRepeat{Do}{do}{while}
@@
 \begin{small}
-\Input{%An abstract domain $A$, 
-A program $\mathcal{S}$, a
-  propagation heuristics $H$, a decision heuristics $\mathit{decide}$,
-and a conflict analysis heuristics $\mathit{conflict\_analysis}$.\rmcmt{calling
-them heuristics is not a good idea since these are key phases of the cdcl
-algorithm. Each phase may have a heuristics for some sub phase but the algorithm
-of each phase still remians a procedure and not a heuristic}}
+\Input{A program in the form of a set of abstract transformers $\abstransset$.}
 \Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
-$\mathcal{T} \leftarrow \langle\rangle$ \;
-$\mathcal{R} \leftarrow \emptyset$ \;
-$\mathit{result} \leftarrow \mathit{deduce}(S,\mathcal{T},\mathcal{R}, H)$ \;
+$\trail \leftarrow \langle\rangle$ \;
+$\reasons \leftarrow []$ \;
+$\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
 \lIf{$\mathit{result}$ = \conflict} {
@@
 \lIf{$\mathit{result}$ = \sat} {
-  \return \unsafe}
-  $d \leftarrow \mathit{decide}(\mathit{abs}(\mathcal{T}))$ \;
-%  $v \leftarrow v \meet d$ \; 
-  $\mathcal{T} \leftarrow \mathcal{T} \cdot d$ \; 
-  $\mathcal{R}[|\mathcal{T}|] \leftarrow \top$ \;
-  $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$\;
+  \return (\unsafe,$\abs(\trail)$)}
+  $d \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
+  $\trail \leftarrow \trail \cdot d$ \; 
+  $\reasons[|\trail|] \leftarrow \top$ \;
+  $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$\;
   \Do{$\mathit{result} = \conflict$} {
-%    learn() \;
-    \lIf{$\neg \mathit{conflict\_analysis}(\mathcal{S},\mathcal{T},\mathcal{R})$} {
+    \lIf{$\neg \analyzeconflict_{\confheur}(\abstransset,\trail,\reasons)$} {
       \return \safe
     }
-    $\mathit{result} \leftarrow \mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R})$ \;
+    $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
   }
@@
 \end{small}
-\caption{Abstract Conflict Driven Clause Learning $ACDCL(A)$ \label{Alg:acdcl}}
+\caption{Abstract Conflict Driven Clause Learning $ACDCL_{\propheur,\decheur,\confheur}(\abstransset)$ \label{Alg:acdcl}}
 \end{algorithm2e}
@@
   (parametrised by the propagation heuristics)}
-$\mathcal{S}$ and propagation, decision, and conflict analysis
-heuristics as inputs.  As internal state, it maintains a propagation
-trail $\mathcal{T}$ and a reason trail \pscmt{map?} $\mathcal{R}$.
+$\abstransset$ and propagation, decision, and conflict analysis
+heuristics as inputs.  
+\rmcmt{Approximation of the concrete transformers in 
+$\abstransset$ are typically available in abstract domain in the 
+form of strongest-post condition or weakest pre-condition. } 
+As internal state, it maintains a propagation
+trail $\trail$ and a reason trail \pscmt{map?} $\reasons$.
 The propagation trail is an initially empty sequence of meet
@@
 and decisions). The reason trail maps \pscmt{certain?} elements of the
-propagation trail to transformers $s\in\mathcal{S}$ that were used to
-derive them. The \emph{abstract value} $\mathit{abs}(\mathcal{T})$
-corresponding to the propagation trail $\mathcal{T}$ is the
+propagation trail to transformers $s\in\abstransset$ that were used to
+derive them. The \emph{abstract value} $\abs(\trail)$
+corresponding to the propagation trail $\trail$ is the
 conjunction of the meet irreducibles on the trail:
-$\mathit{abs}(\mathcal{T})=\bigsqcap_{m \in \mathcal{T}}m$ with
-$\mathit{abs}(\mathcal{T})=\top$ if $\mathcal{T}$ is the empty
+$\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
+$\abs(\trail)=\top$ if $\trail$ is the empty
 sequence $\langle\rangle$.
-\pscmt{trail refinement}
+\pscmt{define trail refinement}
+\pscmt{define concatenation operator $\cdot$}
+\pscmt{explain $\top$ assignment to $\reasons$}
 
 The procedure $deduce$ computes a greatest fixed point over the
-transformers in $\mathcal{S}$ that refines the abstract value,
+transformers in $\abstransset$ that refines the abstract value,
 similar to the BCP step in SAT solvers.  If the result of $deduce$
 returns \textsf{conflict} ($\bot$), the algorithm terminates with
-\textsf{safe}.  Else, the analysis enters into the while($true$) loop
-and makes a new decision by a call to $\mathit{decide}$ which returns
+\textsf{safe}.  Else, the analysis enters into the while($\true$) loop
+and makes a new decision by a call to $\decide$ which returns
 a new meet irreducible $d$ that is consistent with the state of the
@@
   -- then we should either have deduced $\bot$ or
-  $\mathit{abs}(\mathcal{T})$ must be $\gamma$-complete} If $d$
+  $\abs(\trail)$ must be $\gamma$-complete} If $d$
 refines the current abstract value, the corresponding propagation
-and reason trail, $\mathcal{T}$ and $\mathcal{R}$ are updated.
+and reason trail, $\trail$ and $\reasons$ are updated.
 %
@@
 
-The procedure $\mathit{deduce}$ is called next to infer new meet
+The procedure $\deduce$ is called next to infer new meet
 irreducibles taking into account the current decision.  The model search phase
-alternates between the decision and deduction until $\mathit{deduce}$
+alternates between the decision and deduction until $\deduce$
 returns either \textsf{sat} or \textsf{conflict}.
 In the former case, we have found an abstract value that represents 
-models of $\varphi$ and thus we return \textsf{unsafe}.
+models of $\formula$ and thus we return \textsf{unsafe}.
 %
-% has found an abstract value that represents a set of models of $\varphi$ or the
+% has found an abstract value that represents a set of models of $\formula$ or the
 % deduction leads to a \textsf{conflict}.  Checking whether the abstract
-% model concretises to a set of models of $\varphi$; this
+% model concretises to a set of models of $\formula$; this
 % corresponds to a $\gamma$-completeness~\cite{dhk2013-popl} check in
@@
 % \textsf{unsafe} or the current abstraction is insufficient to
-% determine the satisfiability of $\varphi$.
+% determine the satisfiability of $\formula$.
 
 In the latter case the algorithm enters in the
-$\mathit{conflict\_analysis}$ phase to learn the reason for the
+$\analyzeconflict$ phase to learn the reason for the
 conflict.  There can be multiple incomparable reasons for conflict --
@@
 long as the result of deduction is conflicting ($\bot$), that is, the
-abstract valuation $v$ does not abstractly satisfy the formula
+abstract value does not abstractly satisfy the formula
 \pscmt{define}.  If no further backtrack is possible, then the
-algorithm terminates and $\varphi$ is \textsf{safe}. Else, the
+algorithm terminates and $\formula$ is \textsf{safe}. Else, the
 algorithm makes a new decision and the above process is repeated until
@@
 
-%\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\mathcal{T}$?} The state of a ACDCL solver is a tuple 
+%\paragraph{Solver State.}  \pscmt{This paragraph seems redundant. Doesn't this redefine $\trail$?} The state of a ACDCL solver is a tuple 
 %of the form $\langle \mathcal{E}, S \rangle$.  Here, $\mathcal{E}$ 
@@
 %and $t$ is a truth assignment (true or false).  Whereas, a trail 
-%$\mathcal{T}$ in ACDCL contains a sequence of meet irreducibles 
+%$\trail$ in ACDCL contains a sequence of meet irreducibles 
 %inferred by deduction or decision phase where a variable in trail 
@@
 %A current valuation $v$ is a meet of all elements of trail
-%$\mathcal{T}$, such that $v = \top$ when $\mathcal{T}$ is 
-%empty or $v=\underset{i \geq 0 \wedge i \leq |\mathcal{T}|}{\meet m_i}$, where 
-%$m_i \in \mathcal{T}$.  A solver is in conflict if some clauses in
-%$\mathcal{S}$ is not abstractly satisfied by $v$.
+%$\trail$, such that $v = \top$ when $\trail$ is 
+%empty or $v=\underset{i \geq 0 \wedge i \leq |\trail|}{\meet m_i}$, where 
+%$m_i \in \trail$.  A solver is in conflict if some clauses in
+%$\abstransset$ is not abstractly satisfied by $v$.
 %}

### 2016-10-21 deductions.tex
macrofying identifiers up to 5.1

@@
 \DontPrintSemicolon
-\SetKwFunction{deduce}{deduce}
-\SetKwFunction{print}{print}
 \SetKw{return}{return}
-\SetKwFunction{continue}{continue}
-\SetKwFunction{assign}{assign}
 \SetKwData{sat}{sat}
@@
 \begin{small}
-\Input{%Abstract domain $A$
-Program $\mathcal{S}$,
-propagation trail $\mathcal{T}$, reason trail $\mathcal{R}$,
-propagation heuristics $H$}
+\Input{A program in the form of a set of abstract transformers $\abstransset$,
+a propagation trail $\trail$, and a reason trail $\reasons$.}
 \Output{\sat or \conflict or \unknown}
-$\mathit{worklist} \leftarrow H.\mathit{initialize\_worklist}(\mathcal{S})$ \;
+$\worklist \leftarrow \initworklist_{\propheur}(\abstransset)$ \;
 \While{$!\mathit{worklist.empty}()$} 
 {
-  $(s,\mathit{live\_vars}) \leftarrow \mathit{worklist.pop}()$ \;
-  $v \leftarrow ded(s, \mathit{espan}(\mathit{live\_vars}), \mathit{abs}(\mathcal{T}))$\;
+%  $(\abstransel,\livevars) \leftarrow \mathit{worklist.pop}()$ \;
+  $\abstransel{\subdomain} \leftarrow \mathit{worklist.pop}()$ \;
+%  $v \leftarrow ded(s, \makesubdomain(\livevars), \abs(\trail))$\;
+  $v \leftarrow \abstransel{\subdomain}(\abs(\trail))$\;
   \uIf{$v = \bot$} {
-    $\mathcal{R}[\bot] \leftarrow s$ \;
+    $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \;
 %\pscmt{how is this used in the conflict
-%    analysis? wouldn't it be more uniform to add $\bot$ to the $\mathcal{T}$?}
-%    \rmcmt{UIP uses the transformer stored at $\mathcal{R}[\bot]$ to derive the
+%    analysis? wouldn't it be more uniform to add $\bot$ to the $\trail$?}
+%    \rmcmt{UIP uses the transformer stored at $\reasons[\bot]$ to derive the
 %    cut. No, we dont want to make the trail inconsistent by adding $\bot$}
@@
   {
-    $\mathcal{T} \leftarrow \mathcal{T} \cdot \mathit{decomp}(v)$ \; 
+    $\trail \leftarrow \trail \cdot \decomp(v)$ \; 
 %    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
-    $\mathcal{R}[|\mathcal{T}|] \leftarrow s$ \;
-    $H.\mathit{update\_worklist}(v, \mathit{worklist})$ \; 
+    $\reasons[|\trail|] \leftarrow s$ \;
+    $\updateworklist_{\propheur}(v, \worklist)$ \; 
   }
 }
-\lIf{$\mathit{abs}(\mathcal{T})$ is $\gamma$-complete} {
+\lIf{$\abs(\trail)$ is $\gamma$-complete} {
   \return \sat;
@@
 \end{small}
-\caption{Abstract Model Search $\mathit{deduce}(\mathcal{S},\mathcal{T},\mathcal{R},H)$ \label{Alg:ms}}
+\caption{Abstract Model Search $\mathit{deduce}_{\propheur}(\abstransset,\trail,\reasons)$ \label{Alg:ms}}
 \end{algorithm2e}
@@
 %-------------------------------------------------------------------------------
-\subsection{Abstract Transformers} \label{sec:abst}
+\subsection{Abstract Deduction Transformers} \label{sec:abst}
 %-------------------------------------------------------------------------------
@@
 We thus define a specialised variant of the transformer to compute
-deductions w.r.t.\ a given subdomain $L\subseteq A$; $\bot,\top \in L$.
-\[\llbracket s \rrbracket^\sharp(a,L)=\bigsqcup \{a'\mid a'\in L, a' \sqsubseteq a, a'\models s\}\]
-We have the property $\llbracket s \rrbracket^\sharp(a)\sqsubseteq \llbracket s \rrbracket^\sharp(a,L)$.
+deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$.
+A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$.
+\[\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\models \constraint\})\]
+For $\subdomain=\domain$, the abstract transformer behaves in the same way as the one defined in Section~\ref{sec:domains}.
 %
+The restriction to a subdomain makes it less precise. Therefore, we have the property $\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq \abstrans{\domain}{\constraint}^\subdomain(\absval)$.
+%
+\pscmt{explain $\makesubdomain(Vars)$}
 For example, for
-$a=(0\leq y \leq 1 \wedge 5\leq z)$ we have
-$\llbracket x=y+z\rrbracket^\sharp(a,\{x\})=(x\geq 6)$, whereas
-$\llbracket x=y+z\rrbracket^\sharp(a,\{y,z\})=\top$.
+$\absval=(0\leq y \leq 1 \wedge 5\leq z)$ we have
+$\abstrans{\domain}{x=y+z}^{\makesubdomain(\{x\})}(\absval)=(x\geq 6)$, whereas
+$\abstrans{\domain}{x=y+z}^{\makesubdomain(\{y,z\})}=\top$.
 %
-If $L=A$ then intuitively $a$ is intersected with the constraint $s$.
-Yet, with the help of $L$ we can also control whether to perform
-forward (right-hand side to left-hand side, $L=\{x\}$ in the above
-example) or backward (lhs to rhs, $L=\{y,z\}$) propagation.
+The advantage of this definition is that with the help of $\subdomain$
+we can now control how we propagate within our constraint system.
+%
+For the example above, $\subdomain=\makesubdomain(\{x\})$ performs a
+right-hand side to left-hand side propagation and hence emulates a
+forward analysis on the input program; whereas with
+$\subdomain=\makesubdomain(\{y,z\})$) we emulate a backward analysis.
+%
+We call the propagation with $\subdomain=\domain$ \emph{multi-way
+  propagation}, which simultaneously performs forward and backward
+propagation.
 
@@
 search procedure. \pscmt{The propagation heuristics $H$ provides two
-  functions $\mathit{initialize\_worklist}$ and
-  $\mathit{update\_worklist}$. Describe them for the different
+  functions $\initworklist$ and
+  $\updateworklist$. Describe them for the different
   heuristics (already done for initialize, TBD for update...}
@@
 passed to the abstract domain, if the domain infers a new meet irreducible 
-$\mathit{meet\_irrd}$, it is added to $\mathcal{T}$ \pscmt{It usually infers several ones.}.  Additionally, the reason 
-trail $\mathcal{R}$ \pscmt{define} is updated with the transformer $s$ as the reason for inferring
+$\mathit{meet\_irrd}$, it is added to $\trail$ \pscmt{It usually infers several ones.}.  Additionally, the reason 
+trail $\reasons$ \pscmt{define} is updated with the transformer $s$ as the reason for inferring
 $\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 

### 2016-10-21 domains.tex
macrofying identifiers up to 5.1

@@
-\section{Program Model and Abstract Domain}
+\section{Program Model and Abstract Domain}\label{sec:domains}
 This section introduces the program model and identifies properties of abstract

### 2016-10-22 algorithm.tex
moved back from bitbucket to github

@@
   \return (\unsafe,$\abs(\trail)$)}
-  $d \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
-  $\trail \leftarrow \trail \cdot d$ \; 
+  $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
+  $\trail \leftarrow \trail \cdot \decisionvar$ \; 
   $\reasons[|\trail|] \leftarrow \top$ \;
@@
 irreducibles inferred by the abstract model search phase (deductions
-and decisions). The reason trail maps \pscmt{certain?} elements of the
+and decisions). The reason trail maps the elements of the
 propagation trail to transformers $s\in\abstransset$ that were used to
@@
 sequence $\langle\rangle$.
-\pscmt{define trail refinement}
-\pscmt{define concatenation operator $\cdot$}
-\pscmt{explain $\top$ assignment to $\reasons$}
 
-The procedure $deduce$ computes a greatest fixed point over the
+The procedure $\deduce$ computes a greatest fixed point over the
 transformers in $\abstransset$ that refines the abstract value,
-similar to the BCP step in SAT solvers.  If the result of $deduce$
+similar to the BCP step in SAT solvers.  If the result of $\deduce$
 returns \textsf{conflict} ($\bot$), the algorithm terminates with
 \textsf{safe}.  Else, the analysis enters into the while($\true$) loop
-and makes a new decision by a call to $\decide$ which returns
-a new meet irreducible $d$ that is consistent with the state of the
-solver. \pscmt{what happens if we cannot make any decision any more?
-  -- then we should either have deduced $\bot$ or
-  $\abs(\trail)$ must be $\gamma$-complete} If $d$
-refines the current abstract value, the corresponding propagation
-and reason trail, $\trail$ and $\reasons$ are updated.
+and makes a new decision by a call to $\decide$ which returns a new
+meet irreducible $\decisionvar$.
+%
+\pscmt{what happens if we cannot make any decision any more?
+  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
+  be $\gamma$-complete} 
+%
+We concatenate $\decisionvar$ to the
+trail $\trail$.  The decision $\decisionvar$ therefore refines the current
+abstract value $\abs(\trail)$ represented by the trail.
+% WE ExPLAIN THAT LATER
+%However, $\decisionvar$ must be consistent with the state of the
+%solver, i.e. $\abs(\trail\cdot \decisionvar)\neq \bot$.
 %
@@
 intervals for variables.
+%
+We set the corresponding entry in the reason trail $\reasons$ to $\top$
+to mark it as a decision.
 % so the analysis jumps under a 
@@
 
-The procedure $\deduce$ is called next to infer new meet
-irreducibles taking into account the current decision.  The model search phase
-alternates between the decision and deduction until $\deduce$
-returns either \textsf{sat} or \textsf{conflict}.
-In the former case, we have found an abstract value that represents 
-models of $\formula$ and thus we return \textsf{unsafe}.
+The procedure $\deduce$ is called next to infer new meet irreducibles
+taking into account the current decision.  The model search phase
+alternates between the decision and deduction until $\deduce$ returns
+either \textsf{sat} or \textsf{conflict}.  In the former case, we have
+found an abstract value that represents models of $\formula$, which
+are counterexamples to the safety problem, and thus we return
+\textsf{unsafe}.
 %
@@
 
-In the latter case the algorithm enters in the
-$\analyzeconflict$ phase to learn the reason for the
-conflict.  There can be multiple incomparable reasons for conflict --
-based on the choice of Unique Implication Point (UIP)~\cite{cdcl}.
-ACDCL heuristically chooses one reason.  A learnt clause must include
-asserting cuts which guarantees derivation of new meet irreducibles
-after backtracking. The clause learning and backtracking continues as
-long as the result of deduction is conflicting ($\bot$), that is, the
-abstract value does not abstractly satisfy the formula
-\pscmt{define}.  If no further backtrack is possible, then the
-algorithm terminates and $\formula$ is \textsf{safe}. Else, the
-algorithm makes a new decision and the above process is repeated until
-a real counterexample \pscmt{is this a model?} is obtained or the
+In the latter case the algorithm enters in the $\analyzeconflict$
+phase to learn the reason for the conflict.   There can be multiple
+incomparable reasons for conflict \pscmt{do you have to explain this here?} -- based on the choice of Unique
+Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically chooses one
+reason.  A learnt clause must include asserting cuts which guarantees
+derivation of new meet irreducibles after backtracking. The clause
+learning and backtracking continues as long as the result of deduction
+is conflicting ($\bot$), that is, the abstract value does not
+abstractly satisfy the formula \pscmt{define}.  If no further
+backtrack is possible, then the algorithm terminates and $\formula$ is
+\textsf{safe}. Else, the algorithm makes a new decision and the above
+process is repeated until $\deduce$ returns \textsf{sat} or the
 algorithm backtracks to decision level 0 after a conflict in which

### 2016-10-22 background.tex
moved back from bitbucket to github

--- 2016-10-19 background.tex
@@
-\section{Related Work}
-The work of~\cite{sas13} present an instantiation of ACDCL as a decision 
-procedure for floating-point arithmetic that uses interval constraint 
-propagation.  The use of ACDCL as program analysis is proposed
-in~\cite{sas12,tacas12}.  The analysis operate over the abstract domain of 
-floating-point and machine-integer.  Thus, it is restricted to bounds 
-checking for numerical programs.  A similar technique that lifts DPLL(T) 
-to programs is Satisfiability Modulo Path Programs (SMPP). SMPP enumerates 
-path programs using a SAT formula which are then verified using abstract 
-interpretation.  The lifting of CDCL to first-order theory is proposed 
+\paragraph{Related Work}
+The work of~\cite{sas13,DBLP:journals/fmsd/BrainDGHK14} present an
+instantiation of ACDCL as a decision procedure for floating-point
+arithmetic that uses interval constraint propagation.  
+%
+The use of
+ACDCL as program analysis is proposed in~\cite{tacas12}.  The analysis
+operates over the abstract domain of floating-point numbers and
+machine integers.  Thus, it is restricted to bounds checking for
+numerical programs.  
+%
+A similar technique that lifts DPLL(T) to
+programs is Satisfiability Modulo Path Programs (SMPP) \cite{SMPP}. SMPP
+enumerates path programs using a SAT formula which are then verified
+using abstract interpretation.  
+%
+The lifting of CDCL to first-order theory is proposed
 in~\cite{dpll,cp09} and natural domain SMT framework~\cite{ndsmt}.
-\Omit{
-operates on a fixed first-order partial assignment lattice structure, 
-where first-order variables are mapped to domain values, similar to constants lattice 
-in program analysis.  
-}
-However, unlike previous works that operate on a fixed first-order 
-lattice structure, ACDCL can be instantiated with different 
-abstract domains.  This involves model search and learning in 
-abstract lattices.     
-\Omit {
-The abstract lattice in natural domain SMT does 
-not have complementable meet irreducibles, and therefore does not 
-support generalized clause learning~\cite{sas}.  On the other hand, 
-most abstract lattice have complementation property, thus enabling 
-ACDCL to perform generalized clause learning.
-}
+\Omit{ operates on a fixed first-order partial assignment lattice
+  structure, where first-order variables are mapped to domain values,
+  similar to constants lattice in program analysis.  } However, unlike
+previous works that operate on a fixed first-order lattice structure,
+ACDCL can be instantiated with different abstract domains.  This
+involves model search and learning in abstract lattices.  \Omit { The
+  abstract lattice in natural domain SMT does not have complementable
+  meet irreducibles, and therefore does not support generalized clause
+  learning~\cite{sas}.  On the other hand, most abstract lattice have
+  complementation property, thus enabling ACDCL to perform generalized
+  clause learning.  } 
+%
+\cite{DBLP:journals/fmsd/BrainDGHK14} presents an instantiation of ACDCL for heap-manipulating programs, but uses a simplistic conflict analysis.
+%
+\cite{DBLP:conf/vmcai/PelleauMTB13} describes a constraint solving algorithm using relational numerical domains, but performs model search only.
+%
+\cite{DBLP:conf/esop/MineBR16} proposes an algorithm inspired by constraint solvers for inferring disjunctive invariants using the interval domain.
+% is presented in . They
+%use a relational domain, but focus on the model search, whereas the
+%conflict analysis is simplistic and restarts the algorithm with the
+%learnt clause instead of backtracking.
+%}

### 2016-10-22 conclusion.tex
moved back from bitbucket to github

--- 2016-10-20 conclusion.tex
@@
 In this paper, we present a procedure for lifting the model search 
-in satisfiability solvers to relational abstract domains.  We explain 
-specific properties of abstract domain elements that are necessary to infer 
-domain-specific deductions.  We present decision heuristics that takes 
-into account the high-level structure of programs and the underlying 
-expressivity of the abstract domains.  Our conflict analysis procedure 
-learns abstract transformer following UIP computation over relational 
-elements.  We instantiate our algorithms as a program analyzer to determine
-safety of C programs.  Experimental evaluation shows that ACDCL is able 
-to suitably exploit the expressivity of richer abstract domains as well 
-as precision of CDCL architecture.
+in satisfiability solvers to relational abstract domains.  
+%We explain specific properties of abstract domain elements that are necessary to infer domain-specific deductions.  
+We present a parameterised abstract transformer that allows us to
+flexibly implement various propagation heuristics while taking care of
+the closure operation required in relational abstract domains.
+%
+Our decision heuristics takes into account the high-level structure of
+programs and the underlying expressivity of the abstract domains.  Our
+conflict analysis procedure learns abstract transformers following a
+UIP computation over relational domain elements.  We instantiate our
+algorithms as a program analyser to determine safety of C programs.
+%
+Experimental evaluation shows that ACDCL is able to suitably exploit
+the expressivity of richer abstract domains as well as precision of
+CDCL architecture.
 
-Generalizations of conflict reasons for relational domains 
-and non-relational domain such as trace-based abstrac domains is a 
+\Omit{
+Generalisations of conflict reasons for relational domains 
+and non-relational domain such as trace-based abstract domains is a 
 possible direction of future work.  We believe that domain refinement 
@@
 
-\Omit{
 Experimental evaluation on different benchmark sets shows that

### 2016-10-22 contributions.tex
moved back from bitbucket to github

--- 2016-10-20 contributions.tex
@@
 lifts the CDCL algorithm to relational numerical abstract domains.   
-We have implemented our technique in 2LS verification tool.  
+We have implemented our technique in the 2LS verification tool.  
 \Omit{
@@
 generalizes the Boolean Constraint Propagation (BCP) procedure to
-compute a fixed point over a relational domain.  Our abstract 
-conflict analysis algorithm present a non-trivial lifting of the 
-Unique Implication Point (UIP) procedure to relational abstract 
-domains in order to learn a generalized reason for conflicts.
+compute a fixed point over a relational domain.
 
-\item  Experiments indicate two order of magnitude reduction 
+\item We propose several domain-specific decision heuristics that 
+exploit the expressivity of the underlying abstract domain to make 
+effective decisions.
+
+\item We present a non-trivial lifting of the conflict analysis procedure 
+to relational abstract domains in order to learn a generalised reason 
+for conflicts.
+
+\item  Experiments indicate a two order of magnitude reduction 
 in the number of decisions, propagations, and conflicts compared 
-to a SAT solver while being more precise compared to an abstract interpreter.  
+to a SAT solver while being more precise than an abstract interpreter.  
 \Omit {

### 2016-10-22 decision.tex
moved back from bitbucket to github

@@
 \subsection{Decision}
-A decision $\mathcal{D}$ is a meet irreducible that refines the 
-current abstract valuation.  For example, a decision in interval 
-domain can be of the form $\langle x R c \rangle$ where 
-$R \in \{\leq, \geq\}$, $c$ is the bound.  A decision in 
-relational domain such as octagon is of the form $ax - by \leq c$, 
-where $x$ and $y$ are variables, $a,b \in \{-1,0,1\}$ are co-efficients, 
-and $c \in \mathbb{R}\cup\{\infty\}$ is the bound of the inequality.  
-That is, a decision can specify relation between variables, for example 
+A decision $\decisionvar$ is a meet irreducible that refines the
+current abstract value $\abs(\trail)$.  Note that meet irreducibles in
+relational domains may involve several variables.
+%
+% For example, a decision in interval 
+%domain can be of the form $\langle x R c \rangle$ where 
+%$R \in \{\leq, \geq\}$, $c$ is the bound.  A decision in 
+%relational domain such as octagon is of the form $ax - by \leq c$, 
+%where $x$ and $y$ are variables, $a,b \in \{-1,0,1\}$ are co-efficients, 
+%and $c \in \mathbb{R}\cup\{\infty\}$ is the bound of the inequality.  
+%
+Hence, a decision can specify a relation between variables, for example 
 $\{x \leq y)$.  A decision must always be consistent with respect to 
-the trail $\mathcal{T}$, that is, it must not contradict with the elements 
-in the trail.  A new decision always increases the decision level by one.
+the trail $\trail$, i.e.\ $\abs(\trail\cdot \decisionvar)\neq \bot$.  
+A new decision always increases the decision level by one.
 
-A singleton meet irreducible $s$ in an abstract domain $A$ is an element 
-whose lower bound and upper bound are the same.  For example, 
-$\langle x:[1,1] \rangle$ is a singleton meet irreducible in interval domain.  
-For an octagon domain, $\langle 1 \geq x-y \leq 1 \rangle$ is a singleton meet 
-irreducible for the template $(x-y)$, though the concretisation of individual 
-template variables ($x$, $y$) may not be singletons.
+We call a \emph{singleton} a meet irreducible which does represent a
+valid decision.
+%
+For template polyhedra, singletons are the meet irreducibles
+corresponding to a pair of rows $\vec{c}_1,\vec{c}_2$ in matrix
+$\mat{C}$ with $\vec{c}_1\vec{\numvar}=-\vec{c}_2\vec{\numvar}$
+(i.e.\ \emph{matching} rows) and the corresponding bounds
+$\numabsval_1=-\numabsval_1$.
+%
+A singleton in the interval domain corresponds to a singleton interval
+such as $x\leq 1 \wedge -x\leq -1$ For the octagon domain, $1 \leq x-y
+\leq 1$ is a singleton.
+%
+We cannot make decisions for template rows of singletons because there
+is no choice left. Similarly, in a SAT solver, we cannot make a
+decision on a variable that is already assigned $\true$ or $\false$.
+%
+Note that for relational domains singletons do not necessarily
+concretise to singleton sets of concrete values for the variables
+involved.
 
-Given a set of variables $\mathcal{V}$ and a current abstract value $v$, 
-a decision phase heuristically returns a non-singleton meet irreducible 
-over a set of branching variables $\{B\} \subseteq \mathcal{V}$, a 
-bound $c$, and a polarity ($\leq$ or $\geq$).  For interval domain, $|B|=1$ 
-since an interval meet irreducible is defined over a single variable.  For 
-octagons, $|B|=2$.  A decision adds a new meet irreducible $m$ to the trail.  
-Subsequently, the new state of the solver is defined over $m$ and corresponding 
-label information $s=decision$, which is described below. 
-\[decide: \quad (\mathcal{E},S) \rightarrow (\mathcal{E}(m,s),S) \]
+Given 
+%set of variables $\vars$ \pscmt{where from?} and a 
+the current abstract value $\abs(\trail)$, $\decide$ heuristically
+returns a non-singleton meet irreducible.
+%over a set of branching
+%variables $\{B\} \subseteq \mathcal{V}$, a bound $c$, and a polarity
+%($\leq$ or $\geq$).  For interval domain, $|B|=1$ since an interval
+%meet irreducible is defined over a single variable.  For octagons,
+%$|B|=2$.  A decision adds a new meet irreducible $m$ to the trail.
+%Subsequently, the new state of the solver is defined over $m$ and
+%corresponding label information $s=decision$, which is described
+%below.
+%\[decide: \quad (\mathcal{E},S) \rightarrow (\mathcal{E}(m,s),S) \]
 
@@
 
-ACDCL supports several decision heuristics namely, {\em ordered}, 
-{\em longest-range}, {\em random} and 
+We implemented several decision heuristics namely, {\em ordered}, 
+{\em longest-range}, \pscmt{are these two combined?} {\em random} and the
 {\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
@@
 before choosing meet irreducibles with numerical variables.  
-The ordering heuristic is similar to trace partitioning~\cite{toplas07}.
+The ordering heuristic is similar to trace partitioning~\cite{toplas07}. \pscmt{what is this?}
 
-The {\em longest-range} heuristics simply keeps track of the
-interval $[l,u]$ range ($r=u-l$) of a meet irreducible and chooses 
-a meet irreducible with the longest range value.  This ensures a 
-fairness policy in selecting a variable since it guarantees that 
-the intervals of meet irreducibles are uniformly restricted.  
+The {\em longest-range} heuristics simply keeps track of the bounds
+$numabsval_l,numabsval_u$ of matching template rows $\numabsval_l\leq
+\vec{c}\vec{x}\leq \numabsval_u$, picks the one with the longest range
+$\numabsval_u-\numabsval_l$, and randomly returns the meet irreducible
+$\vec{c}\vec{x}\leq
+\lfloor\frac{\numabsval_l+\numabsval_u}{2}\rfloor$ or its
+complement. This ensures a fairness policy in selecting a variable
+since it guarantees that the intervals of meet irreducibles are
+uniformly restricted.
 %
 The {\em random} decision heuristics arbitrarily picks a meet irreducible  
-for making decision. 
+for making decision. \pscmt{uniformly?}
 %
@@
 %
-Whereas, a {\em berkmin} decision heuristics is inspired 
+The {\em berkmin} decision heuristics is inspired 
 from decision heuristic used in Berkmin~\cite{eugoldberg07} SAT solver.  
-The berkmin heuristic is currently implemented for interval constraints only.  
-The heuristic keeps track of the activity of an interval meet irreducible 
-that participate in conflict clauses. 
+The Berkmin heuristics %is currently implemented for interval constraints only.  
+%The heuristic 
+keeps track of the activity of %an interval 
+meet irreducibles that participate in conflict clauses. 
+Based on the most active meet irreducible, ranges a split similar to the {\em longest-range} heuristics.
 \Omit {
@@
 appear in conflict clauses.  The set of conflict clauses is 
-organized chronologically with the top clause 
+organised chronologically with the top clause 
 as the one deduced in the last.  A branching variable is chosen among the 
@@
 %
-A bound of a meet irreducible is heuristically chosen to be an approximation of the 
-arithmetic average of the current bounds.  However, the polarity ($\leq$ or $\geq$) 
-of a meet irreducible is chosen randomly.
+%A bound of a meet irreducible is heuristically chosen to be an
+%approximation of the arithmetic average of the current bounds.
+%However, the polarity ($\leq$ or $\geq$) of a meet irreducible is
+%chosen randomly.

### 2016-10-22 deductions.tex
moved back from bitbucket to github

@@
 {
-%  $(\abstransel,\livevars) \leftarrow \mathit{worklist.pop}()$ \;
   $\abstransel{\subdomain} \leftarrow \mathit{worklist.pop}()$ \; 
-%  $v \leftarrow ded(s, \makesubdomain(\livevars), \abs(\trail))$\;
-  $v \leftarrow \abstransel{\subdomain}(\abs(\trail))$\;
-  \uIf{$v = \bot$} {
-    $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \;
-%\pscmt{how is this used in the conflict
-%    analysis? wouldn't it be more uniform to add $\bot$ to the $\trail$?}
-%    \rmcmt{UIP uses the transformer stored at $\reasons[\bot]$ to derive the
-%    cut. No, we dont want to make the trail inconsistent by adding $\bot$}
+  \rmcmt{how do we get the $\subdomain$ at this point ?}
+  $\newdeductions \leftarrow \onlynew(\abstransel{\subdomain}(\abs(\trail)))$\;
+  \uIf{$\newdeductions = \bot$} {
+    $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \rmcmt{shouldn't this be
+    just $\abstransset$} \;
     $\mathit{worklist.clear}()$ \;
-%\pscmt{why can't you clean up the worklist during conflict analysis? it seems inefficient to throw everything away...}
-%    \rmcmt{When a bottom is deduced, the direction of search
-%    changes--so the worklist is populated with new set of statement and must be
-%    flushed out} 
     \return \conflict \;
@@
   {
-    $\trail \leftarrow \trail \cdot \decomp(v)$ \; 
-%    $v \leftarrow v \meet \mathit{meet\_irrd}$ \; 
-    $\reasons[|\trail|] \leftarrow s$ \;
-    $\updateworklist_{\propheur}(v, \worklist)$ \; 
+%    $\newdeductions=\decomp(\absval)\setminus\decomp(\abs(\trail))$\;
+    $\trail \leftarrow \trail \cdot \decomp(\newdeductions)$ \; 
+    $\reasons[|\trail|] \leftarrow \abstransel{\subdomain}$ \;
+    $\updateworklist_{\propheur}(\newdeductions, \worklist)$ \; 
   }
@@
 %
-We thus define a specialised variant of the transformer to compute
+We thus define a specialised variant of the abstract transformer to compute
 deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$.
-A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$.
-\[\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\models \constraint\})\]
+A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$
+\begin{equation}\label{eq:at2}
+\abstrans{\domain}{\constraint}^\subdomain(\absval)=\absval\meet_\domain \alpha_\subdomain(\{\val\mid \val\in\gamma_\domain(\absval), \val\models \constraint\})
+\end{equation}
 For $\subdomain=\domain$, the abstract transformer behaves in the same way as the one defined in Section~\ref{sec:domains}.
+
+%We construct the subdomain from a set of variables $\vars$ such that $\subdomain_\domain(\vars)$ is the set of meet irreducibles $\in \domain$ that contain at last one variable in $\vars$ and most one variable that is not in $\vars$.
+\pscmt{split into two parts}
 %
-The restriction to a subdomain makes it less precise. Therefore, we have the property $\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq \abstrans{\domain}{\constraint}^\subdomain(\absval)$.
+%For example, $\subdomain=\domain[\{y\}]$ being the octagons domain over variables $\{x,y,z\}$ contains all octagonal meet irreducibles involving $\{y\}$, and the pairs $\{x,y\}$ and $\{y,z\}$, but not the meet irreducibles\pscmt{wrong} over $\{x\}$, $\{z\}$ and~$\{x,z\}$.
+
+The first advantage of this definition is that with the help of $\subdomain$
+we can control how we perform propagation using our abstract transformers.
 %
-\pscmt{explain $\makesubdomain(Vars)$}
-For example, for
-$\absval=(0\leq y \leq 1 \wedge 5\leq z)$ we have
-$\abstrans{\domain}{x=y+z}^{\makesubdomain(\{x\})}(\absval)=(x\geq 6)$, whereas
-$\abstrans{\domain}{x=y+z}^{\makesubdomain(\{y,z\})}=\top$.
+For example, for $\absval=(0\leq y \leq 1 \wedge 5\leq z)$, we have
+$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{x\}]}(\absval)=\absval\meet(x\geq
+6)$. For $x=y+z$, this performs a right-hand side to left-hand side propagation and
+hence emulates a forward analysis, whereas
+$\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{y,z\}]}=\absval\meet\top$
+emulates a backward analysis.
+
+We call the propagation with arbitrary subdomains,
+e.g. $\subdomain=\domain$, \emph{multi-way propagation}, which is able
+to simultaneously perform forward and backward propagation.
 %
-The advantage of this definition is that with the help of $\subdomain$
-we can now control how we propagate within our constraint system.
+Note that the restriction to a subdomain makes the transformer less
+precise. Therefore, we have the property
+$\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq
+\abstrans{\domain}{\constraint}^\subdomain(\absval)$.
+However, the choice of a subdomain makes the deductions more efficient.
+
+The second advantage of Eq.~(\ref{eq:at2}) is that we can compute the
+\emph{closure} operation for relational domains in a lazy manner. To
+this end, we construct a $\subdomain=\makesubdomain_\domain(\vars)$
+which is sufficient to perform one step of the closure operation each
+time the abstract transformer is applied.
 %
-For the example above, $\subdomain=\makesubdomain(\{x\})$ performs a
-right-hand side to left-hand side propagation and hence emulates a
-forward analysis on the input program; whereas with
-$\subdomain=\makesubdomain(\{y,z\})$) we emulate a backward analysis.
-%
-We call the propagation with $\subdomain=\domain$ \emph{multi-way
-  propagation}, which simultaneously performs forward and backward
-propagation.
+For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
+$\vars=\{y\}$. \rmcmt{A binary} octagon \rmcmt{element} relate at 
+most two variables. Thus it is sufficient to consider the subdomain
+$\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
+which will compute the one-step transitive relations of $y$ with each
+of the other variables. If this indeed affects $x$ and $z$ then their
+relations will be inferred in a later step of the propagation
+algorithm, thus gradually computing the closure. Note that this might
+increase the number of propagation steps, but it reduces the size of
+the domain used in the abstract transformer.
 
@@
 
-\pscmt{give an example for $ded$, link it to the reason trail}
-
-\pscmt{define $ded$ for all statements, connect this to worklist}
-
-\pscmt{link livevars to $L$ -- the subdomain spanned by livevars}
-
 %-------------------------------------------------------------------------------
@@
 %
-\rmcmt{Talk about lazy closure operation}
 
-\pscmt{
-Rough and sketchy:
-%
-$espan(V)=$all meet irreducibles $\in A$ that contain at most one variable that is not in $V$.
-%
-$select(S,PV)$ is the set of statements that contain variables in $PV$ where $PV$ is the set of variables in the previous deduction.
-%
-$LV(s)$ are the live variables of a statement $s$, i.e.\ the intersection of $PV$ with the variables in $s$.
-\begin{theorem}
-$$
-\gfp\lambda a. \bigsqcap_{s\in S} ded(s,a,A) = 
-\gfp\lambda a. \bigsqcap_{s\in select(S,PV)} ded(s,a,espan(LV(s)))
-$$
-\end{theorem}
-Proof sketch: see Appendix
-}
+Algorithm~\ref{Alg:ms} presents the deduction phase $\deduce$ in 
+abstract model search procedure.  The input to the algorithm is 
+the set of abstract transformers, a propagation trail and a reason 
+trail.  Additionally, the procedure $\deduce$ is parameterised with 
+a propagation heuristic.  The propagation heuristics provides two 
+functions $\initworklist$ and $\updateworklist$.  Given a set of 
+variables in the \rmcmt{constraint system}, a subdomain is essentially a 
+subset of variables \rmcmt{how we are defining subdomain?} that guides the propagations in forward, 
+backward or multi-way.  For example, the subdomain consist of variables 
+in lhs in forward propagation heuristic.  For backward, the subdomain 
+consist of variables in rhs, whereas the subdomain contains all variables 
+for multi-way propagation.  A {\em worklist} is a queue which contains 
+a set of abstract deduction transformers.  
+The initWorklist initializes the worklist 
+following a heuristics choice for propagation strategies.  A {\em forward} 
+propagation strategy initializes the worklist with abstract deduction 
+transformers that contains constants in the right-hand side. Whereas, a 
+{\em backward} iteration strategy starts with a set of assertions. On the 
+other hand, a {\em multi-way} iteration strategy initializes the worklist 
+with set of all transformers and performs both forward and backward 
+propagations in tandem.  \rmcmt{wrong to say abstract deduction transformer that 
+has lhs, rhs since since it also includes the live variables}.
 
-Algorithm~\ref{Alg:ms} presents the deduction phase in abstract model
-search procedure. \pscmt{The propagation heuristics $H$ provides two
-  functions $\initworklist$ and
-  $\updateworklist$. Describe them for the different
-  heuristics (already done for initialize, TBD for update...}
-A {\em worklist} is a queue which contains \rmcmt{abstract deduction
-transformers}.  A {\em worklist} is initialized following a heuristics 
-choice for propagation strategies. A {\em forward} propagation strategy 
-initializes the worklist with \rmcmt{abstract} transformers that 
-has constants in the right-hand side. Whereas, a {\em backward} 
-iteration strategy starts with a set of assertions. On the other hand, a 
-{\em multi-way} iteration strategy initializes the worklist with set of all 
-transformers and performs both forward and backward propagations in tandem.  
-For every meet irreducible deduction transformer \pscmt{rename} $s \in S$ 
-passed to the abstract domain, if the domain infers a new meet irreducible 
-$\mathit{meet\_irrd}$, it is added to $\trail$ \pscmt{It usually infers several ones.}.  Additionally, the reason 
-trail $\reasons$ \pscmt{define} is updated with the transformer $s$ as the reason for inferring
-$\mathit{meet\_irrd}$. If a conflict is deduced, that is, $\mathit{meet\_irrd} = \bot$ 
-is inferred by the domain, the algorithm terminates with \textsf{conflict}.  However,
-when a fixed-point is reached, the abstract valuation $v$ is checked for set of
-models.  This is determined by checking whether the set of assignments in $v$ is 
+Given a constraint system and a domain, an updateWorklist function adds 
+new abstract deduction transformers to the worklist for those variables in 
+the subdomain that appear in newDeductions.  It also updates the subdomain of 
+existing abstract deduction transformers that are in the worklist.
+
+The function
+$\onlynew(\absval)=\bigsqcap(\decomp(\absval)\setminus\decomp(\abs(\trail)))$
+filters out all known meet irreducibles in the abstract value.
+
+For every abstract deduction transformer in the worklist, 
+the set of new meet irreducibles inferred by the domain is added 
+to the $\trail$.  Additionally, the reason trail $\reasons$ is 
+updated with the corresponding abstract transformer as the reason for inferring 
+those new meet irreducibles.  If a conflict is deduced, 
+%that is, $\mathit{meet\_irrd} = \bot$ is inferred by the domain, 
+the algorithm terminates with \textsf{conflict}.  However,
+when a fixed-point is reached, the abstract value $v$ computed from 
+the trail is checked for set of models.  This is determined by 
+checking whether the set of assignments in $v$ is 
 $\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
-abstract valuation $v$ is abstractly satisfying, the abstract model 
-search terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
+abstract valuation $v$ is abstractly satisfying, the abstract model search 
+terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
 model search procedure makes a new decision.

### 2016-10-22 domains.tex
moved back from bitbucket to github

@@
 %
+We denote $\vars$ the set of variables occurring in the $\constraints$.
 %
@@
 %
-Based on the above program representation, we define a safety equation \pscmt{why ``equation''?} as, 
+Based on the above program representation, we define a safety equation as, 
 $\formula:= \bigwedge_{\constraint\in\constraints} \constraint$. The formula $\formula$ is 
@@
 %In our prototype implementation and for illustrative purposes 
-In this paper we use a product domain $\domain=\booldomain^\booldim\times\reldomain$ where
-$\booldomain$ is the boolean domain
+In this paper we use a product domain $\domain[\vars]=\booldomain^{|\boolvars|}\times\reldomain[\numvars]$ where
+$\booldomain$ is the Boolean domain
 $\langle\{\true,\false,\bot,\top\},\Rightarrow,\wedge,\vee\rangle$, 
-$\booldim$ is the number of boolean variables in the program, 
+$\boolvars$ are the Boolean variables in the program, 
 and $\reldomain$ is a \emph{relational} domain over the 
-numerical (bitvector) variables of respective
+numerical (bitvector) variables $\numvars$ of their respective
 data types (currently, signed and unsigned integers are supported).
@@
 For notational convenience we will denote elements of
-$\booldomain^\booldim$ by their concretisations to propositional formulae,
-For example, the program in Figure~\ref{ssa} has four boolean variables
+$\booldomain^{|\boolvars|}$ by their concretisations to propositional formulae,
+For example, the program in Figure~\ref{ssa} has four Boolean variables
 written as the vector $(g_0,g_1,c,g_2)$; then we will denote an abstract value $(\top, \false, \true, \top) \in \booldomain^4$ as $\neg g_1 \wedge c$.
@@
 An abstract value of the template polyhedra domain~\cite{sriram}
-represents a set $\numconcval$ of values of numerical variables $\vec{\numvar}$. For
+represents a set $\numconcval$ of values of the vector $\vec{\numvar}$ of numerical variables $\numvars$. For
 example, in the program in Figure~\ref{ssa}, we have four numerical
@@
 % \[\gamma[p,q] = \{c \in \mathbb{Z} | p \leq c \leq q\} \qquad \gamma(\bot) = \emptyset \]
-
+%
 \begin{wrapfigure}{r}{4.5cm}
@@
 \scalebox{.65}{\import{figures/}{octagons.pspdftex}}
-\caption{Example of an octagon \pscmt{BUG: should be $x\geq -2$, $y\geq -1$; please rename x to $x_1$ and $y$ to $z$}}\label{octagon}
+\caption{Example of an octagon}\label{octagon}
 \vspace*{-3ex}
 \end{wrapfigure} 
-
-For a program with $N$ variables, the matrix $\mat{C}$ for intervals
+%
+In the interval domain $\intervals[\numvars]$,
+for a program with $N=|\numvars|$ variables, the matrix $\mat{C}$ 
 has $2N$ rows; and hence it generates at most $2N$ inequalities, one
@@
 %
-For octagons, we have at most $2N^2$ inequalities, one for the upper
-and lower bounds of each variable and sums and differences for each
-pair of variables. Figure~\ref{octagon} shows an 
-example octagon and its associated inequalities.
+For octagons $\octagons[\numvars]$, we have at most $2N^2$
+inequalities, one for the upper and lower bounds of each variable and
+sums and differences for each pair of variables. Figure~\ref{octagon}
+shows an example octagon and its associated inequalities.
 
@@
 %
-We will see that our model search algorithm performs closure computations lazily,
-postponing them to the point where they are actually necessary.\rmcmt{are we
-showing this ?}
+%We will see that our model search algorithm performs closure computations lazily,
+%postponing them to the point where they are actually necessary.\rmcmt{are we
+%showing this ?}
 
@@
 The best abstract transformer is 
-\[\abstrans{\domain}{\constraint}(\absval)=\absval\meet \alpha(\{\val\mid \val\models \constraint\})\] where we write 
+\begin{equation}
+\abstrans{\domain}{\constraint}(\absval)=\absval\meet\alpha(\{\val\mid \val\in\gamma(\absval), \val\models \constraint\})
+\end{equation}
+ where we write 
 $\val\models \constraint$ if the concrete value $\val$ satisfies the constraint $\constraint$.
@@
 %A detailed description of abstract transformer is presented in Section~\ref{sec:abst}.
+We denote the set of abstract transformers for a safety equation
+$\formula$ using abstract domain $\domain$ as
+$\abstransset=\{\abstrans{\domain}{\constraint}\mid
+\constraint\in\constraints\}$.
 

### 2016-10-22 example.tex
moved back from bitbucket to github

@@
 %
-\begin{table}
+\begin{table}[!b]
 \begin{center}
@@
 \hline
-Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
+Solver & Decisions & propagations & conflict & conflict literals \pscmt{XXXX} & restarts \\ \hline
 SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline

### 2016-10-22 experiments.tex
moved back from bitbucket to github

--- 2016-10-21 experiments.tex
@@
 ACDCL is instantiated with a product domain of booleans and template 
-polyhedra.  Additionally, the analysis using ACDCL is configured with 
+polyhedra.  We use an interval and octagon templates for our
+experiments.  Additionally, the analysis using ACDCL is configured with 
 a heuristic choice of decision (ordered, random, berkmin), 
@@
 \paragraph {\textbf{Propagation Strategy}}      
-Figure~\ref{prop-dec}(a) presents a comparison between {\em forward} propagation 
-strategy and {\em multi-way} propagation strategy in ACDCL.  The choice of the
-propagation strategy influences the total number of decisions and clause 
-learning iterations.  Hence, the propagation strategy has a significant 
-influence on the run time which can be visualised in Figure~\ref{prop-dec}(a).  
-Compared to forward propagation, the multi-way strategy may take more iterations 
-to reach the fixed-point, but it significantly reduces the total number of 
-decisions and conflicts.  This is attributed to the higher 
-precision of the meet irreducibles inferred by the multi-way strategy which 
-subsequently aids the decision heuristics to make better decisions.
-
-\Omit {
-We observe that the deductions made during chaotic analysis helps the decision 
-heuristics to make better decisions compared to forward analysis.  This is 
-attributed to the precision of deductions inferred by the chaotic analysis.         
-}
+Figure~\ref{prop-dec}(a) presents a comparison between {\em forward}
+propagation strategy and {\em multi-way} propagation strategy in
+ACDCL.  The choice of the propagation strategy influences the total
+number of decisions and clause learning iterations.  Hence, the
+propagation strategy has a significant influence on the run time which
+can be visualised in Figure~\ref{prop-dec}(a).  Compared to forward
+propagation, the multi-way strategy may take more iterations to reach
+the fixed-point, but it significantly reduces the total number of
+decisions and conflicts, in particular for the unsafe benchmarks.
+This is attributed to the higher precision of the meet irreducibles
+inferred by the multi-way strategy which subsequently aids the
+decision heuristics to make better decisions.
+\pscmt{Chaotic $\rightarrow$ multi-way in Fig.6 }
 

### 2016-10-22 learning.tex
experiment graph
learn update
moved back from bitbucket to github

@@
 assignments domain or interval domain, so finding a generalized reason for 
-conflict through UIP computation~\cite{uip} is non-trivial.
+conflict through UIP computation~\cite{uip} is non-trivial.  We formally 
+describe the adaptations of conflict analysis procedure for relational domains. %
+\begin{figure}[t]
+%\caption{An example of Octagon}\label{octagon}
+\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
+\caption{\label{conflict} Abstract Conflict Graph \pscmt{constants to rhs, please!}}
+\end{figure}
 
 \paragraph {\textbf{Abstract Conflict Graph}}
-The first-UIP algorithm in SAT solvers works on a data structure called 
-{\em implication graph} and computes a cut that suffices to produce a 
-conflict.  The trail in ACDCL implicitly encodes a graph structure called 
-{\em abstract conflict graph}, which records the dependencies between 
-deductions made during the abstract model search phase.  Nodes in the graph 
-represent elements of the trail $\mathcal{T}$.  A node can be either a
-{\em decision} node or a {\em deduction} node. Edges can be extracted from
-the transformers in the reason trail.  Incoming arrows to node $n$ 
-indicate that the predecessors of $n$ are sufficient to deduce $n$.  
-A decision node has no incoming edges.  
-For example, consider a formula 
-$\varphi:= (y=x \; \wedge \; y=y+x \; \wedge \; y \leq 0)$ with current 
-abstract valuation $v:= (x-5 \geq 0 \wedge -x+5 \geq 0)$.  
-Figure~\ref{conflict} shows a snapshot of an abstract conflict graph that 
-stores the deductions obtained from abstract model search using the octagon domain.  
-The last deduction conflicts with the constraint $(y \leq 0)$.
+The first-UIP algorithm in SAT solvers works on a data structure
+called {\em implication graph} and computes a cut that suffices to
+produce a conflict.  The propagation and reason trails in ACDCL
+implicitly encode a graph structure called {\em abstract conflict
+  graph}, which records the dependencies between deductions made
+during the abstract model search phase.  Nodes in the graph represent
+elements of the trail $\trail$.  A node can be either a {\em decision}
+node or a {\em deduction} node. Edges can be extracted from the
+transformers in the reason trail $\reasons$.  Incoming arrows to node
+$n$ indicate that the predecessors of $n$ are sufficient to deduce
+$n$.  A decision node has no incoming edges.  For example, consider a
+formula $\formula:= (y{=}x \wedge y{=}y{+}x \wedge y {\leq} 0)$ with
+current abstract valuation $\absval:= (x{-}5 {\geq} 0 \wedge {-}x{+}5 {\geq}
+0)$.  Figure~\ref{conflict} shows a snapshot of an abstract conflict
+graph that stores the deductions obtained from abstract model search
+using the octagon domain.  The last deduction conflicts with the
+meet irreducible $(y \leq 0)$.
 
@@
 used in learning.  An abstract first-UIP algorithm~\cite{DBLP:journals/fmsd/BrainDGHK14} 
-traverses the trail $\mathcal{T}$ starting from the conflict node and 
+traverses the trail $\trail$ starting from the conflict node and 
 computes a cut that suffices to produce a conflict. 
 \Omit {
-contradict with the transformer stored in $\mathcal{R}[\bot]$.  
+contradict with the transformer stored in $\reasons[\bot]$.  
 }
@@
 %    
-\begin{figure}[t]
-%\caption{An example of Octagon}\label{octagon}
-\scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
-\caption{\label{conflict} Abstract Conflict Graph}
-\end{figure} 
-%    
 \paragraph{\textbf{Learning in relational domains}}
-Learning in a propositional solver yields an asserting clause~\cite{cdcl} 
-that expresses the negation of the conflict reasons.  However, we model 
-learning in relational domains as learning a transformer which infers 
-new meet irreducibles using the abstract unit rule called 
-{\em abstract unit transformer (AUnit)}.  We add $\mathit{AUnit}$ to the set of 
-transformers \rmcmt{give symbol} used during model search. $AUnit$ is 
-a generalization of propositional unit rule for numerical domains.  For 
-an abstract lattice $A$ with complementable meet irreducibles and a set 
-of meet irreducibles $C \subseteq A$ of $\varphi$ \pscmt{???} such that $\bigsqcap C$ does not 
-satisfy $\varphi$, $\mathit{AUnit}_C: A \rightarrow A$ is formally defined as follows. 
-\[ AUnit(a) =
+Learning in a propositional solver yields an asserting
+clause~\cite{cdcl} that expresses the negation of the conflict
+reasons.  However, we model learning in relational domains as learning
+a transformer which infers new meet irreducibles using the abstract
+unit rule called {\em abstract unit transformer}.  We add $\aunit$ to
+the set of abstract transformers $\abstransset$ used during model
+search. $\aunit$ is a generalization of propositional unit rule for
+numerical domains.  For an abstract lattice $\domain$ with
+complementable meet irreducibles and a set of meet irreducibles $\conflictset
+\subseteq \domain$ such that $\bigsqcap
+\conflictset$ does not satisfy $\formula$, $\aunit_\conflictset: \domain \rightarrow
+\domain$ is formally defined as follows.
+\[ \aunit(\absval) =
  \begin{cases}
-  \bot       & \quad \text{if } a \sqsubseteq \meet{C} \qquad \qquad \qquad
+  \bot       & \quad \text{if } \absval \sqsubseteq \bigsqcap \conflictset \qquad \qquad \qquad
   \qquad \qquad \qquad \quad (1)\\
-  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C
-  \setminus \{t\}. a  \sqsubseteq t' \qquad \quad \; \; (2) \\
-  \bar{t}    & \quad \text{if } t \in C \; \text{and} \; \forall t' \in C \setminus \{t\}. a
+  \bar{t}    & \quad \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset
+  \setminus \{t\}. \absval  \sqsubseteq t' \qquad \quad \; \; (2) \\
+  \bar{t}    & \quad \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset \setminus \{t\}. \absval
   \not\sqsubseteq t', \; \text{then} \\ 
-             & \quad \forall a' \subseteq decomp(a) \;  \text{and} \; \forall a'' \in a \setminus \{a'\}, \\ 
-             & \quad \meet{a''} \sqsubseteq r' \; \text{and} \; a'
-             \not\sqsubseteq t   \qquad \qquad \qquad \qquad \; \; (3) \\
+             & \quad \forall \absval' \sqsubseteq \absval: \absval'
+             \not\sqsubseteq t \text{ and}\\ 
+             & \quad \absval''=\exclude(\absval,\absval') \text{ and }\absval''\sqsubseteq r' \pscmt{\text{what is r'???}} \qquad  \; \; (3) \\
   \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
@@
 \]
-$\mathit{AUnit}$ returns $\bot$ if $C$ is conflicting following rule (1).  Rule (2) and
-(3) of $\mathit{AUnit}$ infers a valid meet irreducible, which implies that $C$ is unit. 
-However, rule (3) is domain-specific and only applies to interval and octagon
-domain.  For example, let $C = \langle x+z \geq 10,  w \geq 0 \rangle$ be a 
-conflicting element that does not satisfy $\varphi$ and let $a= \langle x+y \geq 2, y+z \geq 5,
-x+z \leq 15, w \geq 1, w \leq 1 \rangle$.  Applying rule (3) of $\mathit{AUnit}_C$, we get a 
-decomposition of $a$ into $a'$ and $a''$ where 
-$a'= \langle x+y \geq 2, y+z \geq 5, x+z \leq 15 \rangle \not\sqsubseteq (x+z \geq 10)$, and 
-$a''= \langle w \geq 1, w \leq 1 \rangle \sqsubseteq (w \geq 0)$.  Then, $\mathit{AUnit}_C(a) = (x+z < 10)$, 
-which constrains the bound of the octagonal constraint.  
-Let us consider another example, where $C = \langle x \geq 2, x \leq 5, y \leq 7 \rangle$
-and let $a = \langle x \geq 3, x \leq 4, y \geq 5, y \leq 6 \rangle$, then $\mathit{AUnit}_C(a) = \bot$ using
-the rule (1), since $a \sqsubseteq C$.  Now, let $a' = \langle x \geq 3, x \leq 4 \rangle$, 
-then $\mathit{AUnit}_C(a') = (y > 7)$ using rule (2), since $a' \sqsubseteq (x \geq 2)$
-and $a' \sqsubseteq (x \leq 5)$.  However, if none of the above rule holds true, $\mathit{AUnit}$ 
-returns $\top$.
+with $\exclude(\absval,\absval')=\bigsqcap(\decomp(\absval) \setminus
+\decomp(\absval'\}))$.  $\aunit$ returns $\bot$ if $\conflictset$ is
+conflicting following rule (1).  Rule (2) and (3) of $\aunit$ infer a
+valid meet irreducible, which implies that $\conflictset$ is unit.
+
+However, rule (3) is domain-specific \pscmt{I don't see anyting
+  specific here...} and only applies to interval and
+octagon\pscmt{relational?}  domain.  For example, let $\conflictset =
+\{ x+z \geq 10, w \geq 0 \}$ be a conflicting element \pscmt{better
+  name???} that does not satisfy $\formula$ and let $\absval= (x+y
+\geq 2\wedge y+z \geq 5\wedge x+z \leq 15\wedge w \geq 1\wedge w \leq
+1)$.  Applying rule (3) of $\aunit_\conflictset$, we get a
+decomposition of $\absval$ into $\absval'$ and $\absval''$ where
+$\wedge'= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15)
+\not\sqsubseteq (x+z \geq 10)$, and $\absval''= (w \geq 1\wedge w \leq
+1) \sqsubseteq (w \geq 0)$.  Then, $\aunit_C(a) = (x+z < 10)$, which
+constrains the bound of the octagonal constraint.
+
+Let us consider another example, where $\conflictset = \{x \geq 2, x
+\leq 5, y \leq 7 \}$ and let $\absval = (x \geq 3\wedge x \leq 4\wedge
+y \geq 5\wedge y \leq 6)$, then $\aunit_\conflictset(\absval) = \bot$
+using the rule (1), since $\absval \sqsubseteq \bigsqcap\conflictset$.  Now,
+let $\absval' = (x \geq 3\wedge x \leq 4)$, then
+$\aunit_\conflictset(\absval') = (y \geq 8)$ using rule (2), since
+$\absval' \sqsubseteq (x \geq 2)$ and $\absval' \sqsubseteq (x \leq
+5)$.  However, if none of the above rule holds true, $\aunit$ returns~$\top$.
 
 \paragraph {\textbf{Backjumping}}
-A backjumping procedure undoes all the assignments in the trail up to 
-a decision level that restores the solver to a non-conflicting state.  
-The backjumping level is defined by the literal of the conflict clause 
-that is closest to the root (decision level 0) where the conflict 
+After adding $\aunit$ to $\abstransset$, we remove all the meet irreducibles from the trail back to a decision level that restores the solver to a
+non-conflicting state.  The backjumping level is defined by the
+meet irreducible of the conflict clause \pscmt{is this $\conflictset$???}
+that is closest to the root (decision level~0) where the conflict
 clause is still unit.  If a conflict clause is globally unit, then the
-backjumping level is the root of the search tree.
+backjumping level is the root of the search tree and
+$\analyzeconflict$ returns $\false$, otherwise $\true$.
 
-The abstract clause learning and backjumping procedures in the abstract 
-conflict graph is stated in terms of the state of ACDCL solver as follows.
-\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
-\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
-\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
-\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
-(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict} \]
+%The abstract clause learning and backjumping procedures in the abstract 
+%conflict graph is stated in terms of the state of ACDCL solver as follows.
+%THIS is already explained above
+%\[AbsLearn: \quad  (\mathcal{E},S) \rightarrow (\mathcal{E},S \wedge
+%\mathcal{L}) \quad \text{if} \; \mathcal{L} \notin S \; \textrm{and}
+%\; (S \wedge \mathcal{L}) \; \text{is not UNSAT} \]
+%\[AbsBackjump: \quad (\mathcal{E}_1(m,s)\mathcal{E}_2,S) \rightarrow
+%(\mathcal{E}_1,S) \quad \text{if} \; (\mathcal{E}_1,S) \; \text{is not in conflict} \]
 

### 2016-10-22 paper.tex
microtype
moved back from bitbucket to github

@@
 %
-\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalization
+\emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
 of the well-known Conflict-Driven Clause Learning (CDCL) algorithm that is
 used in SAT solvers.  ACDCL alternates between a model search, which
-performs overapproximate deduction with constraint propagation, and a
+performs over-approximate deduction with constraint propagation, and a
 conflict analysis, which performs under-approximate abduction with heuristic
@@
 procedure.  Whereas the propositional reasoning used in satisfiability 
-solvers can be rather easily lifted from the boolean lattice to other 
+solvers can be rather easily lifted from the Boolean lattice to other 
 non-relational domains, lifting them to relational domains is 
@@
 relational abstract domains.  We have implemented these algorithms inside
-the 2LS verification tool and instantiated our analyzer over a template 
+the 2LS verification tool and instantiated our analyser over a template 
 polyhedra domain to check safety properties of C programs.  We evaluate 
-the performance of our analyzer in comparison with program verifiers that 
+the performance of our analyser in comparison with program verifiers that 
 uses propositional-level solvers and a static analysis tool.
@@
 %===============================================================================
-\input{background}
+\input{conclusion}
 %===============================================================================
@@
 %===============================================================================
-\input{conclusion}
+\input{background}
 %===============================================================================

### 2016-10-22 contributions.tex
more spell checking

@@
 We have implemented our technique in the 2LS verification tool.
-\Omit{
-To the best of our knowledge, this is the first instantiation of CDCL 
-algorithm with relational abstract domains. 
-}
 
-\item We present an abstract model search algorithm that strictly 
-generalizes the Boolean Constraint Propagation (BCP) procedure to
-compute a fixed point over a relational domain.
+\item We present an abstract model search algorithm 
+%that strictly 
+%generalises the Boolean Constraint Propagation (BCP) procedure 
+to compute a fixed point over a relational domain.
 
-\item We propose several domain-specific decision heuristics that 
+\item We propose domain-specific decision heuristics that 
 exploit the expressivity of the underlying abstract domain to make 
@@
 
-\item We present a non-trivial lifting of the conflict analysis procedure 
+\item We present a lifting of the conflict analysis procedure 
 to relational abstract domains in order to learn a generalised reason 

### 2016-10-22 example.tex
more spell checking

--- 2016-10-20 example.tex
@@
 
-We analyze the program using three different analyzers -- {\em bounded 
+We analyse the program using three different analysers -- {\em bounded 
 model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
 \Omit {
-The SAT solver makes 36 decisions, 1365 boolean constraint propagations 
+The SAT solver makes 36 decisions, 1365 Boolean constraint propagations 
 and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
@@
 imprecision is due to control-flow join at node $n4$.  State-of-the-art 
-abstract interpretation tool, Astr{\'e}e, requires external hints, provided 
+abstract interpretation tool, Astr{\'e}e \cite{se2011}, requires external hints, provided 
 by manually annotating the code with partition directives at $n1$, 
-to prove safety.  The partition directive helps the tool to analyze 
+to prove safety.  The partition directive helps the tool to analyse 
 each program execution paths separately.  To do so, Astr{\'e}e 
@@
 %
-The right-side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
+The right-hand side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
 octagon domain (top) and an interval domain (bottom).  The analysis 
 associates an abstract element for each control location and variable, 
-as shown in the abstract conflict graph in Firgure~\ref{fig:example}.  
+as shown in the abstract conflict graph in Figure~\ref{fig:example}.  
 Octagonal constraints deduced by ACDCL solver are sufficient to prove 
@@
 Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
-ACDCL implements several decision heuritics.  In this case, the 
+ACDCL implements several decision heuristics.  In this case, the 
 branching variable $c$ is chosen by an {\em ordered} decision 
@@
 during fixed point iteration are represented by abstract conflict 
-graph in right-side of Figure~\ref{fig:example}.  Nodes in the abstract 
+graph in right-hand side of Figure~\ref{fig:example}.  Nodes in the abstract 
 conflict graph denotes the corresponding updates to the program 
@@
 
-A clause learning SAT solver would learn the reason for conflict at this 
-point and then backtrack to a level such that the learnt clause is UNIT.  
-Similar to conflict analysis phase in SAT solvers, ACDCL learns that 
-(\(n0: c:=[0,0]\)), that is all error traces must satisfy $(c \neq 1)$ 
-at node $n0$.  The analysis discards all interval constraints that
-lead to the conflict and backtrack to decision level 0.  ACDCL then 
-performs interval analysis with the learnt constraint $(c \neq 1)$ which also 
-leads to a conflict, as shown in right side of Figure~\ref{ssa}.  The 
-analysis cannot backtrack further and therefore it terminates proving 
-that the program is safe.  A DPLL-style solver would perform proof by 
-cases.  However, decision and clause learning are used to avoid case 
-based reasoning and prevents enumeration behavior.
+A clause learning SAT solver would learn the reason for conflict at
+this point and then backtrack to a level such that the learnt clause
+is \emph{unit}.  Similar to conflict analysis phase in SAT solvers,
+ACDCL learns that (\(n0: c:=[0,0]\)), that is all error traces must
+satisfy $(c \neq 1)$ at node $n0$.  The analysis discards all interval
+constraints that lead to the conflict and backtrack to decision level
+0.  ACDCL then performs interval analysis with the learnt constraint
+$(c \neq 1)$ which also leads to a conflict, as shown in right-hand side of
+Figure~\ref{ssa}.  The analysis cannot backtrack further and therefore
+it terminates proving that the program is safe.  A DPLL-style solver
+would perform proof by cases.  However, decision and clause learning
+are used to avoid case-based reasoning and prevents enumeration
+behaviour.
 
@@
 in the number of decisions, propagations, learnt clauses and 
-restrarts.  Compared to abstract interpretation, ACDCL does 
+restarts.  Compared to abstract interpretation, ACDCL does 
 not require external hints to proof the program, thus 

### 2016-10-22 introduction.tex
moved back from bitbucket to github
more spell checking

--- 2016-10-15 introduction.tex
@@
 %
+%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
 Static program analysis based on abstract 
-interpretation~\cite{DBLP:conf/emsoft/Cousot07} has been widely used to
+interpretation~\cite{CC77} has been widely used to
 verify certain classes of properties for safety-critical systems.  A 
@@
 precisely model conjunction, and the join operation over-approximate 
-disjunction. Thus, the precision loss in static analyzers is often 
+disjunction. Thus, the precision loss in static analysers is often 
 due to imprecise join operations. The loss in precision is overcome 
@@
 which is often very expensive since the analysis exhibit case enumeration 
-behavior.
+behaviour. 
+%%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
+On the other hand, the combination of precise conjunction in a partial
+assignments domain with learning allows Conflict Driven Clause
+Learning (CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to
+reason about disjunction without enumerating cases.  Silva
+et. al.~\cite{tacas12, sas12, dhk2013-popl} presented a formal recipe
+for applying abstract interpretation to solve the satisfiability
+problem.  \cite{dhk2013-popl} provide a generalisation of
+CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning}
+(ACDCL), in terms of deduction and abduction transformers,
+fixed-point, interpolation and extrapolation operators on lattices.
+The partial assignments in the CDCL algorithm are an abstract domain,
+unit rule is the best abstract transformer, Boolean Constraint
+Propagation (BCP) computes a greatest fixed-point over the abstract
+domain, decision is dual widening and clause learning can be viewed as
+synthesising an abstract transformer for negation.  
+%Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
+This enables the
+analyser to refine the analysis and prevent enumeration behaviour.
 
-On the other hand, the combination of precise conjunction in partial assignments 
-domain with learning allow Conflict Driven Clause Learning (CDCL)~\cite{cdcl} 
-solvers to reason about disjunction without enumerating cases.  
-Silva et. al.~\cite{tacas12, sas12, dhk2013-popl} presented 
-a formal recipe for applying abstract interpretation~\cite{se2011} to solve the 
-satisfiability problem.  ~\cite{sas12} provide a mathematical generalization 
-of CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning} (ACDCL), 
-in terms of deduction and abduction transformers, fixed-point, interpolation 
-and extrapolation operators on lattices.  The partial assignments in the 
-CDCL algorithm are an abstract domain, unit rule is the best abstract 
-transformer,  Boolean Constraint Propagation (BCP) computes a greatest fixed-point 
-over the abstract domain, decision is dual widening and clause learning can be 
-viewed as synthesizing an abstract transformer for negation.  Thus, ACDCL 
-lifts learning techniques in SAT solver to operate on non-distributive 
-lattice -- thus equipping the analysis with negation. This enables the 
-analyzer to refine the analysis and prevent enumeration behavior.
-
-\rmcmt{relational domains}
-
-The instantiation of ACDCL has so far been limited to interval domains only.  
-In this paper, we identify specific properties of the CDCL algorithm that 
-are necessary to lift the propositional skeleton of CDCL to relational 
-abstract domains such as template polyhedra domains~\cite{sriram} 
-(octagons, zones and equalities).  To this end, we present a characterization 
-of abstract model search procedure and conflict analysis procedure that lift 
-the abstract deduction transformer and the first-UIP algorithm to 
+ACDCL has so far been instantiated as a floating-point decision
+procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser
+that only uses interval constraints~\cite{tacas12}.  Hence, the
+analysis is limited to bounds checking in simple numerical programs.
+However, verification of certain class of programs require more
+expressive domains like octagons or equalities or polyhedra.  These
+domains capture interesting properties of programs which intervals can
+not express.  In this paper, we identify specific properties of the
+CDCL algorithm that are necessary to lift the propositional skeleton
+of CDCL to relational abstract domains such as template polyhedra
+domains~\cite{sriram} (octagons, zones and equalities).
+To this end, we present a characterisation of abstract model search
+procedure and conflict analysis procedure that lift the abstract
+deduction transformer and the conflict analysis procedure to
 relational lattice structures.

### 2016-10-22 algorithm.tex
comments in sec4

@@
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-The algorithm takes a program \pscmt{make this a set of transformers
-  (parametrised by the propagation heuristics)}
-$\abstransset$ and propagation, decision, and conflict analysis
-heuristics as inputs.  
-\rmcmt{Approximation of the concrete transformers in 
-$\abstransset$ are typically available in abstract domain in the 
-form of strongest-post condition or weakest pre-condition. } 
-As internal state, it maintains a propagation
-trail $\trail$ and a reason trail \pscmt{map?} $\reasons$.
+The input to the algorithm is a program in the form of the set of its abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$.
+The algorithm is parametrised by propagation, decision, and conflict analysis
+heuristics.  
+%\rmcmt{Approximation of the concrete transformers in 
+%$\abstransset$ are typically available in abstract domain in the 
+%form of strongest-post condition or weakest pre-condition. } 
+It maintains a propagation
+trail $\trail$ and a reason trail $\reasons$.
 The propagation trail is an initially empty sequence of meet
@@
 and decisions). The reason trail maps the elements of the
-propagation trail to transformers $s\in\abstransset$ that were used to
+propagation trail to transformers $\abstrans\in\abstransset$ that were used to
 derive them. The \emph{abstract value} $\abs(\trail)$
@@
 %
-\pscmt{what happens if we cannot make any decision any more?
-  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
-  be $\gamma$-complete} 
+%\pscmt{what happens if we cannot make any decision any more?
+%  -- then we should either have deduced $\bot$ or $\abs(\trail)$ must
+%  be $\gamma$-complete} 
 %
@@
 phase to learn the reason for the conflict.   There can be multiple
-incomparable reasons for conflict \pscmt{do you have to explain this here?} -- based on the choice of Unique
-Implication Point (UIP)~\cite{cdcl}.  ACDCL heuristically chooses one
-reason.  A learnt clause must include asserting cuts which guarantees
-derivation of new meet irreducibles after backtracking. The clause
-learning and backtracking continues as long as the result of deduction
-is conflicting ($\bot$), that is, the abstract value does not
-abstractly satisfy the formula \pscmt{define}.  If no further
-backtrack is possible, then the algorithm terminates and $\formula$ is
-\textsf{safe}. Else, the algorithm makes a new decision and the above
-process is repeated until $\deduce$ returns \textsf{sat} or the
-algorithm backtracks to decision level 0 after a conflict in which
-case it returns \textsf{safe}.
+incomparable reasons for conflict.
+% \pscmt{do you have to explain this here?} -- based on the choice of Unique
+%Implication Point (UIP)~\cite{cdcl}.  
+ACDCL heuristically chooses one reason and learns it by adding at as
+an abstract transformer to $\abstransset$. Then it backtracks by
+removing the tail of the $\trail$ up to a point where it becomes
+consistent, and continues making deductions from there.
+If $\analyzeconflict$ returns $\false$ then $\formula$ is unsatisfiable
+and we return  \textsf{safe}.
+
+% A learnt clause must include asserting cuts which guarantees
+% derivation of new meet irreducibles after backtracking. The clause
+% learning and backtracking continues as long as the result of deduction
+% is conflicting ($\bot$), that is, the abstract value does not
+% abstractly satisfy the formula \pscmt{define}.  If no further
+% backtrack is possible, then the algorithm terminates and $\formula$ is
+% \textsf{safe}. Else, the algorithm makes a new decision and the above
+% process is repeated until $\deduce$ returns \textsf{sat} or the
+% algorithm backtracks to decision level 0 after a conflict in which
+% case it returns \textsf{safe}.
 

### 2016-10-22 learning.tex
learning text

@@
 \scalebox{.65}{\import{figures/}{conflict_graph.pspdftex}}
-\caption{\label{conflict} Abstract Conflict Graph \pscmt{constants to rhs, please!}}
+\caption{\label{conflict} Abstract Conflict Graph}
 \end{figure}
@@
              \not\sqsubseteq t \text{ and}\\ 
-             & \quad \absval''=\exclude(\absval,\absval') \text{ and }\absval''\sqsubseteq r' \pscmt{\text{what is r'???}} \qquad  \; \; (3) \\
+             & \quad \absval''=\exclude(\absval,\absval') \text{ and
+             }\absval''\sqsubseteq t' \qquad  \; \; (3) \\
   \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
@@
 
-However, rule (3) is domain-specific \pscmt{I don't see anyting
-  specific here...} and only applies to interval and
-octagon\pscmt{relational?}  domain.  For example, let $\conflictset =
-\{ x+z \geq 10, w \geq 0 \}$ be a conflicting element \pscmt{better
-  name???} that does not satisfy $\formula$ and let $\absval= (x+y
+$\aunit$ strictly generalizes the unit rule in SAT solvers.  
+However, rule (3) only applies to interval and octagons domains.  
+For example, let $\conflictset = \{ x+z \geq 10, w \geq 0 \}$ be a conflicting element 
+that does not satisfy $\formula$ and let $\absval= (x+y
 \geq 2\wedge y+z \geq 5\wedge x+z \leq 15\wedge w \geq 1\wedge w \leq
@@
 non-conflicting state.  The backjumping level is defined by the
-meet irreducible of the conflict clause \pscmt{is this $\conflictset$???}
-that is closest to the root (decision level~0) where the conflict
+meet irreducible of the conflict clause that is closest 
+to the root (decision level~0) where the conflict
 clause is still unit.  If a conflict clause is globally unit, then the

### 2016-10-22 deductions.tex
part of deductions

@@
 We thus define a specialised variant of the abstract transformer to compute
-deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$.
+deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$,
+which we call \emph{abstract deduction transformer}.
+%
 A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$
@@
 \end{equation}
-For $\subdomain=\domain$, the abstract transformer behaves in the same way as the one defined in Section~\ref{sec:domains}.
+For $\subdomain=\domain$, the abstract deduction transformer behaves in the same way as the abstract transformer defined in Eq.~\ref{eq:abstrans} in Section~\ref{sec:domains}.
 
@@
 For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
-$\vars=\{y\}$. \rmcmt{A binary} octagon \rmcmt{element} relate at 
+$\vars=\{y\}$. An octagonal inequality relates at 
 most two variables. Thus it is sufficient to consider the subdomain
 $\makesubdomain_\domain(\{y\})=\octagons[\{y\}]\cup\octagons[\{x,y\}]\cup\octagons[\{y,z\}]$,
-which will compute the one-step transitive relations of $y$ with each
-of the other variables. If this indeed affects $x$ and $z$ then their
+which will compute the one-step transitive relations of~$y$ with each
+of the other variables. 
+%
+\Omit{
+%TODO
+Hence, we do not compute the full closure in the full 
+domain $\octagons[\{x,y,z\}]$, but we compute only a single step of the closure when the abstract deduction transformer is applied.
+%
+Only if $\abstrans^\subdomain$ deduces new information about $x$ or $z$
+then the next step of the closure will be 
+}
+If this indeed affects $x$ and~$z$ then their
 relations will be inferred in a later step of the propagation
-algorithm, thus gradually computing the closure. Note that this might
-increase the number of propagation steps, but it reduces the size of
-the domain used in the abstract transformer.
+algorithm, thus gradually computing the closure.
+
+%Note that this might
+%increase the number of propagation steps, but it reduces the size of
+%the subdomain used in the abstract deduction transformer.
 
@@
 trail.  Additionally, the procedure $\deduce$ is parameterised with 
-a propagation heuristic.  The propagation heuristics provides two 
-functions $\initworklist$ and $\updateworklist$.  Given a set of 
-variables in the \rmcmt{constraint system}, a subdomain is essentially a 
-subset of variables \rmcmt{how we are defining subdomain?} that guides the propagations in forward, 
-backward or multi-way.  For example, the subdomain consist of variables 
-in lhs in forward propagation heuristic.  For backward, the subdomain 
-consist of variables in rhs, whereas the subdomain contains all variables 
-for multi-way propagation.  A {\em worklist} is a queue which contains 
-a set of abstract deduction transformers.  
-The initWorklist initializes the worklist 
-following a heuristics choice for propagation strategies.  A {\em forward} 
-propagation strategy initializes the worklist with abstract deduction 
-transformers that contains constants in the right-hand side. Whereas, a 
-{\em backward} iteration strategy starts with a set of assertions. On the 
-other hand, a {\em multi-way} iteration strategy initializes the worklist 
-with set of all transformers and performs both forward and backward 
-propagations in tandem.  \rmcmt{wrong to say abstract deduction transformer that 
-has lhs, rhs since since it also includes the live variables}.
+a propagation heuristic. 
+%
+The algorithm maintains a  {\em worklist}, which is a queue that contains 
+abstract deduction transformers.  
+The propagation heuristics provides two 
+functions $\initworklist$ and $\updateworklist$.
+The order of the elements in the worklist and the subdomains $\subdomain$
+attached to each of the abstract deduction transformers
+determine the propagation strategy (forward, backward, multi-way).
+%
+The {\em forward} propagation strategy initialises the worklist with
+abstract deduction transformers that contains constants in the
+right-hand side and the subdomain is constructed via $\makesubdomain$
+from variables in left-hand side of the equality constraints
+originating from the assignment statements in the program.
+%
+The {\em backward} iteration strategy intialises the worklist with the assertions and the subdomain is constructed from the right-hand side variables.
+%
+The {\em multi-way} iteration strategy initialises the worklist 
+with set of all transformers and each subdomain contains the variables occurring in the respective transformers.
 
-Given a constraint system and a domain, an updateWorklist function adds 
-new abstract deduction transformers to the worklist for those variables in 
-the subdomain that appear in newDeductions.  It also updates the subdomain of 
+Depending on the propagation heuristics $\updateworklist$ adds
+abstract deduction transformers to the worklist that contain variables
+that appear in $\newdeductions$.  It also updates the subdomains of
 existing abstract deduction transformers that are in the worklist.
@@
 $\onlynew(\absval)=\bigsqcap(\decomp(\absval)\setminus\decomp(\abs(\trail)))$
-filters out all known meet irreducibles in the abstract value.
+filters out all meet irreducibles that are already on the trail.
 
+\Omit{
 For every abstract deduction transformer in the worklist, 
@@
 model search procedure makes a new decision.    
+}

### 2016-10-22 decision.tex

@@
 We implemented several decision heuristics namely, {\em ordered}, 
-{\em longest-range}, \pscmt{are these two combined?} {\em random} and the
+{\em longest-range}, {\em random} and the 
 {\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
@@
 before choosing meet irreducibles with numerical variables.  
-The ordering heuristic is similar to trace partitioning~\cite{toplas07}. \pscmt{what is this?}
+The ordered heuristic is similar to trace partitioning~\cite{toplas07}.
 
@@
 The {\em random} decision heuristics arbitrarily picks a meet irreducible  
-for making decision. \pscmt{uniformly?}
+for making decision. %\pscmt{uniformly?}
 %

### 2016-10-22 graphs.tex

@@
       xmin=0.1,xmax=2500, ymin=0.1, ymax=2500, 
-      xlabel=Forward Propagation, ylabel=Chaotic Propagation, 
+      xlabel=Forward Propagation, ylabel=Multi-way Propagation, 
       title={Performance of Propagation Heuristics},
@@
  \addplot table[x=benchmark,y=time] {plotdata/random.dat};
+ \addplot table[x=benchmark,y=time] {plotdata/ordered.dat};
  \addlegendentry{Berkmin}
  \addlegendentry{Random}
+ \addlegendentry{Ordered}
  %\addplot [domain=.1:2500] {};

### 2016-10-22 paper.tex
more spell checking
microtype
sync
enable comments again
nudge abstract

@@
 \usepackage{times}
+\usepackage{microtype}
 \usepackage{array}
@@
 
-\newcommand{\extendedonly}[1]{#1}
-\newcommand{\paperonly}[1]{}
+\newcommand{\extendedonly}[1]{}
+\newcommand{\paperonly}[1]{#1}
 
@@
 \newcommand{\mydef}[1]{\begin{definition}#1\end{definition}}
-\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}
-\newcommand{\pscmt}[1]{{\color{blue}{#1}}}
+\newcommand{\rmcmt}[1]{{\color{magenta}{#1}}}%#1
+\newcommand{\pscmt}[1]{{\color{blue}{#1}}}%#1
 
@@
 %
-In this paper, we present the non-trivial adaptations of the model search
-and conflict analysis algorithms in order to enable the use of
-relational abstract domains.  We have implemented these algorithms inside
-the 2LS verification tool and instantiated our analyser over a template 
-polyhedra domain to check safety properties of C programs.  We evaluate 
-the performance of our analyser in comparison with program verifiers that 
-uses propositional-level solvers and a static analysis tool.
+In this paper, we present the non-trivial liftings of the model search and
+conflict analysis algorithms for relational abstract domains.  We have
+implemented these algorithms inside the 2LS verification tool and
+instantiated our analyser over a template polyhedra domain to check safety
+properties of C programs.  We evaluate the performance of our analyser in
+comparison with program verifiers that uses propositional-level solvers and
+a static analysis tool.
 %
@@
 }
-two orders of magnitude reduction in the number of 
-decisions and conflict analysis compared to a satisfiability 
-solver while being more precise compared to an abstract interpreter. 
+a reduction of two orders of magnitude in the number of 
+decisions and conflict analyses compared to a satisfiability 
+solver and higher precision compared to an abstract interpreter. 
 \Omit{

### 2016-10-22 paper.tex
nudge abstract

@@
 of the well-known Conflict-Driven Clause Learning (CDCL) algorithm that is
-used in SAT solvers.  ACDCL alternates between a model search, which
+used in SAT solvers.
+\Omit {  ACDCL alternates between a model search, which
 performs over-approximate deduction with constraint propagation, and a
@@
 choice.
+}
 %
@@
 solvers can be rather easily lifted from the Boolean lattice to other 
-non-relational domains, lifting them to relational domains is 
+non-relational domains, application to relational domains is 
 challenging because of relational properties of deductions.

### 2016-10-22 deductions.tex

@@
   $\abstransel{\subdomain} \leftarrow \mathit{worklist.pop}()$ \; 
-  \rmcmt{how do we get the $\subdomain$ at this point ?}
   $\newdeductions \leftarrow \onlynew(\abstransel{\subdomain}(\abs(\trail)))$\;
   \uIf{$\newdeductions = \bot$} {
-    $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \rmcmt{shouldn't this be
-    just $\abstransset$} \;
+    $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \;
     $\mathit{worklist.clear}()$ \;
@@
 {\em decisions} to heuristically guess a value for an unassigned 
-literal.  The unit rule overapproximates a model transformer and deduction 
-computes a greatest fixed point over the partial assignments
-domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
-that computes a greatest fixed point over meet irreducible deduction 
-transformer in $S$ \pscmt{???}.  
+literal.  
+%The unit rule overapproximates a model transformer and deduction 
+%computes a greatest fixed point over the partial assignments
+%domain~\cite{dhk2013-popl}.  We present an abstract model search procedure 
+%that computes a greatest fixed point over meet irreducible deduction 
+%transformer in $S$.  
 %
@@
 
-To make our algorithm efficient, we have to focus abstract
+To make our algorithm efficient, we have to focus the abstract
 transformers on performing only the minimally necessary
@@
 
-Depending on the propagation heuristics $\updateworklist$ adds
+Depending on the propagation heuristics, $\updateworklist$ adds
 abstract deduction transformers to the worklist that contain variables
-that appear in $\newdeductions$.  It also updates the subdomains of
-existing abstract deduction transformers that are in the worklist.
+that appear in $\newdeductions$, and it updates the subdomains of
+the abstract deduction transformers in the worklist 
+to include the variables in $\newdeductions$.
 
@@
 $\onlynew(\absval)=\bigsqcap(\decomp(\absval)\setminus\decomp(\abs(\trail)))$
-filters out all meet irreducibles that are already on the trail.
+is used to filter out all meet irreducibles that are already on the trail
+in order to obtain only new deductions when applying the abstract deduction transformer.
 
-\Omit{
-For every abstract deduction transformer in the worklist, 
-the set of new meet irreducibles inferred by the domain is added 
-to the $\trail$.  Additionally, the reason trail $\reasons$ is 
-updated with the corresponding abstract transformer as the reason for inferring 
-those new meet irreducibles.  If a conflict is deduced, 
-%that is, $\mathit{meet\_irrd} = \bot$ is inferred by the domain, 
-the algorithm terminates with \textsf{conflict}.  However,
-when a fixed-point is reached, the abstract value $v$ computed from 
-the trail is checked for set of models.  This is determined by 
-checking whether the set of assignments in $v$ is 
-$\gamma$-complete~\cite{dhk2013-popl}. If $v$ is $\gamma$-complete, that is, the 
-abstract valuation $v$ is abstractly satisfying, the abstract model search 
-terminates with \textsf{sat}.  Otherwise, the algorithm returns \textsf{unknown} and the 
+If $\bot$ is deduced we return \textsf{conflict}.
+Otherwise, when eventually a fixed point has been reached, i.e.\ the worklist is empty, then the abstract value $\abs(\trail)$ is checked whether it is 
+$\gamma$-complete~\cite{dhk2013-popl}. It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
+Otherwise, the algorithm returns \textsf{unknown} and the 
 model search procedure makes a new decision.
-}

### 2016-10-22 paper.tex
nudge abstract
added dkcmt

@@
 \newcommand{\pscmt}[1]{{\color{blue}{#1}}}%#1
+\newcommand{\dkcmt}[1]{{\color{red!70!black}{#1}}}%#1
 
@@
 \emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
-of the well-known Conflict-Driven Clause Learning (CDCL) algorithm that is
-used in SAT solvers.
+of the well-known CDCL algorithm that is used in SAT solvers.
+%
 \Omit {  ACDCL alternates between a model search, which
@@
 instantiated our analyser over a template polyhedra domain to check safety
-properties of C programs.  We evaluate the performance of our analyser in
+properties of C programs.  We~evaluate the performance of our analyser in
 comparison with program verifiers that uses propositional-level solvers and
@@
 %
-Experimental evidence suggest 
 \Omit {
@@
 }
-a reduction of two orders of magnitude in the number of 
+We observe a reduction of two orders of magnitude in the number of 
 decisions and conflict analyses compared to a satisfiability 

### 2016-10-22 experiments.tex
right align numbers
nice symbol names
cleanup experiments

@@
 \section{Experimental Results}
-We have implemented the instantiation of ACDCL on relational domains
-in 2LS verification tool~\cite{2ls}.  We have used the abstract domains
+
+We have implemented the instantiation of ACDCL on relational domains in the
+2LS verification tool~\cite{2ls}.  We have used the abstract domains
 provided by 2LS (intervals, octagons, polyhedra and equalities) for our
-experiments. Although the performance of the abstract domain implementations 
-in 2LS is not competitive with that of APRON~\cite{apron} library, but 
-the domain implementation in 2LS can handle all C operators 
-(including bitwise operations) out of the box and support precise
-complementation of meet irreducibles, which are necessary for conflict-driven 
-learning. Our tool and benchmarks are available online\footnote{http://www.cprover.org/acdcl/}.
+experiments.  The performance of the abstract domain implementations in 2LS
+is not competitive with that of the APRON~\cite{apron} library; however, the
+domain implementation in 2LS handles all C operators (including bitwise
+operations) and supports precise complementation of meet irreducibles, which
+is necessary for conflict-driven learning.  Our tool and benchmarks are
+available online\footnote{http://www.cprover.org/acdcl/} to enable other
+researchers to reproduce our results.
 
-We verified a total of 66 benchmarks in ANSI-C, which are derived from 
-(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
-cycle accurate software models of hardware circuits automatically 
-generated using v2c~\cite{mtk2016}, (c) controller code generated 
-from Simulink (d) several hand-crafted benchmarks for equivalence 
-checking and bounded loop analysis.  All bounded loops are completely 
-unrolled before analysis.  We compare our tool with a state-of-the-art 
-SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
-static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
-Astr{\'e}e is configured with range of abstract domains which includes 
-interval, bit-field, congruence, trace partitioning and relational 
-domains (octagons, polyhedra, zones, equalities, filter).  Whereas, 
-ACDCL is instantiated with a product domain of booleans and template 
-polyhedra.  We use an interval and octagon templates for our
-experiments.  Additionally, the analysis using ACDCL is configured with 
-a heuristic choice of decision (ordered, random, berkmin), 
-propagation (forward and multi-way) and conflict-analysis (learning UIP,
-DPLL-style).  The timeout for our experiments is set to ~1 hour. 
+We verified a total of 66 ANSI-C benchmarks, which are derived from (a)~the
+bit-vector regression category in SV-COMP'16, (b)~bit-precise and
+cycle-accurate software models of hardware circuits automatically generated
+using v2c~\cite{mtk2016}, (c)~controller code generated from Simulink
+(d)~several hand-crafted benchmarks for equivalence checking and bounded
+loop analysis.  All bounded loops are completely unrolled before analysis. 
+We compare our tool with a state-of-the-art SAT-based bounded model checker
+CBMC~\cite{cbmc} and a commercial static analysis tool, Astr{\'e}e. 
+\dkcmt{INSERT VERSION NUMBERS OF CBMC AND ASTREE.} CBMC uses MiniSAT in the
+backend.  Astr{\'e}e uses a range of abstract domains, which includes
+interval, bit-field, congruence, trace partitioning and relational domains
+(octagons, polyhedra, zones, equalities, filter).  ACDCL is instantiated
+with a product domain of booleans and template polyhedra.  We use the
+interval and octagon templates for our experiments.  ACDCL is configured
+with a heuristic choice of decision (ordered, random, berkmin), propagation
+(forward and multi-way) and conflict-analysis (learning UIP, DPLL-style). 
+The timeout for our experiments is set to one hour.
 %
@@
 }
+
 \paragraph {\textbf{Precision and Efficiency of Analysis}}
-Figure~\ref{fig:results} presents a comparison of the analysis 
-using CBMC and ACDCL.  Figure~\ref{fig:results}(a) clearly shows that 
-the SAT solver made significantly more decisions compared to ACDCL 
-for all the benchmarks.  Whereas, the extreme right points below the 
-diagonal in Figure~\ref{fig:results}(b) shows the propagations in SAT solver  
-is maximum for the benchmarks that exhibit relational behavior.  The 
-cumulative statistics corresponding to each phase of the algorithm 
-for all 66 benchmarks is presented in Table~\ref{result}.  
-Experimental result shows a reduction of at least two orders of 
-magnitude in the number of decisions, propagations and conflicts 
-compared to analyis using CBMC.    
 %
+Figure~\ref{fig:results} presents a comparison of the analyses using CBMC
+and ACDCL.  Figure~\ref{fig:results}(a) clearly shows that the SAT solver
+made significantly more decisions compared to ACDCL for all the benchmarks. 
+The points on the extreme right below the diagonal in
+Figure~\ref{fig:results}(b) show that the number of propagations in the SAT
+solver is maximal for the benchmarks that exhibit relational behavior. 
+Cumulative statistics corresponding to each phase of the algorithm for all
+66 benchmarks is given in Table~\ref{result}.  We observe a reduction of at
+least two orders of magnitude in the number of decisions, propagations and
+conflicts compared to analyis using CBMC.
+
 \begin{table}[t]
@@
 {
-\begin{tabular}{l|l|l|l|l|l}
+\begin{tabular}{l|r|r|r|r|r}
 \hline
@@
 \end{table}
-%
-Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
-without any restarts.  The solver is restarted in 50 cases to avoid spending 
-too much time in branches that do not easily lead to a satisfying assignment or 
-stronger clause learning.  In contrast, the analysis using ACDCL solved all
-66 benchmarks without any restarts.  This is attributed to decision 
-heuristics which exploits the high-level structure of the program combined 
-with the stronger deduction and clause learning mechanisms aided by the richer 
-abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
-but the analysis using Astr{\'e}e shows a high degree of imprecision.  
-Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.
+
+Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16
+benchmarks without any restarts.  The solver is restarted in 50 cases to
+avoid spending too much time in ``hopeless'' branches.  By contrast, the
+analysis using ACDCL solved all 66 benchmarks without any restarts.  This is
+attributed to our decision heuristics, which exploit the high-level
+structure of the program combined with the stronger deduction and clause
+learning mechanisms aided by the richer abstract domains.  On the other
+hand, Ast{\'e}e is often faster than ACDCL, but the analysis using
+Astr{\'e}e shows a high degree of imprecision.  Astr{\'e}e reported 15 false
+alarms for the 41 safe benchmarks.
 
 \paragraph {\textbf{Propagation Strategy}}
+%
 Figure~\ref{prop-dec}(a) presents a comparison between {\em forward}
-propagation strategy and {\em multi-way} propagation strategy in
-ACDCL.  The choice of the propagation strategy influences the total
-number of decisions and clause learning iterations.  Hence, the
-propagation strategy has a significant influence on the run time which
-can be visualised in Figure~\ref{prop-dec}(a).  Compared to forward
-propagation, the multi-way strategy may take more iterations to reach
-the fixed-point, but it significantly reduces the total number of
-decisions and conflicts, in particular for the unsafe benchmarks.
-This is attributed to the higher precision of the meet irreducibles
-inferred by the multi-way strategy which subsequently aids the
-decision heuristics to make better decisions.
-\pscmt{Chaotic $\rightarrow$ multi-way in Fig.6 }
+propagation strategy and {\em multi-way} propagation strategy in ACDCL.  The
+choice of the propagation strategy influences the total number of decisions
+and clause learning iterations.  Hence, the propagation strategy has a
+significant influence on the run time which can be visualised in
+Figure~\ref{prop-dec}(a).  Compared to forward propagation, the multi-way
+strategy may take more iterations to reach the fixed-point, but it
+significantly reduces the total number of decisions and conflicts, in
+particular for the unsafe benchmarks.  This is attributed to the higher
+precision of the meet irreducibles inferred by the multi-way strategy which
+subsequently aids the decision heuristics to make better decisions.
 
-\paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
-visualization of the performance of different decision heuristics in ACDCL.  
-Note that the runtimes for all decision heuristics are obtained with multi-way 
-propagtion strategy.  The runtimes are very close, but we observe some key
-characteristics of these heuristics.  Berkmin heuristic performs consistently 
-well for most safe benchmarks and all bit-vector category benchmarks.  Whereas, 
-the ordered heuristic performs better for programs with conditional branches 
-since it prioritises decisions on meet irreducibles that appear in conditional 
+\paragraph {\textbf{Decision Heuristics}}
+%
+Figure~\ref{prop-dec}(b) presents a visualization of the performance of
+different decision heuristics in ACDCL.  Note that the runtimes for all
+decision heuristics are obtained with multi-way propagtion strategy.  The
+runtimes are very close, but we observe some key characteristics of these
+heuristics.  Berkmin heuristic performs consistently well for most safe
+benchmarks and all bit-vector category benchmarks.  Whereas, the ordered
+heuristic performs better for programs with conditional branches since it
+prioritises decisions on meet irreducibles that appear in conditional
 branches over meet irreducibles that involve numerical variables.  Whereas,
-the runtimes for random heuristic is marginally higher than the other two.  This 
-suggest that domain-specific decision heuristics are important for ACDCL.
+the runtimes for random heuristic is marginally higher than the other two. 
+This suggest that domain-specific decision heuristics are important for
+ACDCL.
 
@@
 
-\paragraph {\textbf{Learning}} Learning has a significant influence on
-the runtime of ACDCL.  We compare the UIP-based learning technique with 
-an analysis that performs chronoligical backtracking without learning. 
-The effect of UIP computation allows ACDCL to backtrack non-chronologically 
-and guide the model search with a learnt transformer.  However, the DPLL-style 
-analysis exhibits case-enumeration behavior and could not finish within the time 
-bound for 20\% of our benchmarks.
+\paragraph {\textbf{Learning}}
+%
+Learning has a significant influence on the runtime of ACDCL.  We compare
+the UIP-based learning technique with an analysis that performs
+chronoligical backtracking without learning.  The effect of UIP computation
+allows ACDCL to backtrack non-chronologically and guide the model search
+with a learnt transformer.  However, the DPLL-style analysis exhibits
+case-enumeration behavior and could not finish within the time bound for
+20\% of our benchmarks.
 

### 2016-10-22 deductions.tex

@@
 \emph{closure} operation for relational domains in a lazy manner. To
-this end, we construct a $\subdomain=\makesubdomain_\domain(\vars)$
+this end, we construct a $\subdomain=\makesubdomain_\domain(\subvars)$
 which is sufficient to perform one step of the closure operation each
@@
 For example, let us consider $\domain=\octagons[\{x,y,z\}]$ and
-$\vars=\{y\}$. An octagonal inequality relates at 
+$\subvars=\{y\}$. An octagonal inequality relates at 
 most two variables. Thus it is sufficient to consider the subdomain

### 2016-10-22 graphs.tex
caps
caption

@@
 \begin{tikzpicture}[scale=0.75]
-	\begin{loglogaxis} [xmin=0.1,xmax=4000, ymin=0.1, ymax=4000, xlabel= SAT (decision),
-			ylabel=ACDCL (decision), 
+	\begin{loglogaxis} [xmin=0.1,xmax=4000, ymin=0.1, ymax=4000, xlabel= SAT (Decision),
+			ylabel=ACDCL (Decision), 
 			legend pos = north west,
@@
 \caption{\label{fig:results}
-Comparison between SAT and ACDCL 
-}
+Comparison of the number of decisions and propagations performed by
+propositional SAT and ACDCL}
 \end{figure}

### 2016-10-22 background.tex
nudge
fix

@@
 programs is Satisfiability Modulo Path Programs (SMPP) \cite{SMPP}. SMPP
-enumerates path programs using a SAT formula which are then verified
+enumerates program paths using a SAT formula, which are then verified
 using abstract interpretation.  
 %
-The lifting of CDCL to first-order theory is proposed
-in~\cite{dpll,cp09} and natural domain SMT framework~\cite{ndsmt}.
+The lifting of CDCL to first-order theories is proposed
+in~\cite{dpll,cp09,ndsmt}.
+%
 \Omit{ operates on a fixed first-order partial assignment lattice
@@
 %
-\cite{DBLP:journals/fmsd/BrainDGHK14} presents an instantiation of ACDCL for heap-manipulating programs, but uses a simplistic conflict analysis.
+\cite{DBLP:journals/fmsd/BrainDGHK14} presents an instantiation of ACDCL for
+heap-manipulating programs, but uses a simplistic conflict analysis.
 %
-\cite{DBLP:conf/vmcai/PelleauMTB13} describes a constraint solving algorithm using relational numerical domains, but performs model search only.
+\cite{DBLP:conf/vmcai/PelleauMTB13} describes a constraint solving algorithm
+using relational numerical domains, but performs model search only.
 %
-\cite{DBLP:conf/esop/MineBR16} proposes an algorithm inspired by constraint solvers for inferring disjunctive invariants using the interval domain.
+\cite{DBLP:conf/esop/MineBR16} proposes an algorithm inspired by constraint
+solvers for inferring disjunctive invariants using the interval domain.
+%
 % is presented in . They

### 2016-10-22 deductions.tex
improved explanation in 5.1

@@
 %
-\Omit{
-%TODO
-Hence, we do not compute the full closure in the full 
-domain $\octagons[\{x,y,z\}]$, but we compute only a single step of the closure when the abstract deduction transformer is applied.
+Hence, when the abstract deduction transformer is applied we do not
+compute the full closure in the full domain,
+but we compute only a single step of the closure in a restricted
+domain, which makes single deduction steps more efficient.
 %
-Only if $\abstrans^\subdomain$ deduces new information about $x$ or $z$
-then the next step of the closure will be 
-}
-If this indeed affects $x$ and~$z$ then their
-relations will be inferred in a later step of the propagation
-algorithm, thus gradually computing the closure.
+If $\abstransel{\subdomain}$ deduces new information about $x$ or $z$
+then the next step of the closure will be computed by the worklist
+mechanism of Algorithm~\ref{Alg:ms} that we describe next.
+
+%If this indeed affects $x$ and~$z$ then their
+%relations will be inferred in a later step of the propagation
+%algorithm, thus gradually computing the closure.
 

### 2016-10-22 conclusion.tex

@@
 %
-Experimental evaluation shows that ACDCL is able to suitably exploit
-the expressivity of richer abstract domains as well as precision of
-CDCL architecture.
+Experimental evaluation shows that ACDCL is able to suitably exploit the
+expressivity of richer abstract domains as well as the precision of CDCL
+architecture.
 

### 2016-10-22 example.tex
right align numbers
nudge Fig 1

@@
 \section{Motivating Example}
-%
+
 \begin{figure}[t]
-\scriptsize
-\begin{tabular}{l|l|l}
+\scriptsize\centering
+\begin{tabular}{c|c|c}
 \hline
@@
 {
-  assume(v>=0 && v<=N);
+  assume(v>=0 &&
+         v<=N);
   int x; 
@@
 %\vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{example.pspdftex}}
+\scalebox{.52}{\import{figures/}{example.pspdftex}}
 %\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
@@
 &
-\begin{minipage}{4.50cm}
+\begin{minipage}{5.6cm}
 \centering
 \vspace*{0.3cm}
-\scalebox{.45}{\import{figures/}{acdl_run.pspdftex}}
+\scalebox{.5}{\import{figures/}{acdl_run.pspdftex}}
 %\caption{A control flow graph for P, P' and abstract conflict graph (ACFG) \label{fig:filter}}
@@
 \end{figure}
+
 Let us consider a simple program in Figure~\ref{fig:example} that performs 

### 2016-10-22 example.tex
clean up sec 2

@@
 
-Let us consider a simple program in Figure~\ref{fig:example} that performs 
-multiplication of two machine integers and checks that the result of the 
-multiplication is always positive for an integer input $v$.  To avoid 
-overflow error, we assume $v$ has an upper bound $N$.  The control 
-flow graph is shown in the middle column of Figure~\ref{fig:example}.
+Let us consider a simple program in Figure~\ref{fig:example}, which performs
+multiplication of machine integers and checks that the result is always
+positive.  To avoid overflow error, we assume that the input $v$ has an
+upper bound~$N$.  The control-flow graph is given in the middle column of
+Figure~\ref{fig:example}.
 
@@
 model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
+%
 \Omit {
@@
 }
-Standard BMC tools convert the program to a bit-vector equation and pass 
-them to a SAT solver.  However, standard forward interval analysis is 
-too imprecise to verify the safety of the above program.  The 
-imprecision is due to control-flow join at node $n4$.  State-of-the-art 
-abstract interpretation tool, Astr{\'e}e \cite{se2011}, requires external hints, provided 
-by manually annotating the code with partition directives at $n1$, 
-to prove safety.  The partition directive helps the tool to analyse 
-each program execution paths separately.  To do so, Astr{\'e}e 
-requires a product of trace partitioning domain and an interval domain 
-to determine safety.  However, a relational domain such as octagons is 
-able to prove the program without any partition.  
-\Omit {
-In general, the imprecision is either intended by the tool because such high 
-precision analysis is normally not required for runtime error 
-analysis or the imprecision is unavoidable due to the complexity 
-of the application under analysis.  
+%
+Standard forward interval analysis is too imprecise to verify the
+safety of the above program.  The imprecision is due to control-flow join at
+node $n4$.  A state-of-the-art abstract interpretation tool,
+Astr{\'e}e~\cite{se2011}, requires external hints, provided by manually
+annotating the code with partition directives at $n1$, to prove safety.  The
+partition directive tells the tool to analyse each program execution path
+separately.  To determine safety, Astr{\'e}e requires a product of a trace
+partitioning domain and an interval domain or a relational domain
+such as octagons.
+%
+\Omit { In general, the imprecision is either intended by the tool because such
+high precision analysis is normally not required for runtime error analysis
+or the imprecision is unavoidable due to the complexity of the application
+under analysis.
 }
-%
+
 \begin{table}[!b]
@@
 \end{center}
-\caption{Solver statistics}
+\caption{Solver statistics for Example given in Fig.~\ref{fig:example}}
 \label{solver}
 \end{table}
-%
-The right-hand side of Figure~\ref{fig:example} shows an analysis using ACDCL with 
-octagon domain (top) and an interval domain (bottom).  The analysis 
-associates an abstract element for each control location and variable, 
-as shown in the abstract conflict graph in Figure~\ref{fig:example}.  
-Octagonal constraints deduced by ACDCL solver are sufficient to prove 
-safety without any decision or clause learning.  However, unlike Ast{\'e}e, 
-an automatic interval analysis is also sufficient to prove safety with 
-only a single decision.  Interval constraints generated from forward 
-analysis in the initial deduction phase infers that 
-$\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.
+
+BMC converts the program into a bit-vector equation and passes that to a SAT
+solver.
+
+The right-hand side of Figure~\ref{fig:example} illustrates an analysis
+using ACDCL with the octagon domain (top) and the interval domain (bottom). 
+The analysis associates an abstract element with each control location and
+variable.  The octagons deduced by ACDCL are sufficient to prove safety
+without any decision or clause learning.  However, unlike Ast{\'e}e, an
+interval analysis is also sufficient to prove safety.  The intervals
+generated by forward analysis in the initial deduction phase are $\langle
+x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.
 

### 2016-10-22 introduction.tex
line breaks
intro

@@
 \section{Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
-Static program analysis based on abstract 
-interpretation~\cite{CC77} has been widely used to
-verify certain classes of properties for safety-critical systems.  A 
-standard static analyses~\cite{se2011} for program safety aims to compute 
-invariants as fixed-points of abstract transformers.  Most lattice structure 
-used in static analysis have meet and join operations.  The meet operation 
-precisely model conjunction, and the join operation over-approximate 
-disjunction. Thus, the precision loss in static analysers is often 
-due to imprecise join operations. The loss in precision is overcome 
-by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration 
-behaviour. 
+Static program analysis based on abstract interpretation~\cite{CC77} has
+been widely used to verify certain classes of properties for safety-critical
+systems.  A standard static analyses~\cite{se2011} for program safety aims
+to compute invariants as fixed-points of abstract transformers.  Most
+lattice structure used in static analysis have meet and join operations. 
+The meet operation precisely model conjunction, and the join operation
+over-approximate disjunction.  Thus, the precision loss in static analysers
+is often due to imprecise join operations.  The loss in precision is
+overcome by using richer abstract domains or equipping the analysis with
+disjunction -- which is often very expensive since the analysis exhibit case
+enumeration behaviour.
+%
 %%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
+%
 On the other hand, the combination of precise conjunction in a partial
-assignments domain with learning allows Conflict Driven Clause
-Learning (CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to
-reason about disjunction without enumerating cases.  Silva
-et. al.~\cite{tacas12, sas12, dhk2013-popl} presented a formal recipe
-for applying abstract interpretation to solve the satisfiability
-problem.  \cite{dhk2013-popl} provide a generalisation of
-CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning}
-(ACDCL), in terms of deduction and abduction transformers,
-fixed-point, interpolation and extrapolation operators on lattices.
-The partial assignments in the CDCL algorithm are an abstract domain,
-unit rule is the best abstract transformer, Boolean Constraint
-Propagation (BCP) computes a greatest fixed-point over the abstract
-domain, decision is dual widening and clause learning can be viewed as
-synthesising an abstract transformer for negation.  
+assignments domain with learning allows Conflict Driven Clause Learning
+(CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to reason about
+disjunction without enumerating cases.  Silva et al.~\cite{tacas12, sas12,
+dhk2013-popl} presented a formal recipe for applying abstract interpretation
+to solve the satisfiability problem.  \cite{dhk2013-popl} provide a
+generalisation of CDCL algorithm called, {\em Abstract Conflict Driven
+Clause Learning} (ACDCL), in terms of deduction and abduction transformers,
+fixed-point, interpolation and extrapolation operators on lattices.  The
+partial assignments in the CDCL algorithm are an abstract domain, unit rule
+is the best abstract transformer, Boolean Constraint Propagation (BCP)
+computes a greatest fixed-point over the abstract domain, decision is dual
+widening and clause learning can be viewed as synthesising an abstract
+transformer for negation.
+%
 %Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
@@
 ACDCL has so far been instantiated as a floating-point decision
-procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser
-that only uses interval constraints~\cite{tacas12}.  Hence, the
-analysis is limited to bounds checking in simple numerical programs.
-However, verification of certain class of programs require more
-expressive domains like octagons or equalities or polyhedra.  These
-domains capture interesting properties of programs which intervals can
-not express.  In this paper, we identify specific properties of the
-CDCL algorithm that are necessary to lift the propositional skeleton
-of CDCL to relational abstract domains such as template polyhedra
-domains~\cite{sriram} (octagons, zones and equalities).
-To this end, we present a characterisation of abstract model search
-procedure and conflict analysis procedure that lift the abstract
-deduction transformer and the conflict analysis procedure to
-relational lattice structures.
+procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser that
+only uses interval constraints~\cite{tacas12}.  Hence, the analysis is
+limited to bounds checking in simple numerical programs.  However,
+verification of certain class of programs require more expressive domains
+like octagons or equalities or polyhedra.  These domains capture interesting
+properties of programs which intervals can not express.  In this paper, we
+identify specific properties of the CDCL algorithm that are necessary to
+lift the propositional skeleton of CDCL to relational abstract domains such
+as template polyhedra domains~\cite{sriram} (octagons, zones and
+equalities).  
+\Omit {
+To this end, we present a characterisation of abstract model
+search procedure and conflict analysis procedure that lift the abstract
+deduction transformer and the conflict analysis procedure to relational
+lattice structures.
+}

### 2016-10-22 domains.tex
comments in sec4
equation numbering
nice symbol names
comment in section 3

@@
 $\begin{array}{rcl}
-\conctransset:= (g_0 = (0 \leq v\leq N)) \\
+\constraints:= (g_0 = (0 \leq v\leq N)) \\
 (g_1 = (g_0 \wedge c)) \wedge (x_0 = v) \wedge \\
@@
 techniques~\cite{sriram,GS07b,BJKS15} for computing the domain
-operations, such as $\meet$, $\sqcup$, etc. In our implementation we
+operations, such as \rmcmt{meet($\meet$), join($\sqcup$)}, etc. In our implementation we
 use the strategy iteration approach available in 2LS~\cite{BJKS15}.
@@
 The best abstract transformer is 
-\begin{equation}
+\begin{equation}\label{eq:abstrans}
 \abstrans{\domain}{\constraint}(\absval)=\absval\meet\alpha(\{\val\mid \val\in\gamma(\absval), \val\models \constraint\})
@@
 structure $A$ is an element with the following property.
-\[\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top \]  
+\begin{equation}
+\forall m_1, m_2 \in A: m_1 \meet m_2 = m \implies (m = m_1 \lor m = m_2), m \neq \top  
+\end{equation}
 %\rmcmt{mention minimum complementable element}
@@
 
-\pscmt{check}
+\Omit {
 \begin{definition}{(Abstract Valuation)} An {\em abstract valuation} is a
@@
 \end{definition}
+}
 

### 2016-10-22 introduction.tex
edited intro

--- 2016-10-22 introduction.tex
@@
 \section{Introduction}
+
+%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
 %
-%%%%%%%%%%%%%%%%%%%%%%% AI %%%%%%%%%%%%%%%%%%%%%%%%
-Static program analysis based on abstract 
-interpretation~\cite{CC77} has been widely used to
-verify certain classes of properties for safety-critical systems.  A 
-standard static analyses~\cite{se2011} for program safety aims to compute 
-invariants as fixed-points of abstract transformers.  Most lattice structure 
-used in static analysis have meet and join operations.  The meet operation 
-precisely model conjunction, and the join operation over-approximate 
-disjunction. Thus, the precision loss in static analysers is often 
-due to imprecise join operations. The loss in precision is overcome 
-by using richer abstract domains or equipping the analysis with disjunction --
-which is often very expensive since the analysis exhibit case enumeration 
-behaviour. 
+Static program analysis based on abstract interpretation~\cite{CC77} has
+been widely used to verify of properties of safety-critical systems. 
+Standard static analyses~\cite{se2011} aim to compute program invariants as
+fixed-points of abstract transformers.  Abstract states are chosen from a
+lattice that has meet and join operations.  The meet precisely models set
+intersection (or conjunction, taking a logical view), and the join
+over-approximates set union (or disjunction).  The over-approximation in the
+join operation is a key cause of loss of precision, which yields false
+alarms.  Standard means to address false alarms in static analysis are
+richer abstract domains or equipping the analysis with more precise set
+unions -- which is often very expensive since the analysis exhibits case
+enumeration behaviour.
+
 %%%%%%%%%%%%%%%%%%%%%%% ACDCL %%%%%%%%%%%%%%%%%%%%%
-On the other hand, the combination of precise conjunction in a partial
-assignments domain with learning allows Conflict Driven Clause
-Learning (CDCL)~\cite{cdcl} based satisfiability (SAT) solvers to
-reason about disjunction without enumerating cases.  Silva
-et. al.~\cite{tacas12, sas12, dhk2013-popl} presented a formal recipe
-for applying abstract interpretation to solve the satisfiability
-problem.  \cite{dhk2013-popl} provide a generalisation of
-CDCL algorithm called, {\em Abstract Conflict Driven Clause Learning}
-(ACDCL), in terms of deduction and abduction transformers,
-fixed-point, interpolation and extrapolation operators on lattices.
-The partial assignments in the CDCL algorithm are an abstract domain,
-unit rule is the best abstract transformer, Boolean Constraint
-Propagation (BCP) computes a greatest fixed-point over the abstract
-domain, decision is dual widening and clause learning can be viewed as
-synthesising an abstract transformer for negation.  
-%Thus, ACDCL lifts learning techniques in SAT solver to operate on a non-distributive lattice -- thus equipping the analysis with negation. 
-This enables the
-analyser to refine the analysis and prevent enumeration behaviour.
+%
+On the other hand, propositional SAT solvers are known to be highly
+effective for solving complex propositional formulae, and are frequently
+able to avoid enumerating the exponential number of cases.  The impressive
+performance of modern solvers is credited to well-tuned decision heuristics
+and clause learning algorithms in modern solvers.  Collectively, these
+algorithms are referred to as \emph{Conflict Driven Clause Learning}
+(CDCL)~\cite{cdcl}. The idea to lift these algorithms from the Boolean
+domain to richer, arithmetic domains suggests itself.
 
-ACDCL has so far been instantiated as a floating-point decision
-procedure~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser
-that only uses interval constraints~\cite{tacas12}.  Hence, the
-analysis is limited to bounds checking in simple numerical programs.
-However, verification of certain class of programs require more
-expressive domains like octagons or equalities or polyhedra.  These
-domains capture interesting properties of programs which intervals can
-not express.  In this paper, we identify specific properties of the
-CDCL algorithm that are necessary to lift the propositional skeleton
-of CDCL to relational abstract domains such as template polyhedra
-domains~\cite{sriram} (octagons, zones and equalities).
-To this end, we present a characterisation of abstract model search
-procedure and conflict analysis procedure that lift the abstract
-deduction transformer and the conflict analysis procedure to
-relational lattice structures.
+Abstract Conflict Driven Clause Learning} (ACDCL)~\cite{dhk2013-popl} is one
+such generalisation of CDCL.  The partial satisfying assignments in CDCL
+algorithm are generalized to be chosen from a given abstract domain.  The
+unit rule is the best abstract transformer, Boolean Constraint Propagation
+(BCP) computes a greatest fixed-point over the abstract domain, decision is
+dual widening and clause learning can be viewed as synthesising an abstract
+transformer for negation.  This enables an ACDCL-based analyser to refine an
+imprecise analysis and thus prevents enumeration behaviour.  ACDCL has so
+far been instantiated as a decision procedure for floating-point
+arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser that
+reasons using interval~\cite{tacas12}.  Hence, the existing ACDCL instances
+are limited to bounds checking in numerical programs.
+
+Evidently, the verification of many programs requires more expressive
+domains like octagons or equalities or polyhedra.  These domains capture
+interesting properties of programs that intervals can not express.  In this
+paper, we identify specific modifications to the CDCL algorithm that are
+necessary to lift propositional CDCL to relational abstract domains such as
+octagons, zones and equalities over numerical variables.  To this end, we
+present a characterisation of an abstract model search procedure and a
+conflict analysis procedure that lift the abstract deduction transformer and
+the conflict analysis procedure to relational lattice structures.

### 2016-10-22 introduction.tex

@@
 necessary to lift propositional CDCL to relational abstract domains such as
-octagons, zones and equalities over numerical variables.  To this end, we
+octagons, zones and equalities over numerical variables.
+\Omit { To this end, we
 present a characterisation of an abstract model search procedure and a
@@
 the conflict analysis procedure to relational lattice structures.
+}

### 2016-10-22 experiments.tex
version number

@@
+%================  
+\input{graphs}
+%================
+
 \section{Experimental Results}
@@
 We compare our tool with a state-of-the-art SAT-based bounded model checker
-CBMC~\cite{cbmc} and a commercial static analysis tool, Astr{\'e}e. 
-\dkcmt{INSERT VERSION NUMBERS OF CBMC AND ASTREE.} CBMC uses MiniSAT in the
-backend.  Astr{\'e}e uses a range of abstract domains, which includes
-interval, bit-field, congruence, trace partitioning and relational domains
+CBMC (version 5.5)~\cite{cbmc} and a commercial static analysis tool, 
+Astr{\'e}e (version 14.10). 
+CBMC uses MiniSAT~2.2.1 in the backend.  Astr{\'e}e uses a 
+range of abstract domains, which includes interval, 
+bit-field, congruence, trace partitioning and relational domains
 (octagons, polyhedra, zones, equalities, filter).  ACDCL is instantiated
@@
 20\% of our benchmarks.
-
-%================  
-\input{graphs}
-%================

### 2016-10-22 experiments.tex
solver name

--- 2016-10-22 experiments.tex
@@
 Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
-SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
+MiniSat & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
 ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  

### 2016-10-22 example.tex

@@
 Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
-SAT & 233 & 36436 & 162 & 2604 & 2 \\ \hline
+MiniSat & 233 & 36436 & 162 & 2604 & 2 \\ \hline
 ACDCL (Interval) & 1 & 17 & 1 & 1 & 0 \\ \hline

### 2016-10-22 example.tex
edits Sec 2

@@
 multiplication of machine integers and checks that the result is always
-positive.  To avoid overflow error, we assume that the input $v$ has an
-upper bound~$N$.  The control-flow graph is given in the middle column of
-Figure~\ref{fig:example}.
+positive.  To avoid overflow, we assume that the input $v$ has an upper
+bound~$N$.  The control-flow graph is given in the middle column of
+Figure~\ref{fig:example}.  We apply three different analysis techniques:
+1)~{\em bounded model checking (BMC)}, 2)~{\em abstract interpretation} and
+3)~{\em ACDCL}.
 
-We analyse the program using three different analysers -- {\em bounded 
-model checking (BMC)}, {\em abstract interpretation} and {\em ACDCL}.    
 %

### 2016-10-22 example.tex
edits Sec 2

@@
 \begin{figure}[t]
-\scriptsize\centering
+\centering
 \begin{tabular}{c|c|c}
@@
 \hline
+\scriptsize
 \begin{lstlisting}[mathescape=true,language=C]
@@
 
-Let us consider a simple program in Figure~\ref{fig:example}, which performs
-multiplication of machine integers and checks that the result is always
-positive.  To avoid overflow, we assume that the input $v$ has an upper
-bound~$N$.  The control-flow graph is given in the middle column of
+Let us consider the simple program in Figure~\ref{fig:example}, which
+squares a machine integer and checks that the result is always positive.  To
+avoid overflow, we assume that the input $v$ has an upper bound~$N$.  The
+control-flow graph is given in the middle column of
 Figure~\ref{fig:example}.  We apply three different analysis techniques:
-1)~{\em bounded model checking (BMC)}, 2)~{\em abstract interpretation} and
+1)~{\em abstract interpretation},
+1)~{\em bounded model checking (BMC)} and
 3)~{\em ACDCL}.
 
-%
-\Omit {
-The SAT solver makes 36 decisions, 1365 Boolean constraint propagations 
-and performs 5 conflict analysis with 65 conflict literals.  Moreover, 
-the solver performs a restart, presumably to recover from bad decisions.  
-}
-%
-Standard forward interval analysis is too imprecise to verify the
-safety of the above program.  The imprecision is due to control-flow join at
-node $n4$.  A state-of-the-art abstract interpretation tool,
+A standard forward interval analysis is too imprecise to verify the safety
+of the above program.  The imprecision is owed to the control-flow join at
+node~$n4$.  A state-of-the-art abstract interpretation tool,
 Astr{\'e}e~\cite{se2011}, requires external hints, provided by manually
@@
 separately.  To determine safety, Astr{\'e}e requires a product of a trace
-partitioning domain and an interval domain or a relational domain
-such as octagons.
+partitioning domain and an interval domain or a relational domain such as
+octagons.
 %
@@
 
-BMC converts the program into a bit-vector equation and passes that to a SAT
-solver.
+BMC converts the program into a bit-vector equation and passes that to a
+CDCL-based SAT solver.  We will contast the analysis performed by CDCL with
+that of ACDCL.  The right-hand side of Figure~\ref{fig:example} illustrates
+an analysis using ACDCL with the octagon domain (top) and the interval
+domain (bottom). The analysis associates an abstract element with each
+control location and variable.  The octagons deduced by ACDCL are sufficient
+to prove safety without any decision or clause learning, as in the case of
+Ast{\'e}e.
 
-The right-hand side of Figure~\ref{fig:example} illustrates an analysis
-using ACDCL with the octagon domain (top) and the interval domain (bottom). 
-The analysis associates an abstract element with each control location and
-variable.  The octagons deduced by ACDCL are sufficient to prove safety
-without any decision or clause learning.  However, unlike Ast{\'e}e, an
-interval analysis is also sufficient to prove safety.  The intervals
-generated by forward analysis in the initial deduction phase are $\langle
-x:[-5,5], z:[-25,25] \rangle$.  Clearly, ACDCL cannot prove safety.
+The benefit of ACDCL on this example becomes apparent using the interval
+domain.  The intervals generated by forward analysis in the initial
+deduction phase are $\langle x:[-5,5], z:[-25,25] \rangle$.  Clearly, these
+do not prove safety.  Hence, ACDCL makes a heuristic decision to refine the
+analysis.  Assume we decide that $c$ is $[1,1]$.  Using deduction, ACDCL
+determines that the interval of $x$ is then constrained to $[0,5]$.  The
+deductions made during fixed point iteration are represented by an abstract
+conflict graph, given in right-hand side of Figure~\ref{fig:example}.  Nodes
+in the abstract conflict graph denote updates to the program variables in
+the CFG.  The interval analysis concludes that $(\text{Error}:\bot)$, that
+is, the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, the program is
+{\em safe} for the case $c:=[1,1]$.
 
-Hence, ACDCL makes a decision $c:=[1,1]$ to refine the analysis.  
-ACDCL implements several decision heuristics.  In this case, the 
-branching variable $c$ is chosen by an {\em ordered} decision 
-heuristics that select variables appearing in conditional branches 
-first before choosing other program variables. 
-The decision constrains the interval of $x:=[0,5]$.  The deductions made 
-during fixed point iteration are represented by abstract conflict 
-graph in right-hand side of Figure~\ref{fig:example}.  Nodes in the abstract 
-conflict graph denotes the corresponding updates to the program 
-variables in the CFG.  Interval analysis concludes that $(\text{Error}:\bot)$, that 
-is the decision $c:=[1,1]$ leads to a {\em conflict}.  Thus, 
-the program is {\em safe} for $c:=[1,1]$.
+At this point, a clause learning propositional SAT solver learns a reason
+for the conflict and then backtracks to a level such that the learnt clause
+is \emph{unit}.  Similarly, ACDCL learns that \(n0: c:=[0,0]\), that is, all
+error traces must satisfy $(c \neq 1)$ at node~$n0$.  The analysis discards
+all interval constraints that lead to the conflict and backtracks to
+decision level~0.  ACDCL then performs interval analysis with the learnt
+constraint $(c \neq 1)$, which also leads to a conflict, as shown in
+right-hand side of Figure~\ref{ssa}.  The analysis cannot backtrack further
+and therefore it terminates, proving that the program is safe.  Thus,
+decision and clause learning are used to avoid case-based reasoning and
+prevent enumeration behaviour.
 
-A clause learning SAT solver would learn the reason for conflict at
-this point and then backtrack to a level such that the learnt clause
-is \emph{unit}.  Similar to conflict analysis phase in SAT solvers,
-ACDCL learns that (\(n0: c:=[0,0]\)), that is all error traces must
-satisfy $(c \neq 1)$ at node $n0$.  The analysis discards all interval
-constraints that lead to the conflict and backtrack to decision level
-0.  ACDCL then performs interval analysis with the learnt constraint
-$(c \neq 1)$ which also leads to a conflict, as shown in right-hand side of
-Figure~\ref{ssa}.  The analysis cannot backtrack further and therefore
-it terminates proving that the program is safe.  A DPLL-style solver
-would perform proof by cases.  However, decision and clause learning
-are used to avoid case-based reasoning and prevents enumeration
-behaviour.
-
-For $N=46000$, table~\ref{solver} shows the statistics from 
-MiniSAT~\cite{minisat} solver for analysis using BMC, and 
-analysis using ACDCL with interval and octagon domains.  
-Compared to a SAT solver, there is a significant reduction 
-in the number of decisions, propagations, learnt clauses and 
-restarts.  Compared to abstract interpretation, ACDCL does 
-not require external hints to proof the program, thus 
-automatically performing program and property driven trace 
-partitioning to generate proofs using decision and clause learning.
+Table~\ref{solver} compares statistics for the MiniSAT~\cite{minisat}
+solver for $N=46000$ for the analysis using BMC, and the analysis using
+ACDCL with the interval and octagon domains.  Compared to the propositional
+SAT solver, there is a significant reduction in the number of decisions,
+propagations, learnt clauses and restarts.  Compared to abstract
+interpretation, ACDCL does not require external hints to prove the program,
+thus automatically performing program and property-driven trace partitioning
+to generate proofs using heuristic decisions and clause learning.

### 2016-10-22 deductions.tex
deduction text

@@
 \end{equation}
-For $\subdomain=\domain$, the abstract deduction transformer behaves in the same way as the abstract transformer defined in Eq.~\ref{eq:abstrans} in Section~\ref{sec:domains}.
+For $\subdomain=\domain$, the abstract deduction transformer behaves 
+in the same way as the abstract transformer defined in Eq.~\ref{eq:abstrans} 
+in Section~\ref{sec:domains}.
 
@@
 
-The first advantage of this definition is that with the help of $\subdomain$
-we can control how we perform propagation using our abstract transformers.
+%The first advantage of this definition is that 
+\rmcmt{This parameterisation of the abstract deduction transformer with a $\subdomain$ 
+allow us to guide the propagation in {\em forward}, {\em backward} or {\em
+multi-way} direction.}
+%using our abstract transformers.
+%we can control how we perform 
 %
@@
 
-The second advantage of Eq.~(\ref{eq:at2}) is that we can compute the
+\paragraph {\textbf{Lazy closure computation}} 
+Computing closure for relational domains, such as octagons is extremely
+expensive~\cite{pldi15}.  An advantage of our formalism in 
+Eq.~(\ref{eq:at2}) is that we can compute the
 \emph{closure} operation for relational domains in a lazy manner. To

### 2016-10-22 algorithm.tex
section 5.1

@@
 \Input{A program in the form of a set of abstract transformers $\abstransset$.}
-\Output{The status \safe or \unsafe, and a counterexample if \unsafe.}
+\Output{The status \safe or \unsafe, \rmcmt{and a counterexample if \unsafe.}}
 $\trail \leftarrow \langle\rangle$ \;
@@
 Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-The input to the algorithm is a program in the form of the set of its abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$.
-The algorithm is parametrised by propagation, decision, and conflict analysis
-heuristics.  
+The input to the algorithm is a program in the form of the set of its 
+abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$.  
+The algorithm is parametrised by propagation \rmcmt{$(\propheur)$}, decision
+$(\decheur)$, and conflict analysis heuristics $(\confheur)$.  
 %\rmcmt{Approximation of the concrete transformers in 
@@
 %form of strongest-post condition or weakest pre-condition. } 
-It maintains a propagation
-trail $\trail$ and a reason trail $\reasons$.
-The propagation trail is an initially empty sequence of meet
-irreducibles inferred by the abstract model search phase (deductions
-and decisions). The reason trail maps the elements of the
-propagation trail to transformers $\abstrans\in\abstransset$ that were used to
-derive them. The \emph{abstract value} $\abs(\trail)$
-corresponding to the propagation trail $\trail$ is the
-conjunction of the meet irreducibles on the trail:
+The algorithm begins with an empty propagation trail $\trail$ and 
+a reason trail $\reasons$.
+%The propagation trail is initialized with an empty sequence.  
+The propagation trail stores all meet irreducibles inferred by 
+the abstract model search phase (deductions and decisions).  
+The reason trail maps the elements of the propagation trail to 
+transformers $\abstrans\in\abstransset$ that were used to
+derive them. 
+%
+\begin{definition} 
+The \emph{abstract value} $\abs(\trail)$ corresponding to 
+the propagation trail $\trail$ is the conjunction of the 
+meet irreducibles on the trail:
 $\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
-$\abs(\trail)=\top$ if $\trail$ is the empty
-sequence $\langle\rangle$.
-
+$\abs(\trail)=\top$ if $\trail$ is the empty sequence $\langle\rangle$.
+\end{definition}
+%
 The procedure $\deduce$ computes a greatest fixed point over the
@@
 We concatenate $\decisionvar$ to the
-trail $\trail$.  The decision $\decisionvar$ therefore refines the current
-abstract value $\abs(\trail)$ represented by the trail.
+trail $\trail$.  The decision $\decisionvar$ therefore refines \rmcmt{not
+defined} the current abstract value $\abs(\trail)$ represented by the trail. 
+\rmcmt{a decision is added to the trail only if it refines the trail.}
 % WE ExPLAIN THAT LATER
@@
 We set the corresponding entry in the reason trail $\reasons$ to $\top$
-to mark it as a decision.
+\rmcmt{to mark it as a decision.}
 % so the analysis jumps under a 
@@
 found an abstract value that represents models of $\formula$, which
-are counterexamples to the safety problem, and thus we return
+are counterexamples to the safety problem, and thus \rmcmt{we or ACDCL} return
 \textsf{unsafe}.
@@
 %Implication Point (UIP)~\cite{cdcl}.  
-ACDCL heuristically chooses one reason and learns it by adding at as
-an abstract transformer to $\abstransset$. Then it backtracks by
-removing the tail of the $\trail$ up to a point where it becomes
-consistent, and continues making deductions from there.
-If $\analyzeconflict$ returns $\false$ then $\formula$ is unsatisfiable
-and we return  \textsf{safe}.
+ACDCL heuristically chooses one reason $\conflictset$ and learns it 
+by adding it as an abstract transformer to $\abstransset$. The algorithm 
+backtracks by removing the content of $\trail$ up to a point where it does not 
+conflict with $\conflictset$.  ACDCL then performs deductions with the learnt 
+transformer.  If $\analyzeconflict$ returns $\false$ then no further
+backtracking is possible.  Thus, the $\formula$ is unsatisfiable
+and \rmcmt{we or ACDCL} return \textsf{safe}.
 

### 2016-10-22 deductions.tex
section 5.1

@@
 %-------------------------------------------------------------------------------
-
+\Omit{
 To make our algorithm efficient, we have to focus the abstract
@@
 work. 
+}
+\rmcmt{The key considerations for building an abstract transformer are 
+precision and efficiency.  An efficient transformer is less precise, 
+and vice-versa.  Our propagation strategy provides a domain-dependant 
+heuristic choice to control these variants.}
 %
@@
 %
-A subdomain contains a chosen subset of the elements in $\domain$ including $\bot$ and $\top$
+A subdomain contains a chosen subset of the elements in $\domain$ including
+$\bot$ and $\top$, \rmcmt{which forms a lattice structure}.  An abstract 
+deduction transformer is formally defined as follows. 
 \begin{equation}\label{eq:at2}
@@
 \end{equation}
-For $\subdomain=\domain$, the abstract deduction transformer behaves 
-in the same way as the abstract transformer defined in Eq.~\ref{eq:abstrans} 
+For $\subdomain=\domain$, the abstract deduction transformer is identical 
+to the abstract transformer defined in Eq.~\ref{eq:abstrans} 
 in Section~\ref{sec:domains}.
@@
 %The first advantage of this definition is that 
-\rmcmt{This parameterisation of the abstract deduction transformer with a $\subdomain$ 
-allow us to guide the propagation in {\em forward}, {\em backward} or {\em
-multi-way} direction.}
+\rmcmt{
+This parameterisation of the abstract deduction transformer with a subdomain 
+$\subdomain$ allow us to guide the propagation in {\em forward}, {\em backward} 
+or {\em multi-way} direction.
+}
 %using our abstract transformers.
@@
 %
-Note that the restriction to a subdomain makes the transformer less
-precise. Therefore, we have the property
+\Omit{
+Note that a restriction to a subdomain makes a transformer less
+precise.
+}
+Note that a restricted subdomain makes a transformer less precise, but more
+efficient.  On the contrary, an unrestricted subdomain make a transformer more 
+precise, but less efficient. Therefore, we have the property
 $\abstrans{\domain}{\constraint}^\domain(\absval)\sqsubseteq
 \abstrans{\domain}{\constraint}^\subdomain(\absval)$.
-However, the choice of a subdomain makes the deductions more efficient.
+%However, the choice of a subdomain makes the deductions more efficient.
+\rmcmt{Thus, a choice of subdomain helps to fine-tune the deduction inferred from a 
+propagation phase.}
 

### 2016-10-22 deductions.tex
sync

@@
   $\newdeductions \leftarrow \onlynew(\abstransel{\subdomain}(\abs(\trail)))$\;
+  \rmcmt{can we avoid onlynew or atleast change the name?}
   \uIf{$\newdeductions = \bot$} {
@@
 expensive~\cite{pldi15}.  An advantage of our formalism in 
-Eq.~(\ref{eq:at2}) is that we can compute the
-\emph{closure} operation for relational domains in a lazy manner. To
-this end, we construct a $\subdomain=\makesubdomain_\domain(\subvars)$
-which is sufficient to perform one step of the closure operation each
-time the abstract transformer is applied.
+Eq.~(\ref{eq:at2}) is that
+%we can compute the
+the \emph{closure} operation for relational domains can be computed 
+in a lazy manner. To this end, we construct a subdomain 
+$\subdomain=\makesubdomain_\domain(\subvars)$ which is sufficient 
+to perform one step of the closure operation each
+time the abstract \rmcmt{deduction} transformer is applied.
 %
@@
 %
+\rmcmt{Only if a abstract deduction transformer subsequently makes new deductions 
+on $x$ or $z$, then the next step of the closure will be computed through 
+the subdomain $\octagons[\{x,z\}]$.}
+\Omit{
 Hence, when the abstract deduction transformer is applied we do not
@@
 domain, which makes single deduction steps more efficient.
+}
+\rmcmt{Hence, an application of abstract deduction transformer does not 
+compute the full closure in the full domain, but compute only a 
+single step of the closure in a restricted domain, which makes each 
+deduction step more efficient.  Thus, we delay the closure operation 
+until the point where it is absolutely necessary to compute it, thereby 
+increasing the efficiency of the deductions.}
 %
+\Omit{
 If $\abstransel{\subdomain}$ deduces new information about $x$ or $z$
@@
 mechanism of Algorithm~\ref{Alg:ms} that we describe next.
+}
+%
 
@@
 trail.  Additionally, the procedure $\deduce$ is parameterised with 
-a propagation heuristic. 
+a propagation heuristic ($\propheur$). 
 %
 The algorithm maintains a  {\em worklist}, which is a queue that contains 
-abstract deduction transformers.  
-The propagation heuristics provides two 
+abstract deduction transformers.  The propagation heuristics provides two 
 functions $\initworklist$ and $\updateworklist$.
 The order of the elements in the worklist and the subdomains $\subdomain$
-attached to each of the abstract deduction transformers
-determine the propagation strategy (forward, backward, multi-way).
+attached \rmcmt{or associated} to each abstract deduction transformers 
+\rmcmt{($\abstransel{\subdomain}$)} determine the propagation strategy 
+(forward, backward, multi-way).
 %
@@
 %
-The {\em backward} iteration strategy intialises the worklist with the assertions and the subdomain is constructed from the right-hand side variables.
+The {\em backward} propagation strategy intialises the worklist 
+with the assertions and the subdomain is constructed from the 
+right-hand side variables.
 %
-The {\em multi-way} iteration strategy initialises the worklist 
-with set of all transformers and each subdomain contains the variables occurring in the respective transformers.
-
+The {\em multi-way} propagation strategy initialises the worklist 
+with set of all transformers and each \rmcmt{each or the corresponding ?} 
+subdomains contains the variables occurring in the respective transformers.
+%
 Depending on the propagation heuristics, $\updateworklist$ adds
@@
 to include the variables in $\newdeductions$.
-
+%
 The function
@@
 in order to obtain only new deductions when applying the abstract deduction transformer.
-
+%
+\Omit{
 If $\bot$ is deduced we return \textsf{conflict}.
@@
 model search procedure makes a new decision.    
+}
+\rmcmt{If $\abstransel{\subdomain}$ deduce $\bot$, then 
+the procedure $\mathit{deduce}$ returns \textsf{conflict}.
+Otherwise, when a fixed point is reached, i.e.\ 
+the worklist is empty, then the abstract value $\abs(\trail)$ is checked 
+for $\gamma$-completeness~\cite{dhk2013-popl}.  If it is indeed 
+$\gamma$-complete, then $\mathit{deduce}$ returns \sat.  Otherwise, the 
+algorithm returns \textsf{unknown} and the abstract model search 
+procedure makes a new decision.    
+}

### 2016-10-22 introduction.tex

@@
 
-Abstract Conflict Driven Clause Learning} (ACDCL)~\cite{dhk2013-popl} is one
-such generalisation of CDCL.  The partial satisfying assignments in CDCL
-algorithm are generalized to be chosen from a given abstract domain.  The
-unit rule is the best abstract transformer, Boolean Constraint Propagation
-(BCP) computes a greatest fixed-point over the abstract domain, decision is
-dual widening and clause learning can be viewed as synthesising an abstract
-transformer for negation.  This enables an ACDCL-based analyser to refine an
-imprecise analysis and thus prevents enumeration behaviour.  ACDCL has so
-far been instantiated as a decision procedure for floating-point
-arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14} and a program analyser that
-reasons using interval~\cite{tacas12}.  Hence, the existing ACDCL instances
-are limited to bounds checking in numerical programs.
+\emph{Abstract Conflict Driven Clause Learning} (ACDCL)~\cite{dhk2013-popl}
+is one such generalisation of CDCL.  The partial satisfying assignments in
+CDCL algorithm are generalized to be chosen from a given abstract domain. 
+The unit rule is the best abstract transformer, Boolean Constraint
+Propagation (BCP) computes a greatest fixed-point over the abstract domain,
+decision is dual widening and clause learning can be viewed as synthesising
+an abstract transformer for negation.  This enables an ACDCL-based analyser
+to refine an imprecise analysis and thus prevents enumeration behaviour. 
+ACDCL has so far been instantiated as a decision procedure for
+floating-point arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14} and a
+program analyser that reasons using intervals~\cite{tacas12}.  Hence, the
+existing ACDCL instances are limited to bounds checking in numerical
+programs.
 

### 2016-10-22 decision.tex
sync
decision text

@@
 %
-Hence, a decision can specify a relation between variables, for example 
+Hence, a decision can specify relation between variables, for example 
 $\{x \leq y)$.  A decision must always be consistent with respect to 
@@
 
-We call a \emph{singleton} a meet irreducible which does represent a
-valid decision.
+\rmcmt{We call a \emph{singleton} a meet irreducible which does represent a
+valid decision ??}.
 %
@@
 
-We implemented several decision heuristics namely, {\em ordered}, 
+We have implemented several decision heuristics namely, {\em ordered}, 
 {\em longest-range}, {\em random} and the 
@@
 The {\em longest-range} heuristics simply keeps track of the bounds
-$numabsval_l,numabsval_u$ of matching template rows $\numabsval_l\leq
+$\numabsval_l,\numabsval_u$ of matching template rows $\numabsval_l\leq
 \vec{c}\vec{x}\leq \numabsval_u$, picks the one with the longest range
@@
 The {\em random} decision heuristics arbitrarily picks a meet irreducible  
-for making decision. %\pscmt{uniformly?}
+for making decision. \pscmt{uniformly?}
 %

### 2016-10-22 example.tex
edits Sec 2
fixed enum

@@
 1)~{\em abstract interpretation},
-1)~{\em bounded model checking (BMC)} and
+2)~{\em bounded model checking (BMC)} and
 3)~{\em ACDCL}.
@@
 for the conflict and then backtracks to a level such that the learnt clause
-is \emph{unit}.  Similarly, ACDCL learns that \(n0: c:=[0,0]\), that is, all
-error traces must satisfy $(c \neq 1)$ at node~$n0$.  The analysis discards
-all interval constraints that lead to the conflict and backtracks to
-decision level~0.  ACDCL then performs interval analysis with the learnt
+is \emph{unit}.  Similarly, ACDCL learns that $c:=[0,0]$ at node $n0$, that
+is, all error traces must satisfy $c \neq 1$ at node~$n0$.  The analysis
+discards all interval constraints that lead to the conflict and backtracks
+to decision level~0.  ACDCL then performs interval analysis with the learnt
 constraint $(c \neq 1)$, which also leads to a conflict, as shown in

### 2016-10-22 deductions.tex
meeting changes

@@
     $\reasons[|\trail|] \leftarrow \abstransel{\subdomain}$ \;
-    $\updateworklist_{\propheur}(\newdeductions, \worklist)$ \; 
+    $\updateworklist_{\propheur}(\newdeductions, \worklist)$ \rmcmt{shouldn't
+    this be $\abstransel{\subdomain}$}\; 
   }
@@
 \rmcmt{The key considerations for building an abstract transformer are 
-precision and efficiency.  An efficient transformer is less precise, 
-and vice-versa.  Our propagation strategy provides a domain-dependant 
-heuristic choice to control these variants.}
+precision and efficiency.  An precise transformer is usually less efficient, 
+and vice-versa.  We provide a parameterised abstract transformer that allows us to 
+control precision and efficiency through propagation heuristics. }
 %
@@
 A subdomain contains a chosen subset of the elements in $\domain$ including
-$\bot$ and $\top$, \rmcmt{which forms a lattice structure}.  An abstract 
+$\bot$ and $\top$, \rmcmt{which forms a lattice}.  An abstract 
 deduction transformer is formally defined as follows. 
@@
 to the abstract transformer defined in Eq.~\ref{eq:abstrans} 
-in Section~\ref{sec:domains}.
+in Section~\ref{sec:domains}.  We denote $\abstrans{\domain}{\constraint}^\subdomain$
+as $\abstransel{\subdomain}$.
 
@@
 %However, the choice of a subdomain makes the deductions more efficient.
-\rmcmt{Thus, a choice of subdomain helps to fine-tune the deduction inferred from a 
-propagation phase.}
+\rmcmt{Thus, the choice of a subdomain helps to fine-tune the deductions 
+during the propagation phase.}
 
@@
 in a lazy manner. To this end, we construct a subdomain 
-$\subdomain=\makesubdomain_\domain(\subvars)$ which is sufficient 
-to perform one step of the closure operation each
-time the abstract \rmcmt{deduction} transformer is applied.
+$\subdomain=\makesubdomain_\domain(\subvars)$ for $\abstransel{\subdomain}$,
+which is sufficient to perform one step of the closure operation when 
+$\abstransel{\subdomain}$ is applied.
 %
@@
 }
-\rmcmt{Hence, an application of abstract deduction transformer does not 
+Hence, an application of abstract deduction transformer does not 
 compute the full closure in the full domain, but compute only a 
@@
 deduction step more efficient.  Thus, we delay the closure operation 
-until the point where it is absolutely necessary to compute it, thereby 
-increasing the efficiency of the deductions.}
+until the point where it is absolutely necessary. 
 %
@@
 functions $\initworklist$ and $\updateworklist$.
-The order of the elements in the worklist and the subdomains $\subdomain$
-attached \rmcmt{or associated} to each abstract deduction transformers 
-\rmcmt{($\abstransel{\subdomain}$)} determine the propagation strategy 
-(forward, backward, multi-way).
+The order of the elements in the worklist and the subdomain $\subdomain$ 
+associated with each abstract deduction transformer ($\abstransel{\subdomain}$)} 
+determine the propagation strategy (forward, backward, multi-way).
 %
@@
 The {\em multi-way} propagation strategy initialises the worklist 
-with set of all transformers and each \rmcmt{each or the corresponding ?} 
-subdomains contains the variables occurring in the respective transformers.
+with the set of all transformers and the corresponding subdomains 
+contains the variables occurring in the respective transformers.
 %
@@
 $\gamma$-complete~\cite{dhk2013-popl}. It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
-Otherwise, the algorithm returns \textsf{unknown} and the 
-model search procedure makes a new decision.    
+Otherwise, the algorithm returns \textsf{unknown} and ACDCL makes a new decision.    
 }
-\rmcmt{If $\abstransel{\subdomain}$ deduce $\bot$, then 
+If $\abstransel{\subdomain}$ deduce $\bot$, then 
 the procedure $\mathit{deduce}$ returns \textsf{conflict}.
@@
 for $\gamma$-completeness~\cite{dhk2013-popl}.  If it is indeed 
-$\gamma$-complete, then $\mathit{deduce}$ returns \sat.  Otherwise, the 
-algorithm returns \textsf{unknown} and the abstract model search 
-procedure makes a new decision.    
-}
+$\gamma$-complete, then $\mathit{deduce}$ returns \textsf{sat}.  Otherwise, the 
+algorithm returns \textsf{unknown} and ACDCL makes a new decision.

### 2016-10-22 decision.tex
meeting changes
some compressing

@@
 %
-Hence, a decision can specify relation between variables, for example 
+Hence, a decision can specify a relation between variables, for example 
 $\{x \leq y)$.  A decision must always be consistent with respect to 
 the trail $\trail$, i.e.\ $\abs(\trail\cdot \decisionvar)\neq \bot$.  
-A new decision always increases the decision level by one.
+A new decision increases the decision level by one.
 
@@
 (i.e.\ \emph{matching} rows) and the corresponding bounds
-$\numabsval_1=-\numabsval_1$.
+$\numabsval_1=-\numabsval_2$.
 %
@@
 involved.
-
+%
 Given 
@@
 The {\em random} decision heuristics arbitrarily picks a meet irreducible  
-for making decision. \pscmt{uniformly?}
+for making decision. 
 %

### 2016-10-22 deductions.tex
some compressing

--- 2016-10-22 deductions.tex
@@
   $\abstransel{\subdomain} \leftarrow \mathit{worklist.pop}()$ \; 
-  $\newdeductions \leftarrow \onlynew(\abstransel{\subdomain}(\abs(\trail)))$\;
-  \rmcmt{can we avoid onlynew or atleast change the name?}
-  \uIf{$\newdeductions = \bot$} {
+  $\absval \leftarrow \abstransel{\subdomain}(\abs(\trail))$\;
+  \uIf{$\absval = \bot$} {
     $\reasons[\bot] \leftarrow \abstransel{\subdomain}$ \;
@@
   {
+    $\newdeductions=\onlynew(\absval)$\;
 %    $\newdeductions=\decomp(\absval)\setminus\decomp(\abs(\trail))$\;
@@
     $\reasons[|\trail|] \leftarrow \abstransel{\subdomain}$ \;
-    $\updateworklist_{\propheur}(\newdeductions, \worklist)$ \; 
+    $\updateworklist_{\propheur}(\worklist, \newdeductions, \abstransel{\subdomain},  \abstransset)$ \; 
   }
@@
 For $\subdomain=\domain$, the abstract deduction transformer is identical 
-to the abstract transformer defined in Eq.~\ref{eq:abstrans} 
+to the abstract transformer defined in Eq.~(\ref{eq:abstrans})
 in Section~\ref{sec:domains}.
@@
 Otherwise, when eventually a fixed point has been reached, i.e.\ the worklist is empty, then the abstract value $\abs(\trail)$ is checked whether it is 
-$\gamma$-complete~\cite{dhk2013-popl}. It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
+$\gamma$-complete~\cite{dhk2013-popl}. 
+%
+\begin{definition} 
+\pscmt{reformulate} An abstract value $\absval$ is $\gamma$-complete if all concrete values in 
+$\gamma(\absval)$ satisfy $\formula$.  
+\end{definition}
+%
+It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
 Otherwise, the algorithm returns \textsf{unknown} and the 

### 2016-10-22 paper.tex
typo
some compressing

@@
 \institute{University of Oxford, UK \and University of Sussex, UK \\
-\email{\{rajdeep.mukherjee,kroening,tom.melham\}@cs.ox.ac.uk},
-\email{p.schrammel@sussex.ac.uk}}
+%\email{\{rajdeep.mukherjee,kroening,tom.melham\}@cs.ox.ac.uk},
+%\email{p.schrammel@sussex.ac.uk}
+}
 
@@
 as intervals to improve the performance of a floating-point decision
-procedure.  Whereas the propositional reasoning used in satisfiability 
+procedure.  Whereas the propositional reasoning used in SAT 
 solvers can be rather easily lifted from the Boolean lattice to other 
@@
 properties of C programs.  We~evaluate the performance of our analyser in
-comparison with program verifiers that uses propositional-level solvers and
+comparison with program verifiers that use propositional-level solvers and
 a static analysis tool.
@@
 We observe a reduction of two orders of magnitude in the number of 
-decisions and conflict analyses compared to a satisfiability 
+decisions and conflict analyses compared to a SAT 
 solver and higher precision compared to an abstract interpreter. 

### 2016-10-22 deductions.tex

--- 2016-10-22 deductions.tex
@@
     $\reasons[|\trail|] \leftarrow \abstransel{\subdomain}$ \;
-    $\updateworklist_{\propheur}(\newdeductions, \worklist)$ \rmcmt{shouldn't
-    this be $\abstransel{\subdomain}$}\; 
+    $\updateworklist_{\propheur}(\worklist, \newdeductions, \abstransel{\subdomain},  \abstransset)$ \; 
   }
@@
 For $\subdomain=\domain$, the abstract deduction transformer is identical 
-to the abstract transformer defined in Eq.~\ref{eq:abstrans} 
+to the abstract transformer defined in Eq.~(\ref{eq:abstrans})
 in Section~\ref{sec:domains}.  We denote $\abstrans{\domain}{\constraint}^\subdomain$
@@
 The order of the elements in the worklist and the subdomain $\subdomain$ 
-associated with each abstract deduction transformer ($\abstransel{\subdomain}$)} 
+associated with each abstract deduction transformer ($\abstransel{\subdomain}$) 
 determine the propagation strategy (forward, backward, multi-way).
@@
 Otherwise, when eventually a fixed point has been reached, i.e.\ the worklist is empty, then the abstract value $\abs(\trail)$ is checked whether it is 
-$\gamma$-complete~\cite{dhk2013-popl}. It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
+$\gamma$-complete~\cite{dhk2013-popl}. 
+%
+\begin{definition} 
+\pscmt{reformulate} An abstract value $\absval$ is $\gamma$-complete if all concrete values in 
+$\gamma(\absval)$ satisfy $\formula$.  
+\end{definition}
+%
+It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
 Otherwise, the algorithm returns \textsf{unknown} and ACDCL makes a new decision.    

### 2016-10-22 algorithm.tex
section 5 revised
minor
sync
some compressing
algorithm text

@@
 \Input{A program in the form of a set of abstract transformers $\abstransset$.}
-\Output{The status \safe or \unsafe, \rmcmt{and a counterexample if \unsafe.}}
+\Output{The status \safe or \unsafe.}
 $\trail \leftarrow \langle\rangle$ \;
@@
 \lIf{$\mathit{result}$ = \sat} {
-  \return (\unsafe,$\abs(\trail)$)}
+  \return (\unsafe)}
   $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
@@
 The input to the algorithm is a program in the form of the set of its 
-abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$.  
-The algorithm is parametrised by propagation \rmcmt{$(\propheur)$}, decision
+abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$. \pscmt{define} 
+The algorithm is parametrised by propagation $(\propheur)$, decision
 $(\decheur)$, and conflict analysis heuristics $(\confheur)$.  
@@
 The reason trail maps the elements of the propagation trail to 
-transformers $\abstrans\in\abstransset$ that were used to
+transformers $\abstransel{}\in\abstransset$ that were used to
 derive them. 
@@
 defined} the current abstract value $\abs(\trail)$ represented by the trail. 
-\rmcmt{a decision is added to the trail only if it refines the trail.}
+%\rmcmt{a decision is added to the trail only if it refines the trail.}
 % WE ExPLAIN THAT LATER
@@
 We set the corresponding entry in the reason trail $\reasons$ to $\top$ 
-\rmcmt{to mark it as a decision.}
+to mark it as a decision.
 % so the analysis jumps under a 
@@
 found an abstract value that represents models of $\formula$, which
-are counterexamples to the safety problem, and thus \rmcmt{we or ACDCL} return
+are counterexamples to the safety problem, and thus ACDCL return
 \textsf{unsafe}.
@@
 backtracking is possible.  Thus, the $\formula$ is unsatisfiable
-and \rmcmt{we or ACDCL} return \textsf{safe}.
+and ACDCL return \textsf{safe}.
 

### 2016-10-22 algorithm.tex
fixes in sec4

@@
 \Input{A program in the form of a set of abstract transformers $\abstransset$.}
-\Output{The status \safe or \unsafe.}
+\Output{The status \safe or \unsafe. %and a counterexample if \unsafe.
+}
 $\trail \leftarrow \langle\rangle$ \;
@@
 \lIf{$\mathit{result}$ = \sat} {
-  \return (\unsafe)}
+  \return \unsafe}
   $\decisionvar \leftarrow \decide_{\decheur}(\abs(\trail))$ \;
@@
 The input to the algorithm is a program in the form of the set of its
-abstract transformers $\abstransset$ w.r.t.\ an abstract domain~$\domain$. \pscmt{define} 
-The algorithm is parametrised by propagation $(\propheur)$, decision
-$(\decheur)$, and conflict analysis heuristics $(\confheur)$.  
+abstract transformers
+$\abstransset=\{\abstrans{\domain}{\sigma}|\sigma\in\Sigma\}$
+w.r.t.\ an abstract domain~$\domain$.  The algorithm is parametrised
+by propagation $(\propheur)$, decision $(\decheur)$, and conflict
+analysis heuristics $(\confheur)$.
 %\rmcmt{Approximation of the concrete transformers in 
@@
 We concatenate $\decisionvar$ to the
-trail $\trail$.  The decision $\decisionvar$ therefore refines \rmcmt{not
-defined} the current abstract value $\abs(\trail)$ represented by the trail. 
-%\rmcmt{a decision is added to the trail only if it refines the trail.}
+trail $\trail$.  The decision $\decisionvar$ refines  the current abstract value $\abs(\trail)$ represented by the trail, i.e.\ $\abs(\trail\cdot\decisionvar)\sqsubseteq \abs(\trail)$.
 % WE ExPLAIN THAT LATER

### 2016-10-22 deductions.tex

@@
 }
-\rmcmt{The key considerations for building an abstract transformer are 
+The key considerations for building an abstract transformer are 
 precision and efficiency.  An precise transformer is usually less efficient, 
 and vice-versa.  We provide a parameterised abstract transformer that allows us to 
-control precision and efficiency through propagation heuristics. }
+control precision and efficiency through propagation heuristics. 
 %
@@
 A subdomain contains a chosen subset of the elements in $\domain$ including
-$\bot$ and $\top$, \rmcmt{which forms a lattice}.  An abstract 
+$\bot$ and $\top$, which forms a lattice.  An abstract 
 deduction transformer is formally defined as follows. 
@@
 %The first advantage of this definition is that 
-\rmcmt{
 This parameterisation of the abstract deduction transformer with a subdomain 
@@
 or {\em multi-way} direction.
-}
 %using our abstract transformers.
@@
 %However, the choice of a subdomain makes the deductions more efficient.
-\rmcmt{Thus, the choice of a subdomain helps to fine-tune the deductions 
-during the propagation phase.}
+Thus, the choice of a subdomain helps to fine-tune the deductions 
+during the propagation phase.
 
@@
 %
-\rmcmt{Only if a abstract deduction transformer subsequently makes new deductions 
+Only if a abstract deduction transformer subsequently makes new deductions 
 on $x$ or $z$, then the next step of the closure will be computed through 
-the subdomain $\octagons[\{x,z\}]$.}
+the subdomain $\octagons[\{x,z\}]$.
 \Omit{

### 2016-10-22 domains.tex
meet irreducible
fixes sec5

@@
 techniques~\cite{sriram,GS07b,BJKS15} for computing the domain
-operations, such as \rmcmt{meet($\meet$), join($\sqcup$)}, etc. In our implementation we
+operations, such as meet ($\meet$), join ($\sqcup$), etc. In our implementation we
 use the strategy iteration approach available in 2LS~\cite{BJKS15}.

### 2016-10-22 deductions.tex

@@
 }
-\lIf{$\abs(\trail)$ is $\gamma$-complete} {
+\lIf{$\abstransset$ is $\gamma$-complete at $\abs(\trail)$} {
   \return \sat;
@@
 %
-\begin{definition} 
-\pscmt{reformulate} An abstract value $\absval$ is $\gamma$-complete if all concrete values in 
-$\gamma(\absval)$ satisfy $\formula$.  
-\end{definition}
-%
 It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
@@
 Otherwise, when a fixed point is reached, i.e.\ 
-the worklist is empty, then the abstract value $\abs(\trail)$ is checked 
-for $\gamma$-completeness~\cite{dhk2013-popl}.  If it is indeed 
+the worklist is empty, then we check whether
+the abstract transformers $\abstransset$ are $\gamma$-complete~\cite{dhk2013-popl} for the current abstract value $\abs(\trail)$.
+%
+Intuitively, this means that all concrete values in 
+$\gamma(\absval)$ satisfy $\formula$.  
+%
+If it is indeed 
 $\gamma$-complete, then $\mathit{deduce}$ returns \textsf{sat}.  Otherwise, the 

### 2016-10-22 decision.tex
singletons

@@
 
-\rmcmt{We call a \emph{singleton} a meet irreducible which does represent a
-valid decision ??}.
+We call a meet irreducible that does not represent a valid decision a
+\emph{singleton} meet irreducible.  This is similar to 
+a variable in a SAT solver that is already assigned $\true$ or $\false$,
+and thus we cannot make a decision.
 %
@@
 %
-We cannot make decisions for template rows of singletons because there
-is no choice left. Similarly, in a SAT solver, we cannot make a
-decision on a variable that is already assigned $\true$ or $\false$.
+%We cannot make decisions for template rows of singletons because there
+%is no choice left. 
 %

### 2016-10-22 algorithm.tex
algorithm 1
a few bad tricks to make it fit

@@
 }
-$\trail \leftarrow \langle\rangle$ \;
-$\reasons \leftarrow []$ \;
+$\trail \leftarrow \langle\rangle$, $\reasons \leftarrow []$ \;
 $\mathit{result} \leftarrow \deduce_{\propheur}(\abstransset,\trail,\reasons)$ \;
@@
 %\subsection{Algorithm for Abstract Conflict Driven Clause Learning}
-Algorithm~\ref{Alg:acdcl} presents an overview of the ACDCL algorithm.
-The input to the algorithm is a program in the form of the set of its
-abstract transformers
+% presents an overview of the ACDCL algorithm.
+The input to the ACDCL algorithm (Algorithm~\ref{Alg:acdcl}) is a
+program in the form of the set of its abstract transformers
 $\abstransset=\{\abstrans{\domain}{\sigma}|\sigma\in\Sigma\}$
@@
 %form of strongest-post condition or weakest pre-condition. } 
-The algorithm begins with an empty propagation trail $\trail$ and 
+The algorithm maintains a propagation trail $\trail$ and 
 a reason trail $\reasons$.
@@
 %
-The procedure $\deduce$ computes a greatest fixed point over the
+The algorithm begins with an empty $\trail$, a empty $\reasons$ and an
+abstract value $\top$.  The procedure $\deduce$ (detailed in
+Section~\ref{sec:deduce}) computes a greatest fixed point over the
 transformers in $\abstransset$ that refines the abstract value,
@@
 \textsf{safe}.  Else, the analysis enters into the while($\true$) loop
-and makes a new decision by a call to $\decide$ which returns a new
-meet irreducible $\decisionvar$.
+and makes a new decision by a call to $\decide$ (see
+Section~\ref{sec:decide}) which returns a new meet irreducible
+$\decisionvar$.
 %
@@
 %{\em dual widening} \pscmt{what do we learn?}.  
-
+%
 The procedure $\deduce$ is called next to infer new meet irreducibles
@@
 In the latter case the algorithm enters in the $\analyzeconflict$ 
-phase to learn the reason for the conflict.   There can be multiple
+phase (see Section~\ref{sec:conflict}) to learn the reason for the conflict.   There can be multiple
 incomparable reasons for conflict.

### 2016-10-22 domains.tex
sync
a few bad tricks to make it fit

@@
 %-------------------------------------------------------------------------------
-\paragraph{Template Polyhedra Abstract Domain}
+\paragraph{\textbf{Template Polyhedra Abstract Domain}}
 %
@@
 \begin{wrapfigure}{r}{4.5cm}
-\vspace*{-3ex}
+\vspace*{-7ex}
 \scalebox{.65}{\import{figures/}{octagons.pspdftex}}
 \caption{Example of an octagon}\label{octagon}
-\vspace*{-3ex}
+\vspace*{-5ex}
 \end{wrapfigure} 
@@
 %-------------------------------------------------------------------------------
-\paragraph{Abstract Transformers}
+\paragraph{\textbf{Abstract Transformers}}
 %
@@
 complements, but they can be represented as intersections of
-complementable half-spaces, each of which has a complement. 
+half-spaces, each of which has a complement. 
 %
@@
 \end{equation}
-%\rmcmt{mention minimum complementable element}
+Meet irreducibles in ACDCL correspond to the concept of literals in a SAT solver.
 \end{definition}

### 2016-10-23 background.tex

@@
-\paragraph{Related Work}
+\section{Related Work}
 The work of~\cite{sas13,DBLP:journals/fmsd/BrainDGHK14} present an
@@
 \cite{DBLP:conf/esop/MineBR16} proposes an algorithm inspired by constraint
-solvers for inferring disjunctive invariants using the interval domain.
+solvers for inferring disjunctive invariants using the intervals.
 %

### 2016-10-23 example.tex

@@
 domain (bottom). The analysis associates an abstract element with each
-control location and variable.  The octagons deduced by ACDCL are sufficient
+control location.  The octagons deduced by ACDCL are sufficient
 to prove safety without any decision or clause learning, as in the case of
-Ast{\'e}e.
+Astr{\'e}e.
 
@@
 {\em safe} for the case $c:=[1,1]$.
-
+%
 At this point, a clause learning propositional SAT solver learns a reason

### 2016-10-23 introduction.tex

@@
 richer abstract domains or equipping the analysis with more precise set
-unions -- which is often very expensive since the analysis exhibits case
+unions -- which is often expensive since the analysis exhibits case
 enumeration behaviour.

### 2016-10-23 learning.tex
formatting of aunit

@@
-\section{Abstract Conflict Analysis}
+\section{Abstract Conflict Analysis}\label{sec:conflict}
 Conflict analysis in a SAT solver can be seen as abductive 
@@
 \begin{figure}[t]
+\vspace*{-10ex} %BAD TRICK
 %\caption{An example of Octagon}\label{octagon}
@@
 \paragraph {\textbf{Lifting UIP to relational domains}}
-An UIP is a special node in the abstract conflict graph such that 
+An UIP is a node in the abstract conflict graph such that 
 any path from the last decision node to the conflict node must pass 
-through it.  Unlike a SAT solver, a UIP for relational domain may 
+through it.  Unlike a SAT solver, a UIP for a relational domain may 
 involve meet irreducibles that contain the same variables.  The 
-first-UIP is a unique node closest to the conflict, and the last UIP 
+first UIP is a unique node closest to the conflict, and the last UIP 
 is the decision node itself.  Computing UIPs ensures asserting
@@
 \[ \aunit(\absval) =
- \begin{cases}
-  \bot       & \quad \text{if } \absval \sqsubseteq \bigsqcap \conflictset \qquad \qquad \qquad
-  \qquad \qquad \qquad \quad (1)\\
-  \bar{t}    & \quad \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset
-  \setminus \{t\}. \absval  \sqsubseteq t' \qquad \quad \; \; (2) \\
-  \bar{t}    & \quad \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset \setminus \{t\}. \absval
+ \left\{\begin{array}{l@{\quad}l@{\qquad}l}
+  \bot       & \text{if } \absval \sqsubseteq \bigsqcap \conflictset & (1)\\
+  \bar{t}    & \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset
+  \setminus \{t\}. \absval  \sqsubseteq t' & (2) \\
+  \bar{t}    & \text{if } t \in \conflictset \; \text{and} \; \forall t' \in \conflictset \setminus \{t\}. \absval
   \not\sqsubseteq t', \; \text{then} \\ 
              & \quad \forall \absval' \sqsubseteq \absval: \absval'
-             \not\sqsubseteq t \text{ and}\\ 
+             \not\sqsubseteq t \text{ and} \\ 
              & \quad \absval''=\exclude(\absval,\absval') \text{ and
-             }\absval''\sqsubseteq t' \qquad  \; \; (3) \\
-  \top & \quad \text{otherwise} \qquad \qquad \qquad \qquad \qquad \qquad \quad (4) \\
- \end{cases}
+             }\absval''\sqsubseteq t' & (3) \\
+  \top & \text{otherwise} & (4) \\
+ \end{array}\right.
 \]
@@
 $\absval' \sqsubseteq (x \geq 2)$ and $\absval' \sqsubseteq (x \leq
-5)$.  However, if none of the above rule holds true, $\aunit$ returns~$\top$.
+5)$.  However, if none of the above rules hold true, $\aunit$ returns~$\top$.
 

### 2016-10-23 example.tex
typo

@@
 BMC converts the program into a bit-vector equation and passes that to a
-CDCL-based SAT solver.  We will contast the analysis performed by CDCL with
+CDCL-based SAT solver.  We will contrast the analysis performed by CDCL with
 that of ACDCL.  The right-hand side of Figure~\ref{fig:example} illustrates
@@
 
-Table~\ref{solver} compares statistics for the MiniSAT~\cite{minisat}
+Table~\ref{solver} compares statistics for the MiniSat~\cite{minisat}
 solver for $N=46000$ for the analysis using BMC, and the analysis using

### 2016-10-23 deductions.tex
typo

@@
-\section{Abstract Model Search}
+\section{Abstract Model Search}\label{sec:deduce}
 %
@@
 determine the propagation strategy (forward, backward, multi-way).
-%
+
 The {\em forward} propagation strategy initialises the worklist with
@@
 %
-The {\em backward} propagation strategy intialises the worklist 
+The {\em backward} propagation strategy initialises the worklist 
 with the assertions and the subdomain is constructed from the 
@@
 It is $\gamma$-complete if all concrete values in $\gamma(\abs(\trail))$ satisfy $\formula$.
-Otherwise, the algorithm returns \textsf{unknown} and ACDCL makes a new decision.    
-}
-If $\abstransel{\subdomain}$ deduce $\bot$, then 
+Otherwise, the algorithm returns \textsf{unknown} and ACDCL makes a new decision.}
+
+If $\abstransel{\subdomain}$ deduces $\bot$, then 
 the procedure $\mathit{deduce}$ returns \textsf{conflict}.

### 2016-10-23 domains.tex
Spelling

@@
 %
-The meet irreducibles in the boolean domain $\booldomain$ for a variable $x$ are 
+The meet irreducibles in the Boolean domain $\booldomain$ for a variable $x$ are 
 $x$ and $\neg x$. The meet 
@@
 The precise complementation property of partial assignment lattice can
-be generalized to other lattice structures.  Most numerical abstract
+be generalised to other lattice structures.  Most numerical abstract
 domains, such as intervals, octagons, polyhedra, can be decomposed

### 2016-10-23 experiments.tex
deduction text
Spelling

@@
 (octagons, polyhedra, zones, equalities, filter).  ACDCL is instantiated
-with a product domain of booleans and template polyhedra.  We use the
+with a product domain of Booleans and template polyhedra.  We use the
 interval and octagon templates for our experiments.  ACDCL is configured
@@
 Figure~\ref{fig:results}(b) show that the number of propagations in the SAT
-solver is maximal for the benchmarks that exhibit relational behavior. 
+solver is maximal for the benchmarks that exhibit relational behaviour. 
 Cumulative statistics corresponding to each phase of the algorithm for all
@@
 least two orders of magnitude in the number of decisions, propagations and
-conflicts compared to analyis using CBMC.
+conflicts compared to analysis using CBMC.
 
@@
 \label{result}
+\vspace*{-3ex} %BAD TRICK
 \end{table}
 
-Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16
+Out of 66 benchmarks, CBMC with MiniSat backend could only prove 16
 benchmarks without any restarts.  The solver is restarted in 50 cases to
@@
 %
-Figure~\ref{prop-dec}(b) presents a visualization of the performance of
+Figure~\ref{prop-dec}(b) presents a visualisation of the performance of
 different decision heuristics in ACDCL.  Note that the runtimes for all
-decision heuristics are obtained with multi-way propagtion strategy.  The
+decision heuristics are obtained with multi-way propagation strategy.  The
 runtimes are very close, but we observe some key characteristics of these
@@
 that encountered the maximum number of conflicts to prove safety, 
-thus allowing the heuristics to choose the decison variable among the set of learnt clauses.   
+thus allowing the heuristics to choose the decision variable among the set of learnt clauses.   
 }
@@
 Learning has a significant influence on the runtime of ACDCL.  We compare
-the UIP-based learning technique with an analysis that performs
-chronoligical backtracking without learning.  The effect of UIP computation
-allows ACDCL to backtrack non-chronologically and guide the model search
-with a learnt transformer.  However, the DPLL-style analysis exhibits
-case-enumeration behavior and could not finish within the time bound for
-20\% of our benchmarks.
+the UIP-based learning technique with an analysis that performs DPLL-style 
+analysis.
+%chronoligical backtracking without learning.  
+The effect of UIP computation allows ACDCL to backtrack non-chronologically 
+and guide the model search with a learnt transformer.  However, the DPLL-style 
+analysis exhibits case-enumeration behaviour and could not finish within the 
+time bound for 20\% of our benchmarks.

### 2016-10-23 paper.tex
properly do authors
nudge abstract

@@
-\documentclass[a4paper,conference]{llncs}
+\documentclass[a4paper]{llncs}
 
@@
 \usepackage{dingbat}
+\usepackage{microtype}
 \usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}
@@
 
-\author{Rajdeep Mukherjee\inst{1}\hskip 1em  Peter Schrammel\inst{2}\hskip 1em
-Daniel Kroening\inst{1}\hskip 1em Tom Melham\inst{1}}
+\author{Rajdeep Mukherjee\inst{1} \and Peter Schrammel\inst{2} \and
+Daniel Kroening\inst{1} \and Tom Melham\inst{1}}
 
@@
 as intervals to improve the performance of a floating-point decision
-procedure.  Whereas the propositional reasoning used in SAT 
-solvers can be rather easily lifted from the Boolean lattice to other 
+procedure.  Whereas the propositional reasoning in SAT 
+solvers can be lifted easily from the Boolean lattice to other 
 non-relational domains, application to relational domains is 
-challenging because of relational properties of deductions.
+challenging because of the relational properties of deductions.
 %

### 2016-10-23 introduction.tex
missing article
cannot

@@
 is one such generalisation of CDCL.  The partial satisfying assignments in
-CDCL algorithm are generalized to be chosen from a given abstract domain. 
+the CDCL algorithm are generalized to be chosen from a given abstract domain. 
 The unit rule is the best abstract transformer, Boolean Constraint
@@
 domains like octagons or equalities or polyhedra.  These domains capture
-interesting properties of programs that intervals can not express.  In this
+interesting properties of programs that intervals cannot express.  In this
 paper, we identify specific modifications to the CDCL algorithm that are

### 2016-10-23 example.tex
wording

@@
 A standard forward interval analysis is too imprecise to verify the safety
-of the above program.  The imprecision is owed to the control-flow join at
+of the program.  The imprecision is owed to the control-flow join at
 node~$n4$.  A state-of-the-art abstract interpretation tool,
 Astr{\'e}e~\cite{se2011}, requires external hints, provided by manually
-annotating the code with partition directives at $n1$, to prove safety.  The
-partition directive tells the tool to analyse each program execution path
-separately.  To determine safety, Astr{\'e}e requires a product of a trace
-partitioning domain and an interval domain or a relational domain such as
-octagons.
+annotating the code with partition directives at $n1$, to prove safety using
+intervals.  The partition directive tells the tool to analyse each program
+execution path separately.  Alternatively, Astr{\'e}e can show safety
+using a relational domain such as octagons.
 %

### 2016-10-23 domains.tex
wording

@@
 domain operations.  An example of a closure computation for octagonal 
-inequalities is shown below.
+inequalities is given below.
+%
 \[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
+%
 The closure computes all implied domain constraints.  For octagons,

### 2016-10-23 domains.tex
wording

@@
 %
-Figure~\ref{ssa} shows the SSA constraints for the program in
+Figure~\ref{ssa} gives the SSA constraints for the program in
 Figure~\ref{fig:example}.
@@
 sums and differences for each pair of variables. Figure~\ref{octagon}
-shows an example octagon and its associated inequalities.
+gives an example of an octagon and its associated inequalities.
 
@@
 An abstract transformer $\abstrans{\domain}{\constraint}$ transforms an abstract value $\absval$ through a
-constraint $\constraint$, i.e.\ it \emph{deduces} information from $\absval$ and $\constraint$.
+constraint $\constraint$, i.e., it \emph{deduces} information from $\absval$ and $\constraint$.
 The best abstract transformer is 
@@
 %$\forall \vec{y}\in c: (s[\vec{y}/\vec{\numvar}]
-%\equiv \true)$, i.e.\ the constraint $s$ is valid when evaluated over
+%\equiv \true)$, i.e., the constraint $s$ is valid when evaluated over
 %all values $\vec{y}$ of its variables $\vec{\numvar}$ in the concretisation
@@
 irreducibles in the template polyhedra domain are all elements 
-that concretise to half-spaces, i.e.\ they can be represented 
+that concretise to half-spaces, i.e., they can be represented 
 by a single inequality. For the interval domain, these are 
@@
 iff $\exists \bar{\absval} \in \domain: \neg\gamma(\bar{\absval})=\gamma(\absval)$, 
-i.e.\ there is an element whose negated concretisation equals
+i.e., there is an element whose negated concretisation equals
 the concretisation of $\absval$.

### 2016-10-23 domains.tex
line break

@@
 %
-We denote $\vars$ the set of variables occurring in the $\constraints$.
+We~denote $\vars$ the set of variables occurring in the $\constraints$.
 %
@@
 %A detailed description of abstract transformer is presented in Section~\ref{sec:abst}.
-We denote the set of abstract transformers for a safety equation
+We~denote the set of abstract transformers for a safety equation
 $\formula$ using abstract domain $\domain$ as

### 2016-10-23 introduction.tex
intro and abstract

--- 2016-10-23 introduction.tex
@@
 Static program analysis based on abstract interpretation~\cite{CC77} has
-been widely used to verify of properties of safety-critical systems. 
+been widely used to verify of properties of safety-critical systems~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}.
 Standard static analyses~\cite{se2011} aim to compute program invariants as
@@
 alarms.  Standard means to address false alarms in static analysis are
-richer abstract domains or equipping the analysis with more precise set
+the use of richer abstract domains or equipping the analysis with more precise set
 unions -- which is often expensive since the analysis exhibits case
@@
 performance of modern solvers is credited to well-tuned decision heuristics
-and clause learning algorithms in modern solvers.  Collectively, these
+and clause learning algorithms.  Collectively, these
 algorithms are referred to as \emph{Conflict Driven Clause Learning}

### 2016-10-23 paper.tex
double package
intro and abstract

--- 2016-10-22 paper.tex
@@
 In this paper, we present the non-trivial liftings of the model search and
-conflict analysis algorithms for relational abstract domains.  We have
+conflict analysis algorithms in SAT solver for relational abstract domains.  We have
 implemented these algorithms inside the 2LS verification tool and
@@
 comparison with program verifiers that use propositional-level solvers and
-a static analysis tool.
+a static analyser based on abstract interpretation.
 %

### 2016-10-23 algorithm.tex
commas
line break

@@
 %
-We concatenate $\decisionvar$ to the
-trail $\trail$.  The decision $\decisionvar$ refines  the current abstract value $\abs(\trail)$ represented by the trail, i.e.\ $\abs(\trail\cdot\decisionvar)\sqsubseteq \abs(\trail)$.
+We concatenate $\decisionvar$ to the trail~$\trail$.  The decision
+$\decisionvar$ refines the current abstract value $\abs(\trail)$ represented
+by the trail, i.e., $\abs(\trail\cdot\decisionvar)\sqsubseteq \abs(\trail)$.
+%
 % WE ExPLAIN THAT LATER
 %However, $\decisionvar$ must be consistent with the state of the
-%solver, i.e. $\abs(\trail\cdot \decisionvar)\neq \bot$.
+%solver, i.e., $\abs(\trail\cdot \decisionvar)\neq \bot$.
 %
@@
 conflict with $\conflictset$.  ACDCL then performs deductions with the learnt 
-transformer.  If $\analyzeconflict$ returns $\false$ then no further
+transformer.  If $\analyzeconflict$ returns $\false$, then no further
 backtracking is possible.  Thus, the $\formula$ is unsatisfiable

### 2016-10-23 contributions.tex

@@
 \paragraph{Contributions}
-In this paper, we make the following contributions.
+In this paper, we make the following technical contributions.
+%
 \begin{compactenum}
@@
 %generalises the Boolean Constraint Propagation (BCP) procedure 
-to compute a fixed point over a relational domain.
-
-\item We propose domain-specific decision heuristics that 
-exploit the expressivity of the underlying abstract domain to make 
-effective decisions.
+to compute a fixed point over a relational domain.  We propose 
+domain-specific decision heuristics that exploit the expressivity 
+of the underlying abstract domain to make effective decisions.
 

### 2016-10-23 experiments.tex

@@
 runtimes are very close, but we observe some key characteristics of these
-heuristics.  Berkmin heuristic performs consistently well for most safe
-benchmarks and all bit-vector category benchmarks.  Whereas, the ordered
+heuristics.  The Berkmin heuristic performs consistently well for most safe
+benchmarks and all bit-vector category benchmarks.  By contrast, the ordered
 heuristic performs better for programs with conditional branches since it
 prioritises decisions on meet irreducibles that appear in conditional
-branches over meet irreducibles that involve numerical variables.  Whereas,
-the runtimes for random heuristic is marginally higher than the other two. 
-This suggest that domain-specific decision heuristics are important for
-ACDCL.
+branches over meet irreducibles that involve numerical variables.  The
+runtimes for the random heuristic are marginally higher than for the other
+two.  This suggest that domain-specific decision heuristics are important
+for ACDCL.
 

### 2016-10-23 learning.tex

@@
 }
-For our example, there exist multiple incomaparable reasons for conflict,
+For our example, there exist multiple incomparable reasons for conflict,
 marked as {\em cut0, cut1}, in Figure~\ref{conflict}.  Here, cut0 is the first UIP.  
@@
 can be multiple incomparable reasons for a conflict. But conflict analysis
-procedure choses one that is asserting.
+procedure chooses one that is asserting.
 

### 2016-10-23 paper.tex

--- 2016-10-23 paper.tex
@@
 \usepackage{dingbat}
-\usepackage{microtype}
 \usepackage[algo2e,linesnumbered,ruled,lined]{algorithm2e}

### 2016-10-23 conclusion.tex
proof-reading all section

--- 2016-10-22 conclusion.tex
@@
 %
-Our decision heuristics takes into account the high-level structure of
+Our decision heuristics take into account the high-level structure of
 programs and the underlying expressivity of the abstract domains.  Our
@@
 %
-Experimental evaluation shows that ACDCL is able to suitably exploit
-the expressivity of richer abstract domains as well as precision of
+Experimental evaluation shows that ACDCL is able to suitably exploit the
+expressivity of richer abstract domains as well as the precision of the 
 CDCL architecture.

### 2016-10-23 decision.tex
proof-reading all section

@@
-\subsection{Decision}
+\subsection{Decision}\label{sec:decide}
 A decision $\decisionvar$ is a meet irreducible that refines the
@@
 Hence, a decision can specify a relation between variables, for example 
-$\{x \leq y)$.  A decision must always be consistent with respect to 
+$(x \leq y)$.  A decision must always be consistent with respect to 
 the trail $\trail$, i.e.\ $\abs(\trail\cdot \decisionvar)\neq \bot$.  
@@
 A singleton in the interval domain corresponds to a singleton interval
-such as $x\leq 1 \wedge -x\leq -1$ For the octagon domain, $1 \leq x-y
+such as $x\leq 1 \wedge -x\leq -1$.  For the octagon domain, $1 \leq x-y
 \leq 1$ is a singleton.
@@
 
-We have implemented several decision heuristics namely, {\em ordered}, 
+We have implemented several decision heuristics, namely, {\em ordered}, 
 {\em longest-range}, {\em random} and the 
@@
 The ordered heuristic is similar to trace partitioning~\cite{toplas07}.
-
+%
 The {\em longest-range} heuristics simply keeps track of the bounds
@@
 %
-The {\em berkmin} decision heuristics is inspired 
-from decision heuristic used in Berkmin~\cite{eugoldberg07} SAT solver.  
-The Berkmin heuristics %is currently implemented for interval constraints only.  
+The {\em berkmin} decision heuristic is inspired by the 
+decision heuristic used in the Berkmin~\cite{eugoldberg07} SAT solver.  
+The Berkmin heuristic %is currently implemented for interval constraints only.  
 %The heuristic 
@@
 meet irreducibles that participate in conflict clauses. 
-Based on the most active meet irreducible, ranges a split similar to the {\em longest-range} heuristics.
+Based on the most active meet irreducible, ranges are split 
+similar to the {\em longest-range} heuristics.
 \Omit {

### 2016-10-23 deductions.tex
proof-reading all section

@@
 \lIf{$\abstransset$ is $\gamma$-complete at $\abs(\trail)$} {
-  \return \sat;
+  \return \sat \;
 }
@@
 deductions w.r.t.\ a given subdomain $\subdomain\subseteq \domain$,
-which we call \emph{abstract deduction transformer}.
+which we call the \emph{abstract deduction transformer}.
 %
@@
 The {\em forward} propagation strategy initialises the worklist with
-abstract deduction transformers that contains constants in the
+abstract deduction transformers that contain constants in the
 right-hand side and the subdomain is constructed via $\makesubdomain$
@@
 abstract deduction transformers to the worklist that contain variables
-that appear in $\newdeductions$, and it updates the subdomains of
+that appear in $\newdeductions$, and updates the subdomains of
 the abstract deduction transformers in the worklist 
@@
 the procedure $\mathit{deduce}$ returns \textsf{conflict}.
-Otherwise, when a fixed point is reached, i.e.\ 
+Otherwise, when a fixed-point is reached, i.e.\ 
 the worklist is empty, then we check whether

### 2016-10-23 domains.tex
proof-reading all section

@@
 %
-We~denote $\vars$ the set of variables occurring in the $\constraints$.
+We~denote $\vars$ the set of variables occurring in $\constraints$.
 %
@@
 for a program with $N=|\numvars|$ variables, the matrix $\mat{C}$ 
-has $2N$ rows; and hence it generates at most $2N$ inequalities, one
+has $2N$ rows. Hence, it generates at most $2N$ inequalities, one
 for the upper and lower bounds of each variable.
@@
 elements that help to guide the model search away from the
-conflicting region of the search space.  Complementation operator 
+conflicting region of the search space.  The complementation operator 
 in abstract domains is different from the notion of precise complements. 
@@
 %
-The precise complementation property of partial assignment lattice can
+The precise complementation property of a partial assignment lattice can
 be generalised to other lattice structures.  Most numerical abstract
@@
 Standard abstract interpretation does not require a complementation
-operator. Hence, abstract domain libraries, such as APRON~\cite{apron}
+operator. Hence, abstract domain libraries, such as APRON~\cite{apron},
 do not provide it.  However, it can be implemented with the help of a

### 2016-10-23 example.tex
proof-reading all section

@@
 \hline
-Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
+Solver & Decisions & propagations & conflicts & conflict literals & restarts \\ \hline
 MiniSat & 233 & 36436 & 162 & 2604 & 2 \\ \hline
@@
 determines that the interval of $x$ is then constrained to $[0,5]$.  The
-deductions made during fixed point iteration are represented by an abstract
+deductions made during fixed-point iteration are represented by an abstract
 conflict graph, given in right-hand side of Figure~\ref{fig:example}.  Nodes
@@
 to decision level~0.  ACDCL then performs interval analysis with the learnt
-constraint $(c \neq 1)$, which also leads to a conflict, as shown in
+constraint $(c \neq 1)$, which also leads to a conflict, as shown on 
 right-hand side of Figure~\ref{ssa}.  The analysis cannot backtrack further
-and therefore it terminates, proving that the program is safe.  Thus,
+and therefore terminates, proving that the program is safe.  Thus,
 decision and clause learning are used to avoid case-based reasoning and

### 2016-10-23 experiments.tex
proof-reading all section

@@
 loop analysis.  All bounded loops are completely unrolled before analysis. 
-We compare our tool with a state-of-the-art SAT-based bounded model checker
+We compare our tool with the state-of-the-art SAT-based bounded model checker
 CBMC (version 5.5)~\cite{cbmc} and a commercial static analysis tool, 
@@
 Cumulative statistics corresponding to each phase of the algorithm for all
-66 benchmarks is given in Table~\ref{result}.  We observe a reduction of at
+66 benchmarks are given in Table~\ref{result}.  We observe a reduction of at
 least two orders of magnitude in the number of decisions, propagations and
@@
 \hline
-Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
+Solver & Decisions & propagations & conflicts & conflict literals & restarts \\ \hline
 MiniSat & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
@@
 learning mechanisms aided by the richer abstract domains.  On the other
-hand, Ast{\'e}e is often faster than ACDCL, but the analysis using
+hand, Astr{\'e}e is often faster than ACDCL, but the analysis using
 Astr{\'e}e shows a high degree of imprecision.  Astr{\'e}e reported 15 false
@@
 %
-Figure~\ref{prop-dec}(a) presents a comparison between {\em forward}
-propagation strategy and {\em multi-way} propagation strategy in ACDCL.  The
+Figure~\ref{prop-dec}(a) presents a comparison between the {\em forward}
+propagation strategy and the {\em multi-way} propagation strategy in ACDCL.  The
 choice of the propagation strategy influences the total number of decisions
 and clause learning iterations.  Hence, the propagation strategy has a
-significant influence on the run time which can be visualised in
+significant influence on the runtime which can be visualised in
 Figure~\ref{prop-dec}(a).  Compared to forward propagation, the multi-way
@@
 runtimes for the random heuristic are marginally higher than for the other
-two.  This suggest that domain-specific decision heuristics are important
+two.  This suggests that domain-specific decision heuristics are important
 for ACDCL.

### 2016-10-23 introduction.tex
proof-reading all section

@@
 Static program analysis based on abstract interpretation~\cite{CC77} has
-been widely used to verify of properties of safety-critical systems~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}.
+been widely used to verify properties of safety-critical systems~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}.
 Standard static analyses~\cite{se2011} aim to compute program invariants as

### 2016-10-23 learning.tex
proof-reading all section

@@
 with $\exclude(\absval,\absval')=\bigsqcap(\decomp(\absval) \setminus
-\decomp(\absval'\}))$.  $\aunit$ returns $\bot$ if $\conflictset$ is
+\decomp(\absval'))$.  $\aunit$ returns $\bot$ if $\conflictset$ is
 conflicting following rule (1).  Rule (2) and (3) of $\aunit$ infer a
@@
 decomposition of $\absval$ into $\absval'$ and $\absval''$ where
-$\wedge'= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15)
+$\absval'= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15)
 \not\sqsubseteq (x+z \geq 10)$, and $\absval''= (w \geq 1\wedge w \leq

### 2016-10-23 paper.tex
proof-reading all section

@@
 properties of C programs.  We~evaluate the performance of our analyser in
-comparison with program verifiers that use propositional-level solvers and
+comparison with program verifiers that use propositional solvers and
 a static analyser based on abstract interpretation.

### 2016-10-23 learning.tex
section 6

@@
 
-$\aunit$ strictly generalizes the unit rule in SAT solvers.  
-However, rule (3) only applies to interval and octagons domains.  
-For example, let $\conflictset = \{ x+z \geq 10, w \geq 0 \}$ be a conflicting element 
-that does not satisfy $\formula$ and let $\absval= (x+y
-\geq 2\wedge y+z \geq 5\wedge x+z \leq 15\wedge w \geq 1\wedge w \leq
-1)$.  Applying rule (3) of $\aunit_\conflictset$, we get a
-decomposition of $\absval$ into $\absval'$ and $\absval''$ where
-$\absval'= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15)
-\not\sqsubseteq (x+z \geq 10)$, and $\absval''= (w \geq 1\wedge w \leq
-1) \sqsubseteq (w \geq 0)$.  Then, $\aunit_C(a) = (x+z < 10)$, which
-constrains the bound of the octagonal constraint.
-
-Let us consider another example, where $\conflictset = \{x \geq 2, x
+%$\aunit$ strictly generalizes the unit rule in SAT solvers.  
+Let us consider an example, where $\conflictset = \{x \geq 2, x
 \leq 5, y \leq 7 \}$ and let $\absval = (x \geq 3\wedge x \leq 4\wedge
@@
 $\absval' \sqsubseteq (x \geq 2)$ and $\absval' \sqsubseteq (x \leq
-5)$.  However, if none of the above rules hold true, $\aunit$ returns~$\top$.
+5)$.
+
+However, rule (3) only applies to interval and octagons domains.  
+For example, let $\conflictset = \{ x+z \geq 10, w \geq 0 \}$ 
+be a conflicting element that does not satisfy $\formula$ 
+and let $\absval= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15\wedge w \geq 1\wedge w \leq
+1)$.  Applying rule (3) of $\aunit_\conflictset$, we get a
+decomposition of $\absval$ into $\absval'$ and $\absval''$ where
+$\absval'= (x+y \geq 2\wedge y+z \geq 5\wedge x+z \leq 15)
+\not\sqsubseteq (x+z \geq 10)$, and $\absval''= (w \geq 1\wedge w \leq
+1) \sqsubseteq (w \geq 0)$.  Then, $\aunit_C(a) = (x+z < 10)$, which
+constrains the bound of the octagonal inequalities in $absval$, thereby 
+refining $\absval = (x+y \geq 2\wedge y+z \geq 5 \wedge x+z \leq 9 
+\wedge w \geq 1\wedge w \leq)$.  If none of the above rules hold true, 
+$\aunit$ returns~$\top$.
 
 \paragraph {\textbf{Backjumping}}
-After adding $\aunit$ to $\abstransset$, we remove all the meet irreducibles from the trail back to a decision level that restores the solver to a
+%After adding $\aunit$ to $\abstransset$, we 
+A backjumping procedure remove all the meet irreducibles from 
+the trail up to a decision level that restores the analysis to a
 non-conflicting state.  The backjumping level is defined by the
-meet irreducible of the conflict clause that is closest 
+meet irreducibles of the conflict clause that is closest 
 to the root (decision level~0) where the conflict
@@
 backjumping level is the root of the search tree and
-$\analyzeconflict$ returns $\false$, otherwise $\true$.
+$\analyzeconflict$ returns $\false$, otherwise returns $\true$.
 

### 2016-10-23 domains.tex
domain text

@@
 gives an example of an octagon and its associated inequalities.
-
+\Omit{
 The interval domain is a non-relational domain because a single
-inequality only contains a single variable.
+inequality only contains a single variable.  The octagon domain, 
+however, is relational.  
+}
 %
-The octagon domain, however, is relational.  A fundamental difference
-between relational and non-relational domains is that relational
-domains require the computation of a \emph{closure} in order to obtain
-a normal form that is usually required for precise and efficient 
+A fundamental difference between relational and 
+non-relational domains is that the relational
+domains require the computation of a \emph{closure} 
+in order to obtain a normal form that is usually required for precise 
+%and efficient 
 domain operations.  An example of a closure computation for octagonal 
-inequalities is given below.
-%
-\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
+inequalities is given by
+$\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z
+\leq 5) \wedge (x-z \leq 9))$.
+%below.
+%\[\mathit{closure}((x-y \leq 4) \wedge (y-z \leq 5))=((x-y \leq 4) \wedge (y-z \leq 5) \wedge (x-z \leq 9)) \]  
 %

### 2016-10-23 domains.tex
section 3 updates

@@
 %
-In the interval domain $\intervals[\numvars]$,
-for a program with $N=|\numvars|$ variables, the matrix $\mat{C}$ 
-has $2N$ rows. Hence, it generates at most $2N$ inequalities, one
+%In the interval domain $\intervals[\numvars]$,
+For a program with $N=|\numvars|$ variables, 
+the matrix $\mat{C}$ for the interval domain 
+$\intervals[\numvars]$, has $2N$ rows. Hence, it 
+generates at most $2N$ inequalities, one
 for the upper and lower bounds of each variable.
@@
 The closure computes all implied domain constraints.  For octagons,
-the closure operation is very expensive (cubic complexity in the
-number of program variables).
+closure is the most critical and expensive operator (cubic complexity in the
+number of program variables)~\cite{pldi15}.
 %

### 2016-10-23 introduction.tex
adjusted within 15 pages

@@
 %
-Static program analysis based on abstract interpretation~\cite{CC77} has
-been widely used to verify properties of safety-critical systems~\cite{DBLP:conf/pldi/BlanchetCCFMMMR03}.
+Static program analysis based on abstract interpretation~\cite{CC77, 
+DBLP:conf/pldi/BlanchetCCFMMMR03} has been widely used to 
+verify properties of safety-critical systems.
 Standard static analyses~\cite{se2011} aim to compute program invariants as
@@
 alarms.  Standard means to address false alarms in static analysis are 
-the use of richer abstract domains or equipping the analysis with more precise set
+richer abstract domains or equipping the analysis with more precise set
 unions -- which is often expensive since the analysis exhibits case

### 2016-10-23 paper.tex
adjusted within 15 pages

@@
 properties of C programs.  We~evaluate the performance of our analyser in
-comparison with program verifiers that use propositional solvers and
-a static analyser based on abstract interpretation.
+comparison with a program verifier that use propositional solvers and
+an abstract interpretation tool.
+%static analyser based on abstract interpretation.
 %
@@
 %===============================================================================
-
+\newpage
 \bibliographystyle{splncs03}

### 2016-10-23 conclusion.tex
conclusion

@@
 In this paper, we present a procedure for lifting the model search 
-in satisfiability solvers to relational abstract domains.  
+and conflict analysis procedures in satisfiability solvers for 
+relational abstract domains.  
 %We explain specific properties of abstract domain elements that are necessary to infer domain-specific deductions.  
 We present a parameterised abstract transformer that allows us to
-flexibly implement various propagation heuristics while taking care of
-the closure operation required in relational abstract domains.
+%flexibly implement various propagation heuristics while taking care of
+flexibly guide the deductions in the abstract domain while performing lazy 
+closure operations in the relational domains for efficiency reasons.
 %

### 2016-10-23 paper.tex
Abstract.

@@
 \emph{Abstract Conflict-Driven Clause Learning} (ACDCL) is a generalisation
-of the well-known CDCL algorithm that is used in SAT solvers.
+of the well-known CDCL algorithm used in SAT solvers.
 %
@@
 %
-ACDCL has been successfully used with non-relational abstract domains such
-as intervals to improve the performance of a floating-point decision
-procedure.  Whereas the propositional reasoning in SAT 
+It has been successfully used with non-relational abstract domains, such
+as intervals, to improve the performance of a floating-point decision
+procedure.  Whereas the propositional reasoning of SAT 
 solvers can be lifted easily from the Boolean lattice to other 
@@
 %
-In this paper, we present the non-trivial liftings of the model search and
-conflict analysis algorithms in SAT solver for relational abstract domains.  We have
+In this paper, we present non-trivial liftings of the model search and
+conflict analysis algorithms of SAT solvers to relational abstract domains.  We have
 implemented these algorithms inside the 2LS verification tool and
-instantiated our analyser over a template polyhedra domain to check safety
-properties of C programs.  We~evaluate the performance of our analyser in
-comparison with a program verifier that use propositional solvers and
+instantiated our analyser with a template polyhedra domain to check safety
+properties of C programs.  We~evaluate the performance of our analyser 
+compared to a program verifier that uses propositional solvers and
 an abstract interpretation tool.

### 2016-10-23 introduction.tex
Tweak introduction.

@@
 over-approximates set union (or disjunction).  The over-approximation in the
-join operation is a key cause of loss of precision, which yields false
+join operation loses precision, which can yield false
 alarms.  Standard means to address false alarms in static analysis are 
@@
 On the other hand, propositional SAT solvers are known to be highly
-effective for solving complex propositional formulae, and are frequently
+effective for solving complex propositional formulae, and are often
 able to avoid enumerating the exponential number of cases.  The impressive
@@
 algorithms are referred to as \emph{Conflict Driven Clause Learning}
-(CDCL)~\cite{cdcl}. The idea to lift these algorithms from the Boolean
-domain to richer, arithmetic domains suggests itself.
+(CDCL)~\cite{cdcl}. An obvious goal is to lift these algorithms from the Boolean
+domain to richer, arithmetic domains.
 
@@
 Propagation (BCP) computes a greatest fixed-point over the abstract domain,
-decision is dual widening and clause learning can be viewed as synthesising
+decision is dual widening, and clause learning can be viewed as synthesising
 an abstract transformer for negation.  This enables an ACDCL-based analyser
 to refine an imprecise analysis and thus prevents enumeration behaviour. 
-ACDCL has so far been instantiated as a decision procedure for
+To date, ACDCL has been instantiated as a decision procedure for
 floating-point arithmetic~\cite{DBLP:journals/fmsd/BrainDGHK14} and a
-program analyser that reasons using intervals~\cite{tacas12}.  Hence, the
+program analyser that reasons using intervals~\cite{tacas12}.  So the
 existing ACDCL instances are limited to bounds checking in numerical
@@
 Evidently, the verification of many programs requires more expressive
-domains like octagons or equalities or polyhedra.  These domains capture
+domains, such as octagons, equalities, or polyhedra.  These domains capture
 interesting properties of programs that intervals cannot express.  In this

### 2016-10-23 contributions.tex
proof-reading all section
Sketch

@@
 %generalises the Boolean Constraint Propagation (BCP) procedure 
-to compute a fixed point over a relational domain.  We propose 
+that computes a fixed-point over a relational domain.  We propose 
 domain-specific decision heuristics that exploit the expressivity 
@@
 
-\item We present a lifting of the conflict analysis procedure 
+\item We sketch a lifting of the conflict analysis procedure 
 to relational abstract domains in order to learn a generalised reason 

### 2016-10-23 learning.tex
Section 6 intro paragraph.

@@
 \section{Abstract Conflict Analysis}\label{sec:conflict}
+
 Conflict analysis in a SAT solver can be seen as abductive 
@@
 }
-Since relational abstract domains have more complex structure than the partial 
+Relational abstract domains have more complex structure than the partial 
 assignments domain or interval domain, so finding a generalized reason for 
-conflict through UIP computation~\cite{uip} is non-trivial.  We formally 
-describe the adaptations of conflict analysis procedure for relational domains. %
+conflict through UIP computation~\cite{uip} is non-trivial.  To make this paper self-contained, 
+we breifly explain the adaptations that have to be made to the conflict analysis procedure for relational domains. %
 \begin{figure}[t]

### 2016-10-23 graphs.tex
a few bad tricks to make it fit
figure 6 title

@@
 \caption{\label{fig:results}
-Comparison of the number of decisions and propagations performed by
-propositional SAT and ACDCL}
+Comparison between SAT and ACDCL: number of decisions and propagations}
 \end{figure}
@@
 \caption{\label{prop-dec}
-Effect of Propagation Strategies and Decision Heuristics 
+Effect of Propagation Heuristics and Decision Heuristics in ACDCL
 }

### 2016-10-23 learning.tex
Typo.
proof read learning

@@
 conflict through UIP computation~\cite{uip} is non-trivial.  To make this paper self-contained, 
-we breifly explain the adaptations that have to be made to the conflict analysis procedure for relational domains. %
+we briefly explain the adaptations that have to be made to the conflict analysis procedure for relational domains. %
 \begin{figure}[t]
@@
 \paragraph {\textbf{Abstract Conflict Graph}}
-The first-UIP algorithm in SAT solvers works on a data structure
+The first-UIP~\cite{uip,cdcl} algorithm in SAT solvers works on a data structure
 called {\em implication graph} and computes a cut that suffices to
@@
 graph that stores the deductions obtained from abstract model search
-using the octagon domain.  The last deduction conflicts with the
-meet irreducible $(y \leq 0)$.
+using the octagon domain.  The last deduction conflicts with the constraint 
+$(y \leq 0)$.
 
@@
 using the rule (1), since $\absval \sqsubseteq \bigsqcap\conflictset$.  Now,
-let $\absval' = (x \geq 3\wedge x \leq 4)$, then
-$\aunit_\conflictset(\absval') = (y \geq 8)$ using rule (2), since
-$\absval' \sqsubseteq (x \geq 2)$ and $\absval' \sqsubseteq (x \leq
-5)$.
+consider another abstract value $\absval = (x \geq 3\wedge x \leq 4)$, then
+$\aunit_\conflictset(\absval) = (y \geq 8)$ using rule (2), since
+$\absval \sqsubseteq (x \geq 2)$ and $\absval \sqsubseteq (x \leq 5)$.
 
-However, rule (3) only applies to interval and octagons domains.  
+However, rule (3) of $\aunit$ is specific to interval and octagon domains 
+and is not applicable in the propositional case.  
 For example, let $\conflictset = \{ x+z \geq 10, w \geq 0 \}$ 
@@
 1) \sqsubseteq (w \geq 0)$.  Then, $\aunit_C(a) = (x+z < 10)$, which
-constrains the bound of the octagonal inequalities in $absval$, thereby 
+constrains the bound of the octagonal inequalities in $\absval$, thereby 
 refining $\absval = (x+y \geq 2\wedge y+z \geq 5 \wedge x+z \leq 9 
@@
 backjumping level is the root of the search tree and
-$\analyzeconflict$ returns $\false$, otherwise returns $\true$.
+$\analyzeconflict$ returns $\false$, otherwise it returns $\true$.
 

### 2016-10-23 experiments.tex
experiment and background proof read

@@
 CBMC (version 5.5)~\cite{cbmc} and a commercial static analysis tool, 
-Astr{\'e}e (version 14.10). 
+Astr{\'e}e (version 14.10)~\cite{astree}. 
 CBMC uses MiniSAT~2.2.1 in the backend.  Astr{\'e}e uses a 
@@
 \hline
-Solver & Decisions & propagations & conflicts & conflict literals & restarts \\ \hline
+Solver & decisions & propagations & conflicts & conflict literals & restarts \\ \hline
 MiniSat & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
@@
 branches over meet irreducibles that involve numerical variables.  The
-runtimes for the random heuristic are marginally higher than for the other
+runtimes for the random heuristic are marginally higher than the other
 two.  This suggests that domain-specific decision heuristics are important

### 2016-10-23 experiments.tex
nudge

@@
 is not competitive with that of the APRON~\cite{apron} library; however, the
-domain implementation in 2LS handles all C operators (including bitwise
+domain implementation in 2LS handles all C operators (including bit-wise
 operations) and supports precise complementation of meet irreducibles, which

### 2016-10-23 domains.tex
adjusted within 15 pages
line break

@@
 We~denote $\vars$ the set of variables occurring in $\constraints$.  
+Based on the above program representation, we define a safety equation as, 
+$\formula:= \bigwedge_{\constraint\in\constraints} \constraint$. The formula $\formula$ is 
+unsatisfiable if and only if the program is safe.
 %
@@
 %
-Based on the above program representation, we define a safety equation as, 
-$\formula:= \bigwedge_{\constraint\in\constraints} \constraint$. The formula $\formula$ is 
-unsatisfiable if and only if the program is safe.
 
@@
 %
-A fundamental difference between relational and 
+A~fundamental difference between relational and 
 non-relational domains is that the relational
@@
 \end{equation}
-Meet irreducibles in ACDCL correspond to the concept of literals in a SAT solver.
 \end{definition}
 %
+%Meet irreducibles in ACDCL correspond to the concept of literals in a SAT solver.
 The meet irreducibles in the Boolean domain $\booldomain$ for a variable $x$ are 

### 2016-10-23 decision.tex
decision proof read

@@
 \emph{singleton} meet irreducible.  This is similar to 
-a variable in a SAT solver that is already assigned $\true$ or $\false$,
-and thus we cannot make a decision.
+a literal in a SAT solver that is already assigned $\true$ or $\false$,
+and thus cannot participate in a decision.
 %
@@
 {\em longest-range}, {\em random} and the 
-{\em Berkmin}~\cite{eugoldberg07} decision heuristics.  
-The {\em ordered} decision heuristics creates an ordering among non-singleton 
+{\em Berkmin}~\cite{eugoldberg07} decision heuristic.  
+The {\em ordered} decision heuristic creates an ordering among non-singleton 
 meet irreducibles, thereby making decisions on meet irreducibles that involve 
@@
 before choosing meet irreducibles with numerical variables.  
-The ordered heuristic is similar to trace partitioning~\cite{toplas07}.
+%The ordered heuristic gives an effect of trace partitioning~\cite{toplas07}.
 %
-The {\em longest-range} heuristics simply keeps track of the bounds
+The {\em longest-range} heuristic simply keeps track of the bounds
 $\numabsval_l,\numabsval_u$ of matching template rows $\numabsval_l\leq

### 2016-10-23 deductions.tex
section 5 proof-read

@@
 The key considerations for building an abstract transformer are 
-precision and efficiency.  An precise transformer is usually less efficient, 
+precision and efficiency.  A precise transformer is usually less efficient, 
 and vice-versa.  We provide a parameterised abstract transformer that allows us to 
-control precision and efficiency through propagation heuristics. 
+control the precision and efficiency of deductions through the application of 
+different propagation heuristics. 
 %
@@
 in Section~\ref{sec:domains}.  We denote $\abstrans{\domain}{\constraint}^\subdomain$
-as $\abstransel{\subdomain}$.
+as $\abstransel{\subdomain}$ in Algorithm~\ref{Alg:ms}.
 
@@
 %
-For example, for $\absval=(0\leq y \leq 1 \wedge 5\leq z)$, we have
+Consider an example, where $\absval=(0\leq y \leq 1 \wedge 5\leq z)$, we have
 $\abstrans{\intervals[\{x,y,z\}]}{x=y+z}^{\intervals[\{x\}]}(\absval)=\absval\meet(x\geq
@@
 \paragraph {\textbf{Lazy closure computation}} 
-Computing closure for relational domains, such as octagons is extremely
-expensive~\cite{pldi15}.  An advantage of our formalism in 
-Eq.~(\ref{eq:at2}) is that
+Computing closure for relational domains, such as octagons is an 
+extremely expensive operation~\cite{pldi15}.  An advantage of our 
+formalism in Eq.~(\ref{eq:at2}) is that
 %we can compute the
@@
 deduction step more efficient.  Thus, we delay the closure operation 
-until the point where it is absolutely necessary. 
+until the point where it is absolutely necessary.  This makes our deductions 
+in relational domain more efficient.  
 %

### 2016-10-23 deductions.tex
section 5 check

@@
 abstract deduction transformers that contain constants in the
-right-hand side and the subdomain is constructed via $\makesubdomain$
-from variables in left-hand side of the equality constraints
+right-hand side and the subdomain is constructed via $\makesubdomain_\domain$
+from variables in the left-hand side of the equality constraints
 originating from the assignment statements in the program.
@@
 Intuitively, this means that all concrete values in 
-$\gamma(\absval)$ satisfy $\formula$.  
+$\gamma(\abs(\trail))$ satisfy $\formula$.  
 %

### 2016-10-23 algorithm.tex
proof-reading all section
section 4 proof-read

@@
 The input to the ACDCL algorithm (Algorithm~\ref{Alg:acdcl}) is a
-program in the form of the set of its abstract transformers
+program in the form of a set of abstract transformers
 $\abstransset=\{\abstrans{\domain}{\sigma}|\sigma\in\Sigma\}$
@@
 $\abs(\trail)=\bigsqcap_{m \in \trail}m$ with
-$\abs(\trail)=\top$ if $\trail$ is the empty sequence $\langle\rangle$.
+$\abs(\trail)=\top$ if $\trail$ is an empty sequence $\langle\rangle$.
 \end{definition}
@@
 The algorithm begins with an empty $\trail$, a empty $\reasons$ and an
-abstract value $\top$.  The procedure $\deduce$ (detailed in
-Section~\ref{sec:deduce}) computes a greatest fixed point over the
+abstract value $\top$.  The procedure $\deduce$ (details in
+Section~\ref{sec:deduce}) computes a greatest fixed-point over the
 transformers in $\abstransset$ that refines the abstract value,
@@
 The procedure $\deduce$ is called next to infer new meet irreducibles
-taking into account the current decision.  The model search phase
+based on the current decision.  The model search phase
 alternates between the decision and deduction until $\deduce$ returns
@@
 ACDCL heuristically chooses one reason $\conflictset$ and learns it 
-by adding it as an abstract transformer to $\abstransset$. The algorithm 
+by adding it as an abstract transformer to $\abstransset$. The analysis 
 backtracks by removing the content of $\trail$ up to a point where it does not 
@@
 transformer.  If $\analyzeconflict$ returns $\false$, then no further
-backtracking is possible.  Thus, the $\formula$ is unsatisfiable
+backtracking is possible.  Thus, the safety equation $\formula$ is unsatisfiable
 and ACDCL return \textsf{safe}.

### 2016-10-23 paper.tex
section 4 proof-read

@@
 We observe a reduction of two orders of magnitude in the number of 
-decisions and conflict analyses compared to a SAT 
+decisions, propagations, and conflicts compared to a SAT 
 solver and higher precision compared to an abstract interpreter. 

### 2016-10-23 example.tex
Tweak example.
figure 1

@@
 
-Let us consider the simple program in Figure~\ref{fig:example}, which
+Consider the simple program in Figure~\ref{fig:example}, which
 squares a machine integer and checks that the result is always positive.  To
@@
 \hline
-Solver & Decisions & propagations & conflicts & conflict literals & restarts \\ \hline
+Solver & decisions & propagations & conflicts & conflict literals & restarts \\ \hline
 MiniSat & 233 & 36436 & 162 & 2604 & 2 \\ \hline

### 2016-10-24 domains.tex
final proof-read

@@
 %
-In this paper, we consider \emph{bounded programs} with safety
+We consider \emph{bounded programs} with safety
 properties given as assertions $\assertions$ in the program.
@@
 loops and recursions a finite number of times. The result of this
-transformation is represented as set
+transformation is represented by a set
 $\constraints=\program\cup\{\neg \bigwedge_{\assertion\in\assertions} \assertion\}$
-where $\program$ are the statements of the transformed program in the
+where $\program$ contains the statements of the transformed program in the
 form of constraints as typically obtained by translating the program
@@
 %
-We instantiate $\reldomain$ with the template polyhedra domain.
+We instantiate $\reldomain$ with the template polyhedra domain~\cite{sriram}.
 
 For notational convenience we will denote elements of
-$\booldomain^{|\boolvars|}$ by their concretisations to propositional formulae,
+$\booldomain^{|\boolvars|}$ by their concretisations to propositional formulae.  
 For example, the program in Figure~\ref{ssa} has four Boolean variables
-written as the vector $(g_0,g_1,c,g_2)$; then we will denote an abstract value $(\top, \false, \true, \top) \in \booldomain^4$ as $\neg g_1 \wedge c$.
+written as the vector $(g_0,g_1,c,g_2)$. Thus, we will denote an abstract 
+value $(\top, \false, \true, \top) \in \booldomain^4$ as $\neg g_1 \wedge c$.
 
@@
 techniques~\cite{sriram,GS07b,BJKS15} for computing the domain
-operations, such as meet ($\meet$), join ($\sqcup$), etc. In our implementation we
-use the strategy iteration approach available in 2LS~\cite{BJKS15}.
+operations, such as meet ($\meet$), join ($\sqcup$), etc in the template
+polyhedra domain. In our implementation, we use the strategy 
+iteration approach available in 2LS~\cite{BJKS15}.
 %
@@
 non-singleton element of the partial assignment domain can be 
-decomposed into a set of precisely complementable singleton elements.  
-This property of the partial assignment domain is required to learn 
+decomposed into a set of precisely complementable singleton
+elements~\cite{dhk2013-popl}.  This property of the partial 
+assignment domain is required to learn 
 elements that help to guide the model search away from the

### 2016-10-24 graphs.tex
ordered Fig6b data by 'random' runtimes, no lines, more readable marks

@@
 \begin{tikzpicture}[scale=0.75]
+\pgfplotscreateplotcyclelist{markstyles}{%
+mark=diamond\\%
+mark=square*,red\\%
+mark=triangle*,blue\\%
+}
 	\begin{axis} [
   title={Performance of Decision Heuristics}, ymode = log,
-  xmin=0, xmax=70, 
-  ymin=0, ymax=1500, 
+  xmin=0, xmax=70, only marks,
+  ymin=0, ymax=1600, 
   xlabel={Benchmarks},
@@
   ytick={}, %{0, 0.5, 1.0, 10, 50, 100, 200, 500, 700, 1000, 1500},
-	legend pos=north west,
+	legend pos=north west, mark size=1pt, cycle list name=markstyles,
  ]
